---
title: "HBN Final Sample Selection"
author: "Audrey Luo"
output:
  html_document:
    code_folding: show
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(purrr)
library(stringr)
library(tidyr)
```

# Initial Participant List
```{r}
#list of RBC ids with processed, non-variant diffusion MRI data, T1, and fieldmaps: N = 2063
participants <- read.csv("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/HBN_NonVariantDWI_participantlist.csv") # generated by /cbica/projects/luo_wm_dev/code/Generate_input/4_QC_scripts/HBN_identify_variants.ipynb
colnames(participants) <- c("sub") 

# exclude participants not in pod2 data
# tsv from https://fcp-indi.s3.amazonaws.com/index.html#data/Projects/HBN/BIDS_curated/derivatives/qsiprep/)
# exclude participants that had didn't have preprocessed, BIDS-compliant dMRI data (based on HBN-POD2 curation)
pod2 <- read.delim("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/HBN-POD2.tsv")
pod2 <- pod2 %>% rename(sub=subject_id)
participants.final <- inner_join(participants, pod2[,c('sub', 'expert_qc_score', 'xgb_qsiprep_qc_score', 'xgb_qc_score', 'dl_qc_score', 'site_variant')], by = "sub")  # remove 24 participants from sample of 2063 due to not being BIDS-compliant dMRI data, remaining N = 2039 

# exclude if had ACBD sequence
pod2exclude_abcd <- participants.final %>% filter(str_detect(site_variant, "ABCD"))
participants.final <- participants.final[!participants.final$sub %in% pod2exclude_abcd$sub,] %>% as.data.frame() # remove 15 participants from sample of 2039 due to ABCD sequence, remaining N = 2024

# exclude staten island (1.5 T)
dwi.qc <- read.csv("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/HBN_DWI_QCmetrics.csv") #diffusion scan acquisition parameters and quality metrics, generated by /cbica/projects/luo_wm_dev/code/Generate_input/4_QC_scripts/diffusion_qcmetrics.ipynb
dwi.qc <- dwi.qc %>% rename(sub=subject_id) %>% rename(site=session_id) # format site variable
dwi.qc$site <- gsub("ses-", "", dwi.qc$site)
staten_island <- dwi.qc$sub[which(dwi.qc$site=="HBNsiteSI")]
participants.final <- participants.final[!participants.final$sub %in% staten_island,] %>% as.data.frame() # remove 269 participants from sample of 2024 for having 1.5T data, remaining N = 1755

# Actual starting N = 1755 (3T non-variant diffusion MRI data, T1, and fieldmaps)
```

# Construct Final Sample
```{r}
demographics <- read.delim("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/HBN_Demographics.tsv") # copied over from network_replication
demographics <- demographics %>% rename(sub=participant_id)
demographics <- demographics[demographics$sub %in% participants.final$sub,] #demographic, medical, and clinical information for just the N = 1755 participants with non-variant diffusion MRI data
```

## Health History Exclusion
```{r}
#no additional medical exclusion in HBN
```

## T1 Quality Exclusion
```{r}
# participants to exclude based on manual QA: exclude T1 based on `tri.nary_score_without_rater.2`==0
# note that trinary score == 2 means good, == 0 means bad, == 1 means in between
T1_QA <- read.csv("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/T1_QA_HBN_NathaliaEsper_20230630.csv")
T1_QA <- T1_QA %>% rename(sub=ID)
T1_QA <- T1_QA %>% mutate(t1Exclude = ifelse(tri.nary_score_without_rater.2 == 0, 1, 0))
T1_qualityexclude <- T1_QA %>% filter(t1Exclude==1) %>% select(sub)

participants.final <- participants.final[!participants.final$sub %in% T1_qualityexclude$sub,] %>% as.data.frame() #remove 72 participants from the sample of 1755, remaining N = 1683
```

## Diffusion Acquisition Exclusion
```{r}
dwi.qc <- read.csv("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/HBN_DWI_QCmetrics.csv") #diffusion scan acquisition parameters and quality metrics, generated by /cbica/projects/luo_wm_dev/code/Generate_input/4_QC_scripts/diffusion_qcmetrics.ipynb
dwi.qc <- dwi.qc %>% rename(sub=subject_id) %>% rename(site=session_id) # format site variable
dwi.qc$site <- gsub("ses-", "", dwi.qc$site)
demographics <- merge(demographics, dwi.qc[,c("sub", "site")], by='sub') # add site info to demographics
dwi.qc <- dwi.qc[dwi.qc$sub %in% participants.final$sub,] # QC info for just the N = 1683 participants  
```

```{r}
acquisitionexclude <- dwi.qc %>% filter(raw_num_directions != 129) %>% select(sub) #participants to exclude from the final study sample for missing a diffusion MRI run (none)
participants.final <- participants.final[!participants.final$sub %in% acquisitionexclude$sub,] %>% as.data.frame() #remove 0 participants from the sample, remaining N = 1683
```

## Diffusion Quality Exclusion
```{r}
qualityexclude <- dwi.qc %>% filter(t1_neighbor_corr < 0.7) %>% select(sub) #participants to exclude from the final study sample based on processed diffusion MRI neighborhood correlation
participants.final <- participants.final[!participants.final$sub %in% qualityexclude$sub,] %>% as.data.frame() #remove 52 participants from the sample of 1683 due to diffusion data quality, remaining N = 1631
```

```{r}
# exclude based on expert rater score and CNN-i +q score from HBN-POD2 (CNN-i +q score is the definitive QC score, according to Richie-Halford paper)
pod2exclude <- participants.final %>% filter(expert_qc_score < 0.6 | dl_qc_score < 0.6) 
participants.final <- participants.final[!participants.final$sub %in% pod2exclude$sub,] %>% as.data.frame() # remove 489 participants from sample of 1631 due to failing expert rater, remaining N = 1142
```

## Diffusion Scan Head Motion Exclusion
```{r}
motionexclude <- dwi.qc %>% filter(mean_fd > 1) %>% select(sub) #participants to exclude from the final study sample based on high in-scanner head motion during the diffusion runs. (Note that 8 of these participants were already excluded in previous steps)
participants.final <- participants.final[!participants.final$sub %in% motionexclude$sub,] %>% as.data.frame()  #remove additional 40 participants from the sample of 1142 due to in-scanner head motion, remaining N = 1102
```

## Age Range Exclusion 
```{r}
participants.final <- participants.final[!participants.final$sub %in% demographics$sub[which(is.na(demographics$age))],] %>% as.data.frame()  #remove additional 1 participant from the sample of 1102 due to missing age data, remaining N = 1101
```

## Save Temporary Sample for PyAFQ: Subject List and Demographics
```{r}
# Make temp sample demographics dataframe
demographics.temp <- demographics[demographics$sub %in% participants.final$sub,] %>% select(sub, age, sex, race, site) # combine demographics data 
demographics.temp <- demographics.temp %>% mutate(sex = case_when(
  sex == "Male" ~ "M",
  sex == "Female" ~ "F"
))
# Make temp sample QC metrics dataframe 
dwi.qc.temp <- dwi.qc[dwi.qc$sub %in% participants.final$sub,] %>% select(sub, mean_fd, t1_neighbor_corr) 

sub_id <- participants.final$sub %>% as.data.frame() %>% setNames("sub_id")

write.table(sub_id, "/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/HBN_WMDev_TempSample_1101_age5to22.txt", col.names=T, row.names=F, quote = TRUE)
#write.table(sub_id, "/cbica/projects/luo_wm_dev/input/HBN/subject_list/HBN_tempsubject_list.txt", col.names=F, row.names=F, quote = FALSE)
write.table(sub_id[c(1:3),], "/cbica/projects/luo_wm_dev/input/HBN/subject_list/HBN_subject_list_test.txt", col.names=F, row.names=F, quote = FALSE)
write.csv(demographics.temp, "/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/HBN_WMDev_TempSampleDemographics_N1101_age5to22.csv", quote = F, row.names = F)
write.csv(dwi.qc.temp, "/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/HBN_WMDev_TempSampleQCMetrics_N1101_age5to22.csv", quote = F, row.names = F)
```

## Incomplete Data Processing Exclusion
total excluded for incomplete data processing = 188
```{r}
# load subjects missing Freesurfer data
missing_freesurfer <- read.table("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/missing_freesurfer.txt", header=T)
missing_freesurfer <- missing_freesurfer %>% select(sub_id) 
sub_id <- sub_id[!sub_id$sub_id %in% missing_freesurfer$sub_id,] %>% as.data.frame() %>% setNames("sub_id") #remove 13 participants from the sample of 1101 due to missing Freesurfer data, new N = 1088
write.table(sub_id, "/cbica/projects/luo_wm_dev/input/HBN/subject_list/HBN_subject_list_babs_updated.txt", col.names=T, row.names=F, quote = FALSE)
```

```{r}
# 44 subjects failed qsirecon
# this txt file was made with awk -F, '$8 == "True" { print $1 }' job_status.csv > ~/input/HBN/sample_selection_files/subs_failed_babs.txt
failed_babs <- read.table("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/subs_failed_babs.txt", header=F)
sub_id <- sub_id[!sub_id$sub_id %in% failed_babs$V1,] %>% as.data.frame() %>% setNames("sub_id") # remove 44 participants from sample of 1088 due to failing qsiprep/qsirecon, new N = 1044
```

After running ACT + pyAFQ, many subjects are missing one or more tracts. We can't harmonize a subject's data if they are missing ANY tracts. To keep exclusion criteria consistent across datasets, we are excluding people who are missing bundles that we need for this study.

Note that there are some bundles that are OKAY if they are missing because we are excluding some from analyses:
- cingulum cingulate
- uncinate
- anterior thalamic radiation
```{r}
# load the tract profiles of subjects who are missing at least 1 tract
missing_tracts <- fread("/cbica/projects/luo_wm_dev/input/HBN/derivatives/tract_profiles/all_subjects/missing_tract_profiles_nocovbat.tsv")
# load full tract profiles of everyone (except for the subjects in missing_tracts)
HBN_all <- fread("/cbica/projects/luo_wm_dev/input/HBN/derivatives/tract_profiles/all_subjects/collated_tract_profiles_nocovbat_tmp.tsv")
required_tracts <- unique(HBN_all$tractID)

##########################
# missing tract exclusion
##########################
# it's okay if a participant is missing these tracts
allowed_missing_tracts <- c("Cingulate", "Uncinate", "Anterior Thalamic")

# summarize tracts present for each subject
tract_summary <- missing_tracts %>%
  distinct(sub, tractID) %>%
  group_by(sub) %>%
  summarize(tracts_present = list(tractID), .groups = "drop")

# identify missing tracts for each subject
subjects_missing_tracts <- tract_summary %>%
  mutate(missing_tracts = map(tracts_present, ~ setdiff(required_tracts, .x))) %>%
  select(sub, missing_tracts)
 
# have one row per missing tract
missing_tracts_long <- subjects_missing_tracts %>%
  unnest(cols = c(missing_tracts))

missing_tracts_count <- missing_tracts_long %>%
  group_by(missing_tracts) %>%  # Group by missing tracts
  summarise(count = n()) 

# identify subjects to exclude
# A subject should be excluded if they are missing any REQUIRED tracts (aka tracts not in "allowed_missing_tracts")
# determine the subjects are missing needed tracts  
unacceptable_missing_tracts <- missing_tracts_long %>%
  filter(!str_detect(missing_tracts, paste(allowed_missing_tracts, collapse = "|")))

# get unique subjects who have any unacceptable missing tracts
subjects_to_exclude <- unacceptable_missing_tracts %>%
  distinct(sub)

# count the number of excluded subjects
num_excluded_subjects <- nrow(subjects_to_exclude)
cat("Number of subjects excluded:", num_excluded_subjects, "\n")

#subjects_to_exclude <- subjects_to_exclude %>% rename(sub_id=sub)
write.table(subjects_to_exclude, "/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/subs_missing_tracts.txt", col.names=T, row.names=F, quote = TRUE)

sub_id <- sub_id[!sub_id$sub_id %in% subjects_to_exclude$sub,] %>% as.data.frame() %>% setNames("sub_id") # remove 131 participants from sample of 1044 due to failing qsiprep/qsirecon, new N = 913

```

```{r}
# add back subjects that are missing non-important tracts. save HBN tract profiles
# Identify subjects missing only non-essential tracts (tracts in allowed_missing_tracts)
# = identify subjects who are not in the exclusion list
subjects_with_allowed_missing_tracts <- missing_tracts_long %>%
  filter(sub %notin% subjects_to_exclude$sub) %>%
  filter(str_detect(missing_tracts, paste(allowed_missing_tracts, collapse = "|")))

# get sub ids
subjects_with_allowed_missing_tracts <- subjects_with_allowed_missing_tracts %>%
  distinct(sub)

# add these subjects' data back to the full tract profiles data (should be 52 participants added back)
HBN_all_with_allowed <- bind_rows(HBN_all, missing_tracts %>% filter(sub %in% subjects_with_allowed_missing_tracts$sub))
HBN_all_with_allowed <- HBN_all_with_allowed %>%
  distinct()

# remove the tracts we don't need
HBN_all_with_allowed <- HBN_all_with_allowed %>% filter(!str_detect(tractID, paste(allowed_missing_tracts, collapse = "|")))

write.table(HBN_all_with_allowed, "/cbica/projects/luo_wm_dev/input/HBN/derivatives/tract_profiles/all_subjects/collated_tract_profiles_nocovbat.tsv", row.names = FALSE)
```

## Save Final Sample: Subject List, Demographics, QC measures
```{r}
# Make final sample demographics dataframe
demographics.final<- demographics.temp[demographics.temp$sub %in% sub_id$sub_id,] %>% select(sub, age, sex, race, site) # combine demographics data 
 
# Make final sample QC metrics dataframe 
dwi.qc.final <- dwi.qc.temp[dwi.qc.temp$sub %in% sub_id$sub_id,] %>% select(sub, mean_fd, t1_neighbor_corr) 

write.table(sub_id, "/cbica/projects/luo_wm_dev/input/HBN/subject_list/final_sample/HBN_WMDev_FinalSample.txt", col.names=T, row.names=F, quote = TRUE)
write.csv(merge(demographics.final, dwi.qc.final), "/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/final_sample/HBN_WMDev_FinalSampleDemoQC.csv", quote = F, row.names = F)
```

## Demographics table
```{r}
final_sample <- read.csv("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/final_sample/HBN_WMDev_FinalSampleDemoQC.csv", header=T)

# demographic variables to summarize
demo_vars <- c("sex", "race")

# function to summarize the demographic variables
summary_list <- map(demo_vars, function(var) {
  final_sample %>%
    group_by(.data[[var]]) %>%
    summarise(
      count = n(),
      percent = (n() / nrow(final_sample)) * 100
    )
})

# summaries for each variable
summary_list

range(final_sample$age)
mean(final_sample$age)
sd(final_sample$age)
```

### temp:
### Subsample with no psychiatric diagnosis / low pfactor
```{r}
final_sample <- read.csv("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/final_sample/HBN_WMDev_FinalSampleDemoQC.csv", header=T)
no_dx <- read.csv("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/HBN_phenotype_no_dx.csv")

no_dx <- no_dx %>% rename(sub = Identifiers, dx = Diagnosis_ClinicianConsensus.NoDX)
no_dx$sub <- paste0("sub-", gsub(",assessment", "",no_dx$sub))
no_dx <- no_dx %>% filter(dx == 1)

intersect(no_dx$sub, final_sample$sub)
length(which(no_dx$sub %in% final_sample$sub))
which(final_sample$sub %in% no_dx$sub)


low_p <- read.table("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/HBN_pfactor.tsv", sep = '\t', header = TRUE)
low_p <- low_p %>% rename(sub = participant_id)
low_p$sub <- paste0("sub-", low_p$sub)

low_p <- low_p[which(low_p$p_factor_mcelroy_harmonized_all_samples <= quantile(low_p$p_factor_mcelroy_harmonized_all_samples, na.rm=T)[2]),]

low_p_subs <- intersect(low_p$sub, final_sample$sub)

# Make final sample demographics dataframe
low_p_demographics.final <- final_sample[final_sample$sub %in% low_p_subs,]   

write.table(low_p_subs, "/cbica/projects/luo_wm_dev/input/HBN/subject_list/HBN_low_p_sample.txt", col.names=T, row.names=F, quote = TRUE)
write.csv(low_p_demographics.final, "/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/HBN_low_p_SampleDemoQC.csv", quote = F, row.names = F)

```

