---
title: "HCPD Final Sample Selection"
author: "Audrey Luo"
output:
  html_document:
    code_folding: show
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(purrr)
library(stringr)
library(tidyr)
```

# Initial Participant List
```{r}
#list of RBC ids with processed, non-variant diffusion MRI data, N = 639  
participants <- read.csv("/cbica/projects/luo_wm_dev/input/HCPD/sample_selection_files/HCPD_NonVariantDWI_participantlist.csv") # generated by /cbica/projects/luo_wm_dev/code/Generate_input/4_QC_scripts/HCPD_IdentifyVariants.ipynb

colnames(participants) <- c("rbcid") 
```

# Construct Final Sample
```{r}
demographics <- read.csv("/cbica/projects/luo_wm_dev/input/HCPD/sample_selection_files/hcpd_demographics.csv") # originally downloaded from /cbica/projects/RBC/Curation/RBC_demo_pheno/data/hcpd_demographics.csv but this file has since been moved. Also available on PennLINC slack.
demographics$rbcid <- gsub("HCD", "sub-", demographics$src_subject_id) 
demographics <- demographics[demographics$rbcid %in% participants$rbcid,] #demographic, medical, and clinical information for just the N = 639 participants with non-variant diffusion MRI data
```

## Health History Exclusion
```{r}
# medhis_2e = cancer/leukemia
# medhis_2p = sickle cell anemia
# ms = multiple sclerosis
# ph_9 = has had a seizure
# cfmh_chd_seizure = has been to the doctor for epilepsy
# cfmh_chd_cerpalsy = has been to the doctor for cerebral palsy
# medhis_6j = has been knocked unconscious

healthexclude <- demographics %>% filter(medhis_2e == 1 | medhis_2p == 1 | ms == 1 | ph_9 == 1 | cfmh_chd_seizure == 1 | cfmh_chd_cerpalsy == 1 | medhis_6j == 1) %>% select(rbcid) #participants to exclude from the final study sample for medical or health history

participants.final <- participants[!participants$rbcid %in% healthexclude$rbcid,] %>% as.data.frame() %>% setNames(c("rbcid")) #remove 7 health exclude participants from the initial sample of 639 individuals with non-variant diffusion MRI data, remaining N = 632
```

## Diffusion Acquisition Exclusion
```{r}
dwi.qc <- read.csv("/cbica/projects/luo_wm_dev/input/HCPD/sample_selection_files/HCPD_DWI_QCmetrics.csv") #diffusion scan acquisition parameters and quality metrics, generated by /cbica/projects/luo_wm_dev/code/Generate_input/4_QC_scripts/diffusion_qcmetrics.ipynb

dwi.qc <- dwi.qc[dwi.qc$subject_id %in% participants$rbcid,] # QC info for just the N=639 participants with non-variant diffusion data
```

```{r}
acquisitionexclude <- dwi.qc %>% filter(raw_num_directions != 398) %>% select(subject_id) #participants to exclude from the final study sample for missing a diffusion MRI run, resulting in 298 or 299 diffusion directions instead of 398

participants.final <- participants.final[!participants.final$rbcid %in% acquisitionexclude$subject_id,] %>% as.data.frame() %>% setNames(c("rbcid")) #remove 8 participants from the sample of 632 health include individuals due to missing gradient directions, remaining N = 624
```

## Diffusion Quality Exclusion
```{r}
qualityexclude <- dwi.qc %>% filter(t1_neighbor_corr < 0.6) %>% select(subject_id) #participants to exclude from the final study sample based on processed diffusion MRI neighborhood correlation

participants.final <- participants.final[!participants.final$rbcid %in% qualityexclude$subject_id,] %>% as.data.frame() %>% setNames(c("rbcid")) #remove 14 participants from the sample of 624 due to diffusion data quality, remaining N = 610
```

## Diffusion Scan Head Motion Exclusion
```{r}
motionexclude <- dwi.qc %>% filter(mean_fd > 1) %>% select(subject_id) #participants to exclude from the final study sample based on high in-scanner head motion during the diffusion runs. (Note that 6 of these participants were already excluded in previous steps)

participants.final <- participants.final[!participants.final$rbcid %in% motionexclude$subject_id,] %>% as.data.frame() %>% setNames(c("rbcid")) #remove 24 participants from the sample of 610 due to in-scanner head motion, remaining N = 586
```

## Age Range Exclusion 
```{r}
# Participants to exclude for being outside PNC age range
ageexclude <- demographics %>% filter(interview_age/12 < 8) %>% select(rbcid)

# Remove 14 participants from the sample of 586 due to young age, remaining N = 572
participants.final <- participants.final[!participants.final$rbcid %in% ageexclude$rbcid,] %>% as.data.frame() %>% setNames(c("sub"))
```

## Save Temporary Sample for PyAFQ: Subject List and Demographics
```{r}
# Make temp sample demographics dataframe
demographics.temp <- demographics[demographics$rbcid %in% participants.final$sub,] %>% select(rbcid, interview_age, sex, race, site) %>% mutate(interview_age = interview_age/12) %>% rename(sub = rbcid, age=interview_age) # combine demographics data 

# Make temp sample QC metrics dataframe 
dwi.qc.temp <- dwi.qc[dwi.qc$subject_id %in% participants.final$sub,] %>% select(subject_id, mean_fd, t1_neighbor_corr) %>% rename(sub = subject_id) # note that var(t1_neighbor_corr) = 0.001338567 and var(mean_fd) = 0.02042722
 
sub_id <- demographics.temp$sub %>% as.data.frame() %>% setNames("sub_id") # want the same ordering as demographics df

write.table(sub_id, "/cbica/projects/luo_wm_dev/input/HCPD/sample_selection_files/HCPD_WMDev_TempSample_N572_age8to22.txt", col.names=T, row.names=F, quote = TRUE)
#write.table(sub_id, "/cbica/projects/luo_wm_dev/input/HCPD/subject_list/HCPD_tempsubject_list.txt", col.names=F, row.names=F, quote = FALSE)
write.table(sub_id, "/cbica/projects/luo_wm_dev/input/HCPD/subject_list/HCPD_subject_list_babs.txt", col.names=T, row.names=F, quote = FALSE)
write.table(sub_id[c(1:3),], "/cbica/projects/luo_wm_dev/input/HCPD/subject_list/HCPD_subject_list_test.txt", col.names=F, row.names=F, quote = FALSE)
write.csv(demographics.temp, "/cbica/projects/luo_wm_dev/input/HCPD/sample_selection_files/HCPD_WMDev_TempSampleDemographics_N572_age8to22.csv", quote = F, row.names = F)
write.csv(dwi.qc.temp, "/cbica/projects/luo_wm_dev/input/HCPD/sample_selection_files/HCPD_WMDev_TempSampleQCMetrics_N572_age8to22.csv", quote = F, row.names = F)
```

## Incomplete Data Processing Exclusion
total excluded for incomplete data processing = 4
```{r}
# 3 subjects failed qsirecon
fail_qsirecon <- read.table("/cbica/projects/luo_wm_dev/input/HCPD/sample_selection_files/failed_qsirecon.txt", sep = ",")
sub_id <- sub_id[!sub_id$sub_id %in% fail_qsirecon$V1,] %>% as.data.frame() %>% setNames("sub_id") #remove 3 participants from the sample of 572 due to failure to complete qsirecon, final N = 569
```

After running ACT + pyAFQ, many subjects are missing one or more tracts. We can't harmonize a subject's data if they are missing ANY tracts. To keep exclusion criteria consistent across datasets, we are excluding people who are missing bundles that we need for this study.

Note that there are some bundles that are OKAY if they are missing because we are excluding some from analyses:
- cingulum cingulate
- uncinate
- anterior thalamic radiation
```{r}
# load the tract profiles of subjects who are missing at least 1 tract
missing_tracts <- fread("/cbica/projects/luo_wm_dev/input/HCPD/derivatives/tract_profiles/all_subjects/missing_tract_profiles_nocovbat.tsv")
# load full tract profiles of everyone (except for the subjects in missing_tracts)
HCPD_all <- fread("/cbica/projects/luo_wm_dev/input/HCPD/derivatives/tract_profiles/all_subjects/collated_tract_profiles_nocovbat_tmp.tsv")
required_tracts <- unique(HCPD_all$tractID)

##########################
# missing tract exclusion
##########################
# it's okay if a participant is missing these tracts
#allowed_missing_tracts <- c("Cingulate", "Uncinate", "Anterior Thalamic")
allowed_missing_tracts <- c("Cingulate", "Anterior Thalamic")

# summarize tracts present for each subject
tract_summary <- missing_tracts %>%
  distinct(sub, tractID) %>%
  group_by(sub) %>%
  summarize(tracts_present = list(tractID), .groups = "drop")

# identify missing tracts for each subject
subjects_missing_tracts <- tract_summary %>%
  mutate(missing_tracts = map(tracts_present, ~ setdiff(required_tracts, .x))) %>%
  select(sub, missing_tracts)
 
# have one row per missing tract
missing_tracts_long <- subjects_missing_tracts %>%
  unnest(cols = c(missing_tracts))

missing_tracts_count <- missing_tracts_long %>%
  group_by(missing_tracts) %>%  # Group by missing tracts
  summarise(count = n()) 

# identify subjects to exclude
# A subject should be excluded if they are missing any REQUIRED tracts (aka tracts not in "allowed_missing_tracts")
# determine the subjects are missing needed tracts  
unacceptable_missing_tracts <- missing_tracts_long %>%
  filter(!str_detect(missing_tracts, paste(allowed_missing_tracts, collapse = "|")))

# get unique subjects who have any unacceptable missing tracts
subjects_to_exclude <- unacceptable_missing_tracts %>%
  distinct(sub)

# count the number of excluded subjects
num_excluded_subjects <- nrow(subjects_to_exclude)
cat("Number of subjects excluded:", num_excluded_subjects, "\n")

write.table(subjects_to_exclude, "/cbica/projects/luo_wm_dev/input/HCPD/sample_selection_files/subs_missing_tracts.txt", col.names=T, row.names=F, quote = TRUE)

sub_id <- sub_id[!sub_id$sub_id %in% subjects_to_exclude$sub,] %>% as.data.frame() %>% setNames("sub_id") # remove 1 participant from sample of 569 due to failing qsiprep/qsirecon, new N = 568
```

```{r}
# add back subjects that are missing non-important tracts. save HCPD tract profiles
# Identify subjects missing only non-essential tracts (tracts in allowed_missing_tracts)
# = identify subjects who are not in the exclusion list
subjects_with_allowed_missing_tracts <- missing_tracts_long %>%
  filter(sub %notin% subjects_to_exclude$sub) %>%
  filter(str_detect(missing_tracts, paste(allowed_missing_tracts, collapse = "|")))

# get sub ids
subjects_with_allowed_missing_tracts <- subjects_with_allowed_missing_tracts %>%
  distinct(sub)

# add these subjects' data back to the full tract profiles data (should be 0 participants added back)
HCPD_all_with_allowed <- bind_rows(HCPD_all, missing_tracts %>% filter(sub %in% subjects_with_allowed_missing_tracts$sub))
HCPD_all_with_allowed <- HCPD_all_with_allowed %>%
  distinct()

# remove the tracts we don't need
HCPD_all_with_allowed <- HCPD_all_with_allowed %>% filter(!str_detect(tractID, paste(allowed_missing_tracts, collapse = "|")))

write.table(HCPD_all_with_allowed, "/cbica/projects/luo_wm_dev/input/HCPD/derivatives/tract_profiles/all_subjects/collated_tract_profiles_nocovbat.tsv", row.names = FALSE)
```

## Save Final Sample: Subject List, Demographics, QC measures
```{r}
# Make final sample demographics dataframe
demographics.final <- demographics.temp[demographics.temp$sub %in% sub_id$sub_id,] %>% select(sub, age, sex, race, site) # combine demographics data 
 
# Make final sample QC metrics dataframe 
dwi.qc.final <- dwi.qc.temp[dwi.qc.temp$sub %in% sub_id$sub_id,] %>% select(sub, mean_fd, t1_neighbor_corr) 

write.table(sub_id, "/cbica/projects/luo_wm_dev/input/HCPD/subject_list/final_sample/HCPD_WMDev_FinalSample.txt", col.names=T, row.names=F, quote = TRUE)
write.csv(merge(demographics.final, dwi.qc.final), "/cbica/projects/luo_wm_dev/input/HCPD/sample_selection_files/final_sample/HCPD_WMDev_FinalSampleDemoQC.csv", quote = F, row.names = F)
```

## Demographics table
```{r}
final_sample <- read.csv("/cbica/projects/luo_wm_dev/input/HCPD/sample_selection_files/final_sample/HCPD_WMDev_FinalSampleDemoQC.csv", header=T)

# demographic variables to summarize
demo_vars <- c("sex", "race")

# function to summarize the demographic variables
summary_list <- map(demo_vars, function(var) {
  final_sample %>%
    group_by(.data[[var]]) %>%
    summarise(count = n(),
              percent = (n() / nrow(final_sample)) * 100)})

# summaries for each variable
summary_list

range(final_sample$age)
mean(final_sample$age)
sd(final_sample$age)
```
