---
title: "PNC Final Sample Selection"
author: "Audrey Luo"
output:
  html_document:
    code_folding: show
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(purrr)
library(stringr)
library(tidyr)
```

# Initial Participant List
```{r}
#list of RBC ids with processed, non-variant diffusion MRI data, N = 1368 that has dMRI, T1, and fieldmaps
participants <- read.csv("/cbica/projects/luo_wm_dev/input/PNC/sample_selection_files/PNC_NonVariantDWI_participantlist.csv") # generated by /cbica/projects/luo_wm_dev/code/Generate_input/4_QC_scripts/PNC_identify_variants.ipynb

colnames(participants) <- c("sub") 
```

# Construct Final Sample
```{r}
demographics <- read.csv("/cbica/projects/luo_wm_dev/input/PNC/sample_selection_files/n1601_demographics_go1_20161212.csv") # copied over from network_replication
bblid_scanid <- read.csv("/cbica/projects/luo_wm_dev/input/PNC/sample_selection_files/bblid_scanid_sub.csv")
demographics <- merge(demographics, bblid_scanid, by ="scanid")
demographics <- demographics %>% relocate(bblid, rbcid)
demographics <- dplyr::rename(demographics, sub = rbcid) %>% mutate(age = ageAtScan1/12)
demographics$sub <- paste0("sub-", demographics$sub)
 
demographics <- demographics[demographics$sub %in% participants$sub,] #demographic, medical, and clinical information for just the N = 1368 participants with non-variant diffusion MRI data
```

## Health History Exclusion
```{r}
health_exclude_data <- read.csv("/cbica/projects/luo_wm_dev/input/PNC/sample_selection_files/n1601_health_20170421.csv")
healthexclude <- paste0("sub-", health_exclude_data$bblid[c(which(health_exclude_data$healthExcludev2==1))]) #fyi, bblid in health_exclude_data corresponds to sub in demographics 

participants.final <- demographics[-c(which(demographics$sub %in% healthexclude)),] #remove 118 health exclude participants from the initial sample of 1368 individuals with non-variant diffusion MRI data, remaining N = 1250
```

## T1 Quality Exclusion
```{r}
# participants to exclude based on manual QA
T1_QA <- read.csv("/cbica/projects/luo_wm_dev/input/PNC/sample_selection_files/n1601_t1QaData_20170306.csv")
T1_QA <- merge(T1_QA, bblid_scanid, by ="scanid")
T1_QA <- dplyr::select(T1_QA, -bblid.y)
T1_QA <- T1_QA %>% relocate(rbcid) 
T1_QA <- T1_QA %>% dplyr::rename(sub=rbcid)
T1_QA$sub <- paste0("sub-", T1_QA$sub)

T1_qualityexclude <- T1_QA %>% filter(t1Exclude==1) %>% select(sub)  

participants.final <- participants.final[!participants.final$sub %in% T1_qualityexclude$sub,] %>% as.data.frame()  #remove additional 25 participants from the sample of 1250 for failing T1 QA, remaining N = 1225
```

## Diffusion Acquisition Exclusion
```{r}
dwi.qc <- read.csv("/cbica/projects/luo_wm_dev/input/PNC/sample_selection_files/PNC_DWI_QCmetrics.csv") #diffusion scan acquisition parameters and quality metrics, generated by /cbica/projects/luo_wm_dev/code/Generate_input/4_QC_scripts/diffusion_qcmetrics.ipynb

dwi.qc <- dwi.qc[dwi.qc$subject_id %in% participants$sub,] # QC info for just the N=1368 participants with non-variant diffusion data
```

```{r}
acquisitionexclude <- dwi.qc %>% filter(raw_num_directions != 71) %>% select(subject_id) #participants to exclude from the final study sample for missing a diffusion MRI run, resulting in 35 diffusion directions instead of 71
 
participants.final <- participants.final[!participants.final$sub %in% acquisitionexclude$subject_id,] %>% as.data.frame() #remove 10 participants from the sample of 1225 health include individuals due to missing gradient directions, remaining N = 1215
```

## Diffusion Quality Exclusion
```{r}
#participants to exclude from the final study sample based on processed diffusion MRI neighborhood correlation
qualityexclude <- dwi.qc %>% filter(t1_neighbor_corr < 0.9) %>% select(subject_id) # threshold determined by thalamocortical paper

participants.final <- participants.final[!participants.final$sub %in% qualityexclude$subject_id,] %>% as.data.frame() #remove an additional 57 participants from the sample of 1215 due to diffusion data quality, remaining N = 1158
```

## Diffusion Scan Head Motion Exclusion
```{r}
motionexclude <- dwi.qc %>% filter(mean_fd > 1) %>% select(subject_id) #participants to exclude from the final study sample based on high in-scanner head motion during the diffusion runs.  
 
participants.final <- participants.final[!participants.final$sub %in% motionexclude$subject_id,] %>% as.data.frame()  #remove additional 13 participants from the sample of 1158 due to in-scanner head motion, remaining N = 1145
```

## Age Range Exclusion 
```{r}
# not doing age range exclusion for PNC 
```

## Save Temporary Sample for PyAFQ: Subject List and Demographics
```{r}
# Make temp sample demographics dataframe
demographics.temp <- demographics[demographics$sub %in% participants.final$sub,] %>% mutate(age = ageAtScan1/12) %>% select(sub, age, sex, race) # combine demographics data 

# recode these variables
#1=Male; 2= Female
#1= White; 2= Black Black/African American; 3= US_India/Alaska Native;4=Asian; 5=More Than One Race; 6=Hawaiian/Pacific; 9=Unknown/Unreported 
demographics.temp <- demographics.temp %>% mutate(race = case_when(
      race == "1" ~ "White",
      race == "2" ~ "Black",
      race == "3" ~ "US_India_Alaska",
      race == "4" ~ "Asian",
      race == "5" ~ "More_than_one_race" ,
      race == "6" ~ "Hawaiian_Pacific" ,
      race == "9" ~ "Unknown_Unreported"
    )) %>% mutate(sex = case_when(
      sex == "1" ~ "M", 
      sex == "2" ~ "F"
    ))
# Make temp sample QC metrics dataframe 
dwi.qc.temp <- dwi.qc[dwi.qc$subject_id %in% participants.final$sub,] %>% select(subject_id, mean_fd, t1_neighbor_corr) %>% rename(sub = subject_id) # note that var(dwi.qc.temp$t1_neighbor_corr) = 7.645376e-05 and var(dwi.qc.temp$mean_fd) = 0.02271341
 
sub_id <- demographics.temp$sub %>% as.data.frame() %>% setNames("sub_id") # want the same ordering as demographics df

#write.table(sub_id, "/cbica/projects/luo_wm_dev/input/PNC/subject_list/PNC_tempsubject_list.txt", col.names=F, row.names=F, quote = FALSE)
write.table(sub_id, "/cbica/projects/luo_wm_dev/input/PNC/subject_list/PNC_subject_list_babs.txt", col.names=T, row.names=F, quote = FALSE)
write.table(sub_id[c(1:3),], "/cbica/projects/luo_wm_dev/input/PNC/subject_list/PNC_subject_list_test.txt", col.names=F, row.names=F, quote = FALSE)
write.csv(demographics.temp, "/cbica/projects/luo_wm_dev/input/PNC/sample_selection_files/PNC_WMDev_TempSampleDemographics_N1145_age8to23.csv", quote = F, row.names = F)
write.csv(dwi.qc.temp, "/cbica/projects/luo_wm_dev/input/PNC/sample_selection_files/PNC_WMDev_TempSampleQCMetrics_N1145_age8to23.csv", quote = F, row.names = F)
```

## Incomplete Data Processing Exclusion
total excluded for incomplete data processing = 44
```{r}
# 43 subjects failed qsirecon
# this txt file was made with awk -F, '$8 == "True" { print $1 }' job_status.csv > ~/input/PNC/sample_selection_files/subs_failed_babs.txt
failed_babs <- read.table("/cbica/projects/luo_wm_dev/input/PNC/sample_selection_files/subs_failed_babs.txt", header=F)
sub_id <- sub_id[!sub_id$sub_id %in% failed_babs$V1,] %>% as.data.frame() %>% setNames("sub_id") # remove 43 participants from sample of 1145 due to failing qsiprep/qsirecon, new N = 1102
```

After running ACT + pyAFQ, many subjects are missing one or more tracts. We can't harmonize a subject's data if they are missing ANY tracts. To keep exclusion criteria consistent across datasets, we are excluding people who are missing bundles that we need for this study.

Note that there are some bundles that are OKAY if they are missing because we are excluding some from analyses:
- cingulum cingulate
- uncinate
- anterior thalamic radiation
```{r}
# load the tract profiles of subjects who are missing at least 1 tract
missing_tracts <- fread("/cbica/projects/luo_wm_dev/input/PNC/derivatives/tract_profiles/all_subjects/missing_tract_profiles_nocovbat.tsv")
# load full tract profiles of everyone (except for the subjects in missing_tracts)
PNC_all <- fread("/cbica/projects/luo_wm_dev/input/PNC/derivatives/tract_profiles/all_subjects/collated_tract_profiles_nocovbat_tmp.tsv")
required_tracts <- unique(PNC_all$tractID)

##########################
# missing tract exclusion
##########################
# it's okay if a participant is missing these tracts
#allowed_missing_tracts <- c("Cingulate", "Uncinate", "Anterior Thalamic")
allowed_missing_tracts <- c("Cingulate", "Anterior Thalamic")

# summarize tracts present for each subject
tract_summary <- missing_tracts %>%
  distinct(sub, tractID) %>%
  group_by(sub) %>%
  summarize(tracts_present = list(tractID), .groups = "drop")

# identify missing tracts for each subject
subjects_missing_tracts <- tract_summary %>%
  mutate(missing_tracts = map(tracts_present, ~ setdiff(required_tracts, .x))) %>%
  select(sub, missing_tracts)
 
# have one row per missing tract
missing_tracts_long <- subjects_missing_tracts %>%
  unnest(cols = c(missing_tracts))

missing_tracts_count <- missing_tracts_long %>%
  group_by(missing_tracts) %>%  # Group by missing tracts
  summarise(count = n()) 

# identify subjects to exclude
# A subject should be excluded if they are missing any REQUIRED tracts (aka tracts not in "allowed_missing_tracts")
# determine the subjects are missing needed tracts  
unacceptable_missing_tracts <- missing_tracts_long %>%
  filter(!str_detect(missing_tracts, paste(allowed_missing_tracts, collapse = "|")))

# get unique subjects who have any unacceptable missing tracts
subjects_to_exclude <- unacceptable_missing_tracts %>%
  distinct(sub)

# count the number of excluded subjects
num_excluded_subjects <- nrow(subjects_to_exclude)
cat("Number of subjects excluded:", num_excluded_subjects, "\n")

write.table(subjects_to_exclude, "/cbica/projects/luo_wm_dev/input/PNC/sample_selection_files/subs_missing_tracts.txt", col.names=T, row.names=F, quote = TRUE)

sub_id <- sub_id[!sub_id$sub_id %in% subjects_to_exclude$sub,] %>% as.data.frame() %>% setNames("sub_id") # remove 1 participants from sample of 1102 due to failing qsiprep/qsirecon, new N = 1101

```

```{r}
# add back subjects that are missing non-important tracts. save PNC tract profiles
# Identify subjects missing only non-essential tracts (tracts in allowed_missing_tracts)
# = identify subjects who are not in the exclusion list
subjects_with_allowed_missing_tracts <- missing_tracts_long %>%
  filter(sub %notin% subjects_to_exclude$sub) %>%
  filter(str_detect(missing_tracts, paste(allowed_missing_tracts, collapse = "|")))

# get sub ids
subjects_with_allowed_missing_tracts <- subjects_with_allowed_missing_tracts %>%
  distinct(sub)

# add these subjects' data back to the full tract profiles data (should be 0 participants added back)
PNC_all_with_allowed <- bind_rows(PNC_all, missing_tracts %>% filter(sub %in% subjects_with_allowed_missing_tracts$sub))
PNC_all_with_allowed <- PNC_all_with_allowed %>%
  distinct()

# remove the tracts we don't need
PNC_all_with_allowed <- PNC_all_with_allowed %>% filter(!str_detect(tractID, paste(allowed_missing_tracts, collapse = "|")))

write.table(PNC_all_with_allowed, "/cbica/projects/luo_wm_dev/input/PNC/derivatives/tract_profiles/all_subjects/collated_tract_profiles_nocovbat.tsv", row.names = FALSE)

```

## Save Final Sample: Subject List, Demographics, QC measures
```{r}
# Make final sample demographics dataframe
demographics.final <- demographics.temp[demographics.temp$sub %in% sub_id$sub_id,] %>% select(sub, age, sex, race) # combine demographics data 
 
# Make final sample QC metrics dataframe 
dwi.qc.final <- dwi.qc.temp[dwi.qc.temp$sub %in% sub_id$sub_id,] %>% select(sub, mean_fd, t1_neighbor_corr) 

write.table(sub_id, "/cbica/projects/luo_wm_dev/input/PNC/subject_list/final_sample/PNC_WMDev_FinalSample.txt", col.names=T, row.names=F, quote = TRUE)
write.csv(merge(demographics.final, dwi.qc.final), "/cbica/projects/luo_wm_dev/input/PNC/sample_selection_files/final_sample/PNC_WMDev_FinalSampleDemoQC.csv", quote = F, row.names = F)
```

## Demographics table
```{r}
final_sample <- read.csv("/cbica/projects/luo_wm_dev/input/PNC/sample_selection_files/final_sample/PNC_WMDev_FinalSampleDemoQC.csv", header=T)

# demographic variables to summarize
demo_vars <- c("sex", "race")

# function to summarize the demographic variables
summary_list <- map(demo_vars, function(var) {
  final_sample %>%
    group_by(.data[[var]]) %>%
    summarise(
      count = n(),
      percent = (n() / nrow(final_sample)) * 100
    )
})

# summaries for each variable
summary_list

range(final_sample$age)
mean(final_sample$age)
sd(final_sample$age)
```