---
title: "logistic model scrap for age of maturation"
author: "Audrey Luo"
date: "2024-10-22"
output: html_document
---
 
 
```{r}

plot_diffs <- function(dataset) {
  
  diffs_all <- get(paste0("diffs_", dataset))
  title = gsub("PD", "P-D", dataset) # for HCPD
  diffs_plot <- ggplot(diffs_all, aes(x = mean_SA_diff, y = age_effect_diff, label = bundle_name)) +
  geom_point(size = 5, aes(fill = bundle_name, color = bundle_name), shape = 21, alpha = 0.9) +  # Plot points
  scale_fill_manual(values = aquamarine) +
  scale_color_manual(values = aquamarine) +
  ggrepel::geom_text_repel(position = position_jitter(seed = 3), 
    box.padding = 0.5,  # Increase space around labels
    point.padding = 0.5,  # Increase space around points
    segment.color = "gray",  # Line color connecting point to label
    segment.size = 0.5,  # Line thickness
    max.overlaps = Inf) + # Add labels
  geom_smooth(data = diffs_all, method='lm', se=TRUE, fill=alpha(c("gray70"),.9), col="black") + 
  stat_cor(method = "pearson", label.x = -50, label.y = 10, size = 6) + 
  labs(title = title) + theme_classic() +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
    axis.title = element_blank(),
    axis.text.x = element_text(size = 20),
    axis.text.y = element_text(size = 20))
  
  return(diffs_plot)
}

colormap <- c("#052632","#15CABE","#ABF590","#fffe76")
aquamarine <- colorRampPalette(colormap)(20)


```
 
```{r cars}

test a logistic model:
- for nodes that never matured, code them as 0. For nodes that did mature, code them as 1. Then do a logistic regression. I wonder if I can use NEST for this (since a wald test is used?)
logistic_model <- glm(maturation_status_0_1 ~ SA_axis, family = binomial, data = your_data)
- plotted with help from https://stats.oarc.ucla.edu/r/dae/logit-regression/ 
```{r, fig.height = 5, fig.width = 5}

 
fit_logit_SAaxis <- function(dataset) {
  aggregated_axis <- get(paste0("aggregated_axis_", dataset, "_binary"))
  logistic_model <- glm(maturation_status ~ SA.axis_rank, family = binomial, data = aggregated_axis) # this is group level data so i'm just doing maturation status by SA axis rank
  
  # make sequence of S-A axis rank values for prediction
  sa_axis_seq <- seq(min(aggregated_axis$SA.axis_rank), 
                     max(aggregated_axis$SA.axis_rank), 
                     length.out = 100)
  
  # new data frame for the prediction with the sequence of SA.axis_rank
  newdata <- data.frame(SA.axis_rank = sa_axis_seq)
  
  # get predictions and standard errors from logistic model
  newdata <- cbind(newdata, predict(logistic_model, newdata = newdata, type = "link", se.fit = TRUE))
  
  # back-transform the predicted values and confidence intervals to probabilities
  newdata <- within(newdata, {
    PredictedProb <- plogis(fit)  # Predicted probabilities
    LL <- plogis(fit - (1.96 * se.fit))  # Lower bound of the 95% CI
    UL <- plogis(fit + (1.96 * se.fit))  # Upper bound of the 95% CI
  })
  
  return(list(logistic_model, newdata))
}

 
plot_logit_SAaxis <- function(dataset, newdata, color) {
  aggregated_axis <- get(paste0("aggregated_axis_", dataset, "_binary"))
  title <- gsub("PD", "P-D", dataset)
  plot <- ggplot() +
  geom_point(data = aggregated_axis, 
             aes(x = SA.axis_rank, y = maturation_status, fill = SA.axis_rank),  
             color = "gray", alpha = 0.5, size = 4, shape = 21) +  # Points with gray outline and color fill based on SA.axis_rank
  paletteer::scale_fill_paletteer_c("grDevices::RdYlBu", na.value = "grey100", direction = -1) +  # Correct fill color scale for SA.axis_rank
  geom_ribbon(data = newdata, aes(x = SA.axis_rank, ymin = LL, ymax = UL), fill = color, alpha = 0.5) +  # Confidence interval
  geom_line(data = newdata, aes(x = SA.axis_rank, y = PredictedProb), color = color, size = 1) +  # Predicted probabilities
  labs(x = "S-A Axis Rank", y = "Probability of Maturation", title = title) + 
  theme_classic() + 
  theme(legend.position = "none",
        plot.title = element_text(size = 20, hjust = 0.5),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20)) 
  
  return(plot)
  
}
 
```


```{r sa axis logit, fig.height = 5, fig.width = 5}
aggregated_axis_HCPD_binary <- aggregated_axis_HCPD_binary %>% mutate(maturation_status = ifelse(is.na(regional_mean_ageeffect), 0, 1))
logit_SA_HCPD <- fit_logit_SAaxis("HCPD")
plot_logit_SAaxis("HCPD", logit_SA_HCPD[[2]], color = color1) 


aggregated_axis_HBN_binary <- aggregated_axis_HBN_binary %>% mutate(maturation_status = ifelse(is.na(regional_mean_ageeffect), 0, 1))
logit_SA_HBN <- fit_logit_SAaxis("HBN")
plot_logit_SAaxis("HBN", logit_SA_HBN[[2]], color = color2) 


aggregated_axis_PNC_binary <- aggregated_axis_PNC_binary %>% mutate(maturation_status = ifelse(is.na(regional_mean_ageeffect), 0, 1))
logit_SA_PNC <- fit_logit_SAaxis("PNC")
plot_logit_SAaxis("PNC", logit_SA_PNC[[2]], color = color3) 

```

this shit is real. i just need to prove it with stats.
```{r logit maps, fig.height = 5, fig.width = 5}

agg_ageeffects_HCPD <- agg_ageeffects_HCPD %>% mutate(maturation_status = ifelse(is.na(regional_mean_ageeffect), 0, 1))
agg_ageeffects_HBN <- agg_ageeffects_HBN %>% mutate(maturation_status = ifelse(is.na(regional_mean_ageeffect), 0, 1))
agg_ageeffects_PNC <- agg_ageeffects_PNC %>% mutate(maturation_status = ifelse(is.na(regional_mean_ageeffect), 0, 1))


fit_logit_maps <- function(dataset, map, map_name) {
  agg_ageeffects <- get(paste0("agg_ageeffects_", dataset))
  map <- data.frame(map$data) %>% setNames("map_measure")
  agg_ageeffects_merged <- cbind(agg_ageeffects, map)

  logistic_model <- glm(maturation_status ~ map_measure, family = binomial, data = agg_ageeffects_merged)
  
  # make sequence of map values for prediction
  map_seq <- seq(min(agg_ageeffects_merged$map_measure, na.rm = T), 
                   max(agg_ageeffects_merged$map_measure, na.rm = T), 
                   length.out = 100)
  
  # new data frame for the prediction with the sequence of SA.axis_rank
  newdata <- data.frame(map_measure = map_seq)
  
  # get predictions and standard errors from logistic model
  newdata <- cbind(newdata, predict(logistic_model, newdata = newdata, type = "link", se.fit = TRUE))
  
  # back-transform the predicted values and confidence intervals to probabilities
  newdata <- within(newdata, {
    PredictedProb <- plogis(fit)  # Predicted probabilities
    LL <- plogis(fit - (1.96 * se.fit))  # Lower bound of the 95% CI
    UL <- plogis(fit + (1.96 * se.fit))  # Upper bound of the 95% CI
  })
  
  assign(paste0("agg_ageeffects_", dataset, "_", map_name), agg_ageeffects_merged, envir = .GlobalEnv)
  return(list(logistic_model, newdata))
}

 
 

 
plot_logit_map <- function(dataset, newdata, map_name, color) {
  aggregated_df <- get(paste0("agg_ageeffects_", dataset, "_", map_name))
  title <- gsub("PD", "P-D", dataset)
  plot <- ggplot() +
  geom_point(data = aggregated_df, 
             aes(x = map_measure, y = maturation_status),  
             color = color, fill = color, alpha = 0.5, size = 4, shape = 21) +  # Points with gray outline and color fill based on SA.axis_rank
  geom_ribbon(data = newdata, aes(x = map_measure, ymin = LL, ymax = UL), fill = color, alpha = 0.5) +  # Confidence interval
  geom_line(data = newdata, aes(x = map_measure, y = PredictedProb), color = color, size = 1) +  # Predicted probabilities
  labs(x = "", y = "Probability of Maturation", title = title) + 
  theme_classic() + 
  theme(legend.position = "none",
        plot.title = element_text(size = 20, hjust = 0.5),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20)) 
  
  return(plot)
  
} 
  
  
# t1t2_ageeffect_HCPD
#cor.test(t1t2_mean_HCPD$data, agg_ageeffects$regional_mean_ageeffect, method = 'spearman')   ## 
#cor.test(t1t2_roc_HCPD$data, agg_ageeffects$regional_mean_ageeffect, method = 'spearman') ##
#cor.test(t1t2_maxslopeage_HCPD$data, agg_ageeffects$regional_mean_ageeffect, method = 'spearman') ##
#cor.test(alff_pnc$data, agg_ageeffects$regional_mean_ageeffect, method = 'spearman')  ##
#cor.test(alff_decline_pnc$data, agg_ageeffects$regional_mean_ageeffect, method = 'spearman') 
```



```{r logit maps plots, fig.height = 5, fig.width = 15}
# t1/t2 age effect
logit_HCPD_t1t2ageeffect <- fit_logit_maps("HCPD", t1t2_ageeffect_HCPD, "t1t2_ageeffect") 
logit_HCPD_t1t2ageeffect_plot <- plot_logit_map("HCPD", logit_HCPD_t1t2ageeffect[[2]], "t1t2_ageeffect", color1)

logit_HBN_t1t2ageeffect <- fit_logit_maps("HBN", t1t2_ageeffect_HCPD, "t1t2_ageeffect") 
logit_HBN_t1t2ageeffect_plot <- plot_logit_map("HBN", logit_HBN_t1t2ageeffect[[2]], "t1t2_ageeffect", color2)

logit_PNC_t1t2ageeffect <- fit_logit_maps("PNC", t1t2_ageeffect_HCPD, "t1t2_ageeffect") 
logit_PNC_t1t2ageeffect_plot <- plot_logit_map("PNC", logit_PNC_t1t2ageeffect[[2]], "t1t2_ageeffect", color3)

logit_t1t2ageeffect_final <- ggarrange(logit_HCPD_t1t2ageeffect_plot, logit_HBN_t1t2ageeffect_plot, logit_PNC_t1t2ageeffect_plot, ncol = 3)

x.grob <- textGrob("T1w/T2w Partial R-squared (Age Effect)", 
                   gp=gpar(col="black", fontsize=20))

grid.arrange(arrangeGrob(logit_t1t2ageeffect_final, bottom = x.grob))

# t1t2_mean_HCPD
logit_HCPD_t1t2mean <- fit_logit_maps("HCPD", t1t2_mean_HCPD, "t1t2_mean") 
logit_HCPD_t1t2mean_plot <- plot_logit_map("HCPD", logit_HCPD_t1t2mean[[2]], "t1t2_mean", color1)

logit_HBN_t1t2mean <- fit_logit_maps("HBN", t1t2_mean_HCPD, "t1t2_mean") 
logit_HBN_t1t2mean_plot <- plot_logit_map("HBN", logit_HBN_t1t2mean[[2]], "t1t2_mean", color2)

logit_PNC_t1t2mean <- fit_logit_maps("PNC", t1t2_mean_HCPD, "t1t2_mean") 
logit_PNC_t1t2mean_plot <- plot_logit_map("PNC", logit_PNC_t1t2mean[[2]], "t1t2_mean", color3)

logit_t1t2mean_final <- ggarrange(logit_HCPD_t1t2mean_plot, logit_HBN_t1t2mean_plot, logit_PNC_t1t2mean_plot, ncol = 3)

x.grob <- textGrob("Mean T1w/T2w", 
                   gp=gpar(col="black", fontsize=20))

grid.arrange(arrangeGrob(logit_t1t2mean_final, bottom = x.grob))



# t1t2_roc_HCPD
logit_HCPD_t1t2roc <- fit_logit_maps("HCPD", t1t2_roc_HCPD, "t1t2_roc") 
logit_HCPD_t1t2roc_plot <- plot_logit_map("HCPD", logit_HCPD_t1t2roc[[2]], "t1t2_roc", color1)

logit_HBN_t1t2roc <- fit_logit_maps("HBN", t1t2_roc_HCPD, "t1t2_roc") 
logit_HBN_t1t2roc_plot <- plot_logit_map("HBN", logit_HBN_t1t2roc[[2]], "t1t2_roc", color2)

logit_PNC_t1t2roc <- fit_logit_maps("PNC", t1t2_roc_HCPD, "t1t2_roc") 
logit_PNC_t1t2roc_plot <- plot_logit_map("PNC", logit_PNC_t1t2roc[[2]], "t1t2_roc", color3)

logit_t1t2roc_final <- ggarrange(logit_HCPD_t1t2roc_plot, logit_HBN_t1t2roc_plot, logit_PNC_t1t2roc_plot, ncol = 3)

x.grob <- textGrob("T1w/T2w Annualized Rate of Change", 
                   gp=gpar(col="black", fontsize=20))

grid.arrange(arrangeGrob(logit_t1t2roc_final, bottom = x.grob))


# t1t2_maxslopeage_HCPD
logit_HCPD_t1t2maxslopeage <- fit_logit_maps("HCPD", t1t2_maxslopeage_HCPD, "t1t2_maxslopeage") 
logit_HCPD_t1t2maxslopeage_plot <- plot_logit_map("HCPD", logit_HCPD_t1t2maxslopeage[[2]], "t1t2_maxslopeage", color1)
summary(logit_HCPD_t1t2maxslopeage[[1]])

logit_HBN_t1t2maxslopeage <- fit_logit_maps("HBN", t1t2_maxslopeage_HCPD, "t1t2_maxslopeage") 
logit_HBN_t1t2maxslopeage_plot <- plot_logit_map("HBN", logit_HBN_t1t2maxslopeage[[2]], "t1t2_maxslopeage", color2)
summary(logit_HBN_t1t2maxslopeage[[1]])

logit_PNC_t1t2maxslopeage <- fit_logit_maps("PNC", t1t2_maxslopeage_HCPD, "t1t2_maxslopeage") 
logit_PNC_t1t2maxslopeage_plot <- plot_logit_map("PNC", logit_PNC_t1t2maxslopeage[[2]], "t1t2_maxslopeage", color3)
summary(logit_PNC_t1t2maxslopeage[[1]])

logit_t1t2maxslopeage_final <- ggarrange(logit_HCPD_t1t2maxslopeage_plot, logit_HBN_t1t2maxslopeage_plot, logit_PNC_t1t2maxslopeage_plot, ncol = 3)

x.grob <- textGrob("T1w/T2w Age of Maximum Growth Rate", 
                   gp=gpar(col="black", fontsize=20))

grid.arrange(arrangeGrob(logit_t1t2maxslopeage_final, bottom = x.grob))



# alff_pnc
logit_HCPD_alff_ageeffect <- fit_logit_maps("HCPD", alff_pnc, "alff_ageeffect") 
logit_HCPD_alff_ageeffect_plot <- plot_logit_map("HCPD", logit_HCPD_alff_ageeffect[[2]], "alff_ageeffect", color1)

logit_HBN_alff_ageeffect <- fit_logit_maps("HBN", alff_pnc, "alff_ageeffect") 
logit_HBN_alff_ageeffect_plot <- plot_logit_map("HBN", logit_HBN_alff_ageeffect[[2]], "alff_ageeffect", color2)

logit_PNC_alff_ageeffect <- fit_logit_maps("PNC", alff_pnc, "alff_ageeffect") 
logit_PNC_alff_ageeffect_plot <- plot_logit_map("PNC", logit_PNC_alff_ageeffect[[2]], "alff_ageeffect", color3)

logit_alff_ageeffect_final <- ggarrange(logit_HCPD_alff_ageeffect_plot, logit_HBN_alff_ageeffect_plot, logit_PNC_alff_ageeffect_plot, ncol = 3)

x.grob <- textGrob("ALFF Partial R-squared (Age Effect)", 
                   gp=gpar(col="black", fontsize=20))

grid.arrange(arrangeGrob(logit_alff_ageeffect_final, bottom = x.grob))



# alff_decline_pnc
logit_HCPD_alff_decline <- fit_logit_maps("HCPD", alff_decline_pnc, "alff_decline") 
logit_HCPD_alff_decline_plot <- plot_logit_map("HCPD", logit_HCPD_alff_decline[[2]], "alff_decline", color1)

logit_HBN_alff_decline <- fit_logit_maps("HBN", alff_decline_pnc, "alff_decline") 
logit_HBN_alff_decline_plot <- plot_logit_map("HBN", logit_HBN_alff_decline[[2]], "alff_decline", color2)

logit_PNC_alff_decline <- fit_logit_maps("PNC", alff_decline_pnc, "alff_decline") 
logit_PNC_alff_decline_plot <- plot_logit_map("PNC", logit_PNC_alff_decline[[2]], "alff_decline", color3)

logit_alff_decline_final <- ggarrange(logit_HCPD_alff_decline_plot, logit_HBN_alff_decline_plot, logit_PNC_alff_decline_plot, ncol = 3)

x.grob <- textGrob("ALFF Age of decline", 
                   gp=gpar(col="black", fontsize=20))

grid.arrange(arrangeGrob(logit_alff_decline_final, bottom = x.grob))

```

