---
title: "age of maturation all datasets"
author: "Audrey Luo"
output: html_document
---

```{r}
library(ciftiTools)
#ciftiTools.setOption('wb_path', '/Applications/workbench/')
library(cifti)
library(cowplot)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(gratia) 
library(ggseg)
library(ggsegGlasser)
library(knitr)
library(kableExtra)
library(mgcv)
library(RColorBrewer)
library(scales)
library(stringr)
library(rjson)
library(tidyr)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, include = TRUE) 

font_size <- 20
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1
point_size <- 2

proj_root <- "/cbica/projects/luo_wm_dev/"
input_root <- paste0(proj_root, "input")
output_root <- paste0(proj_root, "output")
png_dir <- "/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/all_datasets/tract_to_region"
local_dir <- "/Users/audluo/Desktop/temp_code"
```

note:
- for IFO, need to use lower threshold - 10%


# brain surface plot of age effects of peripheral white matter for each tract
- first clip 3 nodes from each end of the tract
- for each tract, average the 5 nodes on each end and get a value (age maps are pretty much identical with 10 nodes)
- make a new column in the maps df and for every nonzero thresholded_probability, put that value in 

```{r load glasser maps}
# load glasser labels
glasser_labels <- read.csv("/cbica/projects/luo_wm_dev/atlases/glasser/HCP-MMP1_UniqueRegionList.csv")
glasser_labels$regionID <- c(1:360)
glasser_labels$region <- gsub("7Pl", "7PL", glasser_labels$region)   

# @param depth, A string
load_maps <- function(depth, dataset) {
  
  config_file <- sprintf("%1$s/code/tract_profiles/config/config_%2$s.json", proj_root, dataset)
  config <- fromJSON(paste(readLines(config_file), collapse=""))
  
  vol_to_surf_dir <- paste0(config$data_root, "/derivatives/vol_to_surf")
  group_dir <- paste0(vol_to_surf_dir, "/group")
  

  pattern <- paste0(depth, ".*\\.csv$")

  # load glasser maps for each tract... load for a specific depth?
  glasser_csvs <- list.files(path = group_dir, pattern = pattern, full.names = T)
  tract_names <- lapply(glasser_csvs, function(path) {
    filename <- basename(path)
    sub("_\\d+\\.\\d+_glasser\\.csv$", "", filename)
  })
  tract_names <- unlist(tract_names)
  
  glasser_maps <- lapply(glasser_csvs, read.csv)
  names(glasser_maps) <- tract_names
  
  # merge labels with my maps
  glasser_maps <- lapply(glasser_maps, merge, y = glasser_labels, by = "regionID")
  
  # separate lh and rh maps
  lh_idx <- grep("Left", names(glasser_maps))
  lh_maps <- glasser_maps[lh_idx]
  assign(paste0("lh_maps_", dataset), lh_maps, envir = .GlobalEnv)
  
  rh_idx <- grep("Right", names(glasser_maps))
  rh_maps <- glasser_maps[rh_idx]
  assign(paste0("rh_maps_", dataset), rh_maps, envir = .GlobalEnv)
}

depths = c("1.5")

lapply(depths, load_maps, "HCPD")
lapply(depths, load_maps, "HBN")
lapply(depths, load_maps, "PNC")
 
```

# load age effects / age of maturation
```{r load dev effects files}
scalar = "dti_md"
HCPD_ageeffects <- read.csv(paste0(output_root, "/", "HCPD", "/tract_profiles/GAM/", scalar, "/HCPD_GAM_dev_measures.csv"))
HBN_ageeffects <- read.csv(paste0(output_root, "/", "HBN", "/tract_profiles/GAM/", scalar, "/HBN_GAM_dev_measures.csv"))
PNC_ageeffects <- read.csv(paste0(output_root, "/", "PNC", "/tract_profiles/GAM/", scalar, "/PNC_GAM_dev_measures.csv"))
```

```{r function for formatting age effects}
# make a list of the different df names: [dataset]_GAM_ageeffects_[scalar]
datasets <- c("HCPD", "HBN", "PNC")
ageeffect_df_names <- c()
for (dataset in datasets) {
  ageeffect_df_names <- append(ageeffect_df_names, paste0(dataset, "_ageeffects"))
}

 
# format age effect df's
format_ageeffect <- function(df) {
  df <- get(df)
  # add tractID, tract_label, and nodeID
  df <- df %>% mutate(hemi = ifelse(grepl("Left", tract_node), "Left", 
                                    ifelse(grepl("Right", tract_node), "Right", NA))) %>%
    mutate(tractID = gsub("_[0-9]+", "", tract_node)) %>%
    mutate(tract_label = gsub("Left |Right ", "", gsub("_", " ", gsub("Fronto.", "Fronto-", tractID)))) %>%
    mutate(nodeID = str_extract(tract_node, "[0-9]+")) 
  df$nodeID <- as.numeric(df$nodeID)
  
  # label main orientation
  df <- df %>% 
  mutate(main_orientation = case_when(
    tract_label %in% c("Inferior Fronto-occipital",
                   "Inferior Longitudinal", "Superior Longitudinal") ~ "AP",
    tract_label %in% c("Arcuate") ~ "AP_frontal_temporal",
    tract_label %in% c("Callosum Anterior Frontal", "Callosum Motor", "Callosum Occipital", "Callosum Orbital", 
                       "Callosum Posterior Parietal", "Callosum Superior Frontal", "Callosum Superior Parietal", "Callosum Temporal") ~ "RL",
    tract_label %in% c("Corticospinal", "Posterior Arcuate","Vertical Occipital") ~ "SI",
    TRUE ~ NA_character_
  ))
  df$main_orientation <- as.factor(df$main_orientation)
  df <- df %>% relocate(tract_label, tractID, nodeID, tract_node, hemi, main_orientation)
  return(df)
}

ageeffect_dfs <- lapply(ageeffect_df_names, format_ageeffect)
 
```

```{r fdr correction}
# Function for calculating number of significant nodes (FDR corrected)
# @param df A dataframe with GAM results
sig_nodes <- function(df) {

  df$Anova.age.pvalue.fdr <- p.adjust(df$Anova.smooth.pvalue, method=c("fdr"))
  cat(sprintf("There are %s/%s significant nodes\n", sum(df$Anova.age.pvalue.fdr < 0.05, na.rm=TRUE), nrow(df)))
  df$significant.fdr <- df$Anova.age.pvalue.fdr < 0.05
  df$significant.fdr[df$significant.fdr == TRUE] <- 1
  df$significant.fdr[df$significant.fdr == FALSE] <- 0
  return(df)
}

ageeffect.fdr_dfs <- lapply(ageeffect_dfs, sig_nodes)
names(ageeffect.fdr_dfs) <- ageeffect_df_names

```

```{r function assign age effects to cortical endpoints}
# function for assigning age effects to endpoints: need to do this manually for each tract (but code is applicable to all datasets)

threshold = 0.3

# have to do this manually T_T
make_maps <- function(dataset, bin_num_nodes) {
  
  lh_maps <- get(paste0("lh_maps_", dataset))
  rh_maps <- get(paste0("rh_maps_", dataset))
  deveffects <- get(paste0(dataset, "_deveffects_", bin_num_nodes))
  
  # for tracts with 1 endpoint (CST), extract the age effect for end1 (corresponds to cortical endpoint) 
  CSTL_deveffect <- lh_maps$LeftCorticospinal %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Left_Corticospinal" & node_position == "end1") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end1", NA)
    ) %>%
    relocate(mean_age_effect)
  
  CSTR_deveffect <- rh_maps$RightCorticospinal %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Right_Corticospinal" & node_position == "end1") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end1", NA)
    ) %>%
    relocate(mean_age_effect)
   
  # for tracts with 2 distinct cortical endpoints: ARC, ILF, IFO, SLF
  ARCL_deveffect <- lh_maps$LeftArcuate %>%
    mutate(
      thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>%
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>%
    mutate(
      mean_age_effect = case_when(
        !is.na(thresh_probability) & (str_detect(cortex, "(?i)frontal") | str_detect(cortex, "(?i)motor") | str_detect(cortex, "(?i)Posterior_Opercular")) ~ 
          deveffects %>% filter(tractID == "Left_Arcuate", node_position == "end1") %>% pull(mean_ageeffect),
        !is.na(thresh_probability) & (str_detect(cortex, "(?i)Temp") | str_detect(cortex, "(?i)Auditory")) ~ 
          deveffects %>% filter(tractID == "Left_Arcuate", node_position == "end2") %>% pull(mean_ageeffect),
        TRUE ~ NA_real_
      )
    ) %>%
    mutate(
      end = ifelse(
        !is.na(thresh_probability) & (str_detect(cortex, "(?i)frontal") | str_detect(cortex, "motor") | str_detect(cortex, "Posterior_Opercular")), 
        "end1", 
        ifelse(
          !is.na(thresh_probability) & (str_detect(cortex, "Temp") | str_detect(cortex, "Auditory") ), 
          "end2", 
          NA))
      ) %>%
    relocate(mean_age_effect)
  
  ARCR_deveffect <- rh_maps$RightArcuate %>%
    mutate(
      thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>%
    select(thresh_probability, regionName, region,  cortex) %>%
    arrange(thresh_probability) %>%
    mutate(
      mean_age_effect = case_when(
        !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)frontal") | str_detect(cortex, "(?i)motor") | str_detect(cortex, "(?i)Posterior_Opercular")) ~ 
            deveffects %>% filter(tractID == "Left_Arcuate", node_position == "end1") %>% pull(mean_ageeffect),
        !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)Temp") | str_detect(cortex, "(?i)MT") | str_detect(cortex, "(?i)Auditory")) ~ 
            deveffects %>% filter(tractID == "Left_Arcuate", node_position == "end2") %>% pull(mean_ageeffect),
        TRUE ~ NA_real_
      ),
      end = case_when(
        !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)frontal") | str_detect(cortex, "(?i)motor") | str_detect(cortex, "(?i)Posterior_Opercular")) ~ "end1",
        !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)Temp") | str_detect(cortex, "(?i)MT") | str_detect(cortex, "(?i)Auditory")) ~ "end2",
        TRUE ~ NA_character_
      )
    ) %>%
    relocate(mean_age_effect)
   
  ILFL_deveffect <- lh_maps$LeftInferiorLongitudinal %>%
    mutate(
      thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>%
    select(thresh_probability, regionName, region,  cortex) %>%
    arrange(thresh_probability) %>%
    mutate(
      mean_age_effect = case_when(
        !is.na(thresh_probability) & 
        !str_detect(cortex, "(?i)visual|Inferior_Parietal|Occipital") ~ 
          deveffects %>% filter(tractID == "Left_Inferior_Longitudinal", node_position == "end1") %>% pull(mean_ageeffect),
        !is.na(thresh_probability) & 
        str_detect(cortex, "(?i)visual|Inferior_Parietal|Occipital") ~ 
          deveffects %>% filter(tractID == "Left_Inferior_Longitudinal", node_position == "end2") %>% pull(mean_ageeffect),
        TRUE ~ NA_real_
      ),
      end = case_when(
        !is.na(thresh_probability) & 
        !str_detect(cortex, "(?i)visual|Inferior_Parietal|Occipital") ~ "end1",
        !is.na(thresh_probability) & 
        str_detect(cortex, "(?i)visual|Inferior_Parietal|Occipital") ~ "end2",
        TRUE ~ NA_character_
      )
    ) %>%
    relocate(mean_age_effect)
  
  
  ILFR_deveffect <- rh_maps$RightInferiorLongitudinal %>%
    mutate(
      thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>%
    select(thresh_probability, regionName, region,  cortex) %>%
    arrange(thresh_probability) %>%
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability) & (!str_detect(cortex, "(?i)visual")) & region != "PHA3", 
        deveffects %>%  filter(tractID == "Right_Inferior_Longitudinal", node_position == "end1") %>% pull(mean_ageeffect),
        ifelse(
          !is.na(thresh_probability) & (str_detect(cortex, "(?i)visual")) | region == "PHA3", 
        deveffects %>%  filter(tractID == "Right_Inferior_Longitudinal", node_position == "end2") %>% pull(mean_ageeffect),
          NA))
      ) %>%
    mutate(
      end = ifelse(
        !is.na(thresh_probability) & (!str_detect(cortex, "(?i)visual")) & region != "PHA3", 
        "end1", 
        ifelse(
          !is.na(thresh_probability) & (str_detect(cortex, "(?i)visual")) | region == "PHA3", 
          "end2", 
          NA))
      ) %>%
    relocate(mean_age_effect) %>% arrange(end)
  
  
  IFOL_deveffect <- lh_maps$LeftInferiorFrontooccipital %>%
    mutate(
      thresh_probability = ifelse(probability < 0.1, NA, probability)
    ) %>%
    select(thresh_probability, regionName, region,  cortex) %>%
    arrange(thresh_probability) %>%
    mutate(mean_age_effect = case_when(
      !is.na(thresh_probability) & 
      (str_detect(cortex, "(?i)frontal|Auditory")) ~ 
        deveffects %>% filter(tractID == "Left_Inferior_Fronto.occipital", node_position == "end1") %>% pull(mean_ageeffect),
      !is.na(thresh_probability) & 
      str_detect(cortex, "(?i)visual") ~ 
        deveffects %>% filter(tractID == "Left_Inferior_Fronto.occipital", node_position == "end2") %>% pull(mean_ageeffect),
      TRUE ~ NA_real_)) %>%
    mutate(
      end = ifelse(
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)frontal") | 
         str_detect(cortex, "Auditory")), 
        "end1", 
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)visual")), 
          "end2", 
          NA
        )
      )
    ) %>%
    relocate(mean_age_effect)
   
  IFOR_deveffect <- rh_maps$RightInferiorFrontooccipital %>%
    mutate(
      thresh_probability = ifelse(probability < 0.1, NA, probability)
    ) %>%
    select(thresh_probability, regionName, region,  cortex) %>%
    arrange(thresh_probability) %>%
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)frontal")  |
         str_detect(cortex, "Auditory")), 
        deveffects %>% filter(tractID == "Right_Inferior_Fronto.occipital", node_position == "end1") %>% pull(mean_ageeffect),
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)visual")), 
         deveffects %>% filter(tractID == "Right_Inferior_Fronto.occipital", node_position == "end2") %>% pull(mean_ageeffect),
          NA
        )
      )
    ) %>%
     mutate(
      end = ifelse(
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)frontal") | 
         str_detect(cortex, "Auditory")), 
        "end1", 
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)visual")), 
          "end2", 
          NA
        )
      )
    ) %>%
    relocate(mean_age_effect)
  
  SLFL_deveffect <- lh_maps$LeftSuperiorLongitudinal %>%
    mutate(
      thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>%
    select(thresh_probability, regionName, region,  cortex) %>%
    arrange(thresh_probability) %>%
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)frontal") |
           str_detect(cortex, "(?i)motor") | 
           str_detect(cortex, "(?i)Opercular")), 
        deveffects %>% filter(tractID == "Left_Superior_Longitudinal", node_position == "end1") %>% pull(mean_ageeffect),
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)pariet|(?i)auditory")), 
        deveffects %>% filter(tractID == "Left_Superior_Longitudinal", node_position == "end2") %>% pull(mean_ageeffect),
          NA))) %>%
    mutate(end = ifelse(
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)frontal") |
           str_detect(cortex, "(?i)motor") | 
           str_detect(cortex, "(?i)Opercular")), 
        "end1", 
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)pariet|(?i)auditory")), 
          "end2", 
          NA))) %>%
    relocate(mean_age_effect)
  
  SLFR_deveffect <- rh_maps$RightSuperiorLongitudinal %>%
    mutate(
      thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>%
    select(thresh_probability, regionName, region,  cortex) %>%
    arrange(thresh_probability) %>%
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)frontal") |
           str_detect(cortex, "(?i)motor") | 
           str_detect(cortex, "(?i)Opercular")), 
        deveffects %>% filter(tractID == "Right_Superior_Longitudinal", node_position == "end1") %>% pull(mean_ageeffect),
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)pariet")), 
        deveffects %>% filter(tractID == "Right_Superior_Longitudinal", node_position == "end2") %>% pull(mean_ageeffect),
        NA))) %>%
    mutate(
      end = ifelse(
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)frontal") |
           str_detect(cortex, "(?i)motor") | 
           str_detect(cortex, "(?i)Opercular")), 
        "end1", 
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)pariet")), 
          "end2", 
          NA
        )
      )
    ) %>%
    relocate(mean_age_effect)
    
  # for tracts with 2 cortical endpoints that require some manual work: pARC, VOF 
  pARCL_deveffect <- lh_maps$LeftPosteriorArcuate %>%
    mutate(
      thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>%
    select(thresh_probability, regionName, region,  cortex) %>%
    arrange(thresh_probability) %>%
    mutate(
      mean_age_effect = ifelse( 
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)parietal")),  # might exclude TPOJ1 since its between the two endpoints
        deveffects %>% filter(tractID == "Left_Posterior_Arcuate", node_position == "end1") %>% pull(mean_ageeffect),
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)temporal") |
             str_detect(cortex, "(?i)visual") |
             str_detect(cortex, "(?i)auditory")), 
         deveffects %>% filter(tractID == "Left_Posterior_Arcuate", node_position == "end2") %>% pull(mean_ageeffect), 
          NA))) %>%
    mutate(
      end = ifelse( 
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)parietal")),  # might exclude TPOJ1 since its between the two endpoints
        "end1", 
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)temporal") |
             str_detect(cortex, "(?i)visual") |
             str_detect(cortex, "(?i)auditory")), 
          "end2", 
          NA
        )
      )
    ) %>%
    relocate(mean_age_effect)
  
  pARCR_deveffect <- rh_maps$RightPosteriorArcuate %>%
    mutate(
      thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>%
    select(thresh_probability, regionName, region,  cortex) %>%
    arrange(thresh_probability) %>%
    mutate(
      mean_age_effect = ifelse( 
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)parietal")),  # might exclude TPOJ1 since its between the two endpoints
        deveffects %>% filter(tractID == "Right_Posterior_Arcuate", node_position == "end1") %>% pull(mean_ageeffect), 
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)temporal") |
             str_detect(cortex, "(?i)visual") |
             str_detect(cortex, "(?i)auditory")), 
          deveffects %>% filter(tractID == "Right_Posterior_Arcuate", node_position == "end2") %>% pull(mean_ageeffect), 
          NA
        )
      )
    ) %>%
     mutate(
      end = ifelse( 
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)parietal")),  # might exclude TPOJ1 since its between the two endpoints
        "end1", 
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)temporal") |
             str_detect(cortex, "(?i)visual") |
             str_detect(cortex, "(?i)auditory")), 
          "end2", 
          NA
        )
      )
    ) %>%
    relocate(mean_age_effect)
  
  VOFL_deveffect <- lh_maps$LeftVerticalOccipital %>%
    mutate(
      thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>%
    select(thresh_probability, regionName, region,  cortex) %>%
    arrange(thresh_probability) %>%
    mutate(
      mean_age_effect = ifelse( 
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)dorsal") | 
         str_detect(cortex, "(?i)parietal") |
           region == "LO1" |
           region == "V3CD" |
           region == "V4" | # kind of arbitrarily placing V3 and V4 for now
           region == "MST" |
           region == "MT" | 
           region == "TPOJ3" |
           region == "LO3"),   
        deveffects %>% filter(tractID == "Left_Vertical_Occipital", node_position == "end1") %>% pull(mean_ageeffect), 
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)ventral") | 
           region == "PH" |
           region == "LO2" |
             region == "V3"), 
          deveffects %>% filter(tractID == "Left_Vertical_Occipital", node_position == "end2") %>% pull(mean_ageeffect), 
          NA))) %>%
    mutate(
      end = ifelse( 
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)dorsal") | 
         str_detect(cortex, "(?i)parietal") |
         region == "LO1" |
         region == "V3CD" |
         region == "V4" | # kind of arbitrarily placing V3 and V4 for now
         region == "MST" |
         region == "MT" |
         region == "TPOJ3" |
         region == "LO3"),   
        "end1", 
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)ventral") | 
           region == "PH" |
           region == "LO2" |
             region == "V3" ), 
          "end2", 
          NA
        )
      )
    ) %>%
    relocate(mean_age_effect)
  
  
  VOFR_deveffect <- rh_maps$RightVerticalOccipital %>%
    mutate(
      thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>%
    select(thresh_probability, regionName, region,  cortex) %>%
    arrange(thresh_probability) %>%
    mutate(
      mean_age_effect = ifelse( 
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)dorsal") | 
         str_detect(cortex, "(?i)parietal") |
         region == "LO1" |
         region == "V3CD" |
         region == "V4" | # kind of arbitrarily placing V3 and V4 for now
         region == "LO3" | 
         region == "TPOJ3"),   
         deveffects %>% filter(tractID == "Right_Vertical_Occipital", node_position == "end1") %>% pull(mean_ageeffect), 
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)ventral") | 
           region == "PH" |
           region == "LO2" |
           region == "MST" |
           region == "V3"), 
           deveffects %>% filter(tractID == "Right_Vertical_Occipital", node_position == "end2") %>% pull(mean_ageeffect), NA
        )
      )
    ) %>%
    mutate(
      end = ifelse( 
        !is.na(thresh_probability) & 
        (str_detect(cortex, "(?i)dorsal") |
          str_detect(cortex, "(?i)parietal") |
          region == "LO1" |
          region == "V3CD" |
          region == "V4"| # kind of arbitrarily placing V3 and V4 for now
          region == "LO3" | 
          region == "TPOJ3"),   
        "end1", 
        ifelse(
          !is.na(thresh_probability) & 
          (str_detect(cortex, "(?i)ventral") | 
           region == "PH" |
           region == "LO2" |
           region == "V3" | 
           region == "MST" |
           region == "FST"), 
          "end2", 
          NA
        )
      )
    ) %>%
    relocate(mean_age_effect)
   
  
  # callosum bundles
  COrbL_deveffect <- lh_maps$LeftCallosumOrbital %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Orbital" & node_position == "end2") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end2", NA)
    ) %>%
    relocate(mean_age_effect)
  
  COrbR_deveffect <- rh_maps$RightCallosumOrbital %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Orbital" & node_position == "end1") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end1", NA)
    ) %>%
    relocate(mean_age_effect)
  
  
  CAntFrL_deveffect <- lh_maps$LeftCallosumAnteriorFrontal %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Anterior_Frontal" & node_position == "end2") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end2", NA)
    ) %>%
    relocate(mean_age_effect)
  
  CAntFrR_deveffect <- rh_maps$RightCallosumAnteriorFrontal %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Anterior_Frontal" & node_position == "end1") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end1", NA)
    ) %>%
    relocate(mean_age_effect)
  
  CSupFrL_deveffect <- lh_maps$LeftCallosumSuperiorFrontal %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Superior_Frontal" & node_position == "end2") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end2", NA)
    ) %>%
    relocate(mean_age_effect)
  
  CSupFrR_deveffect <- rh_maps$RightCallosumSuperiorFrontal %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Superior_Frontal" & node_position == "end1") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end1", NA)
    ) %>%
    relocate(mean_age_effect)
  
  CMotL_deveffect <- lh_maps$LeftCallosumMotor %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Motor" & node_position == "end2") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end2", NA)
    ) %>%
    relocate(mean_age_effect)
  
  CMotR_deveffect <- rh_maps$RightCallosumMotor %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Motor" & node_position == "end1") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end1", NA)
    ) %>%
    relocate(mean_age_effect)
  
  CSupParL_deveffect <- lh_maps$LeftCallosumSuperiorParietal %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Superior_Parietal" & node_position == "end2") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end2", NA)
    ) %>%
    relocate(mean_age_effect)
  
  CSupParR_deveffect <- rh_maps$RightCallosumSuperiorParietal %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Superior_Parietal" & node_position == "end1") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end1", NA)
    ) %>%
    relocate(mean_age_effect)
  
  CPostParL_deveffect <- lh_maps$LeftCallosumPosteriorParietal %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Posterior_Parietal" & node_position == "end2") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end2", NA)
    ) %>%
    relocate(mean_age_effect)
  
  CPostParR_deveffect <- rh_maps$RightCallosumPosteriorParietal %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Posterior_Parietal" & node_position == "end1") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end1", NA)
    ) %>%
    relocate(mean_age_effect)
  
  CTempL_deveffect <- lh_maps$LeftCallosumTemporal %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Temporal" & node_position == "end2") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end2", NA)
    ) %>%
    relocate(mean_age_effect)
  
  CTempR_deveffect <- rh_maps$RightCallosumTemporal %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Temporal" & node_position == "end1") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end1", NA)
    ) %>%
    relocate(mean_age_effect)
  
  COccL_deveffect <- lh_maps$LeftCallosumOccipital %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Occipital" & node_position == "end2") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end2", NA)
    ) %>%
    relocate(mean_age_effect)
  
  COccR_deveffect <- rh_maps$RightCallosumOccipital %>% mutate(
    thresh_probability = ifelse(probability < threshold, NA, probability)
    ) %>% 
    select(thresh_probability, regionName, region, cortex) %>%
    arrange(thresh_probability) %>% 
    mutate(
      mean_age_effect = ifelse(
        !is.na(thresh_probability), 
        deveffects %>% filter(tractID == "Callosum_Occipital" & node_position == "end1") %>% pull(mean_ageeffect), NA)
      ) %>% 
     mutate(
      end = ifelse(
        !is.na(thresh_probability), 
        "end1", NA)
    ) %>%
    relocate(mean_age_effect)
  
  assign(paste0("CSTL_deveffect_", dataset), CSTL_deveffect, envir = .GlobalEnv)
  assign(paste0("CSTR_deveffect_", dataset), CSTR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("ARCL_deveffect_", dataset), ARCL_deveffect, envir = .GlobalEnv)
  assign(paste0("ARCR_deveffect_", dataset), ARCR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("IFOL_deveffect_", dataset), IFOL_deveffect, envir = .GlobalEnv)
  assign(paste0("IFOR_deveffect_", dataset), IFOR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("ILFL_deveffect_", dataset), ILFL_deveffect, envir = .GlobalEnv)
  assign(paste0("ILFR_deveffect_", dataset), ILFR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("pARCL_deveffect_", dataset), pARCL_deveffect, envir = .GlobalEnv)
  assign(paste0("pARCR_deveffect_", dataset), pARCR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("SLFL_deveffect_", dataset), SLFL_deveffect, envir = .GlobalEnv)
  assign(paste0("SLFR_deveffect_", dataset), SLFR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("VOFL_deveffect_", dataset), VOFL_deveffect, envir = .GlobalEnv)
  assign(paste0("VOFR_deveffect_", dataset), VOFR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("COrbL_deveffect_", dataset), COrbL_deveffect, envir = .GlobalEnv)
  assign(paste0("COrbR_deveffect_", dataset), COrbR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("CAntFrL_deveffect_", dataset), CAntFrL_deveffect, envir = .GlobalEnv)
  assign(paste0("CAntFrR_deveffect_", dataset), CAntFrR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("CSupFrL_deveffect_", dataset), CSupFrL_deveffect, envir = .GlobalEnv)
  assign(paste0("CSupFrR_deveffect_", dataset), CSupFrR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("CMotL_deveffect_", dataset), CMotL_deveffect, envir = .GlobalEnv)
  assign(paste0("CMotR_deveffect_", dataset), CMotR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("CSupParL_deveffect_", dataset), CSupParL_deveffect, envir = .GlobalEnv)
  assign(paste0("CSupParR_deveffect_", dataset), CSupParR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("CPostParL_deveffect_", dataset), CPostParL_deveffect, envir = .GlobalEnv)
  assign(paste0("CPostParR_deveffect_", dataset), CPostParR_deveffect, envir = .GlobalEnv)
  
  assign(paste0("CTempL_deveffect_", dataset), CTempL_deveffect, envir = .GlobalEnv)
  assign(paste0("CTempR_deveffect_", dataset), CTempR_deveffect, envir = .GlobalEnv)
  
  
  assign(paste0("COccL_deveffect_", dataset), COccL_deveffect, envir = .GlobalEnv)
  assign(paste0("COccR_deveffect_", dataset), COccR_deveffect, envir = .GlobalEnv)
}

 
``` 
 
 
```{r set tract names}
lh_assoc_names <- c("ARCL", "CSTL" ,"ILFL", "IFOL", "SLFL", "pARCL", "VOFL")
rh_assoc_names <- c("ARCR", "CSTR", "ILFR", "IFOR", "SLFR", "pARCR", "VOFR")
lh_callosum_names <- c("COrbL", "CAntFrL" ,"CSupFrL", "CMotL", "CSupParL", "CPostParL", "CTempL", "COccL")
rh_callosum_names <- c("COrbR", "CAntFrR" ,"CSupFrR", "CMotR", "CSupParR", "CPostParR", "CTempR", "COccR")
```

not using bin 10
```{r make age of mat maps bin 10}

HCPD_deveffects_10_agemat <- ageeffect.fdr_dfs$HCPD_ageeffects %>% filter((nodeID < 16 & nodeID > 4) | (nodeID < 95 & nodeID > 85)) %>% mutate(node_position = case_when((nodeID < 16 & nodeID > 4) ~ "end1", (nodeID < 95 & nodeID > 85) ~ "end2")) %>%
  select(tract_label, tractID, nodeID, node_position, hemi, smooth.peak.change, smooth.decrease.offset, smooth.last.change, smooth.slowing.onset) %>%
  group_by(tractID, node_position, hemi) %>% summarise(mean_peak_change = mean(smooth.peak.change, na.rm = T),
                              mean_ageeffect = mean(smooth.decrease.offset, na.rm = T), # mean_ageeffect = mean_age_mat
                              mean_last_change = mean(smooth.last.change, na.rm = T),
                              mean_dev_slowing = mean(smooth.slowing.onset, na.rm = T))

HBN_deveffects_10_agemat <- ageeffect.fdr_dfs$HBN_ageeffects %>% filter((nodeID < 16 & nodeID > 4) | (nodeID < 95 & nodeID > 85)) %>% mutate(node_position = case_when((nodeID < 16 & nodeID > 4) ~ "end1", (nodeID < 95 & nodeID > 85) ~ "end2")) %>%
  select(tract_label, tractID, nodeID, node_position, hemi, smooth.peak.change, smooth.decrease.offset, smooth.last.change, smooth.slowing.onset) %>%
  group_by(tractID, node_position, hemi) %>% summarise(mean_peak_change = mean(smooth.peak.change, na.rm = T),
                              mean_ageeffect = mean(smooth.decrease.offset, na.rm = T),
                              mean_last_change = mean(smooth.last.change, na.rm = T),
                              mean_dev_slowing = mean(smooth.slowing.onset, na.rm = T))
       
 
PNC_deveffects_10_agemat <- ageeffect.fdr_dfs$PNC_ageeffects %>% filter((nodeID < 16 & nodeID > 4) | (nodeID < 95 & nodeID > 85)) %>% mutate(node_position = case_when((nodeID < 16 & nodeID > 4) ~ "end1", (nodeID < 95 & nodeID > 85) ~ "end2")) %>%
  select(tract_label, tractID, nodeID, node_position, hemi, smooth.peak.change, smooth.decrease.offset, smooth.last.change, smooth.slowing.onset) %>%
  group_by(tractID, node_position, hemi) %>% summarise(mean_peak_change = mean(smooth.peak.change, na.rm = T),
                              mean_ageeffect = mean(smooth.decrease.offset, na.rm = T),
                              mean_last_change = mean(smooth.last.change, na.rm = T),
                              mean_dev_slowing = mean(smooth.slowing.onset, na.rm = T))
  
 
make_maps("HCPD", "10_agemat")
make_maps("HBN", "10_agemat")
make_maps("PNC", "10_agemat")

```


```{r make age of mat maps bin 5}

HCPD_deveffects_5_agemat <- ageeffect.fdr_dfs$HCPD_ageeffects %>% filter((nodeID < 11 & nodeID > 4) | (nodeID < 95 & nodeID > 89)) %>% mutate(node_position = case_when((nodeID < 11 & nodeID > 4) ~ "end1", (nodeID < 95 & nodeID > 89) ~ "end2")) %>%
  select(tract_label, tractID, nodeID, node_position, hemi, smooth.peak.change, smooth.decrease.offset, smooth.last.change, smooth.slowing.onset) %>%
  group_by(tractID, node_position, hemi) %>% summarise(mean_peak_change = mean(smooth.peak.change, na.rm = T),
                              mean_ageeffect = mean(smooth.decrease.offset, na.rm = T), # mean_ageeffect = mean_age_mat
                              mean_last_change = mean(smooth.last.change, na.rm = T),
                              mean_dev_slowing = mean(smooth.slowing.onset, na.rm = T))

HBN_deveffects_5_agemat <- ageeffect.fdr_dfs$HBN_ageeffects %>% filter((nodeID < 11 & nodeID > 4) | (nodeID < 95 & nodeID > 89)) %>% mutate(node_position = case_when((nodeID < 11 & nodeID > 4) ~ "end1", (nodeID < 95 & nodeID > 89) ~ "end2")) %>%
  select(tract_label, tractID, nodeID, node_position, hemi, smooth.peak.change, smooth.decrease.offset, smooth.last.change, smooth.slowing.onset) %>%
  group_by(tractID, node_position, hemi) %>% summarise(mean_peak_change = mean(smooth.peak.change, na.rm = T),
                              mean_ageeffect = mean(smooth.decrease.offset, na.rm = T),
                              mean_last_change = mean(smooth.last.change, na.rm = T),
                              mean_dev_slowing = mean(smooth.slowing.onset, na.rm = T))
       
 
PNC_deveffects_5_agemat <- ageeffect.fdr_dfs$PNC_ageeffects %>% filter((nodeID < 11 & nodeID > 4) | (nodeID < 95 & nodeID > 89)) %>% mutate(node_position = case_when((nodeID < 11 & nodeID > 4) ~ "end1", (nodeID < 95 & nodeID > 89) ~ "end2")) %>%
  select(tract_label, tractID, nodeID, node_position, hemi, smooth.peak.change, smooth.decrease.offset, smooth.last.change, smooth.slowing.onset) %>%
  group_by(tractID, node_position, hemi) %>% summarise(mean_peak_change = mean(smooth.peak.change, na.rm = T),
                              mean_ageeffect = mean(smooth.decrease.offset, na.rm = T),
                              mean_last_change = mean(smooth.last.change, na.rm = T),
                              mean_dev_slowing = mean(smooth.slowing.onset, na.rm = T))
  
 
make_maps("HCPD", "5_agemat") # makes [bundle_name]_deveffect_[dataset] for each dataset and bundle. e.g. IFOL_deveffect_HCPD
make_maps("HBN", "5_agemat")
make_maps("PNC", "5_agemat")

```

 
# aggregate age maps
```{r aggregate age maps} 
# create dataframe that combines the age of maturation for all cortical endpoints across tracts (separate for hemisphere). This is done for each dataset.
aggregate_age_maps <- function(dataset) {
  # left hemi
  lh_base_names <- c("COrbL_deveffect_", "CAntFrL_deveffect_", "CSupFrL_deveffect_", 
                "CMotL_deveffect_", "CSupParL_deveffect_", "CPostParL_deveffect_", 
                "CTempL_deveffect_", "COccL_deveffect_", "ARCL_deveffect_", 
                "IFOL_deveffect_", "SLFL_deveffect_", "CSTL_deveffect_", 
                "ILFL_deveffect_", "pARCL_deveffect_", "VOFL_deveffect_")
  lh_dfs <- lapply(lh_base_names, function(name) get(paste0(name, dataset))) # tract_deveffect_dataset (age of maturation maps for each tract from make_maps)
  lh_by_region_all <- do.call(rbind, lh_dfs) %>% group_by(region) %>% 
    summarise(regional_mean_ageeffect = mean(mean_age_effect, na.rm=T))
  lh_by_region_all <- lh_by_region_all %>% mutate(coverage = ifelse(is.na(regional_mean_ageeffect), "no", "yes"))

  # right hemi
  rh_base_names <- c("COrbR_deveffect_", "CAntFrR_deveffect_", "CSupFrR_deveffect_", 
                "CMotR_deveffect_", "CSupParR_deveffect_", "CPostParR_deveffect_", 
                "CTempR_deveffect_", "COccR_deveffect_", "ARCR_deveffect_", 
                "IFOR_deveffect_", "SLFR_deveffect_", "CSTR_deveffect_", 
                "ILFR_deveffect_", "pARCR_deveffect_", "VOFR_deveffect_")

  rh_dfs <- lapply(rh_base_names, function(name) get(paste0(name, dataset)))
  rh_by_region_all <- do.call(rbind, rh_dfs) %>% group_by(region) %>% 
    summarise(regional_mean_ageeffect = mean(mean_age_effect, na.rm=T))
    rh_by_region_all <- rh_by_region_all %>% mutate(coverage = ifelse(is.na(regional_mean_ageeffect), "no", "yes"))

 
  write.csv(lh_by_region_all, sprintf("/cbica/projects/luo_wm_dev/output/%1$s/tract_profiles/tract_to_cortex/lh_agg_age_mat.csv", dataset), row.names=F)
  write.csv(rh_by_region_all, sprintf("/cbica/projects/luo_wm_dev/output/%1$s/tract_profiles/tract_to_cortex/rh_agg_age_mat.csv", dataset), row.names=F)
   
  return(list(lh_by_region_all, rh_by_region_all))
}

region_all_HCPD <- aggregate_age_maps("HCPD")  
lh_by_region_HCPD <- region_all_HCPD[[1]]
rh_by_region_HCPD <- region_all_HCPD[[2]]

region_all_HBN <- aggregate_age_maps("HBN")  
lh_by_region_HBN <- region_all_HBN[[1]]
rh_by_region_HBN <- region_all_HBN[[2]]

region_all_PNC <- aggregate_age_maps("PNC")  
lh_by_region_PNC <- region_all_PNC[[1]]
rh_by_region_PNC <- region_all_PNC[[2]]



# range(lh_by_region_HCPD$regional_mean_ageeffect, na.rm = T)
 
# range(lh_by_region_HCPD$regional_mean_ageeffect, na.rm = T)
# range(rh_by_region_HCPD$regional_mean_ageeffect, na.rm = T)
# range(lh_by_region_HBN$regional_mean_ageeffect, na.rm = T)
# range(rh_by_region_HBN$regional_mean_ageeffect, na.rm = T)
# range(lh_by_region_PNC$regional_mean_ageeffect, na.rm = T)
# range(rh_by_region_PNC$regional_mean_ageeffect, na.rm = T)
# overall range is 15 to 23 or so
```


```{r plot aggregated map, fig.height = 10, fig.width = 10}


plot_aggregated_maps <- function(lh_by_region_df, rh_by_region_df, ylim1, ylim2, dataset) {
  hemi = "left"
  cortical_pos <- c("left lateral", "left medial")
  lh_agg <- ggplot() + 
      geom_brain(data = lh_by_region_df, atlas = glasser, 
                 mapping=aes(fill = regional_mean_ageeffect, colour = coverage, size = I(0.4)), 
                 show.legend=FALSE, 
                 hemi = hemi,
                 position = position_brain(cortical_pos)) + 
    
      paletteer::scale_fill_paletteer_c("grDevices::RdYlBu", na.value = "grey100", limits = c(ylim1, ylim2), oob=squish, direction = -1) +  
      scale_colour_manual(values = c(yes = "gray30", no = "grey84")) + # limits = c(0.16, 0.4)
      theme_void() +
      theme(legend.position = "none",
            legend.key.height = unit(1, 'cm'),
            legend.key.width = unit(2.3, 'cm'),
            legend.margin=margin(0,0,0,0),
            legend.text = element_text(size=24),
            legend.title = element_blank(),
            plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
            plot.title = element_text(size=20, hjust = 0.5)) + guides(
    colour = "none",  # Remove legend for color
    size = "none"     # Remove legend for size if not needed
  )


  hemi = "right"
  cortical_pos <- c("right lateral", "right medial")
  rh_agg <- ggplot() + 
        geom_brain(data = rh_by_region_df, atlas = glasser, 
                   mapping=aes(fill = regional_mean_ageeffect, colour = coverage, size = I(0.4)), 
                   show.legend=FALSE, 
                   hemi = hemi,
                   position = position_brain(cortical_pos)) + 
      
        paletteer::scale_fill_paletteer_c("grDevices::RdYlBu", na.value = "grey100", limits = c(ylim1, ylim2), oob=squish, direction = -1) +  
        scale_colour_manual(values = c(yes = "gray30", no = "grey84")) + 
        theme_void() +
        theme(legend.position = "none",
              legend.key.height = unit(1, 'cm'),
              legend.key.width = unit(2.3, 'cm'),
              legend.margin=margin(0,0,0,0),
              legend.text = element_text(size=24),
              legend.title = element_blank(),
              plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
              plot.title = element_text(size=20, hjust = 0.5))  + guides(
      colour = "none",  # Remove legend for color
      size = "none"     # Remove legend for size if not needed
    )
  
  
  agg_plots <- ggarrange(lh_agg, rh_agg, ncol=1, nrow = 2)
  agg_final <- annotate_figure(agg_plots, top = text_grob(dataset, 
                 color = "black", size = 20, vjust = -0.75)) + theme(plot.margin = unit(c(1, 0, 0, 0), "cm")) 

  return(agg_final)
  
}
```


```{r plot aggregated maps, fig.height = 8, fig.width = 15}

agg_plot_HCPD <- plot_aggregated_maps(lh_by_region_HCPD, rh_by_region_HCPD, ylim1 = 15, ylim2 = 23, dataset = "HCP-D")
#ggsave(paste0(png_dir, "/", "HCPD", "/","dev_aggregated_age_mat.png"), agg_plot_HCPD , height = 10, width = 10, units = "in")

agg_plot_HBN <- plot_aggregated_maps(lh_by_region_HBN, rh_by_region_HBN, ylim1 = 15, ylim2 = 23, dataset = "HBN")
#ggsave(paste0(png_dir, "/", "HBN", "/","dev_aggregated_age_mat.png"), agg_plot_HBN , height = 10, width = 10, units = "in")
 
agg_plot_PNC <- plot_aggregated_maps(lh_by_region_PNC, rh_by_region_PNC, ylim1 = 15, ylim2 = 23, dataset = "PNC")
#ggsave(paste0(png_dir, "/", "PNC", "/","dev_aggregated_age_mat.png"), agg_plot_PNC , height = 10, width = 10, units = "in")

agg_plots <- ggarrange(agg_plot_HCPD, agg_plot_HBN, agg_plot_PNC, ncol = 3)


hemi = "left"
cortical_pos <- c("left lateral", "left medial")
lh_agg <- ggplot() + 
      geom_brain(data = lh_by_region_HCPD, atlas = glasser, 
                 mapping=aes(fill = regional_mean_ageeffect, colour = coverage, size = I(0.4)), 
                 show.legend=TRUE, 
                 hemi = hemi,
                 position = position_brain(cortical_pos)) + 
      paletteer::scale_fill_paletteer_c("grDevices::RdYlBu", na.value = "grey100", limits = c(15, 23), oob=squish, direction = -1) +  
      scale_colour_manual(values = c(yes = "gray30", no = "grey84")) + # limits = c(0.16, 0.4)
      theme_void() +
      theme(legend.position = "bottom",
            legend.key.height = unit(1, 'cm'),
            legend.key.width = unit(2.3, 'cm'),
            legend.margin=margin(0,0,0,0),
            legend.text = element_text(size=24),
            legend.title = element_blank(),
            plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
            plot.title = element_text(size=20, hjust = 0.5)) + guides(
    colour = "none",  # Remove legend for color
    size = "none"     # Remove legend for size if not needed
  )


legend <- get_legend(lh_agg + theme(legend.position = "bottom",
                                              legend.key.height = unit(1, 'cm'),
                                              legend.key.width = unit(2, 'cm'),
                                              legend.margin=margin(-50,0,0,0),
                                              legend.text = element_text(size=20),
                                              legend.title = element_blank()))

legend_title <- textGrob(expression(paste("Age of Maturation")), 
                   gp=gpar(col="black", fontsize=20), hjust = 0.5, vjust = -5) 
legend_with_title <- arrangeGrob(legend, legend_title, ncol = 1, nrow = 2)

agg_plots_with_legend <- ggarrange(agg_plots, legend_with_title, ncol = 1, nrow = 2)
    
agg_final <- annotate_figure(agg_plots_with_legend, top = text_grob("Aggregated Age of Maturation Map Across Tracts", 
                 color = "black", face = "bold", size = 20, vjust = -0.75)) + theme(plot.margin = unit(c(1, 0, 0, 0), "cm")) 
  
agg_final

#ggsave(paste0(png_dir, "/", "all_datasets", "/","dev_aggregated_age_mat.png"), agg_final , height = 8, width = 15, units = "in")


bilat_by_region_HCPD <- rbind(lh_by_region_HCPD, rh_by_region_HCPD)
bilat_by_region_HBN <- rbind(lh_by_region_HBN, rh_by_region_HBN)
bilat_by_region_PNC <- rbind(lh_by_region_PNC, rh_by_region_PNC)

cor.test(bilat_by_region_HCPD$regional_mean_ageeffect, bilat_by_region_HBN$regional_mean_ageeffect) # HCPD-HBN: r = 0.37, pspin = 
cor.test(bilat_by_region_HCPD$regional_mean_ageeffect, bilat_by_region_PNC$regional_mean_ageeffect) # HCPD-PNC: r = 0.80, pspin = 
cor.test(bilat_by_region_HBN$regional_mean_ageeffect, bilat_by_region_PNC$regional_mean_ageeffect) # HBN-PNC: r = 0.44, pspin = 

```
average age of maturation map across datasets
```{r avg age of mat across all dataset, fig.height = 5, fig.width = 8}

lh_HCPD_merge <- lh_by_region_HCPD %>% rename(HCPD_mean_ageeffect = regional_mean_ageeffect)
lh_HBN_merge <- lh_by_region_HBN %>% rename(HBN_mean_ageeffect = regional_mean_ageeffect)
lh_PNC_merge <- lh_by_region_PNC %>% rename(PNC_mean_ageeffect = regional_mean_ageeffect)

merge_temp <- merge(lh_HCPD_merge, lh_HBN_merge, by = "region")
lh_all_datasets_ageMat <- merge(merge_temp, lh_PNC_merge, by = "region")
lh_all_datasets_ageMat <- lh_all_datasets_ageMat %>%
  mutate(regional_mean_ageeffect = rowMeans(select(., HCPD_mean_ageeffect, HBN_mean_ageeffect, PNC_mean_ageeffect), na.rm = TRUE))


rh_HCPD_merge <- rh_by_region_HCPD %>% rename(HCPD_mean_ageeffect = regional_mean_ageeffect)
rh_HBN_merge <- rh_by_region_HBN %>% rename(HBN_mean_ageeffect = regional_mean_ageeffect)
rh_PNC_merge <- rh_by_region_PNC %>% rename(PNC_mean_ageeffect = regional_mean_ageeffect)

merge_temp <- merge(rh_HCPD_merge, rh_HBN_merge, by = "region")
rh_all_datasets_ageMat <- merge(merge_temp, rh_PNC_merge, by = "region")
rh_all_datasets_ageMat <- rh_all_datasets_ageMat %>%
  mutate(regional_mean_ageeffect = rowMeans(select(., HCPD_mean_ageeffect, HBN_mean_ageeffect, PNC_mean_ageeffect), na.rm = TRUE))

plot_aggregated_maps(lh_all_datasets_ageMat, rh_all_datasets_ageMat, ylim1 = 15, ylim2 = 23, dataset = "All Datasets")


```

binarized age of maturation map for each dataset
```{r plot aggregated binary maps, fig.height = 5, fig.width = 15}

lh_by_region_HCPD_binary <- lh_by_region_HCPD %>%
  mutate(regional_mean_ageeffect = case_when(
    regional_mean_ageeffect == max(regional_mean_ageeffect, na.rm = TRUE) ~ 0,
    regional_mean_ageeffect < max(regional_mean_ageeffect, na.rm = TRUE) ~ 1,
    TRUE ~ NA_real_
  )) 

lh_by_region_PNC_binary <- lh_by_region_PNC %>%
  mutate(regional_mean_ageeffect = case_when(
    regional_mean_ageeffect == max(regional_mean_ageeffect, na.rm = TRUE) ~ 0,
    regional_mean_ageeffect < max(regional_mean_ageeffect, na.rm = TRUE) ~ 1,
    TRUE ~ NA_real_
  )) 

lh_by_region_HBN_binary <- lh_by_region_HBN %>%
  mutate(regional_mean_ageeffect = case_when(
    regional_mean_ageeffect == max(regional_mean_ageeffect, na.rm = TRUE) ~ 0,
    regional_mean_ageeffect < max(regional_mean_ageeffect, na.rm = TRUE) ~ 1,
    TRUE ~ NA_real_
  )) 


rh_by_region_HCPD_binary <- rh_by_region_HCPD %>%
  mutate(regional_mean_ageeffect = case_when(
    regional_mean_ageeffect == max(regional_mean_ageeffect, na.rm = TRUE) ~ 0,
    regional_mean_ageeffect < max(regional_mean_ageeffect, na.rm = TRUE) ~ 1,
    TRUE ~ NA_real_
  )) 

rh_by_region_PNC_binary <- rh_by_region_PNC %>%
  mutate(regional_mean_ageeffect = case_when(
    regional_mean_ageeffect == max(regional_mean_ageeffect, na.rm = TRUE) ~ 0,
    regional_mean_ageeffect < max(regional_mean_ageeffect, na.rm = TRUE) ~ 1,
    TRUE ~ NA_real_
  )) 

rh_by_region_HBN_binary <- rh_by_region_HBN %>%
  mutate(regional_mean_ageeffect = case_when(
    regional_mean_ageeffect == max(regional_mean_ageeffect, na.rm = TRUE) ~ 0,
    regional_mean_ageeffect < max(regional_mean_ageeffect, na.rm = TRUE) ~ 1,
    TRUE ~ NA_real_
  )) 

 
color1 = "#C73000FF"
color2 = "#009DA8FF"

plot_aggregated_binary_maps <- function(lh_by_region_df, rh_by_region_df, ylim1, ylim2, dataset) {
  hemi = "left"
  cortical_pos <- c("left lateral", "left medial")
  lh_agg <- ggplot() + 
      geom_brain(data = lh_by_region_df, atlas = glasser, 
                 mapping=aes(fill = regional_mean_ageeffect, colour = coverage, size = I(0.4)), 
                 show.legend=FALSE, 
                 hemi = hemi,
                 position = position_brain(cortical_pos)) + 
    
      scale_fill_manual(values = setNames(c(color1, color2), c("0", "1")), na.value = "white") +   
      scale_colour_manual(values = c(yes = "gray30", no = "grey84")) + # limits = c(0.16, 0.4)
      theme_void() +
      theme(legend.position = "none",
            legend.key.height = unit(1, 'cm'),
            legend.key.width = unit(2.3, 'cm'),
            legend.margin=margin(0,0,0,0),
            legend.text = element_text(size=24),
            legend.title = element_blank(),
            plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
            plot.title = element_text(size=20, hjust = 0.5)) + guides(
    colour = "none",  # Remove legend for color
    size = "none"     # Remove legend for size if not needed
  )


  hemi = "right"
  cortical_pos <- c("right lateral", "right medial")
  rh_agg <- ggplot() + 
        geom_brain(data = rh_by_region_df, atlas = glasser, 
                   mapping=aes(fill = regional_mean_ageeffect, colour = coverage, size = I(0.4)), 
                   show.legend=FALSE, 
                   hemi = hemi,
                   position = position_brain(cortical_pos)) + 
      
       scale_fill_manual(values = setNames(c(color1, color2), c("0", "1")), na.value = "white") +   
        scale_colour_manual(values = c(yes = "gray30", no = "grey84")) + 
        theme_void() +
        theme(legend.position = "bottom",
              legend.key.height = unit(1, 'cm'),
              legend.key.width = unit(2.3, 'cm'),
              legend.margin=margin(0,0,0,0),
              legend.text = element_text(size=24),
              legend.title = element_blank(),
              plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
              plot.title = element_text(size=20, hjust = 0.5))  + guides(
      colour = "none",  # Remove legend for color
      size = "none"     # Remove legend for size if not needed
    )
  
  
  agg_plots <- ggarrange(lh_agg, rh_agg, ncol=1, nrow = 2)
  agg_final <- annotate_figure(agg_plots, top = text_grob(dataset, 
                 color = "black", size = 20, vjust = -0.75)) + theme(plot.margin = unit(c(1, 0, 0, 0), "cm")) 

  return(agg_final)
  
}

  
lh_by_region_HCPD_binary$regional_mean_ageeffect <- as.character(lh_by_region_HCPD_binary$regional_mean_ageeffect)
rh_by_region_HCPD_binary$regional_mean_ageeffect <- as.character(rh_by_region_HCPD_binary$regional_mean_ageeffect)

lh_by_region_PNC_binary$regional_mean_ageeffect <- as.character(lh_by_region_PNC_binary$regional_mean_ageeffect)
rh_by_region_PNC_binary$regional_mean_ageeffect <- as.character(rh_by_region_PNC_binary$regional_mean_ageeffect)


lh_by_region_HBN_binary$regional_mean_ageeffect <- as.character(lh_by_region_HBN_binary$regional_mean_ageeffect)
rh_by_region_HBN_binary$regional_mean_ageeffect <- as.character(rh_by_region_HBN_binary$regional_mean_ageeffect)

agg_binary_plot_HCPD <- plot_aggregated_binary_maps(lh_by_region_HCPD_binary, rh_by_region_HCPD_binary, ylim1 = 15, ylim2 = 23, dataset = "HCP-D")
#ggsave(paste0(png_dir, "/", "HCPD", "/","dev_agg_binaryregated_age_mat.png"), agg_binary_plot_HCPD , height = 10, width = 10, units = "in")

agg_binary_plot_HBN <- plot_aggregated_binary_maps(lh_by_region_HBN_binary, rh_by_region_HBN_binary, ylim1 = 15, ylim2 = 23, dataset = "HBN")
#ggsave(paste0(png_dir, "/", "HBN", "/","dev_agg_binaryregated_age_mat.png"), agg_binary_plot_HBN , height = 10, width = 10, units = "in")
 
agg_binary_plot_PNC <- plot_aggregated_binary_maps(lh_by_region_PNC_binary, rh_by_region_PNC_binary, ylim1 = 15, ylim2 = 23, dataset = "PNC")
#ggsave(paste0(png_dir, "/", "PNC", "/","dev_agg_binaryregated_age_mat.png"), agg_binary_plot_PNC , height = 10, width = 10, units = "in")

agg_binary_plots <- ggarrange(agg_binary_plot_HCPD, agg_binary_plot_HBN, agg_binary_plot_PNC, ncol = 3)
agg_binary_plots

# red = not matured
# blue = matured

cor.test(as.numeric(lh_by_region_HCPD_binary$regional_mean_ageeffect), as.numeric(lh_by_region_HBN_binary$regional_mean_ageeffect))
cor.test(as.numeric(lh_by_region_HCPD_binary$regional_mean_ageeffect), as.numeric(lh_by_region_PNC_binary$regional_mean_ageeffect))
cor.test(as.numeric(lh_by_region_HBN_binary$regional_mean_ageeffect), as.numeric(lh_by_region_PNC_binary$regional_mean_ageeffect))


cor.test(as.numeric(rh_by_region_HCPD_binary$regional_mean_ageeffect), as.numeric(rh_by_region_HBN_binary$regional_mean_ageeffect))
cor.test(as.numeric(rh_by_region_HCPD_binary$regional_mean_ageeffect), as.numeric(rh_by_region_PNC_binary$regional_mean_ageeffect))
cor.test(as.numeric(rh_by_region_HBN_binary$regional_mean_ageeffect), as.numeric(rh_by_region_PNC_binary$regional_mean_ageeffect))
```



# Avg difference in S-A axis rank between tract ends vs. Difference in age effect between tract ends (delta-delta)
```{r load SA axis}
# i'm bringing SA back (yeah)
glasser_SAaxis <- read.csv("/cbica/projects/luo_wm_dev/SAaxis/glasser_SAaxis.csv")
glasser_SAaxis <- glasser_SAaxis %>% select(SA.axis_rank, label)
glasser_SAaxis$regionName <- gsub("_ROI", "", glasser_SAaxis$label)
glasser_SAaxis$regionName <- gsub("^(.)_(.*)$", "\\2_\\1", glasser_SAaxis$region)
```


```{r functions for computing SA rank}

# for each tract's endpoint, compute the mean S-A axis rank
merge_SAaxis <- function(bundle_name, df_to_return = "endpoints_only", dataset, perm_SAaxis = NULL) {
  df <- get(paste0(bundle_name, "_deveffect_", dataset))
  if(is.null(perm_SAaxis)) {
      merged_df <- merge(df, glasser_SAaxis,  by = "regionName")
  } else if (!is.null(perm_SAaxis)) {
     merged_df <- cbind(df, perm_SAaxis)
  }
  merged_df <- merged_df %>% arrange(thresh_probability)
  
  mean_df <- merged_df %>% group_by(end) %>% summarise(mean_SA = mean(SA.axis_rank, na.rm = T),
                                                       median_SA = median(SA.axis_rank, na.rm = T),
                                                      # sd_SA = sd(SA.axis_rank, na.rm = T),
                                                       age_effect = mean(mean_age_effect)) %>% mutate(bundle_name = bundle_name)
  
  if(df_to_return == "all_regions") {
    return(merged_df)
  } else {
    return(mean_df)
  }
}


compute_endpoint_diffs <- function(bundle_name, lh_all_endpoints, rh_all_endpoints) {
 
  callosum <- c("COrb", "CAntFr", "CSupFr", "CMot", "CSupPar", "CPostPar", "CTemp", "COcc")

  if (any(sapply(callosum, function(x) grepl(x, bundle_name)))) {
    bundle_type = "callosum"
  } else {
    bundle_type = "assoc"
  }
  
  # for assocation bundles, get the difference in age of maturation and diff in mean SA
  if(bundle_type == "assoc") {
    bundle_name_string <- paste0(bundle_name) # ugh R idiosyncracies. 
    if(grepl("L$", bundle_name)) {
      SAranks_endpoints <- lh_all_endpoints %>% filter(bundle_name == bundle_name_string) # can't have bundle_name == bundle_name bc it confuses dplyr
    } else{
      SAranks_endpoints <- rh_all_endpoints  %>% filter(bundle_name == bundle_name_string)
    }
    
    endpoints_diffs <- SAranks_endpoints %>% filter(if_all(everything(), ~ !is.na(.))) %>%
        reframe(mean_SA_diff = mean_SA[end == "end1"] - mean_SA[end == "end2"],
          age_effect_diff = age_effect[end == "end1"] - age_effect[end == "end2"]) %>% 
      mutate(bundle_name = bundle_name)
    
    # for callosum bundles, get the difference in age of maturation and diff in mean SA
  } else if (bundle_type == "callosum") {
    bundle_name_lh =  paste0(substr(bundle_name, 1, nchar(bundle_name) - 1), "L")
    bundle_name_rh =  paste0(substr(bundle_name, 1, nchar(bundle_name) - 1), "R")
    
    SAranks_endpoints1 <- lh_all_endpoints %>% filter(bundle_name == bundle_name_lh)  
    SAranks_endpoints2 <- rh_all_endpoints %>% filter(bundle_name == bundle_name_rh) 
    SAranks_endpoints <- rbind(SAranks_endpoints1, SAranks_endpoints2)
    
    endpoints_diffs <- SAranks_endpoints %>% filter(if_all(everything(), ~ !is.na(.))) %>%
        reframe(mean_SA_diff = mean_SA[end == "end1"] - mean_SA[end == "end2"],
          age_effect_diff = age_effect[end == "end1"] - age_effect[end == "end2"]) %>% 
      mutate(bundle_name = bundle_name)
  }
  
 return(endpoints_diffs)
}

# compute mean SA rank for each cortical endpoint
compute_mean_SA <- function(dataset, SAaxis.perm = NULL) {
  lh_names <- c("ARCL", "CSTL" ,"ILFL", "IFOL", "SLFL", "pARCL", "VOFL", 
                "COrbL", "CAntFrL" ,"CSupFrL", "CMotL", "CSupParL", "CPostParL", "CTempL", "COccL")
  if (is.null(SAaxis.perm)) {
    lh_SAranks <- lapply(lh_names, merge_SAaxis, "all_regions", dataset=dataset)
    lh_SAranks_endpoints <- lapply(lh_names, merge_SAaxis, "endpoints_only", dataset=dataset)
  } else if (!is.null(SAaxis.perm)) {
    lh_SAranks <- lapply(lh_names, merge_SAaxis, perm_SAaxis = data.frame(SAaxis.perm) %>% setNames("SA.axis_rank"), 
                       df_to_return = "all_regions", dataset = dataset)
    lh_SAranks_endpoints <- lapply(lh_names, merge_SAaxis, perm_SAaxis = data.frame(SAaxis.perm) %>% setNames("SA.axis_rank"), 
                                 df_to_return = "endpoints_only", dataset=dataset) 
  }
  lh_SAranks_endpoints <- lapply(lh_SAranks_endpoints, function(df) na.omit(df))
   
  rh_names <- c("ARCR", "CSTR" ,"ILFR", "IFOR", "SLFR", "pARCR", "VOFR", 
                "COrbR", "CAntFrR" ,"CSupFrR", "CMotR", "CSupParR", "CPostParR", "CTempR", "COccR")
  if (is.null(SAaxis.perm)) {
    rh_SAranks <- lapply(rh_names, merge_SAaxis, "all_regions", dataset=dataset)
    rh_SAranks_endpoints <- lapply(rh_names, merge_SAaxis, "endpoints_only", dataset=dataset)
  } else if (!is.null(SAaxis.perm)) {
    rh_SAranks <- lapply(rh_names, merge_SAaxis, perm_SAaxis = data.frame(SAaxis.perm) %>% setNames("SA.axis_rank"), 
                       df_to_return = "all_regions", dataset = dataset)
    rh_SAranks_endpoints <- lapply(rh_names, merge_SAaxis, perm_SAaxis = data.frame(SAaxis.perm) %>% setNames("SA.axis_rank"), 
                                 df_to_return = "endpoints_only", dataset=dataset) 
  }
  rh_SAranks_endpoints <- lapply(rh_SAranks_endpoints, function(df) na.omit(df))
  
  names(lh_SAranks) <- lh_names
  names(rh_SAranks) <- rh_names
  
  names(lh_SAranks_endpoints) <- lh_names
  names(rh_SAranks_endpoints) <- rh_names
   
  lh_all_endpoints <- bind_rows(lh_SAranks_endpoints) %>% filter(!is.na(end))
  rh_all_endpoints <- bind_rows(rh_SAranks_endpoints) %>% filter(!is.na(end))
  
  all_endpoints <- rbind(lh_all_endpoints, rh_all_endpoints) %>% arrange(bundle_name)
  
  return(list(lh_all_endpoints, rh_all_endpoints, all_endpoints))
  
}



plot_meanSA_by_age_mat <- function(dataset, annot_text) {
  all_endpoints <- get(paste0("all_endpoints_", dataset))
  title = gsub("PD", "P-D", dataset) # for HCPD
  SA_plot <- ggplot(all_endpoints, aes(x = mean_SA, y = age_effect, label = bundle_name)) +
  geom_point(aes(fill = mean_SA, color = mean_SA, shape = end), size = 4) + 
    paletteer::scale_fill_paletteer_c("grDevices::RdYlBu", direction = -1, 
                                      limits = c(min(all_endpoints$mean_SA), max(all_endpoints$mean_SA)), oob = squish) + 
    paletteer::scale_color_paletteer_c("grDevices::RdYlBu", direction = -1,
                                       limits = c(min(all_endpoints$mean_SA), max(all_endpoints$mean_SA)), oob = squish) +
    scale_shape_manual(values = c(19,1)) + 
  geom_smooth(data = all_endpoints, method='lm', se=TRUE, fill=alpha(c("gray70"),.9), col="black") + 
  #stat_cor(method = "pearson", label.x = 25, label.y = 26, size = 6) + 
  #ggrepel::geom_text_repel(position = position_jitter(seed = 3)) + # Add labels
    ggrepel::geom_text_repel(
    position = position_jitter(seed = 3), 
    box.padding = 0.5,  # Increase space around labels
    point.padding = 0.5,  # Increase space around points
    segment.color = "gray",  # Line color connecting point to label
    segment.size = 0.5,  # Line thickness
    max.overlaps = Inf 
  )  + 
  annotate(geom="text", x=100, y=26, label=annot_text, color="black", size=5) + 
  labs(title = title) + theme_classic() +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
    axis.title = element_blank(),
    axis.text.x = element_text(size = 20),
    axis.text.y = element_text(size = 20)) + xlim(15, 340) + ylim(14, 26) + guides(shape = guide_legend("Endpoint"), color = FALSE, fill = FALSE) 
  return(SA_plot)
}
```

THE MONEY IS WITH AGE OF MATURATION YALLLLLL
```{r compute SA rank}

all_endpoints_HCPD <- compute_mean_SA("HCPD")  
lh_all_endpoints_HCPD <- all_endpoints_HCPD[[1]]
rh_all_endpoints_HCPD <- all_endpoints_HCPD[[2]]
all_endpoints_HCPD <- all_endpoints_HCPD[[3]]

all_endpoints_HBN <- compute_mean_SA("HBN")  
lh_all_endpoints_HBN <- all_endpoints_HBN[[1]]
rh_all_endpoints_HBN <- all_endpoints_HBN[[2]]
all_endpoints_HBN <- all_endpoints_HBN[[3]]

all_endpoints_PNC <- compute_mean_SA("PNC")  
lh_all_endpoints_PNC <- all_endpoints_PNC[[1]]
rh_all_endpoints_PNC <- all_endpoints_PNC[[2]]
all_endpoints_PNC <- all_endpoints_PNC[[3]]
 
  
```


```{r spin test permute SA axis recompute difference}

source("/cbica/projects/luo_wm_dev/software/spin_test/perm.sphere.p.R")
perm.id.full <- readRDS("/cbica/projects/luo_wm_dev/software/spin_test/rotate_parcellation/glasser.coords_sphericalrotations_N10k.rds")

# spin test: spin the S-A axis then recompute average S-A rank for each end. 
# spun p-value for t-test
# @param SAaxis, vector of S-A axis ranks
perm.sphere.SAaxis <- function(SAaxis, perm.id, dataset, spun_ttest = FALSE, alternative = "greater", var.equal = FALSE) {

  nroi = dim(perm.id)[1]  # number of regions
  nperm = dim(perm.id)[2] # number of permutations
  
  # spin SA axis: permutation of measures
  SAaxis.perm = array(NA,dim=c(nroi,nperm))
  for (r in 1:nperm) {
    for (i in 1:nroi) {
      SAaxis.perm[i,r] = SAaxis[perm.id[i,r]]
    }
  }

  # if just want to compute spun p-value for age of maturation vs mean S-A rank for all cortical endpoints: 
  if (spun_ttest == FALSE) { 
    # empirical correlation
    all_endpoints <- get(paste0("all_endpoints_", dataset))
    rho.emp = cor(all_endpoints$mean_SA, all_endpoints$age_effect, method = "pearson", use="complete.obs")  
    
    # correlation between permuted S-A axis and age of maturation
    rho.null.SAaxis = vector(length=nperm)
    for (r in 1:nperm) {
      print(r)
      # merging of permuted S-A axis to age of maturation df
      perm_mean_SAaxis <- compute_mean_SA(dataset, SAaxis.perm[,r]) # merge permuted S-A rank with age of maturation AND compute mean permuted S-A ranks
      perm_mean_SAaxis <- perm_mean_SAaxis[[3]] # get all_endpoints df
      rho.null.SAaxis[r] = cor(perm_mean_SAaxis$mean_SA, perm_mean_SAaxis$age_effect, method="pearson", use="complete.obs")
    }
     # p-value definition depends on the sign of the empirical correlation
    if (rho.emp>0) {
      p.perm = sum(rho.null.SAaxis>rho.emp)/nperm
    } else { 
      p.perm = sum(rho.null.SAaxis<rho.emp)/nperm
    } 
    return(list(p.perm = p.perm, rho.emp = rho.emp))
    # return p-value for the correlation between age of maturation vs. mean S-A rank for cortical endpoint
    
  } else if(spun_ttest == TRUE) {
   
    all_endpoints <- get(paste0("all_endpoints_", dataset))
    all_endpoints_binary <- all_endpoints
    max_value <- max(all_endpoints_binary$age_effect, na.rm = TRUE)
    all_endpoints_binary$age_effect[all_endpoints_binary$age_effect == max_value] <- NA
    all_endpoints_binary <- all_endpoints_binary %>% mutate(maturation_status = ifelse(is.na(age_effect), 0, 1)) # 0 = not matured, 1 = matured
    
    # empirical correlation
    rho.emp = cor(all_endpoints_binary$mean_SA, all_endpoints_binary$age_effect, method = "spearman", use="complete.obs")  
    
    # correlation between permuted S-A axis and age of maturation
    rho.null.SAaxis = vector(length=nperm)
    for (r in 1:nperm) {
      print(r)
      # merging of permuted S-A axis to age of maturation df
      perm_mean_SAaxis <- compute_mean_SA(dataset, SAaxis.perm[,r]) # merge permuted S-A rank with age of maturation AND compute mean permuted S-A ranks
      perm_mean_SAaxis <- perm_mean_SAaxis[[3]] # get all_endpoints df
      max_value_perm <- max(perm_mean_SAaxis$age_effect, na.rm = TRUE)
      perm_mean_SAaxis$age_effect[perm_mean_SAaxis$age_effect == max_value_perm] <- NA
      rho.null.SAaxis[r] = cor(perm_mean_SAaxis$mean_SA, perm_mean_SAaxis$age_effect, method="spearman", use="complete.obs")
    }
     # p-value definition depends on the sign of the empirical correlation
    if (rho.emp>0) {
      p.perm.cor = sum(rho.null.SAaxis>rho.emp)/nperm
    } else { np
      p.perm.cor = sum(rho.null.SAaxis<rho.emp)/nperm
    } 
    
    # empirical t value
    t.emp <- t.test(all_endpoints_binary$mean_SA[which(all_endpoints_binary$maturation_status == 0)], # compare differences in sa rank
                  all_endpoints_binary$mean_SA[which(all_endpoints_binary$maturation_status == 1)], 
                  alternative = alternative, var.equal = var.equal)$statistic # hypothesis is non-matured tract ends have greater S-A rank
    
    
    # t-test between permuted S-A axis differences and age of maturation differences between endpoints
    t.null.SAaxis = vector(length=nperm)
    for (r in 1:nperm) {
      print(paste0(r, " t-test"))
      # merging of permuted S-A axis to age of maturation df
      perm_mean_SAaxis <- compute_mean_SA(dataset, SAaxis.perm[,r]) # merge permuted S-A rank with age of maturation AND compute mean permuted S-A ranks
      perm_mean_SAaxis <- perm_mean_SAaxis[[3]] # get all_endpoints df
      max_value_perm <- max(perm_mean_SAaxis$age_effect, na.rm = TRUE)
      perm_mean_SAaxis$age_effect[perm_mean_SAaxis$age_effect == max_value_perm] <- NA
      perm_mean_SAaxis <- perm_mean_SAaxis %>% mutate(maturation_status = ifelse(is.na(age_effect), 0, 1)) # 0 = not matured, 1 = matured
       
      # calculate t-test and permuted t-test: do tract ends that have NOT matured between endpoints actually have greater s-a ranks than tract ends that have matured?? 
      t.null.SAaxis[r] <- t.test(perm_mean_SAaxis$mean_SA[which(all_endpoints_binary$maturation_status == 0)], 
                                 perm_mean_SAaxis$mean_SA[which(all_endpoints_binary$maturation_status == 1)], 
                                 alternative = alternative, var.equal = var.equal)$statistic
    }
    
    # p-value definition  
    if (t.emp>0) {
      p.perm.t = sum(t.null.SAaxis>t.emp)/nperm
    } else { 
      p.perm.t= sum(t.null.SAaxis<t.emp)/nperm
    } 
  
    return(list(rho.emp = rho.emp, p.perm.cor = p.perm.cor, t.emp = t.emp, p.perm.t = p.perm.t))
  }
}
 

# spin test for delta-delta: spin the S-A axis then recompute average S-A rank for each end. Then calculate difference in rank for delta-delta analysis. Then compute t.test.
# spun p-value for t-test
# @param SAaxis, vector of S-A axis ranks
perm.sphere.SAaxis_delta <- function(SAaxis, perm.id, dataset, alternative = "less", var.equal = FALSE) {

  nroi = dim(perm.id)[1]  # number of regions
  nperm = dim(perm.id)[2] # number of permutations
  
  # spin SA axis: permutation of measures
  SAaxis.perm = array(NA,dim=c(nroi,nperm))
  for (r in 1:nperm) {
    for (i in 1:nroi) {
      SAaxis.perm[i,r] = SAaxis[perm.id[i,r]]
    }
  }

  # if want to compute spun p-value for the delta-delta plot: 
  
  # empirical t value
  diffs_emp <- get(paste0("diffs_", dataset))
  diffs_emp <- diffs_emp %>% mutate(group = case_when(age_effect_diff < 3 ~ "Small Differences in Age of Maturation",
                                         age_effect_diff > 3 ~ "Large Differences Age of Maturation",
                                         TRUE ~ NA_character_))
  t.emp <- t.test(diffs_emp$mean_SA_diff[which(diffs_emp$group == "Small Differences in Age of Maturation")], # compare differences in sa rank
                  diffs_emp$mean_SA_diff[which(diffs_emp$group == "Large Differences Age of Maturation")], 
                  alternative = alternative, var.equal = var.equal)$statistic 
  
  # t-test between permuted S-A axis differences and age of maturation differences between endpoints
  t.null.SAaxis = vector(length=nperm)
  for (r in 1:nperm) {
    print(r)
    # merging of permuted S-A axis to age of maturation df
    perm_mean_SAaxis <- compute_mean_SA(dataset, SAaxis.perm[,r]) # merge permuted S-A rank with age of maturation AND compute mean permuted S-A ranks
    lh_perm_mean_SAaxis <- perm_mean_SAaxis[[1]]
    rh_perm_mean_SAaxis <- perm_mean_SAaxis[[2]]
    
    # calculate difference in rank
    perm_diffs <- compute_diffs_wrapper(dataset, lh_perm_mean_SAaxis, rh_perm_mean_SAaxis)
    perm_diffs <- perm_diffs %>% mutate(group = case_when(age_effect_diff < 3 ~ "Small Differences in Age of Maturation",
                                         age_effect_diff > 3 ~ "Large Differences Age of Maturation",
                                         TRUE ~ NA_character_))
    
    # calculate t-test and permuted t-test: do tracts with very diff ages of maturation between endpoints actually have different s-a ranks?? 
    t.null.SAaxis[r] <- t.test(perm_diffs$mean_SA_diff[which(perm_diffs$group == "Small Differences in Age of Maturation")], 
                               perm_diffs$mean_SA_diff[which(perm_diffs$group == "Large Differences Age of Maturation")], 
                               alternative = alternative, var.equal = var.equal )$statistic
  }
  

  # p-value definition  
  if (t.emp>0) {
    p.perm = sum(t.null.SAaxis>t.emp)/nperm
  } else { 
    p.perm = sum(t.null.SAaxis<t.emp)/nperm
  } 

  return(list(p.perm = p.perm, t.emp = t.emp))
  # return p-value for the t-test to see if tracts with very diff ages of maturation between endpoints actually have different s-a rank
  
}
 
  
```
 

```{r plot meanSA by age of mat, fig.height = 5, fig.width = 15}

perm.sphere.SAaxis(glasser_SAaxis$SA.axis_rank, perm.id.full, "HCPD") # pspin = 0, r = 0.87 
perm.sphere.SAaxis(glasser_SAaxis$SA.axis_rank, perm.id.full, "HBN") # pspin = NS, r - 0.17
perm.sphere.SAaxis(glasser_SAaxis$SA.axis_rank, perm.id.full, "PNC") # pspin = 0, r = 0.8141278


SA_age_mat_HCPD <- plot_meanSA_by_age_mat("HCPD", expression(paste(italic("r"), " = 0.87, ",  , italic(p[spin]), "< 0.0001")))
SA_age_mat_HBN <- plot_meanSA_by_age_mat("HBN", expression(paste(italic("r"), " = 0.17, ",  , italic(p[spin]), "= N.S.")))
SA_age_mat_PNC <- plot_meanSA_by_age_mat("PNC", expression(paste(italic("r"), " = 0.81, ",  , italic(p[spin]), "< 0.0001")))

SA_age_mat_plot_final <- ggarrange(SA_age_mat_HCPD, SA_age_mat_HBN, SA_age_mat_PNC, ncol = 3, common.legend = T, legend = "bottom")

x.grob <- textGrob("Mean S-A Axis Rank of Cortical Endpoint", 
                   gp=gpar(col="black", fontsize=20))

y.grob <- textGrob("Age of Maturation (Years)", 
                   gp=gpar(col="black", fontsize=20), rot=90)

grid.arrange(arrangeGrob(SA_age_mat_plot_final, left = y.grob, bottom = x.grob))

#ggsave(paste0(png_dir, "/", "all_datasets", "/","SA_age_mat_plot.png"), 
      # grid.arrange(arrangeGrob(SA_age_mat_plot_final, left = y.grob, bottom = x.grob)), height = 5, width = 15, units = "in")



```


```{r plot meanSA by mat status ttest and spearman, fig.height = 5, fig.width = 15}
 
perm.sphere.SAaxis(glasser_SAaxis$SA.axis_rank, spun_ttest = TRUE, perm.id.full, "HCPD") # spearman r = 0.5872033, pspin = 0.0273; t = 9.566846 , pspin = 0
perm.sphere.SAaxis(glasser_SAaxis$SA.axis_rank, spun_ttest = TRUE, perm.id.full, "HBN")  # spearman r = 0.5371881, pspin = 0.0026; t = -0.4384347, pspin = 0.3002
perm.sphere.SAaxis(glasser_SAaxis$SA.axis_rank, spun_ttest = TRUE, perm.id.full, "PNC")  # spearman r = 0.6561538, pspin = 0; t = 5.61031, pspin = 0


 
#ggsave(paste0(png_dir, "/", "all_datasets", "/","SA_age_mat_plot.png"), 
      # grid.arrange(arrangeGrob(SA_age_mat_plot_final, left = y.grob, bottom = x.grob)), height = 5, width = 15, units = "in")



```

???
```{r SA by age of maturation parcel level}

merge_SA_parcel <- function(dataset) {
  lh_by_region <- get(paste0("lh_by_region_", dataset))
  rh_by_region <- get(paste0("rh_by_region_", dataset))
  
  lh_by_region$region <- paste0(lh_by_region$region, "_L")
  rh_by_region$region <- paste0(rh_by_region$region, "_R")
  
  aggregated_df <- rbind(lh_by_region, rh_by_region) %>% mutate(regionName = region)
  aggregated_axis <- merge(glasser_SAaxis, aggregated_df, by = "regionName")
  
  return(aggregated_axis)
}

aggregated_axis_HCPD <- merge_SA_parcel("HCPD") # 360 glasser parcels with mean age maturation and SA rank)
aggregated_axis_HBN <- merge_SA_parcel("HBN")   
aggregated_axis_PNC <- merge_SA_parcel("PNC")   



source("/cbica/projects/luo_wm_dev/software/spin_test/perm.sphere.p.R")
perm.id.full <- readRDS("/cbica/projects/luo_wm_dev/software/spin_test/rotate_parcellation/glasser.coords_sphericalrotations_N10k.rds")

cor.test(aggregated_axis_HCPD$SA.axis_rank, aggregated_axis_HCPD$regional_mean_ageeffect) # r = 0.76, pspin = 0
cor.test(aggregated_axis_HBN$SA.axis_rank, aggregated_axis_HBN$regional_mean_ageeffect) # r = 0.18, pspin = 0.0338
cor.test(aggregated_axis_PNC$SA.axis_rank, aggregated_axis_PNC$regional_mean_ageeffect) # r = 0.72, pspin = 0
 
 
perm.sphere.p(x = aggregated_axis_HCPD$SA.axis_rank, y = aggregated_axis_HCPD$regional_mean_ageeffect, perm.id = perm.id.full, corr.type = "pearson") 
perm.sphere.p(x = aggregated_axis_HBN$SA.axis_rank, y = aggregated_axis_HBN$regional_mean_ageeffect, perm.id = perm.id.full, corr.type = "pearson") 
perm.sphere.p(x = aggregated_axis_PNC$SA.axis_rank, y = aggregated_axis_PNC$regional_mean_ageeffect, perm.id = perm.id.full, corr.type = "pearson") 



# Remove parcels where age of maturation is the max age (meaning that node/parcel never reached maturation in our age window)
aggregated_axis_HCPD_binary <- aggregated_axis_HCPD
max_value <- max(aggregated_axis_HCPD_binary$regional_mean_ageeffect, na.rm = TRUE)
aggregated_axis_HCPD_binary$regional_mean_ageeffect[aggregated_axis_HCPD_binary$regional_mean_ageeffect == max_value] <- NA


aggregated_axis_HBN_binary <-aggregated_axis_HBN
max_value <- max(aggregated_axis_HBN_binary$regional_mean_ageeffect, na.rm = TRUE)
aggregated_axis_HBN_binary$regional_mean_ageeffect[aggregated_axis_HBN_binary$regional_mean_ageeffect == max_value] <- NA


aggregated_axis_PNC_binary <-aggregated_axis_PNC
max_value <- max(aggregated_axis_PNC_binary$regional_mean_ageeffect, na.rm = TRUE)
aggregated_axis_PNC_binary$regional_mean_ageeffect[aggregated_axis_PNC_binary$regional_mean_ageeffect == max_value] <- NA


perm.sphere.p(x = aggregated_axis_HCPD_binary$SA.axis_rank, y = aggregated_axis_HCPD_binary$regional_mean_ageeffect, perm.id = perm.id.full, corr.type = "pearson") 
perm.sphere.p(x = aggregated_axis_HBN_binary$SA.axis_rank, y = aggregated_axis_HBN_binary$regional_mean_ageeffect, perm.id = perm.id.full, corr.type = "pearson") 
perm.sphere.p(x = aggregated_axis_PNC_binary$SA.axis_rank, y = aggregated_axis_PNC_binary$regional_mean_ageeffect, perm.id = perm.id.full, corr.type = "pearson") 


cor.test(aggregated_axis_HCPD_binary$SA.axis_rank, aggregated_axis_HCPD_binary$regional_mean_ageeffect) # r = 0.57, pspin = 0
cor.test(aggregated_axis_HBN_binary$SA.axis_rank, aggregated_axis_HBN_binary$regional_mean_ageeffect) # r = 0.46, pspin = 0.0001
cor.test(aggregated_axis_PNC_binary$SA.axis_rank, aggregated_axis_PNC_binary$regional_mean_ageeffect) # r = 0.68, pspin = 0.0001

```


```{r SA by age of mat parcel plot, fig.height = 5, fig.width = 15}

annot_text_HCPD <- expression(paste(italic("r"), " = 0.76, ",  , italic(p[spin]), "< 0.0001"))
annot_text_HBN <- expression(paste(italic("r"), " = 0.18, ",  , italic(p[spin]), "= 0.0338"))
annot_text_PNC <- expression(paste(italic("r"), " = 0.72, ",  , italic(p[spin]), "< 0.0001"))

annot_text_HCPD <- expression(paste(italic("r"), " = 0.60, ",  , italic(p[spin]), "= 0.0001"))
annot_text_HBN <- expression(paste(italic("r"), " = 0.40, ",  , italic(p[spin]), "< 0.0001"))
annot_text_PNC <- expression(paste(italic("r"), " = 0.73, ",  , italic(p[spin]), "< 0.0001"))

plot_agg_SA_parcel <- function(dataset, annot_text) {
  aggregated_axis <- get(paste0("aggregated_axis_", dataset, "_binary"))
  #aggregated_axis <- get(paste0("aggregated_axis_", dataset))
  title = gsub("PD", "P-D", dataset) # for HCPD
  SA_plot <- ggplot(aggregated_axis, aes(x = SA.axis_rank, y = regional_mean_ageeffect, fill = SA.axis_rank)) +
  geom_point(color = "gray", shape = 21, size=3.5) + 
    paletteer::scale_fill_paletteer_c("grDevices::RdYlBu", direction = -1, 
                                      limits = c(min(aggregated_axis$SA.axis_rank), max(aggregated_axis$SA.axis_rank)), oob = squish) +
  geom_smooth(data = aggregated_axis, method='lm', se=TRUE, fill=alpha(c("gray70"),.9), col="black") + 
  labs(title = title) + theme_classic() +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
    axis.title = element_blank(),
    axis.text.x = element_text(size = 20),
    axis.text.y = element_text(size = 20)) + 
    annotate(geom="text", x=150, y=25, label=annot_text, color="black", size=7) +
    xlim(15, 340) + ylim(14, 26)
  return(SA_plot)
}

 
agg_SA_parcel_HCPD <- plot_agg_SA_parcel("HCPD", annot_text_HCPD)
agg_SA_parcel_HBN <- plot_agg_SA_parcel("HBN", annot_text_HBN)
agg_SA_parcel_PNC <- plot_agg_SA_parcel("PNC", annot_text_PNC)

agg_SA_parcel_plot_final <- ggarrange(agg_SA_parcel_HCPD, agg_SA_parcel_HBN, agg_SA_parcel_PNC, ncol = 3)

x.grob <- textGrob("S-A Axis Rank", 
                   gp=gpar(col="black", fontsize=20))

y.grob <- textGrob("Age of Maturation (Years)", 
                   gp=gpar(col="black", fontsize=20), rot=90)

grid.arrange(arrangeGrob(agg_SA_parcel_plot_final, left = y.grob, bottom = x.grob))

ggsave(paste0(png_dir, "/", "all_datasets", "/","SA_age_mat_plot_parcel_max.png"), 
       grid.arrange(arrangeGrob(SA_age_mat_plot_final, left = y.grob, bottom = x.grob)), height = 5, width = 15, units = "in")
# i dont think i can do a spin test on this very easily though - this is just the mean SA rank of each endpoint correlated with the age of maturation at that endpoint
```

```{r, compute mean diffs, fig.height = 5, fig.width = 15}
# compute mean difference in S-A rank between the 2 cortical endpoints by the difference in age effect
# regular difference

lh_names <- c("ARCL", "CSTL" ,"ILFL", "IFOL", "SLFL", "pARCL", "VOFL", "COrbL", "CAntFrL" ,"CSupFrL", "CMotL", "CSupParL", "CPostParL", "CTempL", "COccL")
rh_names <- c("ARCR", "CSTR" ,"ILFR", "IFOR", "SLFR", "pARCR", "VOFR", "COrbR", "CAntFrR" ,"CSupFrR", "CMotR", "CSupParR", "CPostParR", "CTempR", "COccR")

compute_diffs_wrapper <- function(dataset, lh_all_endpoints, rh_all_endpoints) {
  # difference in age of maturation
  lh_diffs <- lapply(lh_names, compute_endpoint_diffs, lh_all_endpoints, rh_all_endpoints)
  rh_diffs <- lapply(rh_names, compute_endpoint_diffs, lh_all_endpoints, rh_all_endpoints)
  names(lh_diffs) <- lh_names
  names(rh_diffs) <- rh_names
   
 
  lh_diffs_all <- bind_rows(lh_diffs) %>%
    filter(if_all(everything(), ~ !is.na(.)))
  rh_diffs_all <- bind_rows(rh_diffs) %>%
    filter(if_all(everything(), ~ !is.na(.)))
  diffs_all <- rbind(lh_diffs_all, rh_diffs_all)
  diffs_all <- diffs_all %>% filter(!grepl("^C.*R$", bundle_name)) # remove redundant callosum bundles (keep just left hemi callosum bundles to show difference between rh and lh endpoints)
  diffs_all$bundle_name <- gsub("^(C.*)L$", "\\1", diffs_all$bundle_name) # rename callosum bundles to NOT have hemisphere
  # make a column to label IFO and ILF, and label Callosum Bundles and Association Bundles - for plotting
  diffs_all <- diffs_all %>% mutate(ifo_ilf_label = case_when(bundle_name == "IFOL" ~ "Left Inf. Fronto-occ.",
                                                              bundle_name == "IFOR" ~ "Right Inf. Fronto-occ.",
                                                              bundle_name == "ILFL" ~ "Left Inf. Long.",
                                                              bundle_name == "ILFR" ~ "Right Inf. Long.",
                                                              TRUE ~ NA_character_)) %>%
                            mutate(label = case_when(str_detect(bundle_name, "^C") ~ "Callosum", TRUE ~ "Association"))
  
  return(diffs_all)

}

diffs_HCPD <- compute_diffs_wrapper("HCPD", lh_all_endpoints_HCPD, rh_all_endpoints_HCPD)  
diffs_HBN <- compute_diffs_wrapper("HBN", lh_all_endpoints_HBN, rh_all_endpoints_HBN)   
diffs_PNC <- compute_diffs_wrapper("PNC", lh_all_endpoints_PNC, rh_all_endpoints_PNC)  

color1 = "#0A595E"
color2 = "#24CEB9"

plot_diffs <- function(dataset, color1, color2) {
  
  diffs_all <- get(paste0("diffs_", dataset))
  title = gsub("PD", "P-D", dataset) # for HCPD
  
  diffs_plot <- ggplot(diffs_all, aes(x = mean_SA_diff, y = age_effect_diff, label = ifo_ilf_label)) +
  geom_point(size = 5, aes(fill = label, color = label), shape = 21, alpha = 0.9) +  # Plot points
  scale_fill_manual(values = c("Callosum" = color1, "Association" = color2)) +
    scale_color_manual(values = c("Callosum" = color1, "Association" = color2)) +
  ggrepel::geom_text_repel(position = position_jitter(seed = 3), 
    box.padding = 0.8,  # Increase space around labels
    point.padding = 0.8,  # Increase space around points
    segment.color = "gray",  # Line color connecting point to label
    segment.size = 0.5,
    size = 6) +  
  geom_smooth(data = diffs_all, method='lm', se=TRUE, fill=alpha(c("gray70"),.7), col="black") + 
  stat_cor(method = "pearson", label.x = -50, label.y = 10, size = 6) + 
  labs(title = title) + theme_classic() +
  theme(legend.position = "none",
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_blank(),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20)) + ylim(-5, 10) + xlim(-65, 255)
 
  return(diffs_plot)
}

 
 
perm.sphere.SAaxis(glasser_SAaxis$SA.axis_rank, perm.id.full[,c(1:1000)], "HCPD", compute_diffs = TRUE, corr.type = "pearson") # pspin =  0
perm.sphere.SAaxis(glasser_SAaxis$SA.axis_rank, perm.id.full[,c(1:1000)], "HBN", compute_diffs = TRUE, corr.type = "pearson") # pspin =  NS
perm.sphere.SAaxis(glasser_SAaxis$SA.axis_rank, perm.id.full[,c(1:1000)], "PNC", compute_diffs = TRUE, corr.type = "pearson") # pspin =  0.001


diffs_plot_HCPD <- plot_diffs("HCPD", color1, color2)
diffs_plot_HBN <- plot_diffs("HBN", color1, color2)
diffs_plot_PNC <- plot_diffs("PNC", color1, color2)

diffs_plot_final <- ggarrange(diffs_plot_HCPD, diffs_plot_HBN, diffs_plot_PNC, ncol = 3, common.legend = T, legend = "top")

x.grob <- textGrob("Difference in Mean S-A Rank between Endpoints", 
                   gp=gpar(col="black", fontsize=20))

y.grob <- textGrob("Difference in Age of Maturation (Years) \n between Endpoints", 
                   gp=gpar(col="black", fontsize=20), rot=90)

grid.arrange(arrangeGrob(diffs_plot_final, left = y.grob, bottom = x.grob))

#ggsave(paste0(png_dir, "/","all_datasets", "/","delta_delta_diff.png"), 
 #      grid.arrange(arrangeGrob(diffs_plot_final, left = y.grob, bottom = x.grob)), height = 6, width = 15, units = "in")
 
 
cor.test(diffs_HCPD$mean_SA_diff, diffs_HCPD$age_effect_diff) # r = 0.90, p = 0 
cor.test(diffs_HBN$mean_SA_diff, diffs_HBN$age_effect_diff) # r = 0.015, p = N.S
cor.test(diffs_PNC$mean_SA_diff, diffs_PNC$age_effect_diff) # r = 0.85, p = 0
 


```


```{r ttest scrap}

diffs_HCPD <- diffs_HCPD %>% mutate(SA_group = case_when(mean_SA_diff < 100 ~ "Small Difference in S-A Rank",
                                           mean_SA_diff > 100 ~ "Large Difference in S-A Rank",
                                           TRUE ~ NA_character_))

t.test(diffs_HCPD$age_effect_diff[which(diffs_HCPD$SA_group == "Small Difference in S-A Rank")], diffs_HCPD$age_effect_diff[which(diffs_HCPD$SA_group == "Large Difference in S-A Rank")], )
 


diffs_HBN <- diffs_HBN %>% mutate(SA_group = case_when(mean_SA_diff < 100 ~ "Small Difference in S-A Rank",
                                           mean_SA_diff > 100 ~ "Large Difference in S-A Rank",
                                           TRUE ~ NA_character_))

t.test(diffs_HBN$age_effect_diff[which(diffs_HBN$SA_group == "Small Difference in S-A Rank")], diffs_HBN$age_effect_diff[which(diffs_HBN$SA_group == "Large Difference in S-A Rank")], )
 


diffs_PNC <- diffs_PNC %>% mutate(SA_group = case_when(mean_SA_diff < 100 ~ "Small Difference in S-A Rank",
                                           mean_SA_diff > 100 ~ "Large Difference in S-A Rank",
                                           TRUE ~ NA_character_))

t.test(diffs_PNC$age_effect_diff[which(diffs_PNC$SA_group == "Small Difference in S-A Rank")], diffs_PNC$age_effect_diff[which(diffs_PNC$SA_group == "Large Difference in S-A Rank")])$statistic
 


# spun p-value for t-test
perm.sphere.p.ttest = function(x,y,perm.id, alternative = "less", var.equal = FALSE) {
  
  nroi = dim(perm.id)[1]  # number of regions
  nperm = dim(perm.id)[2] # number of permutations
  
  t.emp = t.test(x,y, alternative = alternative, var.equal = var.equal)$statistic  # empirical t value
  
  # permutation of measures
  x.perm = y.perm = array(NA,dim=c(nroi,nperm))
  for (r in 1:nperm) {
    for (i in 1:nroi) {
      x.perm[i,r] = x[perm.id[i,r]]
      y.perm[i,r] = y[perm.id[i,r]]
    }
  }
  # correlation to unpermuted measures
  t.null.xy = t.null.yx = vector(length=nperm)
  for (r in 1:nperm) {
    t.null.xy[r] = t.test(x.perm[,r],y, alternative = alternative, var.equal = var.equal)$statistic
    t.null.yx[r] = t.test(y.perm[,r],x, alternative = alternative, var.equal = var.equal)$statistic
  }
  
  # p-value definition depends on the sign of the empirical correlation
  if (t.emp>0) {
    p.perm.xy = sum(t.null.xy>t.emp)/nperm
    p.perm.yx = sum(t.null.yx>t.emp)/nperm
  } else { 
    p.perm.xy = sum(t.null.xy<t.emp)/nperm
    p.perm.yx = sum(t.null.yx<t.emp)/nperm
  } 
  
  # return average p-value
  return((p.perm.xy+p.perm.yx)/2)
  
}
 
 

perm.sphere.p.ttest(x = diffs_HCPD$age_effect_diff[which(diffs_HCPD$SA_group == "Small Difference in S-A Rank")], 
                    y = diffs_HCPD$age_effect_diff[which(diffs_HCPD$SA_group == "Large Difference in S-A Rank")], perm.id = perm.id.full, alternative = "two.sided") # p = 0



perm.sphere.p.ttest(x = diffs_HBN$age_effect_diff[which(diffs_HBN$SA_group == "Small Difference in S-A Rank")], 
                    y = diffs_HBN$age_effect_diff[which(diffs_HBN$SA_group == "Large Difference in S-A Rank")], perm.id = perm.id.full, alternative = "two.sided") # 

perm.sphere.p.ttest(x = diffs_PNC$age_effect_diff[which(diffs_PNC$SA_group == "Small Difference in S-A Rank")], 
                    y = diffs_PNC$age_effect_diff[which(diffs_PNC$SA_group == "Large Difference in S-A Rank")], perm.id = perm.id.full) # p = 0
```


t1/t2 downloaded from https://balsa.wustl.edu/study/P2DmK
alff maps downloaded from https://github.com/PennLINC/spatiotemp_dev_plasticity/tree/main/cortical_maps 

```{r compare aggregated map with other maps}
# t1/t2 and alff
t1t2_ageeffect_HCPD <- read_cifti("/cbica/projects/luo_wm_dev/input/cortical_data/t1t2_hcpd/Figure3_T1wT2w_sAge_partial_bayes_r2.pscalar.nii")
t1t2_mean_HCPD <- read_cifti("/cbica/projects/luo_wm_dev/input/cortical_data/t1t2_hcpd/Figure1_group_mean_T1wT2w_myelin.pscalar.nii")
t1t2_roc_HCPD <- read_cifti("/cbica/projects/luo_wm_dev/input/cortical_data/t1t2_hcpd/Figure4_mean_posterior_annualized_roc_T1wT2w.pscalar.nii")
t1t2_maxslopeage_HCPD <- read_cifti("/cbica/projects/luo_wm_dev/input/cortical_data/t1t2_hcpd/Figure6_median_posterior_age_of_max_slope_T1wT2w.pscalar.nii")

alff_pnc <- read_cifti("/cbica/projects/luo_wm_dev/input/cortical_data/alff_pnc/AgeEffect_PartialR2.pscalar.nii")
alff_decline_pnc <- read_cifti("/cbica/projects/luo_wm_dev/input/cortical_data/alff_pnc/AgeofDeclineOnset_FirstNegDeriv.pscalar.nii")
```
 
 
```{r agg glasser label}
# add glasser labels to aggregated age effect maps
combine_glasser <- function(dataset) {
  tolerance <- 1e-6 
  lh_by_region <- get(paste0("lh_by_region_", dataset))
  #lh_max_value <- max(lh_by_region$regional_mean_ageeffect, na.rm = TRUE) # remove max age of maturation -- means node never matured
  #lh_by_region$regional_mean_ageeffect[abs(lh_by_region$regional_mean_ageeffect - lh_max_value) < tolerance] <- NA

  rh_by_region <- get(paste0("rh_by_region_", dataset))
  #rh_max_value <- max(rh_by_region$regional_mean_ageeffect, na.rm = TRUE)
  #rh_by_region$regional_mean_ageeffect[abs(rh_by_region$regional_mean_ageeffect - rh_max_value) < tolerance] <- NA
 
  lh_glasser_labels <- glasser_labels[which(glasser_labels$LR == "L"),]
  lh_by_region$region <- gsub("_[LR]+", "",lh_by_region$region)
  lh_agg_final <- merge(lh_by_region, lh_glasser_labels, by = "region")
  lh_agg_final <- lh_agg_final %>% mutate(label = gsub("(.*)(_L)$", "L_\\1", regionName))  
  lh_agg_final$label <- paste0(lh_agg_final$label, "_ROI")
   
  rh_glasser_labels <- glasser_labels[which(glasser_labels$LR == "R"),]
  rh_by_region$region <- gsub("_[LR]+", "",rh_by_region$region)
  rh_agg_final <- merge(rh_by_region, rh_glasser_labels, by = "region")
  rh_agg_final <- rh_agg_final %>% mutate(label = gsub("(.*)(_R)$", "R_\\1", regionName))  
  rh_agg_final$label <- paste0(rh_agg_final$label, "_ROI")
  
  agg <- rbind(rh_agg_final, lh_agg_final)
  
  agg_ageeffects <- agg %>%
    mutate(label = factor(label, levels = names(t1t2_ageeffect_HCPD$Parcel))) %>%
    arrange(label) %>% mutate(dataset = dataset) 
  
  assign(paste0("agg_ageeffects_", dataset), agg_ageeffects, envir = .GlobalEnv)
  
  #write.csv(agg_ageeffects, sprintf("/cbica/projects/luo_wm_dev/output/%1$s/tract_profiles/tract_to_cortex/%1$s_agg_ageeffects_ordered_all.csv", dataset), row.names = F)

}

combine_glasser("HCPD") # makes agg_ageeffects_HCPD
combine_glasser("HBN")  # makes agg_ageeffects_HBN
combine_glasser("PNC")  # makes agg_ageeffects_PNC
```
 
 
```{r agg glasser label binary}
# add glasser labels to aggregated age effect maps
combine_glasser_binary <- function(dataset) {
  tolerance <- 1e-6 
  lh_by_region <- get(paste0("lh_by_region_", dataset))
  lh_max_value <- max(lh_by_region$regional_mean_ageeffect, na.rm = TRUE) # remove max age of maturation -- means node never matured
  lh_by_region$regional_mean_ageeffect[abs(lh_by_region$regional_mean_ageeffect - lh_max_value) < tolerance] <- NA

  rh_by_region <- get(paste0("rh_by_region_", dataset))
  rh_max_value <- max(rh_by_region$regional_mean_ageeffect, na.rm = TRUE)
  rh_by_region$regional_mean_ageeffect[abs(rh_by_region$regional_mean_ageeffect - rh_max_value) < tolerance] <- NA
 
  lh_glasser_labels <- glasser_labels[which(glasser_labels$LR == "L"),]
  lh_by_region$region <- gsub("_[LR]+", "",lh_by_region$region)
  lh_agg_final <- merge(lh_by_region, lh_glasser_labels, by = "region")
  lh_agg_final <- lh_agg_final %>% mutate(label = gsub("(.*)(_L)$", "L_\\1", regionName))  
  lh_agg_final$label <- paste0(lh_agg_final$label, "_ROI")
   
  rh_glasser_labels <- glasser_labels[which(glasser_labels$LR == "R"),]
  rh_by_region$region <- gsub("_[LR]+", "",rh_by_region$region)
  rh_agg_final <- merge(rh_by_region, rh_glasser_labels, by = "region")
  rh_agg_final <- rh_agg_final %>% mutate(label = gsub("(.*)(_R)$", "R_\\1", regionName))  
  rh_agg_final$label <- paste0(rh_agg_final$label, "_ROI")
  
  agg <- rbind(rh_agg_final, lh_agg_final)
  
  agg_ageeffects <- agg %>%
    mutate(label = factor(label, levels = names(t1t2_ageeffect_HCPD$Parcel))) %>%
    arrange(label) %>% mutate(dataset = dataset) 
  
  assign(paste0("agg_ageeffects_", dataset, "_binary"), agg_ageeffects, envir = .GlobalEnv)
  
  #write.csv(agg_ageeffects, sprintf("/cbica/projects/luo_wm_dev/output/%1$s/tract_profiles/tract_to_cortex/%1$s_agg_ageeffects_ordered_all.csv", dataset), row.names = F)

}

combine_glasser_binary("HCPD") # makes agg_ageeffects_HCPD_binary
combine_glasser_binary("HBN")  # makes agg_ageeffects_HBN_binary
combine_glasser_binary("PNC")  # makes agg_ageeffects_PNC_binary
```
 
```{r spin test}

source("/cbica/projects/luo_wm_dev/software/spin_test/perm.sphere.p.R")

perm.id.full <- readRDS("/cbica/projects/luo_wm_dev/software/spin_test/rotate_parcellation/glasser.coords_sphericalrotations_N10k.rds")
perm.id.1000 <- perm.id.full[,c(1: 1000)]
 
perm.sphere.p(x = agg_ageeffects_HCPD$regional_mean_ageeffect, y = t1t2_ageeffect_HCPD$data, perm.id = perm.id.full, corr.type = "pearson") # 0.0033 
perm.sphere.p(x = agg_ageeffects_HCPD$regional_mean_ageeffect, y = t1t2_mean_HCPD$data, perm.id = perm.id.full, corr.type = "pearson") # 0.00035   
perm.sphere.p(x = agg_ageeffects_HCPD$regional_mean_ageeffect, y = t1t2_roc_HCPD$data, perm.id = perm.id.full, corr.type = "pearson")  # 0.00015
perm.sphere.p(x = agg_ageeffects_HCPD$regional_mean_ageeffect, y = t1t2_maxslopeage_HCPD$data, perm.id = perm.id.full, corr.type = "pearson") # 0.0452
perm.sphere.p(x = agg_ageeffects_HCPD$regional_mean_ageeffect, y = alff_pnc$data, perm.id = perm.id.full, corr.type = "pearson")  # 0.12225 
perm.sphere.p(x = agg_ageeffects_HCPD$regional_mean_ageeffect, y = alff_decline_pnc$data, perm.id = perm.id.full, corr.type = "pearson")  # 0.1367
 
perm.sphere.p(x = agg_ageeffects_HBN$regional_mean_ageeffect, y = t1t2_ageeffect_HCPD$data, perm.id = perm.id.full, corr.type = "pearson") # 0.3652   
perm.sphere.p(x = agg_ageeffects_HBN$regional_mean_ageeffect, y = t1t2_mean_HCPD$data, perm.id = perm.id.full, corr.type = "pearson")  # 0.47365
perm.sphere.p(x = agg_ageeffects_HBN$regional_mean_ageeffect, y = t1t2_roc_HCPD$data, perm.id = perm.id.full, corr.type = "pearson") #  0.3955
perm.sphere.p(x = agg_ageeffects_HBN$regional_mean_ageeffect, y = t1t2_maxslopeage_HCPD$data, perm.id = perm.id.full, corr.type = "pearson")  # 0.07095
perm.sphere.p(x = agg_ageeffects_HBN$regional_mean_ageeffect, y = alff_pnc$data, perm.id = perm.id.full, corr.type = "pearson")  # 0.3258
perm.sphere.p(x = agg_ageeffects_HBN$regional_mean_ageeffect, y = alff_decline_pnc$data, perm.id = perm.id.full, corr.type = "pearson") # 0.0436

perm.sphere.p(x = agg_ageeffects_PNC$regional_mean_ageeffect, y = t1t2_ageeffect_HCPD$data, perm.id = perm.id.full, corr.type = "pearson")  # 0.0068  
perm.sphere.p(x = agg_ageeffects_PNC$regional_mean_ageeffect, y = t1t2_mean_HCPD$data, perm.id = perm.id.full, corr.type = "pearson")  # 0.01445
perm.sphere.p(x = agg_ageeffects_PNC$regional_mean_ageeffect, y = t1t2_roc_HCPD$data, perm.id = perm.id.full, corr.type = "pearson")   # 1e-04
perm.sphere.p(x = agg_ageeffects_PNC$regional_mean_ageeffect, y = t1t2_maxslopeage_HCPD$data, perm.id = perm.id.full, corr.type = "pearson")  # 0.0011
perm.sphere.p(x = agg_ageeffects_PNC$regional_mean_ageeffect, y = alff_pnc$data, perm.id = perm.id.full, corr.type = "pearson")   # 0.00895
perm.sphere.p(x = agg_ageeffects_PNC$regional_mean_ageeffect, y = alff_decline_pnc$data, perm.id = perm.id.full, corr.type = "pearson")   # 0.0053
 


# pvalues are pre-fdr correction
cor.test(t1t2_ageeffect_HCPD$data, agg_ageeffects_HCPD$regional_mean_ageeffect, method = 'pearson')  ##
cor.test(t1t2_ageeffect_HCPD$data, agg_ageeffects_HBN$regional_mean_ageeffect, method = 'pearson')
cor.test(t1t2_ageeffect_HCPD$data, agg_ageeffects_PNC$regional_mean_ageeffect, method = 'pearson')
# age of maturation vs. t1/t2 partial rsq
# hcpd: r = -0.63, pspin = 0.0033 
# hbn: r = -0.16, N.S.
# pnc: r = -0.61, pspin = 0.0068  


#cor.test(t1t2_mean_HCPD$data, agg_ageeffects$regional_mean_ageeffect, method = 'pearson')   ## 
cor.test(t1t2_mean_HCPD$data, agg_ageeffects_HCPD$regional_mean_ageeffect, method = 'pearson')  
cor.test(t1t2_mean_HCPD$data, agg_ageeffects_HBN$regional_mean_ageeffect, method = 'pearson')  
cor.test(t1t2_mean_HCPD$data, agg_ageeffects_PNC$regional_mean_ageeffect, method = 'pearson')  
# age of maturation vs. t1/t2 group mean
# hcpd: r = -0.61, pspin = 0.00035  
# hbn: r = 0.03, N.S.
# pnc: r = -0.49, pspin = 0.01445

#cor.test(t1t2_roc_HCPD$data, agg_ageeffects$regional_mean_ageeffect, method = 'pearson') ##
cor.test(t1t2_roc_HCPD$data, agg_ageeffects_HCPD$regional_mean_ageeffect, method = 'pearson') 
cor.test(t1t2_roc_HCPD$data, agg_ageeffects_HBN$regional_mean_ageeffect, method = 'pearson') 
cor.test(t1t2_roc_HCPD$data, agg_ageeffects_PNC$regional_mean_ageeffect, method = 'pearson') 
# age of maturation vs. t1/t2 mean posterior annualized rate of change
# hcpd: r = -0.60, pspin = 0.00015
# hbn: r = -0.09, N.S.
# pnc: r = -0.59, pspin = 1e-04

#cor.test(t1t2_maxslopeage_HCPD$data, agg_ageeffects$regional_mean_ageeffect, method = 'pearson') ##
cor.test(t1t2_maxslopeage_HCPD$data, agg_ageeffects_HCPD$regional_mean_ageeffect, method = 'pearson')
cor.test(t1t2_maxslopeage_HCPD$data, agg_ageeffects_HBN$regional_mean_ageeffect, method = 'pearson')
cor.test(t1t2_maxslopeage_HCPD$data, agg_ageeffects_PNC$regional_mean_ageeffect, method = 'pearson')
# age of maturation vs t1/t2 age of max slope
# hcpd: r = 0.54, pspin = 0.0452
# hbn: r = 0.47, pspin = 0.07095
# pnc: r = 0.63, pspin = 0.0011

#cor.test(alff_pnc$data, agg_ageeffects$regional_mean_ageeffect, method = 'pearson')  ##
cor.test(alff_pnc$data, agg_ageeffects_HCPD$regional_mean_ageeffect, method = 'pearson') 
cor.test(alff_pnc$data, agg_ageeffects_HBN$regional_mean_ageeffect, method = 'pearson') 
cor.test(alff_pnc$data, agg_ageeffects_PNC$regional_mean_ageeffect, method = 'pearson') 
# age of maturation vs. alff age effect
# hcpd: r = 0.38, pspin = N.S
# hbn: r = 0.17, N.S.
# pnc: r = 0.52, pspin = 0.00895

#cor.test(alff_decline_pnc$data, agg_ageeffects$regional_mean_ageeffect, method = 'pearson')  
cor.test(alff_decline_pnc$data, agg_ageeffects_HCPD$regional_mean_ageeffect, method = 'pearson') 
cor.test(alff_decline_pnc$data, agg_ageeffects_HBN$regional_mean_ageeffect, method = 'pearson') 
cor.test(alff_decline_pnc$data, agg_ageeffects_PNC$regional_mean_ageeffect, method = 'pearson') 
# age of maturation vs. alff age of decline
# hcpd: r = 0.43, pspin = N.S.
# hbn: r = 0.49, pspin = 0.0436
# pnc: r = 0.59, pspin = 0.0053

```



```{r to delete t1t2 and alff, fig.width = 15, fig.height = 5}

color1 = "#3FB8AFFF"
color2 = "#FF9E9DFF"
color3 = "#cc0468"

plot_annot_maps <- function(dataset, map, annot_text, annot_x, annot_y, ylim1, ylim2) {
  map <- get(map)$data # i.e. t1t2_ageeffect_HCPD
  map <- data.frame(map) %>% setNames("map_measure")
  agg_ageeffects <- get(paste0("agg_ageeffects_", dataset))
  agg_map <- cbind(agg_ageeffects, map)
  
  title = gsub("PD", "P-D", dataset) # for HCPD
  
  if(dataset == "HCPD") {
    color = color1
  } else if (dataset=="HBN") {
    color = color2
  } else {
    color = color3
  }
  
  agg_plot <- ggplot(agg_map, aes(x = regional_mean_ageeffect, y = map_measure)) +
  geom_point(size = 5, fill = color, color = color, shape = 21, alpha = 0.6) +  
  geom_smooth(data = agg_map, method='lm', se=TRUE, fill=alpha(c("gray70"),.7), col="black") + 
  #stat_cor(method = "spearman", size = 6) + 
  labs(title = title) + theme_classic() +
  theme(legend.position = "none",
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_blank(),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20)) +
  annotate(geom="text", x=annot_x, y=annot_y, label=annot_text, color="black", size=7) + xlim(15, 23) + ylim(ylim1, ylim2)
 
  return(agg_plot)
}

 
HCPD_t1t2_ageeffect <- plot_annot_maps("HCPD", "t1t2_ageeffect_HCPD", 
                                       expression(paste(italic("r"), " = -0.71, ",  , italic(p[spin]), "= 0.00045")),
                                       annot_x = 18, annot_y = 0.35, ylim1 = 0, ylim2 = 0.35) 
HCPD_t1t2_mean <- plot_annot_maps("HCPD", "t1t2_mean_HCPD", 
                                       expression(paste(italic("r"), " = -0.67, ",  , italic(p[spin]), "= 0.0003")),
                                       annot_x = 18, annot_y = 2, ylim1 = 1.3, ylim2 = 2) 
HCPD_t1t2_roc <- plot_annot_maps("HCPD", "t1t2_roc_HCPD", 
                                       expression(paste(italic("r"), " = -0.72, ",  , italic(p[spin]), "= 0.0001")),
                                       annot_x = 18, annot_y = 0.023, ylim1 = -0.0018, ylim2 = 0.025) 
HCPD_t1t2_maxslopeage <- plot_annot_maps("HCPD", "t1t2_maxslopeage_HCPD", 
                                       expression(paste(italic("r"), " = 0.57, ",  , italic(p[spin]), "= 0.05915")),
                                       annot_x = 18, annot_y = 17.6, ylim1 = 8.5, ylim2 = 19) 
HCPD_alff_ageeffect <- plot_annot_maps("HCPD", "alff_pnc", 
                                       expression(paste(italic("r"), " = 0.51, ",  , italic(p[spin]), "= 0.03815")),
                                       annot_x = 18, annot_y = 0.05, ylim1 = -0.15, ylim2 = 0.06) 

HCPD_alff_agedecline <- plot_annot_maps("HCPD", "alff_decline_pnc", 
                                       expression(paste(italic("r"), " = 0.48, ",  , italic(p[spin]), "= 0.117")),
                                       annot_x = 18, annot_y = 20, ylim1 = 8, ylim2 = 23) 

 
  
HBN_t1t2_ageeffect <- plot_annot_maps("HBN", "t1t2_ageeffect_HCPD",
                                      expression(paste(italic("r"), " = -0.1, ",  , italic(p[spin]), "= N.S")),
                                      annot_x = 18, annot_y = 0.35, ylim1 = 0, ylim2 = 0.35) 
HBN_t1t2_mean <- plot_annot_maps("HBN", "t1t2_mean_HCPD", 
                                       expression(paste(italic("r"), " = 0.05, ",  , italic(p[spin]), "= N.S")),
                                       annot_x = 18, annot_y = 2, ylim1 = 1.3, ylim2 = 2) 
HBN_t1t2_roc <- plot_annot_maps("HBN", "t1t2_roc_HCPD", 
                                       expression(paste(italic("r"), " = -0.09, ",  , italic(p[spin]), "= N.S")),
                                       annot_x = 18, annot_y = 0.023, ylim1 =  -0.0018, ylim2 = 0.025) 
HBN_t1t2_maxslopeage <- plot_annot_maps("HBN", "t1t2_maxslopeage_HCPD", 
                                       expression(paste(italic("r"), " = 0.51, ",  , italic(p[spin]), "= 0.05")),
                                       annot_x = 18, annot_y = 17.6, ylim1 = 8.5, ylim2 = 19) 
HBN_alff_ageeffect <- plot_annot_maps("HBN", "alff_pnc", 
                                       expression(paste(italic("r"), " = 0.15, ",  , italic(p[spin]), "= N.S.")),
                                       annot_x = 18, annot_y = 0.05, ylim1 = -0.15, ylim2 = 0.06) 
HBN_alff_agedecline <- plot_annot_maps("HBN", "alff_decline_pnc", 
                                       expression(paste(italic("r"), " = 0.51, ",  , italic(p[spin]), "= 0.03815")),
                                       annot_x = 18, annot_y = 0.05, ylim1 = -0.15, ylim2 = 0.06) 

PNC_t1t2_ageeffect <- plot_annot_maps("PNC", "t1t2_ageeffect_HCPD",
                                      expression(paste(italic("r"), " = -0.64, ",  , italic(p[spin]), "= 0.0047")),
                                      annot_x = 18, annot_y = 0.35, ylim1 = 0, ylim2 = 0.35) 
PNC_t1t2_mean <- plot_annot_maps("PNC", "t1t2_mean_HCPD", 
                                       expression(paste(italic("r"), " = -0.49, ",  , italic(p[spin]), "= 0.0428")),
                                       annot_x = 18, annot_y = 2, ylim1 = 1.3, ylim2 = 2) 
PNC_t1t2_roc <- plot_annot_maps("PNC", "t1t2_roc_HCPD", 
                                       expression(paste(italic("r"), " = -0.68, ",  , italic(p[spin]), "= 0.00015")),
                                       annot_x = 18, annot_y = 0.023, ylim1 = -0.0018, ylim2 = 0.025) 
PNC_t1t2_maxslopeage <- plot_annot_maps("PNC", "t1t2_maxslopeage_HCPD", 
                                       expression(paste(italic("r"), " = 0.66, ",  , italic(p[spin]), "= 0.00275")),
                                       annot_x = 18, annot_y = 17.6, ylim1 = 8.5, ylim2 = 19) 
PNC_alff_ageeffect <- plot_annot_maps("PNC", "alff_pnc", 
                                       expression(paste(italic("r"), " = 0.58, ",  , italic(p[spin]), "= 0.0031")),
                                       annot_x = 18, annot_y = 0.05, ylim1 = -0.15, ylim2 = 0.06) 

PNC_alff_agedecline <- plot_annot_maps("PNC", "alff_decline_pnc", 
                                       expression(paste(italic("r"), " = 0.51, ",  , italic(p[spin]), "= 0.03815")),
                                       annot_x = 18, annot_y = 0.05, ylim1 = -0.15, ylim2 = 0.06) 

x.grob <- textGrob("WM Cortical Endpoint Age of Maturation (Years)", 
                   gp=gpar(col="black", fontsize=20))

y.grob <- textGrob("T1w/T2w Partial R-squared (Age Effect)", 
                   gp=gpar(col="black", fontsize=20), rot=90)
title.grob <- textGrob("T1w/T2w Age Effect", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
t1t2_ageeffect_plots <- ggarrange(HCPD_t1t2_ageeffect, HBN_t1t2_ageeffect, PNC_t1t2_ageeffect, ncol = 3)
grid.arrange(arrangeGrob(t1t2_ageeffect_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","t1t2_ageeffect_plots.png"), grid.arrange(arrangeGrob(t1t2_ageeffect_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 15, units = "in")

y.grob <- textGrob("Mean T1w/T2w", 
                   gp=gpar(col="black", fontsize=20), rot=90)
title.grob <- textGrob("Group T1w/T2w Map", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
t1t2_mean_plots <- ggarrange(HCPD_t1t2_mean, HBN_t1t2_mean, PNC_t1t2_mean, ncol = 3)
grid.arrange(arrangeGrob(t1t2_mean_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","t1t2_mean_plots.png"), grid.arrange(arrangeGrob(t1t2_mean_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 15, units = "in")

y.grob <- textGrob("T1w/T2w Annualized Rate of Change", 
                   gp=gpar(col="black", fontsize=20), rot=90)
title.grob <- textGrob("T1w/T2w Growth Rate", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
t1t2_roc_plots <- ggarrange(HCPD_t1t2_roc, HBN_t1t2_roc, PNC_t1t2_roc, ncol = 3)
grid.arrange(arrangeGrob(t1t2_roc_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","t1t2_roc_plots.png"), grid.arrange(arrangeGrob(t1t2_roc_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 15, units = "in")

y.grob <- textGrob("T1w/T2w Age of \nMaximum Growth Rate", 
                   gp=gpar(col="black", fontsize=20), rot=90)
title.grob <- textGrob("T1w/T2w Age of Peak Growth", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
t1t2_maxslopeage_plots <- ggarrange(HCPD_t1t2_maxslopeage, HBN_t1t2_maxslopeage, PNC_t1t2_maxslopeage, ncol = 3)
grid.arrange(arrangeGrob(t1t2_maxslopeage_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","t1t2_maxslopeage_plots.png"), grid.arrange(arrangeGrob(t1t2_maxslopeage_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 15, units = "in")

y.grob <- textGrob("ALFF Partial R-squared (Age Effect)", 
                   gp=gpar(col="black", fontsize=20), rot=90)
title.grob <- textGrob("ALFF Age Effect", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
alff_ageeffect_plots <- ggarrange(HCPD_alff_ageeffect, HBN_alff_ageeffect, PNC_alff_ageeffect, ncol = 3)
grid.arrange(arrangeGrob(alff_ageeffect_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","alff_ageeffect.png"), grid.arrange(arrangeGrob(alff_ageeffect_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 15, units = "in")

```


```{r correct plot_annot_maps, fig.height = 5, fig.width = 17}

color1 = "#3FB8AFFF"
color2 = "#FF9E9DFF"
color3 = "#cc0468"

plot_annot_maps <- function(dataset, map, annot_text, annot_x, annot_y, xlim1, xlim2) {
  map <- get(map)$data # i.e. t1t2_ageeffect_HCPD
  map <- data.frame(map) %>% setNames("map_measure")
  agg_ageeffects <- get(paste0("agg_ageeffects_", dataset))
  agg_map <- cbind(agg_ageeffects, map)
  
  title = gsub("PD", "P-D", dataset) # for HCPD
  
  if(dataset == "HCPD") {
    color = color1
  } else if (dataset=="HBN") {
    color = color2
  } else {
    color = color3
  }
  
  agg_plot <- ggplot(agg_map, aes(x = map_measure, y = regional_mean_ageeffect)) +
  geom_point(size = 5, fill = color, color = color, shape = 21, alpha = 0.6) +  
  geom_smooth(data = agg_map, method='lm', se=TRUE, fill=alpha(c("gray70"),.7), col="black") + 
  #stat_cor(method = "spearman", size = 6) + 
  labs(title = title) + theme_classic() +
  theme(legend.position = "none",
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_blank(),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20)) +
  annotate(geom="text", x=annot_x, y=annot_y, label=annot_text, color="black", size=7) + ylim(15, 24) + xlim(xlim1, xlim2)
 
  return(agg_plot)
}

 
HCPD_t1t2_ageeffect <- plot_annot_maps("HCPD", "t1t2_ageeffect_HCPD", 
                                       expression(paste(italic("r"), " = -0.63, ",  , italic(p[spin]), "= 0.0033")),
                                       annot_y =  23.5, annot_x = 0.15, xlim1 = 0, xlim2 = 0.35) 
HCPD_t1t2_mean <- plot_annot_maps("HCPD", "t1t2_mean_HCPD", 
                                       expression(paste(italic("r"), " = -0.61, ",  , italic(p[spin]), "= 0.00035")),
                                       annot_y =  23.5, annot_x = 1.6, xlim1 = 1.3, xlim2 = 2) 
HCPD_t1t2_roc <- plot_annot_maps("HCPD", "t1t2_roc_HCPD", 
                                       expression(paste(italic("r"), " = -0.60, ",  , italic(p[spin]), "= 0.00015")),
                                       annot_y =  23.5, annot_x = 0.01, xlim1 = 0, xlim2 = 0.025) 
HCPD_t1t2_maxslopeage <- plot_annot_maps("HCPD", "t1t2_maxslopeage_HCPD", 
                                       expression(paste(italic("r"), " = 0.54, ",  , italic(p[spin]), "= 0.0452")),
                                       annot_y =  23.5, annot_x = 12.5, xlim1 = 8.5, xlim2 = 19) 
HCPD_alff_ageeffect <- plot_annot_maps("HCPD", "alff_pnc", 
                                       expression(paste(italic("r"), " = 0.38, ",  , italic(p[spin]), "= N.S")),
                                       annot_y =  23.5, annot_x = -0.05, xlim1 = -0.15, xlim2 = 0.06) 
HCPD_alff_agedecline <- plot_annot_maps("HCPD", "alff_decline_pnc", 
                                       expression(paste(italic("r"), " = 0.43, ",  , italic(p[spin]), "= N.S")),
                                       annot_x = 12, annot_y = 23.5, xlim1 = 8, xlim2 = 16) 

 
HBN_t1t2_ageeffect <- plot_annot_maps("HBN", "t1t2_ageeffect_HCPD",
                                      expression(paste(italic("r"), " = -0.16, ",  , italic(p[spin]), "= N.S")),
                                      annot_y =  23.5, annot_x = 0.15, xlim1 = 0, xlim2 = 0.35) 
HBN_t1t2_mean <- plot_annot_maps("HBN", "t1t2_mean_HCPD", 
                                       expression(paste(italic("r"), " = 0.03, ",  , italic(p[spin]), "= N.S")),
                                         annot_y =  23.5, annot_x = 1.6, xlim1 = 1.3, xlim2 = 2) 
HBN_t1t2_roc <- plot_annot_maps("HBN", "t1t2_roc_HCPD", 
                                       expression(paste(italic("r"), " = -0.09, ",  , italic(p[spin]), "= N.S")),
                                       annot_y =  23.5, annot_x = 0.01, xlim1 = 0, xlim2 = 0.025) 
HBN_t1t2_maxslopeage <- plot_annot_maps("HBN", "t1t2_maxslopeage_HCPD", 
                                       expression(paste(italic("r"), " = 0.47, ",  , italic(p[spin]), "= 0.0709")),
                                       annot_y =  23.5, annot_x = 12.5, xlim1 = 8.5, xlim2 = 19) 
HBN_alff_ageeffect <- plot_annot_maps("HBN", "alff_pnc", 
                                       expression(paste(italic("r"), " = 0.17, ",  , italic(p[spin]), "= N.S.")),
                                       annot_y =  23.5, annot_x = -0.05, xlim1 = -0.15, xlim2 = 0.06) 
HBN_alff_agedecline <- plot_annot_maps("HBN", "alff_decline_pnc", 
                                       expression(paste(italic("r"), " = 0.49, ",  , italic(p[spin]), "= 0.0436")),
                                       annot_x = 12, annot_y = 23.5, xlim1 = 8, xlim2 = 16) 

 

PNC_t1t2_ageeffect <- plot_annot_maps("PNC", "t1t2_ageeffect_HCPD",
                                      expression(paste(italic("r"), " = -0.61, ",  , italic(p[spin]), "= 0.0068")),
                                      annot_y =  23.5, annot_x = 0.15, xlim1 = 0, xlim2 = 0.35) 
PNC_t1t2_mean <- plot_annot_maps("PNC", "t1t2_mean_HCPD", 
                                       expression(paste(italic("r"), " = -0.49, ",  , italic(p[spin]), "= 0.01445")),
                                       annot_y =  23.5, annot_x = 1.6, xlim1 = 1.3, xlim2 = 2) 
PNC_t1t2_roc <- plot_annot_maps("PNC", "t1t2_roc_HCPD", 
                                       expression(paste(italic("r"), " = -0.59, ",  , italic(p[spin]), "< 0.0001")),
                                       annot_y =  23.5, annot_x = 0.01, xlim1 = 0, xlim2 = 0.025) 
PNC_t1t2_maxslopeage <- plot_annot_maps("PNC", "t1t2_maxslopeage_HCPD", 
                                       expression(paste(italic("r"), " = 0.63, ",  , italic(p[spin]), "= 0.0011")),
                                       annot_y =  23.5, annot_x = 12.5, xlim1 = 8.5, xlim2 = 19) 
PNC_alff_ageeffect <- plot_annot_maps("PNC", "alff_pnc", 
                                       expression(paste(italic("r"), " = 0.52, ",  , italic(p[spin]), "= 0.00895")),
                                       annot_y =  23.5, annot_x = -0.05, xlim1 = -0.15, xlim2 = 0.06) 

PNC_alff_agedecline <- plot_annot_maps("PNC", "alff_decline_pnc", 
                                       expression(paste(italic("r"), " = 0.59, ",  , italic(p[spin]), "= 0.0053")),
                                       annot_x = 12, annot_y = 23.5, xlim1 = 8, xlim2 = 16) 
 


y.grob <- textGrob("WM Cortical Endpoint \nAge of Maturation (Years)", 
                   gp=gpar(col="black", fontsize=20), rot=90)

x.grob <- textGrob("T1w/T2w Partial R-squared (Age Effect)", 
                   gp=gpar(col="black", fontsize=20))
title.grob <- textGrob("T1w/T2w Age Effect", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
t1t2_ageeffect_plots <- ggarrange(HCPD_t1t2_ageeffect, HBN_t1t2_ageeffect, PNC_t1t2_ageeffect, ncol = 3)
grid.arrange(arrangeGrob(t1t2_ageeffect_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","t1t2_ageeffect_plots.png"), grid.arrange(arrangeGrob(t1t2_ageeffect_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 16, units = "in")

x.grob <- textGrob("Mean T1w/T2w", 
                   gp=gpar(col="black", fontsize=20))
title.grob <- textGrob("Group T1w/T2w Map", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
t1t2_mean_plots <- ggarrange(HCPD_t1t2_mean, HBN_t1t2_mean, PNC_t1t2_mean, ncol = 3)
grid.arrange(arrangeGrob(t1t2_mean_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","t1t2_mean_plots.png"), grid.arrange(arrangeGrob(t1t2_mean_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 16, units = "in")

x.grob <- textGrob("T1w/T2w Annualized Rate of Change", 
                   gp=gpar(col="black", fontsize=20))
title.grob <- textGrob("T1w/T2w Growth Rate", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
t1t2_roc_plots <- ggarrange(HCPD_t1t2_roc, HBN_t1t2_roc, PNC_t1t2_roc, ncol = 3)
grid.arrange(arrangeGrob(t1t2_roc_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","t1t2_roc_plots.png"), grid.arrange(arrangeGrob(t1t2_roc_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 16, units = "in")

x.grob <- textGrob("T1w/T2w Age of Maximum Growth Rate", 
                   gp=gpar(col="black", fontsize=20))
title.grob <- textGrob("T1w/T2w Age of Peak Growth", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
t1t2_maxslopeage_plots <- ggarrange(HCPD_t1t2_maxslopeage, HBN_t1t2_maxslopeage, PNC_t1t2_maxslopeage, ncol = 3)
grid.arrange(arrangeGrob(t1t2_maxslopeage_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","t1t2_maxslopeage_plots.png"), grid.arrange(arrangeGrob(t1t2_maxslopeage_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 16, units = "in")

x.grob <- textGrob("ALFF Partial R-squared (Age Effect)", 
                   gp=gpar(col="black", fontsize=20))
title.grob <- textGrob("ALFF Age Effect", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
alff_ageeffect_plots <- ggarrange(HCPD_alff_ageeffect, HBN_alff_ageeffect, PNC_alff_ageeffect, ncol = 3)
grid.arrange(arrangeGrob(alff_ageeffect_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","alff_ageeffect.png"), grid.arrange(arrangeGrob(alff_ageeffect_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 16, units = "in")



x.grob <- textGrob("ALFF Age of Decline", 
                   gp=gpar(col="black", fontsize=20))
title.grob <- textGrob("ALFF Age of Decline", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
alff_agedecline_plots <- ggarrange(HCPD_alff_agedecline, HBN_alff_agedecline, PNC_alff_agedecline, ncol = 3)
grid.arrange(arrangeGrob(alff_agedecline_plots, left = y.grob, bottom = x.grob, top = title.grob))

```


```{r agg_ageeffects binary plot, fig.height = 5, fig.width = 17}

perm.sphere.p(x = agg_ageeffects_HCPD_binary$regional_mean_ageeffect, y = t1t2_ageeffect_HCPD$data, perm.id = perm.id.full, corr.type = "spearman") # N.S   
perm.sphere.p(x = agg_ageeffects_HCPD_binary$regional_mean_ageeffect, y = t1t2_mean_HCPD$data, perm.id = perm.id.full, corr.type = "spearman")  # N.S   
perm.sphere.p(x = agg_ageeffects_HCPD_binary$regional_mean_ageeffect, y = t1t2_roc_HCPD$data, perm.id = perm.id.full, corr.type = "spearman")   
perm.sphere.p(x = agg_ageeffects_HCPD_binary$regional_mean_ageeffect, y = t1t2_maxslopeage_HCPD$data, perm.id = perm.id.full, corr.type = "spearman")   
perm.sphere.p(x = agg_ageeffects_HCPD_binary$regional_mean_ageeffect, y = alff_pnc$data, perm.id = perm.id.full, corr.type = "spearman")   
perm.sphere.p(x = agg_ageeffects_HCPD_binary$regional_mean_ageeffect, y = alff_decline_pnc$data, perm.id = perm.id.full, corr.type = "spearman")   
 
perm.sphere.p(x = agg_ageeffects_HBN_binary$regional_mean_ageeffect, y = t1t2_ageeffect_HCPD$data, perm.id = perm.id.full, corr.type = "spearman")    
perm.sphere.p(x = agg_ageeffects_HBN_binary$regional_mean_ageeffect, y = t1t2_mean_HCPD$data, perm.id = perm.id.full, corr.type = "spearman")  
perm.sphere.p(x = agg_ageeffects_HBN_binary$regional_mean_ageeffect, y = t1t2_roc_HCPD$data, perm.id = perm.id.full, corr.type = "spearman")   
perm.sphere.p(x = agg_ageeffects_HBN_binary$regional_mean_ageeffect, y = t1t2_maxslopeage_HCPD$data, perm.id = perm.id.full, corr.type = "spearman")   
perm.sphere.p(x = agg_ageeffects_HBN_binary$regional_mean_ageeffect, y = alff_pnc$data, perm.id = perm.id.full, corr.type = "spearman")   
perm.sphere.p(x = agg_ageeffects_HBN_binary$regional_mean_ageeffect, y = alff_decline_pnc$data, perm.id = perm.id.full, corr.type = "spearman")   

perm.sphere.p(x = agg_ageeffects_PNC_binary$regional_mean_ageeffect, y = t1t2_ageeffect_HCPD$data, perm.id = perm.id.full, corr.type = "spearman")    
perm.sphere.p(x = agg_ageeffects_PNC_binary$regional_mean_ageeffect, y = t1t2_mean_HCPD$data, perm.id = perm.id.full, corr.type = "spearman")  
perm.sphere.p(x = agg_ageeffects_PNC_binary$regional_mean_ageeffect, y = t1t2_roc_HCPD$data, perm.id = perm.id.full, corr.type = "spearman")   
perm.sphere.p(x = agg_ageeffects_PNC_binary$regional_mean_ageeffect, y = t1t2_maxslopeage_HCPD$data, perm.id = perm.id.full, corr.type = "spearman")   
perm.sphere.p(x = agg_ageeffects_PNC_binary$regional_mean_ageeffect, y = alff_pnc$data, perm.id = perm.id.full, corr.type = "spearman")   
perm.sphere.p(x = agg_ageeffects_PNC_binary$regional_mean_ageeffect, y = alff_decline_pnc$data, perm.id = perm.id.full, corr.type = "spearman")   
 
 
# pvalues are pre-fdr correction
cor.test(t1t2_ageeffect_HCPD$data, agg_ageeffects_HCPD_binary$regional_mean_ageeffect, method = 'spearman')  
cor.test(t1t2_ageeffect_HCPD$data, agg_ageeffects_HBN_binary$regional_mean_ageeffect, method = 'spearman')
cor.test(t1t2_ageeffect_HCPD$data, agg_ageeffects_PNC_binary$regional_mean_ageeffect, method = 'spearman') 
# age of maturation vs. t1/t2 partial rsq
# hcpd: r = 0.088, pspin = N.S
# hbn: r = -0.34, N.S
# pnc: r = -0.33, pspin = N.S.

#cor.test(t1t2_mean_HCPD$data, agg_ageeffects$regional_mean_ageeffect, method = 'spearman')   ## 
cor.test(t1t2_mean_HCPD$data, agg_ageeffects_HCPD_binary$regional_mean_ageeffect, method = 'spearman')  
cor.test(t1t2_mean_HCPD$data, agg_ageeffects_HBN_binary$regional_mean_ageeffect, method = 'spearman')  
cor.test(t1t2_mean_HCPD$data, agg_ageeffects_PNC_binary$regional_mean_ageeffect, method = 'spearman')  
# age of maturation vs. t1/t2 group mean
# hcpd: r = -0.23, pspin = N.S
# hbn: r = -0.36, 0.07955
# pnc: r = -0.48, pspin = 0.02155


#cor.test(t1t2_roc_HCPD$data, agg_ageeffects$regional_mean_ageeffect, method = 'spearman') ##
cor.test(t1t2_roc_HCPD$data, agg_ageeffects_HCPD_binary$regional_mean_ageeffect, method = 'spearman') 
cor.test(t1t2_roc_HCPD$data, agg_ageeffects_HBN_binary$regional_mean_ageeffect, method = 'spearman') 
cor.test(t1t2_roc_HCPD$data, agg_ageeffects_PNC_binary$regional_mean_ageeffect, method = 'spearman') 
# age of maturation vs. t1/t2 mean posterior annualized rate of change
# hcpd: r = -0.13, pspin = N.S
# hbn: r = -0.39,  0.05775
# pnc: r = -0.47, pspin = 0.01925


#cor.test(t1t2_maxslopeage_HCPD$data, agg_ageeffects$regional_mean_ageeffect, method = 'spearman') ##
cor.test(t1t2_maxslopeage_HCPD$data, agg_ageeffects_HCPD_binary$regional_mean_ageeffect, method = 'spearman')
cor.test(t1t2_maxslopeage_HCPD$data, agg_ageeffects_HBN_binary$regional_mean_ageeffect, method = 'spearman')
cor.test(t1t2_maxslopeage_HCPD$data, agg_ageeffects_PNC_binary$regional_mean_ageeffect, method = 'spearman')
# age of maturation vs t1/t2 age of max slope
# hcpd: r = 0.46, pspin = 0.09215
# hbn: r = 0.59, pspin =  0.00085  
# pnc: r = 0.65, pspin = 0.00085


#cor.test(alff_pnc$data, agg_ageeffects$regional_mean_ageeffect, method = 'spearman')  ##
cor.test(alff_pnc$data, agg_ageeffects_HCPD_binary$regional_mean_ageeffect, method = 'spearman') 
cor.test(alff_pnc$data, agg_ageeffects_HBN_binary$regional_mean_ageeffect, method = 'spearman') 
cor.test(alff_pnc$data, agg_ageeffects_PNC_binary$regional_mean_ageeffect, method = 'spearman') 
# age of maturation vs. alff age effect
# hcpd: r = -0.09, pspin = N.S
# hbn: r = 0.26, N.S.
# pnc: r = 0.42, pspin = 0.03975


#cor.test(alff_decline_pnc$data, agg_ageeffects$regional_mean_ageeffect, method = 'spearman')  
cor.test(alff_decline_pnc$data, agg_ageeffects_HCPD_binary$regional_mean_ageeffect, method = 'spearman') 
cor.test(alff_decline_pnc$data, agg_ageeffects_HBN_binary$regional_mean_ageeffect, method = 'spearman') 
cor.test(alff_decline_pnc$data, agg_ageeffects_PNC_binary$regional_mean_ageeffect, method = 'spearman') 
# age of maturation vs. alff age of decline
# hcpd: r = 0.24, pspin = N.S.
# hbn: r = 0.39, pspin = 0.08015
# pnc: r = 0.54, pspin = 0.02025



color1 = "#3FB8AFFF"
color2 = "#FF9E9DFF"
color3 = "#cc0468"

plot_annot_maps <- function(dataset, map, annot_text, annot_x, annot_y, xlim1, xlim2) {
  map <- get(map)$data # i.e. t1t2_ageeffect_HCPD
  map <- data.frame(map) %>% setNames("map_measure")
  agg_ageeffects <- get(paste0("agg_ageeffects_", dataset, "_binary"))
  agg_map <- cbind(agg_ageeffects, map)
  
  title = gsub("PD", "P-D", dataset) # for HCPD
  
  if(dataset == "HCPD") {
    color = color1
  } else if (dataset=="HBN") {
    color = color2
  } else {
    color = color3
  }
  
  agg_plot <- ggplot(agg_map, aes(x = map_measure, y = regional_mean_ageeffect)) +
  geom_point(size = 5, fill = color, color = color, shape = 21, alpha = 0.6) +  
  geom_smooth(data = agg_map, method='lm', se=TRUE, fill=alpha(c("gray70"),.7), col="black") + 
  #stat_cor(method = "spearman", size = 6) + 
  labs(title = title) + theme_classic() +
  theme(legend.position = "none",
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_blank(),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20)) +
  annotate(geom="text", x=annot_x, y=annot_y, label=annot_text, color="black", size=7) + ylim(15, 24) + xlim(xlim1, xlim2)
 
  return(agg_plot)
}

 
HCPD_t1t2_ageeffect <- plot_annot_maps("HCPD", "t1t2_ageeffect_HCPD", 
                                       expression(paste(italic("r"), " = 0.09, ",  , italic(p[spin]), "= N.S")),
                                       annot_y =  23.5, annot_x = 0.15, xlim1 = 0, xlim2 = 0.35) 
HCPD_t1t2_mean <- plot_annot_maps("HCPD", "t1t2_mean_HCPD", 
                                       expression(paste(italic("r"), " = -0.23, ",  , italic(p[spin]), "= N.S")),
                                       annot_y =  23.5, annot_x = 1.6, xlim1 = 1.3, xlim2 = 2) 
HCPD_t1t2_roc <- plot_annot_maps("HCPD", "t1t2_roc_HCPD", 
                                       expression(paste(italic("r"), " = -0.13, ",  , italic(p[spin]), "= N.S")),
                                       annot_y =  23.5, annot_x = 0.01, xlim1 = 0, xlim2 = 0.025) 
HCPD_t1t2_maxslopeage <- plot_annot_maps("HCPD", "t1t2_maxslopeage_HCPD", 
                                       expression(paste(italic("r"), " = 0.46, ",  , italic(p[spin]), "= 0.09215")),
                                       annot_y =  23.5, annot_x = 12.5, xlim1 = 8.5, xlim2 = 19) 
HCPD_alff_ageeffect <- plot_annot_maps("HCPD", "alff_pnc", 
                                       expression(paste(italic("r"), " = -0.09, ",  , italic(p[spin]), "= N.S")),
                                       annot_y =  23.5, annot_x = -0.05, xlim1 = -0.15, xlim2 = 0.06) 
HCPD_alff_agedecline <- plot_annot_maps("HCPD", "alff_decline_pnc", 
                                       expression(paste(italic("r"), " = 0.24, ",  , italic(p[spin]), "= N.S")),
                                       annot_x = 12, annot_y = 23.5, xlim1 = 8, xlim2 = 16) 

 
HBN_t1t2_ageeffect <- plot_annot_maps("HBN", "t1t2_ageeffect_HCPD",
                                      expression(paste(italic("r"), " = -0.34, ",  , italic(p[spin]), "= N.S")),
                                      annot_y =  23.5, annot_x = 0.15, xlim1 = 0, xlim2 = 0.35) 
HBN_t1t2_mean <- plot_annot_maps("HBN", "t1t2_mean_HCPD", 
                                       expression(paste(italic("r"), " = -0.36, ",  , italic(p[spin]), "= 0.07955")),
                                         annot_y =  23.5, annot_x = 1.6, xlim1 = 1.3, xlim2 = 2) 
HBN_t1t2_roc <- plot_annot_maps("HBN", "t1t2_roc_HCPD", 
                                       expression(paste(italic("r"), " =  -0.39, ",  , italic(p[spin]), "= 0.05775")),
                                       annot_y =  23.5, annot_x = 0.01, xlim1 = 0, xlim2 = 0.025) 
HBN_t1t2_maxslopeage <- plot_annot_maps("HBN", "t1t2_maxslopeage_HCPD", 
                                       expression(paste(italic("r"), " =  0.59, ",  , italic(p[spin]), "= 0.00085")),
                                       annot_y =  23.5, annot_x = 12.5, xlim1 = 8.5, xlim2 = 19) 
HBN_alff_ageeffect <- plot_annot_maps("HBN", "alff_pnc", 
                                       expression(paste(italic("r"), " = 0.26, ",  , italic(p[spin]), "= N.S.")),
                                       annot_y =  23.5, annot_x = -0.05, xlim1 = -0.15, xlim2 = 0.06) 
HBN_alff_agedecline <- plot_annot_maps("HBN", "alff_decline_pnc", 
                                       expression(paste(italic("r"), " = 0.39, ",  , italic(p[spin]), "= 0.08015")),
                                       annot_x = 12, annot_y = 23.5, xlim1 = 8, xlim2 = 16) 

 

PNC_t1t2_ageeffect <- plot_annot_maps("PNC", "t1t2_ageeffect_HCPD",
                                      expression(paste(italic("r"), " = -0.33, ",  , italic(p[spin]), "= N.S.")),
                                      annot_y =  23.5, annot_x = 0.15, xlim1 = 0, xlim2 = 0.35) 
PNC_t1t2_mean <- plot_annot_maps("PNC", "t1t2_mean_HCPD", 
                                       expression(paste(italic("r"), " = -0.48, ",  , italic(p[spin]), "= 0.02155")),
                                       annot_y =  23.5, annot_x = 1.6, xlim1 = 1.3, xlim2 = 2) 
PNC_t1t2_roc <- plot_annot_maps("PNC", "t1t2_roc_HCPD", 
                                       expression(paste(italic("r"), " = -0.47, ",  , italic(p[spin]), "= 0.01925")),
                                       annot_y =  23.5, annot_x = 0.01, xlim1 = 0, xlim2 = 0.025) 
PNC_t1t2_maxslopeage <- plot_annot_maps("PNC", "t1t2_maxslopeage_HCPD", 
                                       expression(paste(italic("r"), " = 0.65, ",  , italic(p[spin]), "= 0.00085")),
                                       annot_y =  23.5, annot_x = 12.5, xlim1 = 8.5, xlim2 = 19) 
PNC_alff_ageeffect <- plot_annot_maps("PNC", "alff_pnc", 
                                       expression(paste(italic("r"), " = 0.42, ",  , italic(p[spin]), "= 0.03975")),
                                       annot_y =  23.5, annot_x = -0.05, xlim1 = -0.15, xlim2 = 0.06) 

PNC_alff_agedecline <- plot_annot_maps("PNC", "alff_decline_pnc", 
                                       expression(paste(italic("r"), " = 0.54, ",  , italic(p[spin]), "= 0.02025")),
                                       annot_x = 12, annot_y = 23.5, xlim1 = 8, xlim2 = 16) 
 


y.grob <- textGrob("WM Cortical Endpoint \nAge of Maturation (Years)", 
                   gp=gpar(col="black", fontsize=20), rot=90)

x.grob <- textGrob("T1w/T2w Partial R-squared (Age Effect)", 
                   gp=gpar(col="black", fontsize=20))
title.grob <- textGrob("T1w/T2w Age Effect", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
t1t2_ageeffect_plots <- ggarrange(HCPD_t1t2_ageeffect, HBN_t1t2_ageeffect, PNC_t1t2_ageeffect, ncol = 3)
grid.arrange(arrangeGrob(t1t2_ageeffect_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","t1t2_ageeffect_plots.png"), grid.arrange(arrangeGrob(t1t2_ageeffect_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 16, units = "in")

x.grob <- textGrob("Mean T1w/T2w", 
                   gp=gpar(col="black", fontsize=20))
title.grob <- textGrob("Group T1w/T2w Map", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
t1t2_mean_plots <- ggarrange(HCPD_t1t2_mean, HBN_t1t2_mean, PNC_t1t2_mean, ncol = 3)
grid.arrange(arrangeGrob(t1t2_mean_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","t1t2_mean_plots.png"), grid.arrange(arrangeGrob(t1t2_mean_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 16, units = "in")

x.grob <- textGrob("T1w/T2w Annualized Rate of Change", 
                   gp=gpar(col="black", fontsize=20))
title.grob <- textGrob("T1w/T2w Growth Rate", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
t1t2_roc_plots <- ggarrange(HCPD_t1t2_roc, HBN_t1t2_roc, PNC_t1t2_roc, ncol = 3)
grid.arrange(arrangeGrob(t1t2_roc_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","t1t2_roc_plots.png"), grid.arrange(arrangeGrob(t1t2_roc_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 16, units = "in")

x.grob <- textGrob("T1w/T2w Age of Maximum Growth Rate", 
                   gp=gpar(col="black", fontsize=20))
title.grob <- textGrob("T1w/T2w Age of Peak Growth", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
t1t2_maxslopeage_plots <- ggarrange(HCPD_t1t2_maxslopeage, HBN_t1t2_maxslopeage, PNC_t1t2_maxslopeage, ncol = 3)
grid.arrange(arrangeGrob(t1t2_maxslopeage_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","t1t2_maxslopeage_plots.png"), grid.arrange(arrangeGrob(t1t2_maxslopeage_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 16, units = "in")

x.grob <- textGrob("ALFF Partial R-squared (Age Effect)", 
                   gp=gpar(col="black", fontsize=20))
title.grob <- textGrob("ALFF Age Effect", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
alff_ageeffect_plots <- ggarrange(HCPD_alff_ageeffect, HBN_alff_ageeffect, PNC_alff_ageeffect, ncol = 3)
grid.arrange(arrangeGrob(alff_ageeffect_plots, left = y.grob, bottom = x.grob, top = title.grob))
#ggsave(paste0(png_dir, "/", "all_datasets", "/","alff_ageeffect.png"), grid.arrange(arrangeGrob(alff_ageeffect_plots, left = y.grob, bottom = x.grob, top = title.grob)) , height = 5, width = 16, units = "in")



x.grob <- textGrob("ALFF Age of Decline", 
                   gp=gpar(col="black", fontsize=20))
title.grob <- textGrob("ALFF Age of Decline", 
                   gp=gpar(col="black", fontsize=20, fontface="bold"))
alff_agedecline_plots <- ggarrange(HCPD_alff_agedecline, HBN_alff_agedecline, PNC_alff_agedecline, ncol = 3)
grid.arrange(arrangeGrob(alff_agedecline_plots, left = y.grob, bottom = x.grob, top = title.grob))


```



```{r}
perm.id = perm.id.full

aggregated_axis_HCPD_binary <- aggregated_axis_HCPD_binary %>% mutate(maturation_status = ifelse(is.na(regional_mean_ageeffect), 0, 1))
aggregated_axis_HBN_binary <- aggregated_axis_HBN_binary %>% mutate(maturation_status = ifelse(is.na(regional_mean_ageeffect), 0, 1))
aggregated_axis_PNC_binary <- aggregated_axis_PNC_binary %>% mutate(maturation_status = ifelse(is.na(regional_mean_ageeffect), 0, 1))

# spun p-value for t-test
perm.sphere.p.ttest = function(binary_df, map, perm.id, alternative = "two.sided", var.equal = FALSE) {
  if (!"SA.axis_rank" %in% names(map)) {
    map <- data.frame(map$data) %>% setNames("map_measure")
    agg_ageeffects_merged <- cbind(binary_df, map)
  }  else { # SA axis is already included in my binary df
    agg_ageeffects_merged <- binary_df %>% rename(map_measure = SA.axis_rank)
    map <- data.frame(binary_df$SA.axis_rank) %>% setNames("map_measure")
  }
  

  nroi = dim(perm.id)[1]  # number of regions
  nperm = dim(perm.id)[2] # number of permutations
  
  t.emp = t.test(agg_ageeffects_merged$map_measure[which(agg_ageeffects_merged$maturation_status == 0)], # compare differences in sa rank
                    agg_ageeffects_merged$map_measure[which(agg_ageeffects_merged$maturation_status == 1)], 
                    alternative = alternative, var.equal = var.equal)$statistic 
  
  # permutation of map
  map.perm = array(NA,dim=c(nroi,nperm))
  map <- map$map_measure
  for (r in 1:nperm) {
    for (i in 1:nroi) {
      map.perm[i,r] = map[perm.id[i,r]]
    }
  }
  # t.test using permuted map values
  t.null.map = vector(length=nperm)
  for (r in 1:nperm) {
    map.perm.tobind <- data.frame(map.perm[,r]) %>% setNames("perm_map_measure")
    agg_ageeffects_perm <- cbind(binary_df, map.perm.tobind)
    t.null.map[r] = t.test(agg_ageeffects_perm$perm_map_measure[which(agg_ageeffects_perm$maturation_status == 0)], # compare differences in sa rank
                    agg_ageeffects_perm$perm_map_measure[which(agg_ageeffects_perm$maturation_status == 1)], 
                    alternative = alternative, var.equal = var.equal)$statistic 
  }
  
  # p-value definition depends on the sign of the empirical correlation
  if (t.emp>0) {
    p.perm = sum(t.null.map>t.emp)/nperm
   
  } else { 
    p.perm = sum(t.null.map<t.emp)/nperm
   
  } 
  # return p-value
  return(list(p.perm = p.perm, t.emp = t.emp))
  
}
 
 
perm.sphere.p.ttest(aggregated_axis_HCPD_binary, t1t2_ageeffect_HCPD, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = 0.0353, t = 2.657315 
perm.sphere.p.ttest(aggregated_axis_HCPD_binary, t1t2_mean_HCPD, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS
perm.sphere.p.ttest(aggregated_axis_HCPD_binary, t1t2_roc_HCPD, perm.id.full, alternative = "two.sided", var.equal = FALSE) #  pspin = 0.0275, t = 2.720289 
perm.sphere.p.ttest(aggregated_axis_HCPD_binary, t1t2_maxslopeage_HCPD, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS, t = -1.911331 
perm.sphere.p.ttest(aggregated_axis_HCPD_binary, alff_pnc, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = 0.0645
perm.sphere.p.ttest(aggregated_axis_HCPD_binary, alff_decline_pnc, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = 0.0988
perm.sphere.p.ttest(aggregated_axis_HCPD_binary, glasser_SAaxis, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = 0, t = 16

  

perm.sphere.p.ttest(aggregated_axis_HBN_binary, t1t2_ageeffect_HCPD, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS
perm.sphere.p.ttest(aggregated_axis_HBN_binary, t1t2_mean_HCPD, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS
perm.sphere.p.ttest(aggregated_axis_HBN_binary, t1t2_roc_HCPD, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS
perm.sphere.p.ttest(aggregated_axis_HBN_binary, t1t2_maxslopeage_HCPD, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = 0.0543
perm.sphere.p.ttest(aggregated_axis_HBN_binary, alff_pnc, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS
perm.sphere.p.ttest(aggregated_axis_HBN_binary, alff_decline_pnc, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS
perm.sphere.p.ttest(aggregated_axis_HBN_binary, glasser_SAaxis, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS

 

perm.sphere.p.ttest(aggregated_axis_PNC_binary, t1t2_ageeffect_HCPD, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS
perm.sphere.p.ttest(aggregated_axis_PNC_binary, t1t2_mean_HCPD, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS
perm.sphere.p.ttest(aggregated_axis_PNC_binary, t1t2_roc_HCPD, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS
perm.sphere.p.ttest(aggregated_axis_PNC_binary, t1t2_maxslopeage_HCPD, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS
perm.sphere.p.ttest(aggregated_axis_PNC_binary, alff_pnc, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = 0.0267
perm.sphere.p.ttest(aggregated_axis_PNC_binary, alff_decline_pnc, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = NS
perm.sphere.p.ttest(aggregated_axis_PNC_binary, glasser_SAaxis, perm.id.full, alternative = "two.sided", var.equal = FALSE) # pspin = 0, t= 7.87

 
test <- cbind(aggregated_axis_HCPD_binary, glasser_SAaxis$SA.axis_rank)
 
t.test(aggregated_axis_HCPD_binary$SA.axis_rank[which(aggregated_axis_HCPD_binary$maturation_status == 0)], # compare differences in sa rank
                    aggregated_axis_HCPD_binary$SA.axis_rank[which(aggregated_axis_HCPD_binary$maturation_status == 1)], 
                    alternative = alternative, var.equal = var.equal)

t.test(test$SA.axis_rank[which(test$maturation_status == 0)], # compare differences in sa rank
                    test$SA.axis_rank[which(test$maturation_status == 1)]) 
 
  
cbind(map$map_measure, test$SA.axis_rank)

data.frame(map$SA.axis_rank)
map$SA.axis_rank

data.frame(map$SA.axis_rank) %>% setNames("map_measure")
data.frame(glasser_SAaxis$SA.axis_rank) %>% setNames("map_measure")
```


```{r ttests of maps, fig.height = 5, fig.width = 15}


color1 = "#3FB8AFFF"
color2 = "#FF9E9DFF"
color3 = "#cc0468"
plot_ttest <- function(dataset, map, pvalue_expression, color) {
  df <- get(paste0("aggregated_axis_", dataset, "_binary"))
  
  if (!"SA.axis_rank" %in% names(map)) {
    map <- data.frame(map$data) %>% setNames("map_measure")
    agg_ageeffects_merged <- cbind(df, map)
  }  else { # SA axis is already included in my binary df
    agg_ageeffects_merged <- df %>% rename(map_measure = SA.axis_rank)
    map <- data.frame(df$SA.axis_rank) %>% setNames("map_measure")
  }
  agg_ageeffects_merged <- agg_ageeffects_merged %>% mutate(maturation_status = ifelse(maturation_status == 1, "Matured", "Not Matured"))
  title <- gsub("PD", "P-D", dataset)
  plot <- ggplot(agg_ageeffects_merged, aes(x = as.factor(maturation_status), y = map_measure)) +
  geom_boxplot(alpha = 0.8, outlier.shape = NA, fill = color) +   
  labs(
    title = title,
    subtitle = pvalue_expression,
    x = "Group",
    y = "Value"
  ) +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20),
        plot.title = element_text(size = 20, hjust = 0.5),
        plot.subtitle = element_text(size = 16, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
  
  return(plot)
}

HCPD_ttest_t1t2_ageeffect <- plot_ttest("HCPD", t1t2_ageeffect_HCPD, expression(paste(italic(p[spin]), "= 0.0353")), color1)
HCPD_ttest_t1t2_mean <- plot_ttest("HCPD", t1t2_mean_HCPD, expression(paste(italic(p[spin]), "= NS")), color1)
HCPD_ttest_t1t2_roc <- plot_ttest("HCPD", t1t2_roc_HCPD, expression(paste(italic(p[spin]), "= 0.0275")), color1)
HCPD_ttest_t1t2_maxslopeage <- plot_ttest("HCPD", t1t2_maxslopeage_HCPD, expression(paste(italic(p[spin]), "= NS")), color1)
HCPD_ttest_alff_ageeffect <- plot_ttest("HCPD", alff_pnc, expression(paste(italic(p[spin]), "= 0.0645")), color1)
HCPD_ttest_alff_decline <- plot_ttest("HCPD", alff_decline_pnc, expression(paste(italic(p[spin]), "= 0.0988")), color1)
HCPD_ttest_SAaxis <- plot_ttest("HCPD", glasser_SAaxis, expression(paste(italic(p[spin]), "< 0.0001")), color1)
 


HBN_ttest_t1t2_ageeffect <- plot_ttest("HBN", t1t2_ageeffect_HCPD, expression(paste(italic(p[spin]), "= NS")), color2)
HBN_ttest_t1t2_mean <- plot_ttest("HBN", t1t2_mean_HCPD, expression(paste(italic(p[spin]), "= NS")), color2)
HBN_ttest_t1t2_roc <- plot_ttest("HBN", t1t2_roc_HCPD, expression(paste(italic(p[spin]), "= NS")), color2)
HBN_ttest_t1t2_maxslopeage <- plot_ttest("HBN", t1t2_maxslopeage_HCPD, expression(paste(italic(p[spin]), "= 0.0543")), color2)
HBN_ttest_alff_ageeffect <- plot_ttest("HBN", alff_pnc, expression(paste(italic(p[spin]), "= NS")), color2)
HBN_ttest_alff_decline <- plot_ttest("HBN", alff_decline_pnc, expression(paste(italic(p[spin]), "= NS")), color2)
HBN_ttest_SAaxis <- plot_ttest("HBN", glasser_SAaxis, expression(paste(italic(p[spin]), "< NS")), color2)


PNC_ttest_t1t2_ageeffect <- plot_ttest("PNC", t1t2_ageeffect_HCPD, expression(paste(italic(p[spin]), "= NS")), color3)
PNC_ttest_t1t2_mean <- plot_ttest("PNC", t1t2_mean_HCPD, expression(paste(italic(p[spin]), "= NS")), color3)
PNC_ttest_t1t2_roc <- plot_ttest("PNC", t1t2_roc_HCPD, expression(paste(italic(p[spin]), "= NS")), color3)
PNC_ttest_t1t2_maxslopeage <- plot_ttest("PNC", t1t2_maxslopeage_HCPD, expression(paste(italic(p[spin]), "= NS")), color3)
PNC_ttest_alff_ageeffect <- plot_ttest("PNC", alff_pnc, expression(paste(italic(p[spin]), "= 0.0267")), color3)
PNC_ttest_alff_decline <- plot_ttest("PNC", alff_decline_pnc, expression(paste(italic(p[spin]), "= NS")), color3)
PNC_ttest_SAaxis <- plot_ttest("PNC", glasser_SAaxis, expression(paste(italic(p[spin]), "< 0.0001")), color3)
 

ttest_t1t2ageeffect_final <- ggarrange(HCPD_ttest_t1t2_ageeffect, HBN_ttest_t1t2_ageeffect, PNC_ttest_t1t2_ageeffect, ncol = 3)
y.grob <- textGrob("T1w/T2w Partial R-squared (Age Effect)", 
                   gp=gpar(col="black", fontsize=20), rot = 90)
grid.arrange(arrangeGrob(ttest_t1t2ageeffect_final, left = y.grob))


ttest_t1t2mean_final <- ggarrange(HCPD_ttest_t1t2_mean, HBN_ttest_t1t2_mean, PNC_ttest_t1t2_mean, ncol = 3)
y.grob <- textGrob("Mean T1w/T2w", 
                   gp=gpar(col="black", fontsize=20), rot = 90)

grid.arrange(arrangeGrob(ttest_t1t2mean_final, left = y.grob))


ttest_t1t2roc_final <- ggarrange(HCPD_ttest_t1t2_roc, HBN_ttest_t1t2_roc, PNC_ttest_t1t2_roc, ncol = 3)
y.grob <- textGrob("T1w/T2w Annualized Rate of Change", 
                   gp=gpar(col="black", fontsize=20), rot = 90)
grid.arrange(arrangeGrob(ttest_t1t2roc_final, left = y.grob))



ttest_t1t2maxslopeage_final <- ggarrange(HCPD_ttest_t1t2_maxslopeage, HBN_ttest_t1t2_maxslopeage, PNC_ttest_t1t2_maxslopeage, ncol = 3)
y.grob <- textGrob("T1w/T2w Age of Maximum Growth Rate", 
                   gp=gpar(col="black", fontsize=20), rot = 90)
grid.arrange(arrangeGrob(ttest_t1t2maxslopeage_final, left = y.grob))




ttest_alffageeffect_final <- ggarrange(HCPD_ttest_alff_ageeffect, HBN_ttest_alff_ageeffect, PNC_ttest_alff_ageeffect, ncol = 3)
y.grob <- textGrob("ALFF Partial R-squared (Age Effect)", 
                   gp=gpar(col="black", fontsize=20), rot = 90)

grid.arrange(arrangeGrob(ttest_alffageeffect_final, left = y.grob))



ttest_alff_decline_final <- ggarrange(HCPD_ttest_alff_decline, HBN_ttest_alff_decline, PNC_ttest_alff_decline, ncol = 3)

y.grob <- textGrob("ALFF Age of decline", 
                   gp=gpar(col="black", fontsize=20), rot = 90)

grid.arrange(arrangeGrob(ttest_alff_decline_final, left = y.grob))



ttest_SAaxis_final <- ggarrange(HCPD_ttest_SAaxis, HBN_ttest_SAaxis, PNC_ttest_SAaxis, ncol = 3)

y.grob <- textGrob("S-A Axis Rank", 
                   gp=gpar(col="black", fontsize=20), rot = 90)

grid.arrange(arrangeGrob(ttest_SAaxis_final, left = y.grob))
 
```


```{r delta delta ttest, fig.height = 5, fig.width = 15}
diffs_all <- get(paste0("diffs_", dataset))
plot_ttest_delta_delta <- function(dataset, pvalue_expression, color) {
  diffs_all <- get(paste0("diffs_", dataset))
  
   
  diffs_all <- diffs_all %>% mutate(large_SA_diff = ifelse(is.na(ifo_ilf_label), "Small Difference\n in S-A Rank", "Large Difference\n in S-A Rank"))
  title <- gsub("PD", "P-D", dataset)
  plot <- ggplot(diffs_all, aes(x = as.factor(large_SA_diff), y = age_effect_diff)) +
  geom_boxplot(alpha = 0.8, outlier.shape = NA, fill = color) +   
  labs(
    title = title,
    subtitle = pvalue_expression,
    y = "Difference in Age of Maturation"
  ) +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16),
        plot.title = element_text(size = 16, hjust = 0.5),
        plot.subtitle = element_text(size = 16, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
  
  return(plot)
}


HCPD_delta <- plot_ttest_delta_delta("HCPD", expression(paste(italic(p[spin]), "= 0")), color1)

HBN_delta <- plot_ttest_delta_delta("HBN", expression(paste(italic(p[spin]), "= NS")), color2)

PNC_delta <- plot_ttest_delta_delta("PNC", expression(paste(italic(p[spin]), "= 0.001")), color3)

delta_delta_ttest <- ggarrange(HCPD_delta, HBN_delta,PNC_delta, ncol = 3)

y.grob <- textGrob("Difference in Age of Maturation\n between Endpoints", 
                   gp=gpar(col="black", fontsize=20), rot = 90)

grid.arrange(arrangeGrob(delta_delta_ttest, left = y.grob))
```

