---
title: "Check residuals"
author: "Audrey Luo"
output: 
  html_document:
    highlight: haddock
    code_folding: hide
    number_sections: no
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    number_sections: no
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

library(cowplot)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(gratia)
library(kableExtra)
library(mgcv)
library(RColorBrewer)
library(stringr)
library(tidyr)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, include = TRUE) 

font_size <- 18
theme_set(theme_minimal(base_family = "sans",base_size = font_size))
line_size <- 0.5
point_size <- 2
```


```{r load tract profiles, cache=TRUE}
cohortfile <- read.csv("/cbica/projects/luo_wm_dev/output/HBN/tract_profiles/cohortfiles/dti_fa/dti_fa_cohortfile.csv")
cohortfile <- cohortfile %>% rename(subjectID=sub) %>% select(subjectID, age, sex, mean_fd)

tract_profiles <- fread("/cbica/projects/luo_wm_dev/input/HBN/HBN_tractprofiles/all_subjects/collated_tract_profiles_HBN_reoriented.tsv")

cohortfile <- cohortfile %>% select(subjectID, age, sex, mean_fd)
gam_df <- merge(tract_profiles, cohortfile, by="subjectID")
gam_df <- gam_df %>% mutate(tract_node = paste0(tract_hemi, "_", nodeID))
gam_df$sex <- as.factor(gam_df$sex)
```

```{r take subset of 75 subjects, cache=TRUE}
subjects <- read.table("/cbica/projects/luo_wm_dev/input/HBN/subject_list/HBN_tempsubject_list_N1263.txt")
subjects <- subjects %>% setNames("subjectID")
subject_subset <- subjects$subjectID[c(1:75)]
gam_df <- gam_df %>% filter(subjectID %in% subject_subset) # fit GAMs on a subset of 75 participants
```

```{r functions}
# fit tractwise gams
fit_tractwise_gam_re <- function(tract_name, scalar) {
  df <- gam_df %>% filter(tract_hemi == tract_name)
  df$subjectID <- as.factor(df$subjectID)
  te_gam <- gam(as.formula(sprintf("%s ~ 
  s(age, k = 3, fx = T, bs = 'cr') + 
  s(nodeID, k = 24, fx = F, bs = 'cr') + 
  ti(age, nodeID, k=c(3, 24), fx = F) + 
  sex + mean_fd + s(subjectID, bs = 're')", scalar)), data = df) 
  
  print(summary(te_gam))
  print(k.check(te_gam))
  print(paste(tract_name, "done"))
  
  
  return(te_gam)  # return the model object 
}

fit_tractwise_bam <- function(tract_name, scalar) {
  df <- gam_df %>% filter(tract_hemi == tract_name)
  df$subjectID <- as.factor(df$subjectID)
  te_gam <- bam(as.formula(sprintf("%s ~ 
  s(age, k = 3, fx = T, bs = 'cr') + 
  s(nodeID, k = 24, fx = F, bs = 'cr', by = subjectID) + 
  ti(age, nodeID, k=c(3, 24), fx = F) + 
  + s(subjectID, bs = 're') + sex + mean_fd", scalar)), data = df, method = "fREML", discrete = TRUE) 
  
  print(summary(te_gam))
  print(k.check(te_gam))
  print(paste(tract_name, "done"))
 
  
  return(te_gam)  # return the model object 
}
 


# comparing residuals and generating plots
plot_compare_residuals <- function(tract, num_subjects, model_number) { 
  gam_df_toplot <- gam_df %>% filter(tract_hemi == tract)
  if(model_number == 1) {
    gam_df_toplot$residuals_gam_re <- resid(tractwise_gam_re[[tract]])
    subjects_to_plot <- subject_subset[c(1:num_subjects)]
    plot <- ggplot(gam_df_toplot[gam_df_toplot$subjectID %in% subjects_to_plot, ],
                       aes(x = nodeID, y = residuals_gam_re, color = factor(subjectID))) +
      geom_line() + theme(legend.position = "none",
                        plot.title = element_text(hjust=0.5, size = 12),
                        axis.title.x = element_blank(), 
                        axis.title.y = element_blank()) + 
    labs(title = gsub("_", " ", tract))  
    
  } else if (model_number == 2) {
    gam_df_toplot$residuals_bam <- resid(tractwise_bam[[tract]])
    subjects_to_plot <- subject_subset[c(1:num_subjects)]
    plot <- ggplot(gam_df_toplot[gam_df_toplot$subjectID %in% subjects_to_plot, ],
                       aes(x = nodeID, y = residuals_bam, color = factor(subjectID))) +
      geom_line() + theme(legend.position = "none",
                        plot.title = element_text(hjust=0.5, size = 12),
                        axis.title.x = element_blank(), 
                        axis.title.y = element_blank()) + 
    labs(title = gsub("_", " ", tract))  
  
  }
  
  return(plot)
}

arrange_tracts_hemis <- function(list_figures, text) {
  
  ATR_plots <- ggarrange(list_figures$Left_Anterior_Thalamic_Radiation, list_figures$Right_Anterior_Thalamic_Radiation, nrow = 2)
  
  AP_plots <- ggarrange(list_figures$Left_Cingulum_Cingulate, 
                           list_figures$`Left_Inferior_Fronto-occipital_Fasciculus`, 
                           list_figures$Left_Inferior_Longitudinal_Fasciculus, 
                           list_figures$Left_Superior_Longitudinal_Fasciculus, 
                           list_figures$Right_Cingulum_Cingulate, 
                           list_figures$`Right_Inferior_Fronto-occipital_Fasciculus`, 
                           list_figures$Right_Inferior_Longitudinal_Fasciculus, 
                           list_figures$Right_Superior_Longitudinal_Fasciculus, ncol=4, nrow = 2)
  
  AP_ATR_plots <- ggarrange(ATR_plots, AP_plots, ncol=2, widths = c(1,4)) 
  AP_plots_final <- annotate_figure(AP_ATR_plots, top = text_grob("Anterior - Posterior", 
                 color = "black", face = "bold", size = 12))
  
  
  AP_frontotemp_plots <- ggarrange(list_figures$Left_Arcuate_Fasciculus, list_figures$Left_Uncinate_Fasciculus,
                                   list_figures$Right_Arcuate_Fasciculus, list_figures$Right_Uncinate_Fasciculus, ncol=2, nrow = 2)
  AP_frontotemp_plots_final <- annotate_figure(AP_frontotemp_plots, top = text_grob("Anterior - Posterior (Frontal - Temporal)", 
                 color = "black", face = "bold", size = 12))
  
   
  RL_plots <- ggarrange(list_figures$Forceps_Major, list_figures$Forceps_Minor, 
                           ncol=2, nrow = 1)
  RL_plots_final <- annotate_figure(RL_plots, top = text_grob("Right - Left", 
                 color = "black", face = "bold", size = 12))
  
  CST_plots <- ggarrange(list_figures$Left_Corticospinal_Tract,list_figures$Right_Corticospinal_Tract, ncol=1, nrow = 2)
  SI_plots <- ggarrange(list_figures$Left_Posterior_Arcuate, 
                        list_figures$Left_Vertical_Occipital_Fasciculus, 
                        list_figures$Right_Posterior_Arcuate, 
                        list_figures$Right_Vertical_Occipital_Fasciculus, common.legend=TRUE, legend=c("right"), ncol=2, nrow = 2)
  
  SI_CST_plots <- ggarrange(CST_plots, SI_plots, widths = c(1, 2))
  SI_plots_final <- annotate_figure(SI_CST_plots, top = text_grob("Superior - Inferior", 
                 color = "black", face = "bold", size = 12))
  
  
  tractprofiles_plot <- ggarrange(AP_plots_final, ggarrange(AP_frontotemp_plots_final, RL_plots_final, ncol = 2), SI_plots_final, nrow = 3) + bgcolor("white")    
  
  tractprofiles_plot_final <- annotate_figure(tractprofiles_plot, top = text_grob(text, 
               color = "black", face = "italic", size = 18))
  
  return(tractprofiles_plot_final)
}
```


```{r fit gams, eval=FALSE}
tractwise_gam_re <- lapply(unique(gam_df$tract_hemi), fit_tractwise_gam_re, scalar="dti_md")
names(tractwise_gam_re) <- unique(gam_df$tract_hemi)
saveRDS(tractwise_gam_re, "/cbica/projects/luo_wm_dev/output/HBN/tract_profiles/GAM/dti_md/check_resid_tractwise_gam_re.RData")

tractwise_bam <- lapply(unique(gam_df$tract_hemi), fit_tractwise_bam, scalar="dti_md")
names(tractwise_bam) <- unique(gam_df$tract_hemi)
saveRDS(tractwise_bam, "/cbica/projects/luo_wm_dev/output/HBN/tract_profiles/GAM/dti_md/check_resid_tractwise_bam2.RData")

```

 
```{r load gam models, cache = TRUE, cache.lazy = FALSE}

tractwise_gam_re <- readRDS("/cbica/projects/luo_wm_dev/output/HBN/tract_profiles/GAM/dti_md/check_resid_tractwise_gam_re.RData")
tractwise_bam <- readRDS("/cbica/projects/luo_wm_dev/output/HBN/tract_profiles/GAM/dti_md/check_resid_tractwise_bam2.RData")

```


I fit 2 tractwise models on a subset of HCP-D (full sample = 569; subsetted sample = 75 that represents the full age range)

1) Model 1: include nodeID as a smooth and subjectID as a random effect (most similar to GAM used in Tractable)
`gam(dti_md ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr') + ti(age, nodeID, k=c(3, 24), fx = F) + sex + mean_fd + s(subjectID, bs = 're'), data = df)`

2) Model 2: include `by = subjectID` in the nodeID smooth and also subjectID as a random effect 
`bam(dti_md ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr', by = subjectID) + ti(age, nodeID, k=c(3, 24), fx = F) + s(subjectID, bs = 're') + sex + mean_fd, data = df, method = "fREML", discrete = TRUE)`

 
# Residual plots for Model 1 for each tract (showing 15 subjects only)
- seeing some curved residuals: model may capture the trend of some data points better than others
- not quite centered around zero
```{r residual plot model 1, fig.width=17, fig.height=20}
tracts <- unique(gam_df$tract_hemi)

x.grob <- textGrob("Position on Tract (Node ID)", 
                 gp=gpar(col="black", fontsize=16))
y.grob <- textGrob("Residuals", 
                 gp=gpar(col="black", fontsize=16), rot=90)

space.grob <- textGrob(" ", 
                   gp=gpar(col="black", fontsize=28), rot=90)


plot_15_subs <- lapply(tracts, plot_compare_residuals, 15, model_number = 1)
names(plot_15_subs) <- tracts
 

plot_15_subs_final <- arrange_tracts_hemis(plot_15_subs, "Raw Residual Curves from Model 1: \n s(age, k = 3, fx = T, bs = 'cr') + \ns(nodeID, k = 24, fx = F, bs = 'cr') + \nti(age, nodeID, k=c(3, 24), fx = F) + \ns(subjectID, bs = 're') + sex + mean_fd")
grid.arrange(arrangeGrob(plot_15_subs_final, left = y.grob, bottom = x.grob, right = space.grob))
```


# Residual plots for Model 2 for each tract (showing 15 subjects only)
- seeing flat squiggles centered around zero - good!
- from Jeff: "Only thing I notice is that the boundaries often have sharp oscillations — but that’s a pretty minor issue due to how the smoother is implemented, and definitely not an easy fix. I’d say if y’all’re happy with these fits they should work." 
```{r residual plot model 2, fig.width=17, fig.height=20}
tracts <- unique(gam_df$tract_hemi)

  
x.grob <- textGrob("Position on Tract (Node ID)", 
                 gp=gpar(col="black", fontsize=16))
y.grob <- textGrob("Residuals", 
                 gp=gpar(col="black", fontsize=16), rot=90)

space.grob <- textGrob(" ", 
                   gp=gpar(col="black", fontsize=28), rot=90)


plot_15_subs <- lapply(tracts, plot_compare_residuals, 15, model_number = 2)
names(plot_15_subs) <- tracts
 

plot_15_subs_final <- arrange_tracts_hemis(plot_15_subs, "Smooth Residual Fit from Model 2: \n s(age, k = 3, fx = T, bs = 'cr') + \ns(nodeID, k = 24, fx = F, bs = 'cr', by = subjectID) + \nti(age, nodeID, k=c(3, 24), fx = F) + \ns(subjectID, bs = 're') + sex + mean_fd")
grid.arrange(arrangeGrob(plot_15_subs_final, left = y.grob, bottom = x.grob, right = space.grob))
```
 
 

# ACF plots for each type of model 
 

```{r acf function}
plot_compare_acf <- function(tract_list, ncols = 5, model_number) {
  num_plots <- length(tract_list)
  nrows <- 5
  
  par(mfrow = c(nrows, ncols))  # Set up the grid layout
  
  for (i in 1:num_plots) {
    tract <- tract_list[i]
    if(model_number == 1) {
      acf(resid(tractwise_gam_re[[tract]], type = "response"), main = paste0(tract))
    } else if (model_number == 2) {
      acf(resid(tractwise_bam[[tract]], type = "response"), main = paste0(tract))
    }
  }
  par(mfrow = c(1, 1))   
}
```

## Model 1
- ACFs are pretty high for a pretty long lag time
```{r model 1 acf, fig.height=15, fig.width=15}
plot_compare_acf(unique(gam_df$tract_hemi), model_number = 1)
```

## Model 2
- ACF appears to decrease, looks much better than model 1
```{r model 2 acf, fig.height=15, fig.width=15}
plot_compare_acf(unique(gam_df$tract_hemi), model_number = 2)
```
 
 
```{r test, eval = F, include = F, echo=F}
plot_compare_acf <- function(tract_list, output_file, ncols = 5, file_type = "pdf") {
  num_plots <- length(tract_list)
  nrows <- 5
  
  # Set up the plot device
  if (file_type == "pdf") {
    pdf(output_file)
  } else if (file_type == "png") {
    png(output_file, width = 1200, height = 800, res = 150)
  } else {
    stop("Unsupported file type. Choose 'pdf' or 'png'.")
  }
  
  par(mfrow = c(nrows, ncols))  # Set up the grid layout
  
  
  for (i in 1:num_plots) {
    tract <- tract_list[i]
    acf(resid(tractwise_bam[[tract]], type = "response"), main = paste0(tract))
   
  }
  
  #par(mfrow = c(1, 1))  # Reset to default single plot layout
  dev.off()  # Close the plot device
}


# Example usage to save plots to a PDF
plot_compare_acf(tract_list = unique(gam_df$tract_hemi),
                 output_file = "/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/both_datasets/acf_plots.pdf",
                 ncols = 5,
                 file_type = "pdf")
```
 
 