---
title: "Left Inferior Longitudinal Fasciculus"
author: "Audrey Luo"
output: 
  html_document:
    highlight: haddock
    code_folding: hide
    number_sections: no
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    number_sections: no
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

library(cowplot)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(gratia)
library(kableExtra)
library(mgcv)
library(RColorBrewer)
library(stringr)
library(tidyr)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, include = TRUE) 

font_size <- 18
theme_set(theme_minimal(base_family = "sans",base_size = font_size))
line_size <- 0.5
point_size <- 2
```


```{r load tract profiles, cache=TRUE}
cohortfile <- read.csv("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/cohortfiles/dti_fa/dti_fa_cohortfile.csv")
cohortfile <- cohortfile %>% rename(subjectID=sub) %>% select(subjectID, age, sex, mean_fd)

tract_profiles <- fread("/cbica/projects/luo_wm_dev/input/HCPD/HCPD_tractprofiles/all_subjects/collated_tract_profiles_HCPD_reoriented.tsv")

cohortfile <- cohortfile %>% select(subjectID, age, sex, mean_fd)
gam_df <- merge(tract_profiles, cohortfile, by="subjectID")
gam_df <- gam_df %>% mutate(tract_node = paste0(tract_hemi, "_", nodeID))
gam_df$sex <- as.factor(gam_df$sex)
```

```{r take subset of 75 subjects, cache=TRUE}
subjects <- read.table("/cbica/projects/luo_wm_dev/input/HCPD/subject_list/HCPD_subject_list_N569.txt")
subjects <- subjects %>% setNames("subjectID")
subject_subset <- subjects$subjectID[c(1:75)]
gam_df <- gam_df %>% filter(subjectID %in% subject_subset) # fit GAMs on a subset of 75 participants


gam_df_ilfl <- gam_df %>% filter(tract_hemi == "Left_Inferior_Longitudinal_Fasciculus")
write.csv(gam_df_ilfl, "/Users/audluo/Desktop/temp_code/jeff/gam_df_ilfl.csv")

gam_df <- gam_df_ilfl
```


```{r}
gam_df <- read.csv("/Users/audluo/Desktop/temp_code/jeff/gam_df_ilfl.csv") # this dataframe contains tract profiles data for left inferior longitudinal fasciculus and demographic data for 75 subjects
```

```{r function for fitting tractwise gam}

fit_tractwise_gamm_re <- function(tract_name, scalar) {
  df <- gam_df %>% filter(tract_hemi == tract_name)
  df$subjectID <- as.factor(df$subjectID)
  te_gam <- gamm(as.formula(sprintf("%s ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr') + 
                                  ti(age, nodeID, k=c(3, 24), fx = F) + sex + mean_fd", scalar)), random = list(subjectID=~1), data = df) 
  
  print(summary(te_gam$gam))
  print(k.check(te_gam$gam))
  print(paste(tract_name, "done"))
 
  
  return(te_gam)  # return the model object 
}


fit_tractwise_gam_re <- function(tract_name, scalar) {
  df <- gam_df %>% filter(tract_hemi == tract_name)
  df$subjectID <- as.factor(df$subjectID)
  te_gam <- gam(as.formula(sprintf("%s ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr') + 
                                  ti(age, nodeID, k=c(3, 24), fx = F) + sex + mean_fd + s(subjectID, bs = 're')", scalar)), data = df) 
  
  print(summary(te_gam))
  print(k.check(te_gam))
  print(paste(tract_name, "done"))
  
  
  return(te_gam)  # return the model object 
}

 

# 3) Using bam and using separate smooth function for subjectID
#`bam(dti_md ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr') + ti(age, nodeID, k=c(3, 24), fx = F) + sex + mean_fd + s(nodeID, by = subjectID), data = df, method = "fREML", discrete = TRUE)`

fit_tractwise_bam <- function(tract_name, scalar) {
  df <- gam_df %>% filter(tract_hemi == tract_name)
  df$subjectID <- as.factor(df$subjectID)
  te_gam <- bam(as.formula(sprintf("%s ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr', by = subjectID) + 
                                  ti(age, nodeID, k=c(3, 24), fx = F) + sex + mean_fd", scalar)), data = df, method = "fREML", discrete = TRUE) 
  
  print(summary(te_gam))
  print(k.check(te_gam))
  print(paste(tract_name, "done"))
 
  
  return(te_gam)  # return the model object 
}



```


```{r fit gams, eval=FALSE}
tractwise_gamm_re <- lapply(unique(gam_df$tract_hemi), fit_tractwise_gamm_re, scalar="dti_md")
names(tractwise_gamm_re) <- unique(gam_df$tract_hemi)
saveRDS(tractwise_gamm_re, "check_resid_tractwise_gamm_re.RData")

tractwise_gam_re <- lapply(unique(gam_df$tract_hemi), fit_tractwise_gam_re, scalar="dti_md")
names(tractwise_gam_re) <- unique(gam_df$tract_hemi)
saveRDS(tractwise_gam_re, "check_resid_tractwise_gam_re.RData")
 

tractwise_bam <- lapply(unique(gam_df$tract_hemi), fit_tractwise_bam, scalar="dti_md")
names(tractwise_bam) <- unique(gam_df$tract_hemi)
saveRDS(tractwise_bam, "check_resid_tractwise_bam.RData")
```

 
```{r load gam models, cache = TRUE, cache.lazy = FALSE}
tractwise_gamm_re <- readRDS("/Users/audluo/Desktop/temp_code/jeff/check_resid_tractwise_gamm_re.RData")
tractwise_gam_re <- readRDS("/Users/audluo/Desktop/temp_code/jeff/check_resid_tractwise_gam_re.RData")
#tractwise_bam <- readRDS("/Users/audluo/Desktop/temp_code/jeff/check_resid_tractwise_gam_smooth.RData")

tractwise_bam <- readRDS("/Users/audluo/Desktop/temp_code/jeff/check_resid_tractwise_bam.RData")

```


I fit 3 models for left inferior longitudinal fasciculus on a subset of HCP-D (full sample = 569; subsetted sample = 75 that represents the full age range)
 
 
1) Using gam with subjectID as a random effect:
`gam(dti_md ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr') + ti(age, nodeID, k=c(3, 24), fx = F) + sex + mean_fd + s(subjectID, bs = 're'), data = df)`

2) Using gamm and having subjectID as a random effect (gamm relies on lme, I believe):
`gamm(dti_md ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr') + ti(age, nodeID, k=c(3, 24), fx = F) + sex + mean_fd, random = list(subjectID=~1), data = df)`

3) Using bam and using separate smooth function for subjectID but with k=24
`bam(dti_md ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr', by = subjectID) + ti(age, nodeID, k=c(3, 24), fx = F) + sex + mean_fd, data = df, method = "fREML", discrete = TRUE)`


```{r function for comparing residuals}

plot_compare_residuals <- function(tract, num_subjects) { 
  gam_df_toplot <- gam_df %>% filter(tract_hemi == tract)
  
  # Extract residuals for each model
  gam_df_toplot$residuals_gam <- resid(tractwise_gam_re[[tract]])
  gam_df_toplot$residuals_gamm <- resid(tractwise_gamm_re[[tract]]$gam)
  gam_df_toplot$residuals_bam <- resid(tractwise_bam[[tract]])
  #gam_df_toplot$residuals_bam2 <- resid(tractwise_bam2[[tract]])
  
  subjects_to_plot <- subject_subset[c(1:num_subjects)]
    
  # Plot residuals for each model
  gam_plot <- ggplot(gam_df_toplot[gam_df_toplot$subjectID %in% subjects_to_plot, ], aes(x = nodeID, y = residuals_gam, color = factor(subjectID))) +
    geom_line() + theme(plot.title = element_text(hjust=0.5, size = 16, face="bold"), 
                        axis.title.x = element_blank(), 
                        axis.title.y = element_blank()) + 
    labs(title = "Raw Residual Curves from Model 1 \ngam: s(subjectID, bs = 're')", x = "Node ID", y = "Residuals")  
  
  gamm_plot <- ggplot(gam_df_toplot[gam_df_toplot$subjectID %in% subjects_to_plot, ], aes(x = nodeID, y = residuals_gamm, color = factor(subjectID))) +
    geom_line() + theme(plot.title = element_text(hjust=0.5, size = 16, face="bold"),
                        axis.title.x = element_blank(), 
                        axis.title.y = element_blank()) + 
    labs(title = "Raw Residual Curves from Model 2 \ngamm: random = list(subjectID=~1)", x = "Node ID", y = "Residuals")  
  
  # Plot smooth residuals from the new approach
  bam_plot <- ggplot(gam_df_toplot[gam_df_toplot$subjectID %in% subjects_to_plot, ], aes(x = nodeID, y = residuals_bam, color = factor(subjectID))) +
    geom_line() + theme(plot.title = element_text(hjust=0.5, size = 16, face = "bold"),
                        axis.title.x = element_blank(), 
                        axis.title.y = element_blank()) + 
    labs(title = "Smooth Residual Fit from Model 3 \nbam: s(nodeID, k=24, by = subjectID)", x = "Node ID", y = "Residuals")  
  
  #bam2_plot <- ggplot(gam_df_toplot[gam_df_toplot$subjectID %in% subjects_to_plot, ], aes(x = nodeID, y = residuals_bam2, color = factor(subjectID))) +
   # geom_line() + theme(plot.title = element_text(hjust=0.5, size = 16, face = "bold"),
      #                  axis.title.x = element_blank(), 
    #                    axis.title.y = element_blank()) + 
    #labs(title = "Smooth Residual Fit from Model 4 \nbam: s(nodeID, k=24, by = subjectID)", x = "Node ID", y = "Residuals")
  
  x.grob <- textGrob("Position on Tract (Node ID)", 
                   gp=gpar(col="black", fontsize=16))
  y.grob <- textGrob("Residuals", 
                   gp=gpar(col="black", fontsize=16), rot=90)
  
  z.grob <- textGrob(tract, 
                   gp=gpar(col="black", fontsize=20))
  
  plots <- ggarrange(gam_plot, gamm_plot, bam_plot, ncol=3, common.legend=T, legend = "right") 
  


  return(grid.arrange(arrangeGrob(plots, left = y.grob, bottom = x.grob, top = z.grob)))
}

```

# Residual plots for each type of model for each tract (showing 5 subjects only)
```{r, fig.width=15, fig.height=5}
tracts <- unique(gam_df$tract_hemi)
tracts <- tracts[-c(which(tracts=="Forceps_Major"))]
lapply(tracts, plot_compare_residuals, 5)
```
 

# Residual plots for each type of model for each tract (showing 15 subjects -- idk this more helpful to look at?)
```{r, fig.width=15, fig.height=5}
lapply(tracts, plot_compare_residuals, 15)

```

# ACF plots for each type of model 
```{r, fig.height=3, fig.width=10}

plot_compare_acf <- function(tract) {
  # Define the layout: 2 rows (1 for title, 1 for plots), 3 columns for plots
layout(matrix(c(1, 1, 1, 2, 3, 4), ncol = 3, byrow = TRUE), heights = c(1, 4))

# Plot the title
par(mar = c(0, 0, 1, 0))  # Set margins for the title
plot.new()
text(0.5, 0.5, tract, cex = 2, font = 2)

# Reset margins and set up for 1 row, 3 columns for the plots
par(mar = c(5, 4, 4, 2) + 0.1)

# Plot the ACFs
acf(resid(tractwise_gam_re[[tract]], type = "response"), main = paste0("Model 1 \ngam: s(subjectID, bs = 're')"))
acf(resid(tractwise_gamm_re[[tract]]$gam, type = "response"), main = paste0("Model 2 \ngamm: random = list(subjectID=~1)"))
acf(resid(tractwise_bam[[tract]], type = "response"), main = paste0("Model 3 \nbam: s(nodeID, k=24, by = subjectID)"))
}
 

#plot_compare_acf4 <- function(tract) {
  # Define the layout: 2 rows (1 for title, 1 for plots), 3 columns for plots
#  layout(matrix(c(1, 1, 1, 1, 2, 3, 4, 5), ncol = 4, byrow = TRUE), heights = c(1, 4))
  
  # Plot the title
#  par(mar = c(0, 0, 1, 0))  # Set margins for the title
#  plot.new()
#  text(0.5, 0.5, tract, cex = 2, font = 2)
  
  # Reset margins and set up for 1 row, 4 columns for the plots
#  par(mar = c(5, 4, 4, 2) + 0.1)
  
  # Plot the ACFs
#  acf(resid(tractwise_gam_re[[tract]], type = "response"), main = paste0("Model 1 \ngam: s(subjectID, bs = 're')"))
#  acf(resid(tractwise_gamm_re[[tract]]$gam, type = "response"), main = paste0("Model 2 \ngamm: random = list(subjectID=~1)"))
#  acf(resid(tractwise_bam[[tract]], type = "response"), main = paste0("Model 3 \nbam: s(nodeID, by = subjectID)"))
  #acf(resid(tractwise_bam2[[tract]], type = "response"), main = paste0("Model 4 \nbam: s(nodeID, k=24, by = subjectID)"))

#}
 
```


```{r, fig.height=3, fig.width=8}
lapply(unique(gam_df$tract_hemi), plot_compare_acf)
```

