---
title: 'Covariate Analyses: Sex'
author: "Audrey Luo"
output: html_document
---

```{r setup, include=FALSE}
library(cowplot)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(gratia) 
library(knitr)
library(kableExtra)
library(mgcv)
library(RColorBrewer)
library(scales)
library(stringr)
library(rjson)
library(tidyr)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, include = TRUE) 

font_size <- 20
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1
point_size <- 2

proj_root <- "/cbica/projects/luo_wm_dev/"
input_root <- paste0(proj_root, "input")
output_root <- paste0(proj_root, "output")
png_dir <- "/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/all_datasets/flux/"
local_dir <- "/Users/audluo/Desktop/temp_code"
 
text_size <- 20
default <- theme_set(theme_classic() + 
  theme(
    plot.title = element_text(size = text_size, hjust = 0.5),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_text(size = text_size),
    axis.text.y = element_text(size = text_size)
  ) 
) 
```

# age by sex interaction
```{r load age by sex interaction}
HCPD_sex_int <- read.csv("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_md/HCPD_GAM_agebysex_interaction.csv")
HBN_sex_int <- read.csv("/cbica/projects/luo_wm_dev/output/HBN/tract_profiles/GAM/dti_md/HBN_GAM_agebysex_interaction.csv")
PNC_sex_int <- read.csv("/cbica/projects/luo_wm_dev/output/PNC/tract_profiles/GAM/dti_md/PNC_GAM_agebysex_interaction.csv")
```


```{r fdr sex interaction}
# Function for calculating number of significant nodes (FDR corrected)
# @param df A dataframe with GAM results
sig_nodes <- function(df, pval_var) {

  df[[paste0(pval_var, ".fdr")]] <- p.adjust(df[[pval_var]], method=c("fdr"))
  
  df$significant.fdr <- df[[paste0(pval_var, ".fdr")]] < 0.05
  df$significant.fdr[df$significant.fdr == TRUE] <- 1
  df$significant.fdr[df$significant.fdr == FALSE] <- 0
  
  sigeffect.totaln <- sum(df$significant.fdr) #total number of significant connections 
  sigeffect.percent <- round(sigeffect.totaln/length(df$significant)*100, 2) #percent of significant connections
  cat(sprintf("There are %s/%s nodes that survive FDR (%s percent)\n", sum(df$significant.fdr == 1, na.rm=TRUE), nrow(df), sigeffect.percent))
  return(df)
}
 

sex_int_fdr <- lapply(list(HCPD_sex_int, HBN_sex_int, PNC_sex_int), sig_nodes, pval_var="GAM.int.pvalue")
names(sex_int_fdr) <- c("HCPD", "HBN", "PNC")
# HCPD: 63/2200 = 2.86% have significant age*sex interaction
# HBN: 0/2200 = 0% have significant age*sex interaction
# PNC: 957/2200 = 43.5% have significant age*sex interaction  
```

# sex main effect
```{r load sex main}
HCPD_sex_main <- read.csv("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_md/HCPD_GAM_sex_maineffects.csv")
HBN_sex_main <- read.csv("/cbica/projects/luo_wm_dev/output/HBN/tract_profiles/GAM/dti_md/HBN_GAM_sex_maineffects.csv")
PNC_sex_main <- read.csv("/cbica/projects/luo_wm_dev/output/PNC/tract_profiles/GAM/dti_md/PNC_GAM_sex_maineffects.csv")
```

```{r fdr sex main}
sex_main_fdr <- lapply(list(HCPD_sex_main, HBN_sex_main, PNC_sex_main), sig_nodes, pval_var="GAM.cov.pvalue")
names(sex_main_fdr) <- c("HCPD", "HBN", "PNC")
# HCPD: 773/2200 = 35.14% have significant main effect of sex 
# HBN: 1470/2200 = 66.82% have significant main effect of sex 
# PNC: 802/2200 = 36.45% have significant main effect of sex 

range(sex_main_fdr$HCPD$GAM.cov.tvalue)
range(sex_main_fdr$HBN$GAM.cov.tvalue)
range(sex_main_fdr$PNC$GAM.cov.tvalue)

```
```{r format sex effect dfs}
# format sex effect df's
format_sexeffect <- function(df) {
  # add tractID, tract_label, and nodeID
  df <- df %>% mutate(hemi = ifelse(grepl("Left", tract_node), "Left", 
                                    ifelse(grepl("Right", tract_node), "Right", NA))) %>%
    mutate(tractID = gsub("_[0-9]+", "", tract_node)) %>%
    mutate(tract_label = gsub("Left |Right ", "", gsub("_", " ", gsub("Fronto.", "Fronto-", tractID)))) %>%
    mutate(nodeID = str_extract(tract_node, "[0-9]+")) 
  df$nodeID <- as.numeric(df$nodeID)
  
  # label main orientation
  df <- df %>% 
  mutate(main_orientation = case_when(
    tract_label %in% c("Inferior Fronto-occipital",
                   "Inferior Longitudinal", "Superior Longitudinal") ~ "AP",
    tract_label %in% c("Arcuate") ~ "AP_frontal_temporal",
    tract_label %in% c("Callosum Anterior Frontal", "Callosum Motor", "Callosum Occipital", "Callosum Orbital", 
                       "Callosum Posterior Parietal", "Callosum Superior Frontal", "Callosum Superior Parietal", "Callosum Temporal") ~ "RL",
    tract_label %in% c("Corticospinal", "Posterior Arcuate","Vertical Occipital") ~ "SI",
    TRUE ~ NA_character_
  ))
  df$main_orientation <- as.factor(df$main_orientation)
  df <- df %>% relocate(tract_label, tractID, nodeID, tract_node, hemi, main_orientation)
  return(df)
}

sexeffect_dfs <- lapply(sex_main_fdr, format_sexeffect)
```

1) sex effect plot: GAM.cov.tvalue in y-axis, node position for each tract in x-axis
```{r plot sex effect function}

# function for plotting sex effect
plot_sex_effect_clipEnds <- function(tract, scalar, sexeffect_measure, color1, color2, color3, clipEnds, ylim1, ylim2) {
  # sexeffect_measure = GAM.smooth.AdjRsq or GAM.smooth.partialR2
    HCPD <- sexeffect_dfs$HCPD %>% filter(tract_label == {{tract}}) %>% filter(nodeID > (clipEnds-1) & nodeID < (99-clipEnds+1)) %>%
      mutate(Dataset = "HCPD")
    HBN <- sexeffect_dfs$HBN %>% filter(tract_label == {{tract}}) %>% filter(nodeID > (clipEnds-1) & nodeID < (99-clipEnds+1)) %>%
      mutate(Dataset = "HBN")
    PNC <- sexeffect_dfs$PNC %>% filter(tract_label == {{tract}}) %>% filter(nodeID > (clipEnds-1) & nodeID < (99-clipEnds+1)) %>%
      mutate(Dataset = "PNC")
   
    # NA out color/fill aes if adj rsq = 0 or if p-value doesn't survive FDR correction (makes the color gray)
    if (sexeffect_measure == "GAM.smooth.AdjRsq" | sexeffect_measure == "GAM.smooth.partialR2") {
      includes_zero_HCPD <- which(HCPD[[sexeffect_measure]]==0 | HCPD$GAM.cov.pvalue.fdr > 0.05)
      includes_zero_HBN <- which(HBN[[sexeffect_measure]]==0 | HBN$GAM.cov.pvalue.fdr > 0.05)
      includes_zero_PNC <- which(PNC[[sexeffect_measure]]==0 | PNC$GAM.cov.pvalue.fdr > 0.05)
    } else {
      includes_zero_HCPD <- which(is.na(HCPD[[sexeffect_measure]]) | HCPD[[sexeffect_measure]]==0 | HCPD$GAM.cov.pvalue.fdr > 0.05)
      includes_zero_HBN <- which(is.na(HBN[[sexeffect_measure]]) | HBN[[sexeffect_measure]]==0 | HBN$GAM.cov.pvalue.fdr > 0.05)
      includes_zero_PNC <- which(is.na(PNC[[sexeffect_measure]]) | PNC[[sexeffect_measure]]==0 | PNC$GAM.cov.pvalue.fdr > 0.05)
    }
    
    HCPD$Dataset[includes_zero_HCPD] <- NA
    HBN$Dataset[includes_zero_HBN] <- NA
    PNC$Dataset[includes_zero_PNC] <- NA
    
    # make df to plot
    df <- rbind(HCPD, HBN, PNC)
    
    if(str_detect(tract, "Callosum")) {
      plot <- ggplot(data = df, aes(x = nodeID, y = get(sexeffect_measure), colour = Dataset, fill = Dataset)) +
        geom_point(size=2, alpha = 0.8) + 
        geom_smooth(aes(group = 1), method = "loess", color = "gray40", fill = "gray50", alpha = 0.2, se = FALSE, size = 3) +
        scale_colour_manual(values = c("HCPD" = color1, "HBN" = color2, "PNC" = color3), na.value = "grey50") +
        scale_fill_manual(values = c("HCPD" = color1, "HBN" = color2, "PNC" = color3), na.value = "grey50") + 
        theme(legend.position = "bottom",
          legend.text = element_text(size = 20),
          legend.title = element_text(size = 20),
          legend.box = "vertical",
          axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          plot.title = element_text(size=20, hjust=0.5),
           plot.margin = unit(c(1, 1, 0.2, 0.2), "cm")) + labs(title = tract) + ylim(ylim1, ylim2) +
        guides(shape = guide_legend("Hemi", override.aes = list(alpha = 1, size = 6)),  # Ensure alpha = 1 in legend for shape
         colour = guide_legend(override.aes = list(alpha = 1, size = 6)))
      } else {
        plot <- ggplot(data = df, aes(x = nodeID, y = get(sexeffect_measure), colour = Dataset, 
                                      fill = Dataset, shape = hemi)) +
        geom_point(size=2, alpha = 0.8) + 
        geom_smooth(aes(group = 1), method = "loess", color = "gray40", fill = "gray50", alpha = 0.2, se = FALSE, size = 3) +
        scale_colour_manual(values = c("HCPD" = color1, "HBN" = color2, "PNC" = color3), na.value = "grey50") +
        scale_fill_manual(values = c("HCPD" = color1, "HBN" = color2, "PNC" = color3), na.value = "grey50") +
        scale_shape_manual(values = c(19,1)) + guides(shape = guide_legend("Hemi")) + 
        theme(legend.position = "bottom",
          legend.text = element_text(size = 20),
          legend.title = element_text(size = 20),
          legend.box = "vertical",
          axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          plot.title = element_text(size=20, hjust=0.5),
          plot.margin = unit(c(1, 1, 0.2, 0.2), "cm")) + labs(title = gsub("\\.", "-", tract)) + ylim(ylim1, ylim2) +
           guides(shape = guide_legend("Hemi", override.aes = list(alpha = 1, size = 6)),  # Ensure alpha = 1 in legend for shape
         colour = guide_legend(override.aes = list(alpha = 1, size = 6)))
        

      }
       
    return(plot)
}

arrange_callosum_plots <- function(list_figures, x_axis_grob, grob_hjust) {
  callosum1_plots <- ggarrange(list_figures$`Callosum Orbital` + theme(legend.position = "none"), 
                              list_figures$`Callosum Anterior Frontal` + theme(legend.position = "none"),  
                              list_figures$`Callosum Superior Frontal`+ theme(legend.position = "none"), 
                              list_figures$`Callosum Motor`+ theme(legend.position = "none"), ncol = 1, nrow = 4)
  callosum1_plots <- annotate_figure(callosum1_plots, bottom = text_grob(x_axis_grob, color = "black", size = 20, hjust = grob_hjust))
  
  callosum2_plots <- ggarrange(list_figures$`Callosum Superior Parietal`+ theme(legend.position = "none"), 
                              list_figures$`Callosum Posterior Parietal`+ theme(legend.position = "none"), 
                              list_figures$`Callosum Temporal`+ theme(legend.position = "none"), 
                              list_figures$`Callosum Occipital`+ theme(legend.position = "none"),  ncol = 1, nrow = 4)
  callosum2_plots <- annotate_figure(callosum2_plots, bottom = text_grob(x_axis_grob, color = "black", size = 20, hjust = grob_hjust))
  
  callosum_plots <- ggarrange(callosum1_plots, callosum2_plots)
  legend <- get_legend(list_figures$`Callosum Motor`)
  callosum_plot_final <- ggarrange(callosum_plots, legend, nrow=2, heights = c(2, 0.2))
   return(callosum_plot_final)
}
 
 
arrange_tract_plots <- function(list_figures, x_axis_grob, grob_hjust) {
  AP_plots <- ggarrange(list_figures$`Arcuate` + theme(legend.position = "none"),
                        list_figures$`Inferior Fronto-occipital`+ theme(legend.position = "none"), 
                        list_figures$`Inferior Longitudinal`+ theme(legend.position = "none"), 
                        list_figures$`Superior Longitudinal` + theme(legend.position = "none"), ncol=1, nrow = 4)
  AP_plots_final <- annotate_figure(AP_plots, top = text_grob("Anterior - Posterior", 
                 color = "black", face = "bold", size = 20, hjust = 0.4), 
                 bottom = text_grob(x_axis_grob, color = "black", size = 20, hjust = grob_hjust))
  
   
  SI_plots <- ggarrange(list_figures$`Posterior Arcuate` + theme(legend.position = "none"), 
                        list_figures$`Vertical Occipital` + theme(legend.position = "none"), 
                        list_figures$`Corticospinal` + theme(legend.position = "none"),
                           ncol=1, nrow = 3)
  legend <- get_legend(list_figures$`Posterior Arcuate`)
 
  SI_plots <- annotate_figure(SI_plots, top = text_grob("Superior - Inferior", 
                 color = "black", face = "bold", size = 20, hjust = 0.4),
                 bottom = text_grob(x_axis_grob, color = "black", size = 20, hjust = grob_hjust))

  SI_plots_final <- ggarrange(SI_plots, NULL, legend, heights = c(2.1, 0.67, 0.5))
  tractprofiles_plot_final <- ggarrange(AP_plots_final, SI_plots_final, ncol = 2, widths = c(1, 2))
   
 
   return(tractprofiles_plot_final) 
}
```


```{r set colors}
color1 = "#3FB8AFFF"
color2 = "#FF9E9DFF"
color3 = "#cc0468"
```

```{r run sex effect plotting}
sexeffect_measure = "GAM.cov.tvalue"
tracts = unique(sexeffect_dfs$HCPD$tract_label)
sexeffects_plots <- lapply(tracts, plot_sex_effect_clipEnds, scalar = "dti_md", sexeffect_measure = sexeffect_measure, 
                                color1 = color1, color2 = color2, color3 = color3, clipEnds = 3, ylim2 = 8, ylim1 = -8)
names(sexeffects_plots) <- tracts
```


fyi: "F" is reference levels in model. Positive t-val: MD higher in males. Negative t-val: MD higher in females.
```{r sex effects, fig.height = 20, fig.width = 16}
sexeffects_assoc_plots <- arrange_tract_plots(sexeffects_plots, "Position on Tract (Node ID)", 0.3)
 
y.grob <- textGrob(expression("Sex Effect (t-value)"), 
                   gp=gpar(col="black", fontsize=20), rot=90)

grid.arrange(arrangeGrob(sexeffects_assoc_plots, left = y.grob))

ggsave(paste0(png_dir, "sexeffect_association.png"), grid.arrange(arrangeGrob(sexeffects_assoc_plots, left = y.grob)), height = 20, width = 16, units = "in")
```

```{r dti md tracts sd callosum, fig.height = 20, fig.width = 10}

sexeffects_callosum_plots <- arrange_callosum_plots(sexeffects_plots, "Position on Tract (Node ID)", 0.3)
 
y.grob <- textGrob(expression("Sex Effect (t-value)"), 
                   gp=gpar(col="black", fontsize=20), rot=90)

grid.arrange(arrangeGrob(sexeffects_callosum_plots, left = y.grob))

ggsave(paste0(png_dir, "sexeffect_callosum.png"), grid.arrange(arrangeGrob(sexeffects_callosum_plots, left = y.grob)), height = 20, width = 10, units = "in")
```

2) developmental trajectories separate for females and males. 
  - a: for all nodes for each tract
  - b: do f/m for each end and deep WM. So, end1: f/m, end2: f/m, deep: f/m. where are the differences the greatest? guessing peripheral.
```{r load files for dev traj all nodes}
HCPD_smooths <- read.csv("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_md/HCPD_GAM_agebysex_interaction_fits.csv")
HBN_smooths <- read.csv("/cbica/projects/luo_wm_dev/output/HBN/tract_profiles/GAM/dti_md/HBN_GAM_agebysex_interaction_fits.csv")
PNC_smooths <- read.csv("/cbica/projects/luo_wm_dev/output/PNC/tract_profiles/GAM/dti_md/PNC_GAM_agebysex_interaction_fits.csv")

HCPD.smooths <- format_sexeffect(HCPD_smooths %>% rename(tract_node = tract, age = smooth_var))
HBN.smooths <- format_sexeffect(HBN_smooths %>% rename(tract_node = tract, age = smooth_var))
PNC.smooths <- format_sexeffect(PNC_smooths %>% rename(tract_node = tract, age = smooth_var))
```

```{r dev traj all nodes female}
```

```{r dev traj all nodes male}
# maybe not do bc this is a lot to visualize right now
 
```


```{r dev traj ends and deep}

color1 = "#009DA8FF"
color2 = "#F0BA52FF"
color3 = "#C73000FF"

plot_smooths_binned <- function(tract, df, bin_num_nodes, clipEnds) {
  df$nodeID <- as.numeric(df$nodeID)
  df <- df %>% filter(tract_label == tract) %>% 
    filter(nodeID > (clipEnds-1) & nodeID < (99-clipEnds+1)) %>%
    group_by(tract_node)
  
 
  df_binned <- df %>%
        filter(
          nodeID < (bin_num_nodes + clipEnds) |  # Start of the tract
          nodeID > (99 - bin_num_nodes - clipEnds) |  # End of the tract
         (nodeID > (50-bin_num_nodes) & nodeID <= (50+bin_num_nodes)) # Deep WM nodes: the 30 nodes in the center
        )  %>%
        mutate(
          node_position = case_when(
            nodeID < bin_num_nodes + clipEnds ~ "end1",
            nodeID > 99 - bin_num_nodes - clipEnds ~ "end2",
            (nodeID > (50-bin_num_nodes) & nodeID <= (50+bin_num_nodes)) ~ "deep",
            TRUE ~ NA_character_
          )
        ) %>%
        group_by(node_position, age, sex) %>% # removed hemi
        summarise(
          mean_est = mean(fitted, na.rm = TRUE)
        )
  
  plot <- ggplot(df_binned, aes(x = age, y = mean_est, linetype = sex, group = interaction(sex, node_position))) +
      geom_line(aes(colour = node_position), size = 3.5, alpha = 0.9) + 
      theme_classic() +                         
      theme(text=element_text(size=24),
            axis.text.x = element_text(size=24),
            axis.text.y = element_text(size=24),
            legend.position = "bottom",
            legend.box = "vertical", 
            plot.title = element_text(hjust=0.5, size = 24)) + 
      scale_colour_manual(values = c("end1" = color1, "deep" = color2, "end2" = color3), na.value = "grey50") + 
      labs(x="", y="", title=tract, colour="Tract Region") + guides(linetype=guide_legend(keywidth = 4, keyheight = 1))

   return(plot)
}
```



```{r plot binned smooths}
bin_num_nodes = 10
clipEnds = 3

tracts <- unique(HCPD.smooths$tract_label)

binned_smooths_HCPD <- lapply(tracts, plot_smooths_binned, df = HCPD.smooths, 
                              bin_num_nodes = bin_num_nodes, clipEnds = clipEnds)
names(binned_smooths_HCPD) <- tracts

binned_smooths_HBN <- lapply(tracts, plot_smooths_binned, df = HBN.smooths, 
                             bin_num_nodes = bin_num_nodes, clipEnds = clipEnds)
names(binned_smooths_HBN) <- tracts

binned_smooths_PNC <- lapply(tracts, plot_smooths_binned, df = PNC.smooths, 
                             bin_num_nodes = bin_num_nodes, clipEnds = clipEnds)
names(binned_smooths_PNC) <- tracts

```



```{r arrange binned smooths HCPD, fig.height = 20, fig.width = 20}
y.grob <- textGrob("Fitted DTI MD", 
                   gp=gpar(col="black", fontsize=24), rot=90)

binned_smooths_HCPD_main <- arrange_tract_plots(binned_smooths_HCPD, "Age", -1)
filename <- "fitted_sex"
ggsave(paste0(png_dir, "dti_md_", filename, "_HCPD.png"), grid.arrange(arrangeGrob(binned_smooths_HCPD_main, left = y.grob)), width = 20, height = 20, units = "in")
```

```{r arrange binned smooths callosum HCPD, fig.height = 20, fig.width = 12}
y.grob <- textGrob("Fitted DTI MD", 
                   gp=gpar(col="black", fontsize=24), rot=90)

binned_smooths_HCPD_callosum <- arrange_callosum_plots(binned_smooths_HCPD, "Age", -1)
filename <- "fitted_sex_callosum"
ggsave(paste0(png_dir, "dti_md_", filename, "_HCPD.png"), grid.arrange(arrangeGrob(binned_smooths_HCPD_callosum, left = y.grob)), width = 12, height = 20, units = "in")
```


```{r arrange binned smooths HBN, fig.height = 20, fig.width = 20}
y.grob <- textGrob("Fitted DTI MD", 
                   gp=gpar(col="black", fontsize=24), rot=90)

binned_smooths_HBN_main <- arrange_tract_plots(binned_smooths_HBN, "Age", -1)
filename <- "fitted_sex"
ggsave(paste0(png_dir, "dti_md_", filename, "_HBN.png"), grid.arrange(arrangeGrob(binned_smooths_HBN_main, left = y.grob)), width = 20, height = 20, units = "in")
```

```{r arrange binned smooths callosum HBN, fig.height = 20, fig.width = 12}
y.grob <- textGrob("Fitted DTI MD", 
                   gp=gpar(col="black", fontsize=24), rot=90)

binned_smooths_HBN_callosum <- arrange_callosum_plots(binned_smooths_HBN, "Age", -1)
filename <- "fitted_sex_callosum"
ggsave(paste0(png_dir, "dti_md_", filename, "_HBN.png"), grid.arrange(arrangeGrob(binned_smooths_HBN_callosum, left = y.grob)), width = 12, height = 20, units = "in")
```


```{r arrange binned smooths PNC, fig.height = 20, fig.width = 20}
y.grob <- textGrob("Fitted DTI MD", 
                   gp=gpar(col="black", fontsize=24), rot=90)

binned_smooths_PNC_main <- arrange_tract_plots(binned_smooths_PNC, "Age", -1)
filename <- "fitted_sex"
ggsave(paste0(png_dir, "dti_md_", filename, "_PNC.png"), grid.arrange(arrangeGrob(binned_smooths_PNC_main, left = y.grob)), width = 20, height = 20, units = "in")
```

```{r arrange binned smooths callosum PNC, fig.height = 20, fig.width = 12}
y.grob <- textGrob("Fitted DTI MD", 
                   gp=gpar(col="black", fontsize=24), rot=90)

binned_smooths_PNC_callosum <- arrange_callosum_plots(binned_smooths_PNC, "Age", -1)
filename <- "fitted_sex_callosum"
ggsave(paste0(png_dir, "dti_md_", filename, "_PNC.png"), grid.arrange(arrangeGrob(binned_smooths_PNC_callosum, left = y.grob)), width = 12, height = 20, units = "in")
```