---
title: "Prep data for visualization"
author: "Audrey Luo"
output: html_document
---

```{r setup, include=FALSE}
library(cowplot)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(gratia) 
library(kableExtra)
library(mgcv)
library(RColorBrewer)
library(stringr)
library(tidyr)
library(rjson)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, include = TRUE) 

font_size <- 14
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1
point_size <- 2

input_root <- "/cbica/projects/luo_wm_dev/input"
output_root <- "/cbica/projects/luo_wm_dev/output"
```


# Tract Names and Orientations
https://yeatmanlab.github.io/pyAFQ/explanations/bundle_orientation.html#bundle-orientation 
- Left Anterior Thalamic: anterior to posterior
- Right Anterior Thalamic: anterior to posterior
- Left Cingulum Cingulate: anterior to posterior
- Right Cingulum Cingulate: anterior to posterior
- Left Corticospinal: superior to inferior
- Right Corticospinal: superior to inferior
- Left Inferior Fronto-occipital: anterior to posterior
- Right Inferior Fronto-occipital: anterior to posterior
- Left Inferior Longitudinal: anterior to posterior
- Right Inferior Longitudinal: anterior to posterior
- Left Superior Longitudinal: anterior to posterior
- Right Superior Longitudinal: anterior to posterior
- Left Arcuate: anterior to posterior
- Right Arcuate: anterior to posterior
- Left Uncinate: superior to inferior
- Right Uncinate: superior to inferior
- Forceps Minor: right to left
- Forceps Major: right to left
- Left Posterior Arcuate: superior to inferior
- Right Posterior Arcuate: superior to inferior
- Left Vertical Occipital: superior to inferior
- Right Vertical Occipital: superior to inferior
```{r}
dataset="PNC"
config_data <- fromJSON(file=sprintf("/cbica/projects/luo_wm_dev/code/tract_profiles/config/config_%1$s.json", dataset))
demographics <- read.csv(config_data$demo_qc)
data_root <- config_data$tract_profiles_root

all_subjects <- fread(sprintf("%1$s/all_subjects/collated_tract_profiles_nocovbat_unc_cc.tsv", data_root))
all_subjects$tractID <- gsub("Fronto-occipital", "Fronto.occipital", all_subjects$tractID)
all_subjects <- all_subjects %>% mutate(hemi = ifelse(grepl("Left", tractID), "Left", "Right")) %>% 
  mutate(tract_node = gsub(" ", "_", paste0(tractID, "_", nodeID)))
all_subjects$sub <- as.factor(all_subjects$sub)
all_subjects <- all_subjects %>% 
  mutate(nodeID = str_extract(tract_node, "[0-9]+")) %>%
  mutate(tractID = gsub("_[0-9]+", "", tract_node)) %>%
  mutate(hemi = str_extract(tractID, "Left|Right"))
all_subjects$nodeID <- as.numeric(all_subjects$nodeID)
all_subjects$sub <- as.factor(all_subjects$sub)
all_subjects <- all_subjects %>% select(sub, tract_node, nodeID, tractID, hemi, dti_fa, dti_md) %>% arrange(sub, tractID, nodeID, hemi)
saveRDS(all_subjects, sprintf("%1$s/all_subjects/collated_tract_profiles_final_unc_cc.RData", data_root))

tract_profiles_long <- all_subjects
```


```{r load tract profiles, cache=TRUE}
HBN <- readRDS(sprintf("%1$s/%2$s/derivatives/tract_profiles/all_subjects/collated_tract_profiles_final_unc_cc.RData", input_root, "HBN"))
PNC <- readRDS(sprintf("%1$s/%2$s/derivatives/tract_profiles/all_subjects/collated_tract_profiles_final_unc_cc.RData", input_root, "PNC"))

# format tract profiles df's
format_tract_profiles <- function(df) {
  
  # make tract_label
  df <- df %>% mutate(tract_label = gsub("_", " ", tractID) %>%
                       gsub("Left|Right", "", .) %>%
                       gsub("Fronto.occipital", "Fronto-occipital", .) %>% # change to hyphen for plotting purposes
                       trimws())
  
  
  return(df)
}


#HCPD_tp <- format_tract_profiles(HCPD)
HBN_tp <- format_tract_profiles(HBN)
PNC_tp <- format_tract_profiles(PNC)
   
#saveRDS(HCPD_tp, sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz.RData", output_root, "HCPD"))
#saveRDS(HBN_tp, sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz.RData", output_root, "HBN"))
#saveRDS(PNC_tp, sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz.RData", output_root, "PNC"))

saveRDS(HBN_tp, sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz_unc_cc.RData", output_root, "HBN"))
saveRDS(PNC_tp, sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz_unc_cc.RData", output_root, "PNC"))
```


```{r}

HBN_md <- HBN_tp %>% group_by(sub, tractID) %>% summarise(mean_dti_md = mean(dti_md)) %>% pivot_wider(names_from = "tractID", values_from = "mean_dti_md")
HBN_fa <- HBN_tp %>% group_by(sub, tractID) %>% summarise(mean_dti_fa = mean(dti_fa)) %>% pivot_wider(names_from = "tractID", values_from = "mean_dti_fa")

 

PNC_md <- PNC_tp %>% group_by(sub, tractID) %>% summarise(mean_dti_md = mean(dti_md)) %>% pivot_wider(names_from = "tractID", values_from = "mean_dti_md")
PNC_fa <- PNC_tp %>% group_by(sub, tractID) %>% summarise(mean_dti_fa = mean(dti_fa)) %>% pivot_wider(names_from = "tractID", values_from = "mean_dti_fa")

length(unique(HBN_md$sub))
#write.csv(HBN_md, paste0(output_root, "/HBN/tract_profiles/all_subjects/HBN_tract_average_dti_md.csv"), row.names=F)
#write.csv(HBN_fa, paste0(output_root, "/HBN/tract_profiles/all_subjects/HBN_tract_average_dti_fa.csv"), row.names=F)

#write.csv(PNC_md, paste0(output_root, "/PNC/tract_profiles/all_subjects/PNC_tract_average_dti_md.csv"), row.names=F)
#write.csv(PNC_fa, paste0(output_root, "/PNC/tract_profiles/all_subjects/PNC_tract_average_dti_fa.csv"), row.names=F)

write.csv(HBN_md, paste0(output_root, "/HBN/tract_profiles/all_subjects/HBN_tract_average_dti_md_unc_cc.csv"), row.names=F)
write.csv(HBN_fa, paste0(output_root, "/HBN/tract_profiles/all_subjects/HBN_tract_average_dti_fa_unc_cc.csv"), row.names=F)

write.csv(PNC_md, paste0(output_root, "/PNC/tract_profiles/all_subjects/PNC_tract_average_dti_md_unc_cc.csv"), row.names=F)
write.csv(PNC_fa, paste0(output_root, "/PNC/tract_profiles/all_subjects/PNC_tract_average_dti_fa_unc_cc.csv"), row.names=F)

write.csv(HBN_md, "/Users/audluo/Desktop/HBN_tract_average_dti_md_unc_cc.csv", row.names=F)
write.csv(HBN_fa, "/Users/audluo/Desktop/HBN_tract_average_dti_fa_unc_cc.csv", row.names=F)

write.csv(PNC_md, "/Users/audluo/Desktop/PNC_tract_average_dti_md_unc_cc.csv", row.names=F)
write.csv(PNC_fa, "/Users/audluo/Desktop/PNC_tract_average_dti_fa_unc_cc.csv", row.names=F)
```

