---
title: "Explore baseline tract profiles"
author: "Audrey Luo"
output: html_document
---

```{r setup, include=FALSE}
library(cowplot)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(gratia) 
library(knitr)
library(kableExtra)
library(mgcv)
library(RColorBrewer)
library(stringr)
library(tidyr)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, include = TRUE) 

font_size <- 14
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1
point_size <- 2

input_root <- "/cbica/projects/luo_wm_dev/input"
output_root <- "/cbica/projects/luo_wm_dev/output"
png_dir <- "/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/both_datasets/act_maxbval/"
  
```


# Tract Names and Orientations
https://yeatmanlab.github.io/pyAFQ/explanations/bundle_orientation.html#bundle-orientation 
- All callosum bundles: right to left
- Left Corticospinal: superior to inferior
- Right Corticospinal: superior to inferior
- Left Inferior Fronto-occipital: anterior to posterior
- Right Inferior Fronto-occipital: anterior to posterior
- Left Inferior Longitudinal: anterior to posterior
- Right Inferior Longitudinal: anterior to posterior
- Left Superior Longitudinal: anterior to posterior
- Right Superior Longitudinal: anterior to posterior
- Left Arcuate: anterior to posterior
- Right Arcuate: anterior to posterior
- Forceps Minor: right to left
- Forceps Major: right to left
- Left Posterior Arcuate: superior to inferior
- Right Posterior Arcuate: superior to inferior
- Left Vertical Occipital: superior to inferior
- Right Vertical Occipital: superior to inferior

```{r load tract profiles, cache=TRUE}
HCPD <- readRDS(sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz.RData", output_root, "HCPD"))
HBN <- readRDS(sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz.RData", output_root, "HBN"))
PNC <- readRDS(sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz.RData", output_root, "PNC"))

HCPD <- readRDS(sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz_combat_all_site.RData", output_root, "HCPD"))
HBN <- readRDS(sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz_combat_all_site.RData", output_root, "HBN"))
PNC <- readRDS(sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz_combat_all_site.RData", output_root, "PNC"))

 
```

# plot baseline for each measure: dki_fa, dki_md, dti_fa, dti_md and maybe compare
```{r functions for visualizing baseline tract profiles}


# make summary df's: this creates summary_[dataset]_[scalar] dataframes that include mean, sd, se for each scalar/dataset
make_summary_dfs <- function(scalar, dataset_name, df) {
  # skip PNC when scalar is dki_md or dki_fa
  if (dataset_name == "PNC" && (scalar == "dki_md" || scalar == "dki_fa")) {
    return(NULL)  
  }
  # set variables
  mean_scalar <- paste0("mean_", scalar)
  # Process the given dataset df
  scalar_data <- df %>% select(sub, tractID, tract_label, nodeID, hemi, all_of(scalar))
  sqrt_n <- sqrt(length(unique(scalar_data$sub)))
  setDT(scalar_data)
  summary_data <- scalar_data[, .(
    mean_scalar = mean(.SD[[1]], na.rm = TRUE),
    sd = sd(.SD[[1]], na.rm = TRUE),
    se = sd(.SD[[1]], na.rm = TRUE) / sqrt_n,
    ymin_sd = mean(.SD[[1]], na.rm = TRUE) - sd(.SD[[1]], na.rm = TRUE),
    ymax_sd = mean(.SD[[1]], na.rm = TRUE) + sd(.SD[[1]], na.rm = TRUE),
    ymin_se = mean(.SD[[1]], na.rm = TRUE) - (sd(.SD[[1]], na.rm = TRUE) / sqrt_n),
    ymax_se = mean(.SD[[1]], na.rm = TRUE) + (sd(.SD[[1]], na.rm = TRUE) / sqrt_n)
  ), by = .(tract_label, tractID, nodeID, hemi), .SDcols = scalar]

  setnames(summary_data, "mean_scalar", paste0("mean_", scalar))
  summary_data[, Dataset := dataset_name]
  summary_dataset_scalar <- paste0("summary_", dataset_name, "_", scalar)
  assign(summary_dataset_scalar, summary_data, envir = .GlobalEnv)
}


# function for plotting tractprofiles mean and sd or se for a given scalar
plot_mean_var <- function(tract, scalar, variance_measure, color1, color2, color3, ylim1 = 0, ylim2 = 1) {
 
  mean_scalar <- paste0("mean_", scalar)

  if (scalar == "dti_fa" | scalar == "dti_md") {
    summary_HCPD <- get(paste0("summary_HCPD_", scalar))
    summary_HBN <- get(paste0("summary_HBN_", scalar))
    summary_PNC <- get(paste0("summary_PNC_", scalar))
    
    df_profiles <- rbind(summary_HCPD, summary_HBN, summary_PNC)
    df <- df_profiles %>% filter(tract_label == tract)
    df$Dataset <- factor(df$Dataset, levels = c("PNC", "HCPD", "HBN"))
      # plot
      if(str_detect(tract, "Callosum")) {
        plot <- ggplot(data = df, aes(x = nodeID, y = get(mean_scalar))) +
        geom_ribbon(data = df, aes_string(x = "nodeID", ymin = paste0("ymin_", variance_measure), ymax = paste0("ymax_", variance_measure), fill = "Dataset"), alpha = .3) + 
        geom_line(data = df, aes(x = nodeID, y = get(mean_scalar), color = Dataset),size = 1) +
        scale_fill_manual(values = c("HCPD" = color1, "HBN" = color2, "PNC" = color3)) +
        scale_color_manual(values = c("HCPD" = color1, "HBN" = color2, "PNC" = color3)) +
        theme(legend.position = "none",
              legend.text = element_text(size = 20),
              legend.title = element_text(size = 20),
              axis.text.x = element_text(size = 20),
              axis.text.y = element_text(size = 20),
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=20, hjust=0.5)) + labs(title = tract) + ylim(ylim1, ylim2)
        
      } else {
        plot <- ggplot(data = df, aes(x = nodeID, y = get(mean_scalar), linetype = hemi)) +
        geom_ribbon(data = df, aes_string(x = "nodeID", ymin = paste0("ymin_", variance_measure), ymax = paste0("ymax_", variance_measure), fill = "Dataset"), alpha = .3) + 
        geom_line(data = df, aes(x = nodeID, y = get(mean_scalar), color = Dataset),size = 1) +
          guides(linetype = guide_legend("Hemi", override.aes = list(fill = "white"))) + 
        scale_fill_manual(values = c("HCPD" = color1, "HBN" = color2, "PNC" = color3)) +
        scale_color_manual(values = c("HCPD" = color1, "HBN" = color2, "PNC" = color3)) +
        theme(legend.position = "none",
              legend.text = element_text(size = 20),
              legend.title = element_text(size = 20),
              axis.text.x = element_text(size = 20),
              axis.text.y = element_text(size = 20),
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=20, hjust=0.5)) + labs(title = tract) + ylim(ylim1, ylim2)
      }
    return(plot)
  } else {
    summary_HCPD <- get(paste0("summary_HCPD_", scalar))
    summary_HBN <- get(paste0("summary_HBN_", scalar))
      
    df_profiles <- rbind(summary_HCPD, summary_HBN)
    df <- df_profiles %>% filter(tract_label == tract)
     
    if(str_detect(tract, "Callosum")) {
      plot <- ggplot(data = df, aes(x = nodeID, y = get(mean_scalar))) +
      geom_ribbon(data = df, aes_string(x = "nodeID", ymin = paste0("ymin_", variance_measure), ymax = paste0("ymax_", variance_measure), fill = "Dataset"), alpha = .3) + 
      geom_line(data = df, aes(x = nodeID, y = get(mean_scalar), color = Dataset),size = 1) +
      scale_fill_manual(values = c("HCPD" = color1, "HBN" = color2)) +
      scale_color_manual(values = c("HCPD" = color1, "HBN" = color2)) +
      theme(legend.position = "none",
            legend.text = element_text(size = 20),
            legend.title = element_text(size = 20),
            axis.text.x = element_text(size = 20),
            axis.text.y = element_text(size = 20),
            axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            plot.title = element_text(size=20, hjust=0.5)) + labs(title = tract) + ylim(ylim1, ylim2)
      
    } else {
      plot <- ggplot(data = df, aes(x = nodeID, y = get(mean_scalar), linetype = hemi)) +
      geom_ribbon(data = df, aes_string(x = "nodeID", ymin = paste0("ymin_", variance_measure), ymax = paste0("ymax_", variance_measure), fill = "Dataset"), alpha = .3) + 
      geom_line(data = df, aes(x = nodeID, y = get(mean_scalar), color = Dataset),size = 1) +
        guides(linetype = guide_legend("Hemi",override.aes = list(fill = "white"))) + 
      scale_fill_manual(values = c("HCPD" = color1, "HBN" = color2)) +
      scale_color_manual(values = c("HCPD" = color1, "HBN" = color2)) +
      theme(legend.position = "none",
            legend.text = element_text(size = 20),
            legend.title = element_text(size = 20),
            axis.text.x = element_text(size = 20),
            axis.text.y = element_text(size = 20),
            axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            plot.title = element_text(size=20, hjust=0.5)) + labs(title = tract) + ylim(ylim1, ylim2)
    }
    return(plot)
  }
  
}


arrange_tract_plots <- function(list_figures, title) {

  AP_plots <- ggarrange(list_figures$`Arcuate`,
                        list_figures$`Inferior Fronto-occipital`, 
                        list_figures$`Inferior Longitudinal`, 
                        list_figures$`Superior Longitudinal`, ncol=4, nrow = 1)
  AP_plots_final <- annotate_figure(AP_plots, top = text_grob("Anterior - Posterior", 
                 color = "black", face = "bold", size = 20))
  
   
  SI_plots <- ggarrange(list_figures$`Corticospinal`, list_figures$`Posterior Arcuate`, 
                           list_figures$`Vertical Occipital`, common.legend=TRUE, legend=c("right"),
                           ncol=3, nrow = 1)
  SI_plots_final <- annotate_figure(SI_plots, top = text_grob("Superior - Inferior", 
                 color = "black", face = "bold", size = 20))
  
  
  callosum_plots <- ggarrange(list_figures$`Callosum Anterior Frontal`, 
                           list_figures$`Callosum Motor`, 
                           list_figures$`Callosum Occipital`, 
                           list_figures$`Callosum Orbital`, 
                           list_figures$`Callosum Posterior Parietal`,
                           list_figures$`Callosum Superior Frontal`, 
                           list_figures$`Callosum Superior Parietal`, 
                           list_figures$`Callosum Temporal`, ncol=4, nrow = 2)
  callosum_plots_final <- annotate_figure(callosum_plots, top = text_grob("Right - Left", 
                 color = "black", face = "bold", size = 20))
  
  tractprofiles_plot <- ggarrange(callosum_plots_final, ggarrange(AP_plots_final, SI_plots_final, widths = c(1, 1.2), nrow = 2), heights = c(1, 1.2),nrow=2)
  
  tractprofiles_plot_final <- annotate_figure(tractprofiles_plot, top = text_grob(title, 
               color = "black", face = "italic", size = 20))
   return(tractprofiles_plot_final)
}


 
```

 


```{r make summary dfs}

datasets <- list(HCPD = HCPD, HBN = HBN, PNC = PNC)
#scalars <- c("dki_md", "dti_md", "dki_fa", "dti_fa")
#scalars <- c("dti_md", "dti_fa")
scalars <- c("dti_md")

# create summary dfs
for (dataset_name in names(datasets)) {
  dataset <- datasets[[dataset_name]]
  lapply(scalars, function(scalar) {
    make_summary_dfs(scalar, dataset_name, dataset) # this creates summary_[dataset]_[scalar] dataframes that include mean, sd, se for each scalar
  })
}
 
```

```{r set colors}
color1 = "#3FB8AFFF"
color2 = "#FF9E9DFF"
color3 = "#cc0468"
```

 
## DTI MD
```{r run dti md sd plotting}
variance_measure = "sd"
dti_md_sd_plots <- lapply(unique(HCPD$tract_label), plot_mean_var, scalar = "dti_md", variance_measure = variance_measure, 
                                color1 = color1, color2 = color2, color3 = color3, ylim1 = 0.00058, ylim2 = 0.0012)
names(dti_md_sd_plots) <- unique(HCPD$tract_label)
```

```{r dti md tracts sd, fig.height = 20, fig.width = 20}
dti_md_final <- arrange_tract_plots(dti_md_sd_plots, "DTI MD: Mean +/- 1SD")
x.grob <- textGrob("Position on Tract (Node ID)", 
                   gp=gpar(col="black", fontsize=20))
y.grob <- textGrob("Mean Diffusivity", 
                   gp=gpar(col="black", fontsize=20), rot=90)

grid.arrange(arrangeGrob(dti_md_final, left = y.grob, bottom = x.grob))

#ggsave(paste0(png_dir, "dti_md_baseline_tracts.png"), grid.arrange(arrangeGrob(dti_md_final, left = y.grob, bottom = x.grob)), width = 20, height = 20, units = "in")
```
  
## DTI FA
```{r run dti fa sd plotting}
variance_measure = "sd"
dti_fa_sd_plots <- lapply(unique(HCPD$tract_label), plot_mean_var, scalar = "dti_fa", variance_measure = variance_measure, 
                                color1 = color1, color2 = color2, color3 = color3, ylim1 = 0, ylim2 = 1)
names(dti_fa_sd_plots) <- unique(HCPD$tract_label)
```

```{r dti fa tracts sd, fig.height = 20, fig.width = 20}
dti_fa_final <- arrange_tract_plots(dti_fa_sd_plots, "DTI FA: Mean +/- 1SD")
x.grob <- textGrob("Position on Tract (Node ID)", 
                   gp=gpar(col="black", fontsize=20))
y.grob <- textGrob("Fractional Anisotropy", 
                   gp=gpar(col="black", fontsize=20), rot=90)

grid.arrange(arrangeGrob(dti_fa_final, left = y.grob, bottom = x.grob))

ggsave(paste0(png_dir, "dti_fa_baseline_tracts.png"), grid.arrange(arrangeGrob(dti_fa_final, left = y.grob, bottom = x.grob)), width = 20, height = 20, units = "in")
```

# consistency of measures across datasets

```{r dti md across datasets}
summary_HCPD_dti_md <- summary_HCPD_dti_md %>% arrange(tract_label, nodeID)
summary_HBN_dti_md <- summary_HBN_dti_md %>% arrange(tract_label, nodeID)
summary_PNC_dti_md <- summary_PNC_dti_md %>% arrange(tract_label, nodeID)

# Calculate correlations
cor_HCPD_HBN <- cor.test(summary_HCPD_dti_md$mean_dti_md, summary_HBN_dti_md$mean_dti_md)$estimate
cor_HCPD_PNC <- cor.test(summary_HCPD_dti_md$mean_dti_md, summary_PNC_dti_md$mean_dti_md)$estimate
cor_HBN_PNC <- cor.test(summary_HBN_dti_md$mean_dti_md, summary_PNC_dti_md$mean_dti_md)$estimate

# Organize the data for plotting (lower triangle)
cor_data <- data.frame(
  Dataset1 = c("HCPD", "HCPD", "HBN"),
  Dataset2 = c("HBN", "PNC", "PNC"),
  Correlation = c(cor_HCPD_HBN, cor_HCPD_PNC, cor_HBN_PNC)
)
 
# Create a lower triangle correlation plot
dti_md_corr <- ggplot(cor_data, aes(x = Dataset1, y = Dataset2, fill = Correlation)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation") +
  geom_text(aes(label = round(Correlation, 2)), color = "black", size = 7) +
  theme_minimal() + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(vjust = 1, 
                                   size = 12, hjust = 1),
        axis.text.y = element_text(vjust = 1, 
                                   size = 12, hjust = 1)) +
  coord_fixed() +
  labs(title = "Correlation of DTI MD Across Datasets") +
  scale_y_discrete(limits = rev(levels(factor(cor_data$Dataset2))))  

ggsave(paste0(png_dir, "dti_md_corr.png"), dti_md_corr, width = 5, height = 5, units = "in")

```

 


```{r dti fa across datasets}
# Calculate correlations
cor_HCPD_HBN <- cor.test(summary_HCPD_dti_fa$mean_dti_fa, summary_HBN_dti_fa$mean_dti_fa)$estimate
cor_HCPD_PNC <- cor.test(summary_HCPD_dti_fa$mean_dti_fa, summary_PNC_dti_fa$mean_dti_fa)$estimate
cor_HBN_PNC <- cor.test(summary_HBN_dti_fa$mean_dti_fa, summary_PNC_dti_fa$mean_dti_fa)$estimate

# Organize the data for plotting (lower triangle)
cor_data <- data.frame(
  Dataset1 = c("HCPD", "HCPD", "HBN"),
  Dataset2 = c("HBN", "PNC", "PNC"),
  Correlation = c(cor_HCPD_HBN, cor_HCPD_PNC, cor_HBN_PNC)
)
 
# Create a lower triangle correlation plot
dti_fa_corr <- ggplot(cor_data, aes(x = Dataset1, y = Dataset2, fill = Correlation)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation") +
  geom_text(aes(label = round(Correlation, 2)), color = "black", size = 7) +
  theme_minimal() + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(vjust = 1, 
                                   size = 12, hjust = 1),
        axis.text.y = element_text(vjust = 1, 
                                   size = 12, hjust = 1)) +
  coord_fixed() +
  labs(title = "Correlation of DTI FA Across Datasets") +
  scale_y_discrete(limits = rev(levels(factor(cor_data$Dataset2))))  

ggsave(paste0(png_dir, "dti_fa_corr.png"), dti_fa_corr, width = 5, height = 5, units = "in")

```
 


# Tracts exhibit distinct mean diffusivity profiles

```{r, fig.height = 15, fig.width = 20}
# filter for central nodes
HCPD <- HCPD %>% mutate(Dataset="HCPD")
HBN <- HBN %>% mutate(Dataset="HBN")
PNC <- PNC %>% mutate(Dataset="PNC")
 
central_tract_profiles_HCPD <- HCPD %>% filter(nodeID > 39 & nodeID < 60)
mycolors <- colorRampPalette(paletteer::paletteer_d("PNWColors::Sailboat"))(22)

HCPD_boxplots <- central_tract_profiles_HCPD %>%
  ggplot(aes(x = reorder(tractID, dti_md, fun = mean), y=dti_md, fill=tractID)) +
    geom_boxplot(aes(color = tractID, fill = tractID), alpha=0.6) +
    scale_fill_manual(values = mycolors) + 
    scale_color_manual(values = mycolors) + 
    theme(
      legend.position="none",
      plot.title = element_text(size=18,hjust=0.5),
      axis.text.x = element_text(angle=45, hjust=0.95, size=18),
      axis.text.y = element_text(size=18),
      axis.title.y = element_blank()
    ) +
    ggtitle("HCP-D (N = 568)") +
    xlab("") + ylim(0.0001, 0.001)


central_tract_profiles <- HBN %>% filter(nodeID > 39 & nodeID < 60)
mycolors <- colorRampPalette(paletteer::paletteer_d("PNWColors::Sailboat"))(22)
 
HBN_boxplots <- central_tract_profiles %>%
  ggplot(aes(x = reorder(tractID, dti_md, fun = mean), y=dti_md, fill=tractID)) +
    geom_boxplot(aes(color = tractID, fill = tractID), alpha=0.6) +
    scale_fill_manual(values = mycolors) + 
    scale_color_manual(values = mycolors) + 
    theme(
      legend.position="none",
      plot.title = element_text(size=18, hjust=0.5),
      axis.text.x = element_text(angle=45, hjust=0.95, size=18),
      axis.text.y = element_text(size=18),
      axis.title.y = element_blank()

    ) +
    ggtitle("HBN (N = 913)") +
    xlab("")  + ylim(0.0001, 0.001)

 
 
central_tract_profiles_PNC <- PNC %>% filter(nodeID > 39 & nodeID < 60)
mycolors <- colorRampPalette(paletteer::paletteer_d("PNWColors::Sailboat"))(22)

PNC_boxplots <- central_tract_profiles_PNC %>%
  ggplot(aes(x = reorder(tractID, dti_md, fun = mean), y=dti_md, fill=tractID)) +
    geom_boxplot(aes(color = tractID, fill = tractID), alpha=0.6) +
    scale_fill_manual(values = mycolors) + 
    scale_color_manual(values = mycolors) + 
    theme(
      legend.position="none",
      plot.title = element_text(size=18,hjust=0.5),
      axis.text.x = element_text(angle=45, hjust=0.95, size=18),
      axis.text.y = element_text(size=18),
      axis.title.y = element_blank()
    ) +
    ggtitle("PNC (N = 1101)") +
    xlab("") + ylim(0.0001, 0.001)


 
y.grob <- textGrob("Mean Diffusivity", 
                   gp=gpar(col="black", fontsize=18), rot=90)

 
# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
boxplots_final <- ggarrange(
  ggarrange(HCPD_boxplots, HBN_boxplots, ncol = 2),  # Top row with 2 plots
  ggarrange(NULL, PNC_boxplots, NULL, ncol = 3, widths = c(1, 2, 1)),  # Center the bottom plot
  nrow = 2
)

grid.arrange(arrangeGrob(boxplots_final, left = y.grob))


ggsave(paste0(png_dir, "dti_md_deepWM_boxplots.png"), grid.arrange(arrangeGrob(boxplots_final, left = y.grob)), width = 20, height = 15, units = "in")
```

