---
title: "Explore baseline tract profiles"
author: "Audrey Luo"
output: html_document
---

```{r setup, include=FALSE}
library(cowplot)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(gratia) 
library(knitr)
library(kableExtra)
library(mgcv)
library(RColorBrewer)
library(stringr)
library(tidyr)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, include = TRUE) 

font_size <- 14
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1
point_size <- 2

input_root <- "/cbica/projects/luo_wm_dev/input"
output_root <- "/cbica/projects/luo_wm_dev/output"
png_dir <- "/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/both_datasets/"
  
```


# Tract Names and Orientations
https://yeatmanlab.github.io/pyAFQ/explanations/bundle_orientation.html#bundle-orientation 
- Left Anterior Thalamic: anterior to posterior
- Right Anterior Thalamic: anterior to posterior
- Left Cingulum Cingulate: anterior to posterior
- Right Cingulum Cingulate: anterior to posterior
- Left Corticospinal: superior to inferior
- Right Corticospinal: superior to inferior
- Left Inferior Fronto-occipital: anterior to posterior
- Right Inferior Fronto-occipital: anterior to posterior
- Left Inferior Longitudinal: anterior to posterior
- Right Inferior Longitudinal: anterior to posterior
- Left Superior Longitudinal: anterior to posterior
- Right Superior Longitudinal: anterior to posterior
- Left Arcuate: anterior to posterior
- Right Arcuate: anterior to posterior
- Left Uncinate: superior to inferior
- Right Uncinate: superior to inferior
- Forceps Minor: right to left
- Forceps Major: right to left
- Left Posterior Arcuate: superior to inferior
- Right Posterior Arcuate: superior to inferior
- Left Vertical Occipital: superior to inferior
- Right Vertical Occipital: superior to inferior

```{r load tract profiles, cache=TRUE}
HCPD <- readRDS(sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz.RData", output_root, "HCPD"))
HBN <- readRDS(sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz.RData", output_root, "HBN"))
PNC <- readRDS(sprintf("%1$s/%2$s/tract_profiles/all_subjects/tract_profiles_for_viz.RData", output_root, "PNC"))
```

# plot baseline for each measure: dki_fa, dki_md, dti_fa, dti_md and maybe compare
```{r functions for visualizing baseline tract profiles}


# make summary df's: this creates summary_[dataset]_[scalar] dataframes that include mean, sd, se for each scalar/dataset
make_summary_dfs <- function(scalar, dataset_name, df) {
  # skip PNC when scalar is dki_md or dki_fa
  if (dataset_name == "PNC" && (scalar == "dki_md" || scalar == "dki_fa")) {
    return(NULL)  
  }
  # set variables
  mean_scalar <- paste0("mean_", scalar)
  # Process the given dataset df
  scalar_data <- df %>% select(sub, tractID, tract_label, nodeID, hemi, all_of(scalar))
  sqrt_n <- sqrt(length(unique(scalar_data$sub)))
  setDT(scalar_data)
  summary_data <- scalar_data[, .(
    mean_scalar = mean(.SD[[1]], na.rm = TRUE),
    sd = sd(.SD[[1]], na.rm = TRUE),
    se = sd(.SD[[1]], na.rm = TRUE) / sqrt_n,
    ymin_sd = mean(.SD[[1]], na.rm = TRUE) - sd(.SD[[1]], na.rm = TRUE),
    ymax_sd = mean(.SD[[1]], na.rm = TRUE) + sd(.SD[[1]], na.rm = TRUE),
    ymin_se = mean(.SD[[1]], na.rm = TRUE) - (sd(.SD[[1]], na.rm = TRUE) / sqrt_n),
    ymax_se = mean(.SD[[1]], na.rm = TRUE) + (sd(.SD[[1]], na.rm = TRUE) / sqrt_n)
  ), by = .(tract_label, tractID, nodeID, hemi), .SDcols = scalar]

  setnames(summary_data, "mean_scalar", paste0("mean_", scalar))
  summary_data[, Dataset := dataset_name]
  summary_dataset_scalar <- paste0("summary_", dataset_name, "_", scalar)
  assign(summary_dataset_scalar, summary_data, envir = .GlobalEnv)
}


# function for plotting tractprofiles mean and sd or se for a given scalar
plot_mean_var <- function(tract, scalar, variance_measure, color1, color2, color3, ylim1 = 0, ylim2 = 1) {
 
  mean_scalar <- paste0("mean_", scalar)

  if (scalar == "dti_fa" | scalar == "dti_md") {
    summary_HCPD <- get(paste0("summary_HCPD_", scalar))
    summary_HBN <- get(paste0("summary_HBN_", scalar))
    summary_PNC <- get(paste0("summary_PNC_", scalar))
    
    df_profiles <- rbind(summary_HCPD, summary_HBN, summary_PNC)
    df <- df_profiles %>% filter(tract_label == tract)
 
      # plot
      if(str_detect(tract, "Callosum")) {
        plot <- ggplot(data = df, aes(x = nodeID, y = get(mean_scalar))) +
        geom_ribbon(data = df, aes_string(x = "nodeID", ymin = paste0("ymin_", variance_measure), ymax = paste0("ymax_", variance_measure), fill = "Dataset"), alpha = .3) + 
        geom_line(data = df, aes(x = nodeID, y = get(mean_scalar), color = Dataset),size = 1) +
        scale_fill_manual(values = c("HCPD" = color1, "HBN" = color2, "PNC" = color3)) +
        scale_color_manual(values = c("HCPD" = color1, "HBN" = color2, "PNC" = color3)) +
        theme(legend.position = "none",
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract) + ylim(ylim1, ylim2)
        
      } else {
        plot <- ggplot(data = df, aes(x = nodeID, y = get(mean_scalar), linetype = hemi)) +
        geom_ribbon(data = df, aes_string(x = "nodeID", ymin = paste0("ymin_", variance_measure), ymax = paste0("ymax_", variance_measure), fill = "Dataset"), alpha = .3) + 
        geom_line(data = df, aes(x = nodeID, y = get(mean_scalar), color = Dataset),size = 1) +
          guides(linetype = guide_legend("Hemi", override.aes = list(fill = "white"))) + 
        scale_fill_manual(values = c("HCPD" = color1, "HBN" = color2, "PNC" = color3)) +
        scale_color_manual(values = c("HCPD" = color1, "HBN" = color2, "PNC" = color3)) +
        theme(legend.position = "none",
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract) + ylim(ylim1, ylim2)
      }
    return(plot)
  } else {
    summary_HCPD <- get(paste0("summary_HCPD_", scalar))
    summary_HBN <- get(paste0("summary_HBN_", scalar))
      
    df_profiles <- rbind(summary_HCPD, summary_HBN)
    df <- df_profiles %>% filter(tract_label == tract)
     
    if(str_detect(tract, "Callosum")) {
      plot <- ggplot(data = df, aes(x = nodeID, y = get(mean_scalar))) +
      geom_ribbon(data = df, aes_string(x = "nodeID", ymin = paste0("ymin_", variance_measure), ymax = paste0("ymax_", variance_measure), fill = "Dataset"), alpha = .3) + 
      geom_line(data = df, aes(x = nodeID, y = get(mean_scalar), color = Dataset),size = 1) +
      scale_fill_manual(values = c("HCPD" = color1, "HBN" = color2)) +
      scale_color_manual(values = c("HCPD" = color1, "HBN" = color2)) +
      theme(legend.position = "none",
            axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract) + ylim(ylim1, ylim2)
      
    } else {
      plot <- ggplot(data = df, aes(x = nodeID, y = get(mean_scalar), linetype = hemi)) +
      geom_ribbon(data = df, aes_string(x = "nodeID", ymin = paste0("ymin_", variance_measure), ymax = paste0("ymax_", variance_measure), fill = "Dataset"), alpha = .3) + 
      geom_line(data = df, aes(x = nodeID, y = get(mean_scalar), color = Dataset),size = 1) +
        guides(linetype = guide_legend("Hemi",override.aes = list(fill = "white"))) + 
      scale_fill_manual(values = c("HCPD" = color1, "HBN" = color2)) +
      scale_color_manual(values = c("HCPD" = color1, "HBN" = color2)) +
      theme(legend.position = "none",
            axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract) + ylim(ylim1, ylim2)
    }
    return(plot)
  }
  
}


arrange_tract_plots <- function(list_figures, title) {

  AP_plots <- ggarrange(list_figures$`Anterior Thalamic`, 
                           list_figures$`Cingulum Cingulate`, 
                           list_figures$`Inferior Fronto-occipital`, 
                           list_figures$`Inferior Longitudinal`, 
                           list_figures$`Superior Longitudinal`, ncol=5, nrow = 1)
  AP_plots_final <- annotate_figure(AP_plots, top = text_grob("Anterior - Posterior", 
                 color = "black", face = "bold", size = 14))
  
  AP_frontotemp_plots <- ggarrange(list_figures$`Arcuate`,
                                      list_figures$`Uncinate`, ncol=2, nrow = 1)
  AP_frontotemp_plots_final <- annotate_figure(AP_frontotemp_plots, top = text_grob("Anterior - Posterior (Frontal - Temporal)", 
                 color = "black", face = "bold", size = 14))

  
  SI_plots <- ggarrange(list_figures$`Corticospinal`, list_figures$`Posterior Arcuate`, 
                           list_figures$`Vertical Occipital`, common.legend=TRUE, legend=c("right"),
                           ncol=3, nrow = 1)
  SI_plots_final <- annotate_figure(SI_plots, top = text_grob("Superior - Inferior", 
                 color = "black", face = "bold", size = 14))
  
  
  tractprofiles_plot <- ggarrange(AP_plots_final, ggarrange(AP_frontotemp_plots_final, SI_plots_final, widths = c(1, 1.5)), nrow = 2)
  
  tractprofiles_plot_final <- annotate_figure(tractprofiles_plot, top = text_grob(title, 
               color = "black", face = "italic", size = 18))
   return(tractprofiles_plot_final)
}


arrange_callosum_plots <- function(list_figures, title) {
  callosum_plots <- ggarrange(list_figures$`Callosum Anterior Frontal`, 
                           list_figures$`Callosum Motor`, 
                           list_figures$`Callosum Occipital`, 
                           list_figures$`Callosum Orbital`, 
                           list_figures$`Callosum Posterior Parietal`,
                           list_figures$`Callosum Superior Frontal`, 
                           list_figures$`Callosum Superior Parietal`, 
                           list_figures$`Callosum Temporal`, ncol=4, nrow = 2)
  tractprofiles_plot_final <- annotate_figure(callosum_plots, top = text_grob(title, 
               color = "black", face = "italic", size = 18))
   return(tractprofiles_plot_final)
}
 
```

 


```{r make summary dfs}

datasets <- list(HCPD = HCPD, HBN = HBN, PNC = PNC)
scalars <- c("dki_md", "dti_md", "dki_fa", "dti_fa")

# create summary dfs
for (dataset_name in names(datasets)) {
  dataset <- datasets[[dataset_name]]
  lapply(scalars, function(scalar) {
    make_summary_dfs(scalar, dataset_name, dataset) # this creates summary_[dataset]_[scalar] dataframes that include mean, sd, se for each scalar
  })
}
 
```

```{r set colors}
color1 = "#3FB8AFFF"
color2 = "#FF9E9DFF"
color3 = "#cc0468"
```


## DKI MD
```{r run dki md sd plotting}
variance_measure = "sd"
dki_md_major_sd_plots <- lapply(unique(HCPD$tract_label), plot_mean_var, scalar = "dki_md", variance_measure = variance_measure, 
                                color1 = color1, color2 = color2, color3 = color3, ylim1 = 0.0005, ylim2 = 0.00165)
names(dki_md_major_sd_plots) <- unique(HCPD$tract_label)
```

```{r dki md major tracts sd, fig.height = 10, fig.width = 20}
dki_md_major_final <- arrange_tract_plots(dki_md_major_sd_plots, "DKI MD: Mean +/- 1SD")
x.grob <- textGrob("Position on Tract (Node ID)", 
                   gp=gpar(col="black", fontsize=14))
y.grob <- textGrob("DKI: Mean Diffusivity", 
                   gp=gpar(col="black", fontsize=14), rot=90)

ggsave(paste0(png_dir, "dki_md_main_tracts.png"), grid.arrange(arrangeGrob(dki_md_major_final, left = y.grob, bottom = x.grob)), width = 20, height = 10, units = "in")
```


```{r dki md callosum tracts sd, fig.height = 10, fig.width = 20}
dki_md_callosum_final <- arrange_callosum_plots(dki_md_major_sd_plots, "DKI MD: Mean +/- 1SD")


ggsave(paste0(png_dir, "dki_md_call_tracts.png"), grid.arrange(arrangeGrob(dki_md_callosum_final, left = y.grob, bottom = x.grob)), width = 20, height = 10, units = "in")
```


```{r run dki md se plotting}
variance_measure = "se"
dki_md_major_se_plots <- lapply(unique(HCPD$tract_label), plot_mean_var, scalar = "dki_md", variance_measure = variance_measure, 
                                color1 = color1, color2 = color2, color3 = color3, ylim1 = 0.0005, ylim2 = 0.0015)
names(dki_md_major_se_plots) <- unique(HCPD$tract_label)
```

```{r dki md major tracts se, fig.height = 10, fig.width = 20}
arrange_tract_plots(dki_md_major_se_plots, "DKI MD: Mean +/- 1se")
```

```{r dki md callosum tracts se, fig.height = 10, fig.width = 20}
arrange_callosum_plots(dki_md_major_se_plots, "DKI MD: Mean +/- 1se")
```

## DTI MD
```{r run dti md sd plotting}
variance_measure = "sd"
dti_md_major_sd_plots <- lapply(unique(HCPD$tract_label), plot_mean_var, scalar = "dti_md", variance_measure = variance_measure, 
                                color1 = color1, color2 = color2, color3 = color3, ylim1 = 0.0003, ylim2 = 0.0012)
names(dti_md_major_sd_plots) <- unique(HCPD$tract_label)
```

```{r dti md major tracts sd, fig.height = 10, fig.width = 20}
dti_md_major_final <- arrange_tract_plots(dti_md_major_sd_plots, "DTI MD: Mean +/- 1SD")

y.grob <- textGrob("DTI: Mean Diffusivity", 
                   gp=gpar(col="black", fontsize=14), rot=90)

ggsave(paste0(png_dir, "dti_md_main_tracts.png"), grid.arrange(arrangeGrob(dti_md_major_final, left = y.grob, bottom = x.grob)), width = 20, height = 10, units = "in")
```

```{r dti md callosum tracts sd, fig.height = 10, fig.width = 20}
dti_md_callosum_final <- arrange_callosum_plots(dti_md_major_sd_plots, "DTI MD: Mean +/- 1SD")
ggsave(paste0(png_dir, "dti_md_call_tracts.png"), grid.arrange(arrangeGrob(dti_md_callosum_final, left = y.grob, bottom = x.grob)), width = 20, height = 10, units = "in")
```

## DKI FA
```{r run dki fa sd plotting}
variance_measure = "sd"
dki_fa_major_sd_plots <- lapply(unique(HCPD$tract_label), plot_mean_var, scalar = "dki_fa", variance_measure = variance_measure, color1 = color1, color2 = color2, color3 = color3)
names(dki_fa_major_sd_plots) <- unique(HCPD$tract_label)
```

```{r dki fa major tracts sd, fig.height = 10, fig.width = 20}
dki_fa_major_final <- arrange_tract_plots(dki_fa_major_sd_plots, "DKI FA: Mean +/- 1SD")

y.grob <- textGrob("DKI: Fractional Anisotropy", 
                   gp=gpar(col="black", fontsize=14), rot=90)
ggsave(paste0(png_dir, "dki_fa_main_tracts.png"), grid.arrange(arrangeGrob(dki_fa_major_final, left = y.grob, bottom = x.grob)), width = 20, height = 10, units = "in")

```

```{r dki fa callosum tracts sd, fig.height = 10, fig.width = 20}
dki_fa_callosum_final <- arrange_callosum_plots(dki_fa_major_sd_plots, "DKI FA: Mean +/- 1SD")

ggsave(paste0(png_dir, "dki_fa_call_tracts.png"), grid.arrange(arrangeGrob(dki_fa_callosum_final, left = y.grob, bottom = x.grob)), width = 20, height = 10, units = "in")

```

## DTI FA
```{r run dti fa sd plotting}
variance_measure = "sd"
dti_fa_major_sd_plots <- lapply(unique(HCPD$tract_label), plot_mean_var, scalar = "dti_fa", variance_measure = variance_measure, color1 = color1, color2 = color2, color3 = color3)
names(dti_fa_major_sd_plots) <- unique(HCPD$tract_label)
```

```{r dti fa major tracts sd, fig.height = 10, fig.width = 20}
dti_fa_main_final <- arrange_tract_plots(dti_fa_major_sd_plots, "DTI FA: Mean +/- 1SD")

y.grob <- textGrob("DTI: Fractional Anisotropy", 
                   gp=gpar(col="black", fontsize=14), rot=90)
ggsave(paste0(png_dir, "dti_fa_main_tracts.png"), grid.arrange(arrangeGrob(dti_fa_main_final, left = y.grob, bottom = x.grob)), width = 20, height = 10, units = "in")

```

```{r dti fa callosum tracts sd, fig.height = 10, fig.width = 20}
dti_fa_call_final <- arrange_callosum_plots(dti_fa_major_sd_plots, "DTI FA: Mean +/- 1SD")

ggsave(paste0(png_dir, "dti_fa_call_tracts.png"), grid.arrange(arrangeGrob(dti_fa_call_final, left = y.grob, bottom = x.grob)), width = 20, height = 10, units = "in")

```

# check site differences

```{r function for plotting by site}
# function for plotting tractprofiles mean and sd or se for a given scalar
plot_mean_by_site <- function(tract, dataset, scalar, variance_measure, color1, color2, color3, ylim1 = 0, ylim2 = 1) {
 
  mean_scalar <- paste0("mean_", scalar) 
  df_profiles <- get(paste0("summary_", dataset, "_", scalar, "_site"))
  df <- df_profiles %>% filter(tract_label == tract)
   
  if(str_detect(tract, "Callosum")) {
    plot <- ggplot(data = df, aes(x = nodeID, y = get(mean_scalar)), group = site) +
    geom_ribbon(data = df, aes_string(x = "nodeID", ymin = paste0("ymin_", variance_measure), ymax = paste0("ymax_", variance_measure), fill = "site"), alpha = .3) + 
    geom_line(data = df, aes(x = nodeID, y = get(mean_scalar), color = site),size = 1) +
   
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract) + ylim(ylim1, ylim2)
    
  } else {
    plot <- ggplot(data = df, aes(x = nodeID, y = get(mean_scalar), , group = site)) +
    geom_ribbon(data = df, aes_string(x = "nodeID", ymin = paste0("ymin_", variance_measure), ymax = paste0("ymax_", variance_measure), fill = "site"), alpha = .3) + 
    geom_line(data = df, aes(x = nodeID, y = get(mean_scalar), color = site),size = 1) +
      guides(linetype = guide_legend("Hemi",override.aes = list(fill = "white"))) + 
     
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract) + ylim(ylim1, ylim2)
  }
  return(plot)
  
}
```


```{r function to make summary dfs for site}
make_summary_dfs_site <- function(scalar, dataset_name, df) {
  # set variables
  mean_scalar <- paste0("mean_", scalar)
  # Process the given dataset df
  scalar_data <- df %>% select(sub, tractID, tract_label, nodeID, hemi, all_of(scalar), site)
  sqrt_n <- sqrt(length(unique(scalar_data$sub)))
  setDT(scalar_data)
  summary_data <- scalar_data[, .(
    mean_scalar = mean(.SD[[1]], na.rm = TRUE),
    sd = sd(.SD[[1]], na.rm = TRUE),
    se = sd(.SD[[1]], na.rm = TRUE) / sqrt_n,
    ymin_sd = mean(.SD[[1]], na.rm = TRUE) - sd(.SD[[1]], na.rm = TRUE),
    ymax_sd = mean(.SD[[1]], na.rm = TRUE) + sd(.SD[[1]], na.rm = TRUE),
    ymin_se = mean(.SD[[1]], na.rm = TRUE) - (sd(.SD[[1]], na.rm = TRUE) / sqrt_n),
    ymax_se = mean(.SD[[1]], na.rm = TRUE) + (sd(.SD[[1]], na.rm = TRUE) / sqrt_n)
  ), by = .(tract_label, tractID, nodeID, hemi, site), .SDcols = scalar]

  setnames(summary_data, "mean_scalar", paste0("mean_", scalar))
  summary_data[, Dataset := dataset_name]
  summary_dataset_scalar <- paste0("summary_", dataset_name, "_", scalar, "_site")
  assign(summary_dataset_scalar, summary_data, envir = .GlobalEnv)
}

 
```


```{r make summary dfs for site}

dem_HCPD <- read.csv("/cbica/projects/luo_wm_dev/input/HCPD/sample_selection_files/final_sample/HCPD_WMDev_FinalSampleDemoQC.csv")
setDT(dem_HCPD)
dem_HBN <- read.csv("/cbica/projects/luo_wm_dev/input/HBN/sample_selection_files/final_sample/HBN_WMDev_FinalSampleDemoQC.csv")
setDT(dem_HBN)

HCPD <- merge(HCPD, dem_HCPD[, .(sub, site)])
HBN <- merge(HBN, dem_HBN[, .(sub, site)])

datasets <- list(HCPD = HCPD, HBN = HBN)
scalars <- c("dki_md", "dti_md", "dki_fa", "dti_fa")

# create summary dfs
for (dataset_name in names(datasets)) {
  dataset <- datasets[[dataset_name]]
  lapply(scalars, function(scalar) {
    make_summary_dfs_site(scalar, dataset_name, dataset) # this creates summary_[dataset]_[scalar] dataframes that include mean, sd, se for each scalar
  })
}
 
```

## DKI MD
```{r run dki md major tracts site HCPD}
variance_measure = "sd"
dki_md_site_major_HCPD <- lapply(unique(HCPD$tract_label), plot_mean_by_site, dataset = "HCPD", scalar = "dki_md", variance_measure = variance_measure, ylim1 = 0.0003, ylim2 = 0.0016)
names(dki_md_site_major_HCPD) <- unique(HCPD$tract_label)
```

```{r dki md major tracts site HCPD, fig.height = 10, fig.width = 20}
arrange_tract_plots(dki_md_site_major_HCPD, "HCPD: DKI MD")
```

```{r dki md callosum tracts site HCPD, fig.height = 10, fig.width = 20}
arrange_callosum_plots(dki_md_site_major_HCPD, "HCPD: DKI MD")
```


```{r run dki md major tracts site HBN}
variance_measure = "sd"
dki_md_site_major_HBN <- lapply(unique(HBN$tract_label), plot_mean_by_site, dataset = "HBN", scalar = "dki_md", variance_measure = variance_measure, ylim1 = 0.0003, ylim2 = 0.0016)
names(dki_md_site_major_HBN) <- unique(HBN$tract_label)
```

```{r dki md major tracts site HBN, fig.height = 10, fig.width = 20}
arrange_tract_plots(dki_md_site_major_HBN, "HBN: DKI MD")
```

```{r dki md callosum tracts site HBN, fig.height = 10, fig.width = 20}
arrange_callosum_plots(dki_md_site_major_HBN, "HBN: DKI MD")
```

## DTI MD
```{r run DTI md major tracts site HCPD}
variance_measure = "sd"
dti_md_site_major_HCPD <- lapply(unique(HCPD$tract_label), plot_mean_by_site, dataset = "HCPD", scalar = "dti_md", variance_measure = variance_measure, ylim1 = 0.0003, ylim2 = 0.0016)
names(dti_md_site_major_HCPD) <- unique(HCPD$tract_label)
```

```{r DTI md major tracts site HCPD, fig.height = 10, fig.width = 20}
arrange_tract_plots(dti_md_site_major_HCPD, "HCPD: DTI MD")
```

```{r DTI md callosum tracts site HCPD, fig.height = 10, fig.width = 20}
arrange_callosum_plots(dti_md_site_major_HCPD, "HCPD: DTI MD")
```


```{r run DTI md major tracts site HBN}
variance_measure = "sd"
dti_md_site_major_HBN <- lapply(unique(HBN$tract_label), plot_mean_by_site, dataset = "HBN", scalar = "dti_md", variance_measure = variance_measure, ylim1 = 0.0003, ylim2 = 0.0016)
names(dti_md_site_major_HBN) <- unique(HBN$tract_label)
```

```{r DTI md major tracts site HBN, fig.height = 10, fig.width = 20}
arrange_tract_plots(dti_md_site_major_HBN, "HBN: DTI MD")
```

```{r DTI md callosum tracts site HBN, fig.height = 10, fig.width = 20}
arrange_callosum_plots(dti_md_site_major_HBN, "HBN: DTI MD")
```

## DKI FA
```{r run dki fa major tracts site HCPD}
variance_measure = "sd"
dki_fa_site_major_HCPD <- lapply(unique(HCPD$tract_label), plot_mean_by_site, dataset = "HCPD", scalar = "dki_fa", variance_measure = variance_measure)
names(dki_fa_site_major_HCPD) <- unique(HCPD$tract_label)
```

```{r dki fa major tracts site HCPD, fig.height = 10, fig.width = 20}
arrange_tract_plots(dki_fa_site_major_HCPD, "DKI FA: Mean +/- 1SD")
```

```{r dki fa callosum tracts site HCPD, fig.height = 10, fig.width = 20}
arrange_callosum_plots(dki_fa_site_major_HCPD, "DKI FA: Mean +/- 1SD")
```

```{r run dki fa major tracts site HBN}
variance_measure = "sd"
dki_fa_site_major_HBN <- lapply(unique(HBN$tract_label), plot_mean_by_site, dataset = "HBN", scalar = "dki_fa", variance_measure = variance_measure)
names(dki_fa_site_major_HBN) <- unique(HBN$tract_label)
```

```{r dki fa major tracts site HBN, fig.height = 10, fig.width = 20}
arrange_tract_plots(dki_fa_site_major_HBN, "DKI FA: Mean +/- 1SD")
```

```{r dki fa callosum tracts site HBN, fig.height = 10, fig.width = 20}
arrange_callosum_plots(dki_fa_site_major_HBN, "DKI FA: Mean +/- 1SD")
```

## DTI FA
```{r run dti fa major tracts site HCPD}
variance_measure = "sd"
dti_fa_site_major_HCPD <- lapply(unique(HCPD$tract_label), plot_mean_by_site, dataset = "HCPD", scalar = "dti_fa", variance_measure = variance_measure)
names(dti_fa_site_major_HCPD) <- unique(HCPD$tract_label)
```

```{r dti fa major tracts site HCPD, fig.height = 10, fig.width = 20}
arrange_tract_plots(dti_fa_site_major_HCPD, "DTI FA: Mean +/- 1SD")
```

```{r dti fa callosum tracts site HCPD, fig.height = 10, fig.width = 20}
arrange_callosum_plots(dti_fa_site_major_HCPD, "DTI FA: Mean +/- 1SD")
```

```{r run dti fa major tracts site HBN}
variance_measure = "sd"
dti_fa_site_major_HBN <- lapply(unique(HBN$tract_label), plot_mean_by_site, dataset = "HBN", scalar = "dti_fa", variance_measure = variance_measure)
names(dti_fa_site_major_HBN) <- unique(HBN$tract_label)
```

```{r dti fa major tracts site HBN, fig.height = 10, fig.width = 20}
arrange_tract_plots(dti_fa_site_major_HBN, "DTI FA: Mean +/- 1SD")
```

```{r dti fa callosum tracts site HBN, fig.height = 10, fig.width = 20}
arrange_callosum_plots(dti_fa_site_major_HBN, "DTI FA: Mean +/- 1SD")
```
# consistency of measures across datasets
```{r dki md across datasets}
 
cor_HCPD_HBN <- cor.test(summary_HCPD_dki_md$mean_dki_md, summary_HBN_dki_md$mean_dki_md)$estimate
cor_HCPD_PNC <- NA  # Set to NA
cor_HBN_PNC <- NA   # Set to NA

cor_data <- data.frame(
  Dataset1 = c("HCPD", "HCPD", "HBN"),
  Dataset2 = c("HBN", "PNC", "PNC"),
  Correlation = c(cor_HCPD_HBN, cor_HCPD_PNC, cor_HBN_PNC)
)

dki_md_corr <- ggplot(cor_data, aes(x = Dataset1, y = Dataset2, fill = Correlation)) +
  geom_tile(color = "white", na.rm = TRUE) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation", na.value = "grey90") +
  geom_text(aes(label = ifelse(is.na(Correlation), "NA", round(Correlation, 2))), 
            color = "black", size = 7, na.rm = TRUE) +
  theme_minimal() + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(vjust = 1, 
                                   size = 12, hjust = 1),
        axis.text.y = element_text(vjust = 1, 
                                   size = 12, hjust = 1)) +
  coord_fixed() +
  labs(title = "Correlation of DKI MD Across Datasets") +
  scale_y_discrete(limits = rev(levels(factor(cor_data$Dataset2))))


ggsave(paste0(png_dir, "dki_md_corr.png"), dki_md_corr, width = 20, height = 10, units = "in")

```


```{r dti md across datasets}
# Calculate correlations
cor_HCPD_HBN <- cor.test(summary_HCPD_dti_md$mean_dti_md, summary_HBN_dti_md$mean_dti_md)$estimate
cor_HCPD_PNC <- cor.test(summary_HCPD_dti_md$mean_dti_md, summary_PNC_dti_md$mean_dti_md)$estimate
cor_HBN_PNC <- cor.test(summary_HBN_dti_md$mean_dti_md, summary_PNC_dti_md$mean_dti_md)$estimate

# Organize the data for plotting (lower triangle)
cor_data <- data.frame(
  Dataset1 = c("HCPD", "HCPD", "HBN"),
  Dataset2 = c("HBN", "PNC", "PNC"),
  Correlation = c(cor_HCPD_HBN, cor_HCPD_PNC, cor_HBN_PNC)
)
 
# Create a lower triangle correlation plot
dti_md_corr <- ggplot(cor_data, aes(x = Dataset1, y = Dataset2, fill = Correlation)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation") +
  geom_text(aes(label = round(Correlation, 2)), color = "black", size = 7) +
  theme_minimal() + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(vjust = 1, 
                                   size = 12, hjust = 1),
        axis.text.y = element_text(vjust = 1, 
                                   size = 12, hjust = 1)) +
  coord_fixed() +
  labs(title = "Correlation of DTI MD Across Datasets") +
  scale_y_discrete(limits = rev(levels(factor(cor_data$Dataset2))))  

ggsave(paste0(png_dir, "dti_md_corr.png"), dti_md_corr, width = 20, height = 10, units = "in")

```

```{r dki fa across datasets}
 
cor_HCPD_HBN <- cor.test(summary_HCPD_dki_fa$mean_dki_fa, summary_HBN_dki_fa$mean_dki_fa)$estimate
cor_HCPD_PNC <- NA  # Set to NA
cor_HBN_PNC <- NA   # Set to NA

cor_data <- data.frame(
  Dataset1 = c("HCPD", "HCPD", "HBN"),
  Dataset2 = c("HBN", "PNC", "PNC"),
  Correlation = c(cor_HCPD_HBN, cor_HCPD_PNC, cor_HBN_PNC)
)

dki_fa_corr <- ggplot(cor_data, aes(x = Dataset1, y = Dataset2, fill = Correlation)) +
  geom_tile(color = "white", na.rm = TRUE) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation", na.value = "grey90") +
  geom_text(aes(label = ifelse(is.na(Correlation), "NA", round(Correlation, 2))), 
            color = "black", size = 7, na.rm = TRUE) +
  theme_minimal() + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(vjust = 1, 
                                   size = 12, hjust = 1),
        axis.text.y = element_text(vjust = 1, 
                                   size = 12, hjust = 1)) +
  coord_fixed() +
  labs(title = "Correlation of DKI FA Across Datasets") +
  scale_y_discrete(limits = rev(levels(factor(cor_data$Dataset2))))


ggsave(paste0(png_dir, "dki_fa_corr.png"), dki_fa_corr, width = 20, height = 10, units = "in")

```


```{r dti fa across datasets}
# Calculate correlations
cor_HCPD_HBN <- cor.test(summary_HCPD_dti_fa$mean_dti_fa, summary_HBN_dti_fa$mean_dti_fa)$estimate
cor_HCPD_PNC <- cor.test(summary_HCPD_dti_fa$mean_dti_fa, summary_PNC_dti_fa$mean_dti_fa)$estimate
cor_HBN_PNC <- cor.test(summary_HBN_dti_fa$mean_dti_fa, summary_PNC_dti_fa$mean_dti_fa)$estimate

# Organize the data for plotting (lower triangle)
cor_data <- data.frame(
  Dataset1 = c("HCPD", "HCPD", "HBN"),
  Dataset2 = c("HBN", "PNC", "PNC"),
  Correlation = c(cor_HCPD_HBN, cor_HCPD_PNC, cor_HBN_PNC)
)
 
# Create a lower triangle correlation plot
dti_fa_corr <- ggplot(cor_data, aes(x = Dataset1, y = Dataset2, fill = Correlation)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation") +
  geom_text(aes(label = round(Correlation, 2)), color = "black", size = 7) +
  theme_minimal() + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(vjust = 1, 
                                   size = 12, hjust = 1),
        axis.text.y = element_text(vjust = 1, 
                                   size = 12, hjust = 1)) +
  coord_fixed() +
  labs(title = "Correlation of DTI FA Across Datasets") +
  scale_y_discrete(limits = rev(levels(factor(cor_data$Dataset2))))  

ggsave(paste0(png_dir, "dti_fa_corr.png"), dti_fa_corr, width = 20, height = 10, units = "in")

```

 
```{r}


# correlations
cor_HCPD_md <- cor.test(summary_HCPD_dki_md$mean_dki_md, summary_HCPD_dti_md$mean_dti_md)
cor_HCPD_fa <- cor.test(summary_HCPD_dki_fa$mean_dki_fa, summary_HCPD_dti_fa$mean_dti_fa)

cor_HBN_md <- cor.test(summary_HBN_dki_md$mean_dki_md, summary_HBN_dti_md$mean_dti_md)
cor_HBN_fa <- cor.test(summary_HBN_dki_fa$mean_dki_fa, summary_HBN_dti_fa$mean_dti_fa)

# make tables 
cor_table_HCPD <- data.frame(
  Measure = c("MD", "FA"),
  Correlation = c(cor_HCPD_md$estimate, cor_HCPD_fa$estimate),
  P_Value = c(cor_HCPD_md$p.value, cor_HCPD_fa$p.value)
)

cor_table_HBN <- data.frame(
  Measure = c("MD", "FA"),
  Correlation = c(cor_HBN_md$estimate, cor_HBN_fa$estimate),
  P_Value = c(cor_HBN_md$p.value, cor_HBN_fa$p.value)
)

# display tables 
kable(cor_table_HCPD, caption = "<b style='color:black;'>Correlation Between DKI and DTI Measures (MD and FA) for HCPD</b>", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("condensed"), full_width = FALSE, position = "center", fixed_thead = TRUE, font_size = 16)


```
 
```{r}
kable(cor_table_HBN, caption = "<b style='color:black;'>Correlation Between DKI and DTI Measures (MD and FA) for HBN</b>", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("condensed"), full_width = FALSE, position = "center", fixed_thead = TRUE, font_size = 16)
```

# think about plotting them in an intuitive way...
