---
title: "HCP-D Tract Profiles"
author: "Audrey Luo"
output: 
  html_document:
    highlight: haddock
    code_folding: hide
    number_sections: no
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    number_sections: no
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

library(cowplot)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(gratia)
library(kableExtra)
library(mgcv)
library(RColorBrewer)
library(stringr)
library(tidyr)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, include = TRUE) 

font_size <- 14
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 0.5
point_size <- 2
```

```{r load tract profiles, cache=TRUE}
cohortfile <- read.csv("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/cohortfiles/dti_fa/dti_fa_cohortfile.csv")
all_subjects <- fread("/cbica/projects/luo_wm_dev/input/HCPD/HCPD_tractprofiles/all_subjects/collated_tract_profiles_nocovbat.tsv")
 
 
all_subjects <- all_subjects %>%
  mutate(
    hemi = ifelse(grepl("_L", tractID), "Left", "Right"),
    tractID = gsub("_[LR]", "", tractID),
    tractID = case_when(
      tractID == "ATR" ~ "Anterior Thalamic Radiation",
      tractID == "CGC" ~ "Cingulum Cingulate",
      tractID == "CST" ~ "Corticospinal Tract",
      tractID == "IFO" ~ "Inferior Fronto-occipital Fasciculus",
      tractID == "ILF" ~ "Inferior Longitudinal Fasciculus",
      tractID == "SLF" ~ "Superior Longitudinal Fasciculus",
      tractID == "ARC" ~ "Arcuate Fasciculus",
      tractID == "UNC" ~ "Uncinate Fasciculus",
      tractID == "FA" ~ "Forceps Minor",
      tractID == "FP" ~ "Forceps Major",
      tractID == "pARC" ~ "Posterior Arcuate",
      tractID == "VOF" ~ "Vertical Occipital Fasciculus",
      TRUE ~ tractID
    ),
    tract_hemi = paste(hemi, tractID),
    tract_hemi = gsub("Right Forceps", "Forceps", tract_hemi),
    tract_hemi = str_replace_all(tract_hemi, " ", "_")
  )


all_subjects$hemi[all_subjects$tract_hemi == "Forceps_Minor" | all_subjects$tract_hemi == "Forceps_Major"] <- NA
all_subjects$subjectID <- as.factor(all_subjects$subjectID)
 
cohortfile <- cohortfile %>% rename(subjectID=sub) %>% select(subjectID, age, sex, mean_fd)
```

# Tract Names and Orientations
Note that I used an older version of pyAFQ here, so I had to flip some tracts to ensure consistent orientations. I will rerun tract profiles using most updated pyAFQ once newest qsiprep is released.  


|Abbrev | Full Name                              | Node 0        | Node 99         | Need to flip? | Final Orientation            |
|-------|----------------------------------------|---------------|-----------------|---------------|------------------------------|
| ATR   | anterior thalamic radiation            | A (Frontal)   | P (Thalamus)    | 0             | A - P (Frontal - Thalamus)   |
| CGC   | cingulum cinguluate                    | P             | A               | 1             | A - P                        |
| IFO   | inferior fronto-occipital fasciculus   | P (Occipital) | A (Frontal)     | 1             | A - P (Frontal - Occipital)  |
| ILF   | inferior longitudinal fasciculus       | P (Occipital) | A (Temporal)    | 1             | A - P (Temporal - Occipital) |
| SLF   | superior longitudinal fasciculus       | A             | P               | 0             | A - P                        |
| ARC   | arcuate                                | A (Frontal)   | P (Temporal)    | 0             | A - P (Frontal - Temporal)   |
| UNC   | uncinate fasciculus                    | P (Temporal)  | A (Frontal)     | 1             | A - P (Frontal - Temporal)   |
| FA    | forceps minor                          | R             | L               | 0             | R - L                        |
| FP    | forceps major                          | R             | L               | 0             | R - L                        |
| pARC  | posterior arcuate                      | S             | I               | 0             | S - I                        |
| CST   | corticospinal                          | I (Brainstem) | S (Motor cortex)| 1             | S - I (Motor - Brainstem)    |
| VOF   | vertical occipital fasciculus          | S             | I               | 0             | S - I                        |
 
```{r fix tract orientations}

# fix tract orientations
tract_profiles <- all_subjects %>%
  group_by(tractID) %>%
  mutate(nodeID = ifelse(tractID %in% c("Cingulum Cingulate", "Inferior Fronto-occipital Fasciculus", "Inferior Longitudinal Fasciculus", "Uncinate Fasciculus", "Corticospinal Tract"), max(nodeID) - nodeID, nodeID))

# label main orientation
tract_profiles <- tract_profiles %>% 
  mutate(main_orientation = case_when(
    tractID %in% c("Anterior Thalamic Radiation", "Cingulum Cingulate", "Inferior Fronto-occipital Fasciculus",
                   "Inferior Longitudinal Fasciculus", "Superior Longitudinal Fasciculus") ~ "AP",
    tractID %in% c("Arcuate Fasciculus", "Uncinate Fasciculus") ~ "AP_frontal_temporal",
    tractID %in% c("Forceps Minor", "Forceps Major") ~ "RL",
    tractID %in% c("Corticospinal Tract", "Posterior Arcuate","Vertical Occipital Fasciculus") ~ "SI",
    TRUE ~ NA_character_
  ))

tract_profiles$main_orientation <- as.factor(tract_profiles$main_orientation)


#write.table(tract_profiles, "/cbica/projects/luo_wm_dev/input/HCPD/HCPD_tractprofiles/all_subjects/collated_tract_profiles_reoriented.tsv")

```

# 1. Tract profiles
```{r functions for fitting gams and visualizing tract profiles}

# Func to visualize GAM model outputs
# adapted from https://github.com/bart-larsen/GAMM-Tutorial/blob/master/GAMM_tutorial.Rmd
visualize_model_hemis <- function(modobj, modobj_rh = NULL, both_hemis = F, smooth_var, int_var, group_var, plabels = NULL,check_diagnostics = F,derivative_plot = F, custom_color, custom_color_rh = NULL) { # modobj corresponds to lh if plotting both hemispheres 
  
  this_font_size = font_size*1.25
  
  if (both_hemis == T) { # if want to plot model for left and right tract
    if (is.null(modobj_rh)) {
      stop("Please provide models for both hemispheres")
    } 
    if (any(class(modobj)=="gam")) {
      model_lh <- modobj
      model_rh <- modobj_rh
    } else if (class(modobj$gam)=="gam") {
      model_lh <- modobj$gam
      model_rh <- modobj_rh$gam
    } else {
      stop("Can't find a gam object to plot")
    }
    
    s_lh <- summary(model_lh)
    s_rh <- summary(model_rh)
  
    ## Generate custom line plot
    np <- 100 #number of predicted values - originally 100000 but we're doing only 100 discrete nodes...
    df_lh <- model_lh$model
    df_lh <- df_lh %>% mutate(hemi = "Left")
    df_rh <- model_rh$model
    df_rh <- df_rh %>% mutate(hemi = "Right")
    
    theseVars <- attr(model_lh$terms,"term.labels") # vars are identical for lh and rh
    varClasses <- attr(model_lh$terms,"dataClasses")
    thisResp <- as.character(model_lh$terms[[2]])
     
    # No interaction variable, just produce a single line plot
    thisPred_lh <- data.frame(init = rep(0,np))
    thisPred_rh <- data.frame(init = rep(0,np))
    
    # lh
    for (v in c(1:length(theseVars))) { 
      thisVar_lh <- theseVars[[v]]
      thisClass_lh <- varClasses[thisVar_lh]
      if (thisVar_lh == smooth_var) {
        thisPred_lh[,smooth_var] = seq(min(df_lh[,smooth_var],na.rm = T),max(df_lh[,smooth_var],na.rm = T), length.out = np)
      } else {
        switch (thisClass_lh,
            "numeric" = {thisPred_lh[,thisVar] = median(df_lh[,thisVar])},
            "factor" = {thisPred_lh[,thisVar] = levels(df_lh[,thisVar])[[1]]},
            "ordered" = {thisPred_lh[,thisVar] = levels(df_lh[,thisVar])[[1]]}
              )
      }
    }
    pred_lh <- thisPred_lh %>% select(-init)
    p_lh <-data.frame(predict(model_lh,pred_lh,se.fit = T))
    pred_lh <- cbind(pred_lh,p_lh)
    pred_lh$selo <- pred_lh$fit - 2*pred_lh$se.fit
    pred_lh$sehi <- pred_lh$fit + 2*pred_lh$se.fit
    pred_lh[,group_var] = NA
    pred_lh[,thisResp] = 1
    pred_lh <- pred_lh %>% mutate(hemi = "Left")
     
    # rh
    for (v in c(1:length(theseVars))) { 
      thisVar_rh <- theseVars[[v]]
      thisClass_rh <- varClasses[thisVar_rh]
      if (thisVar_rh == smooth_var) {
        thisPred_rh[,smooth_var] = seq(min(df_rh[,smooth_var],na.rm = T),max(df_rh[,smooth_var],na.rm = T), length.out = np)
      } else {
        switch (thisClass_rh,
            "numeric" = {thisPred_rh[,thisVar] = median(df_rh[,thisVar])},
            "factor" = {thisPred_rh[,thisVar] = levels(df_rh[,thisVar])[[1]]},
            "ordered" = {thisPred_rh[,thisVar] = levels(df_rh)[[1]]}
              )
      }
    }
    pred_rh <- thisPred_rh %>% select(-init)
    p_rh <-data.frame(predict(model_rh,pred_rh,se.fit = T))
    pred_rh <- cbind(pred_rh,p_rh)
    pred_rh$selo <- pred_rh$fit - 2*pred_rh$se.fit
    pred_rh$sehi <- pred_rh$fit + 2*pred_rh$se.fit
    pred_rh[,group_var] = NA
    pred_rh[,thisResp] = 1
    pred_rh <- pred_rh %>% mutate(hemi = "Right")
    
    df <- rbind(df_lh, df_rh)
    df$hemi <- as.factor(df$hemi)
    pred <- rbind(pred_lh, pred_rh)
    pred$hemi <- as.factor(pred$hemi)
    
    p1 <- ggplot(data = df, aes_string(x = smooth_var,y = thisResp, group = "hemi")) +
    geom_ribbon(data = pred,aes_string(x = smooth_var , ymin = "selo",ymax = "sehi", fill = "hemi"),alpha = .5, linetype = 0) + 
    geom_line(data = pred,aes_string(x = smooth_var, y = "fit", color = "hemi"),size = line_size) +
    scale_fill_manual(values = c("Left" = custom_color, "Right" = custom_color_rh)) +
    scale_color_manual(values = c("Left" = custom_color, "Right" = custom_color_rh)) +
    labs(title = plabels) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) # will add labels after plotting all models together
 
  } else {
    if (any(class(modobj)=="gam")) {
      model <- modobj
    } else if (class(modobj$gam)=="gam") {
      model <- modobj$gam
    } else {
      stop("Can't find a gam object to plot")
    }
    s<-summary(model)
  
    ## Generate custom line plot
    np <- 100 #number of predicted values
    df = model$model
  
    theseVars <- attr(model$terms,"term.labels")
    varClasses <- attr(model$terms,"dataClasses")
    thisResp <- as.character(model$terms[[2]])
  
  
    # No interaction variable, just produce a single line plot
    thisPred <- data.frame(init = rep(0,np))
  
    for (v in c(1:length(theseVars))) {
      thisVar <- theseVars[[v]]
      thisClass <- varClasses[thisVar]
      if (thisVar == smooth_var) {
        thisPred[,smooth_var] = seq(min(df[,smooth_var],na.rm = T),max(df[,smooth_var],na.rm = T), length.out = np)
      } else {
        switch (thisClass,
            "numeric" = {thisPred[,thisVar] = median(df[,thisVar])},
            "factor" = {thisPred[,thisVar] = levels(df[,thisVar])[[1]]},
            "ordered" = {thisPred[,thisVar] = levels(df[,thisVar])[[1]]}
              )
      }
    }
    pred <- thisPred %>% select(-init)
    p<-data.frame(predict(model,pred,se.fit = T))
    pred <- cbind(pred,p)
    pred$selo <- pred$fit - 2*pred$se.fit
    pred$sehi <- pred$fit + 2*pred$se.fit
    pred[,group_var] = NA
    pred[,thisResp] = 1
    
    p1 <- ggplot(data = df, aes_string(x = smooth_var,y = thisResp)) +
    geom_ribbon(data = pred,aes_string(x = smooth_var, ymin = "selo", ymax = "sehi"),alpha = .5, linetype = 0, fill = custom_color) +
    geom_line(data = pred,aes_string(x = smooth_var, y = "fit"),size = line_size, color = custom_color) +
    labs(title = plabels) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) # will add labels after plotting all models together
  }
    
  
 
  if (derivative_plot == T) {
    # We will add a bar that shows where the derivative is significant.
    # First make some adjustments to the line plot.
    p1<- p1+theme(text = element_text(size=this_font_size),
                axis.text = element_text(size = this_font_size),
                axis.title.y = element_text(size = this_font_size),
                axis.title.x = element_blank(),
                axis.text.x = element_blank(),
                axis.ticks.x = element_blank(),
                legend.text = element_text(size = this_font_size),
                legend.title = element_text(size = this_font_size),
                axis.title = element_text(size = this_font_size),
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                panel.background = element_rect(fill = "transparent",colour = NA),
                plot.background = element_rect(fill = "transparent",colour = NA),
                plot.margin = unit(c(.2, .2, 0, .2), "cm")) #Top, left,Bottom, right
    scatter <- list(p1)
    
    # Now add the plots using the derivative plotting function
    if (any(grepl(x = row.names(s$s.table),pattern =  ":") & grepl(x=row.names(s$s.table),pattern = int_var))) {
      # Factor levels separately if there is an interaction in the model.
      f<-formula(model) # current formula
      fterms <- terms(f)
      fac <- attr(fterms, "factors")
      idx <- which(as.logical(colSums(fac[grep(x=row.names(fac),pattern = int_var),])))
      new_terms <- drop.terms(fterms, dropx = idx, keep.response = TRUE)
      new_formula <- formula(new_terms) # Formula without any interaction terms in the model.
      
      #add derivative gradients for each level of the factor
      num_levels <- length(levels(df[,int_var]))
      level_colors <- suppressWarnings(RColorBrewer::brewer.pal(num_levels,"Set1")) #use the same palette as the line plot
      plotlist = vector(mode = "list",length = num_levels+1) # we will be building a list of plots
      plotlist[1] = scatter # first the scatter plot
      
      for (fcount in 1:num_levels) {
        this_level <- levels(df[,int_var])[fcount]
        df$subset <- df[,int_var] == this_level
        df$group_var <- df[,group_var]
        this_mod <- gam(formula = new_formula,data = df,subset = subset,random=list(group_var=~1))
        # this_d <- get_derivs_and_plot(modobj = this_mod,smooth_var = smooth_var,low_color = "white",hi_color = level_colors[fcount])
        this_d <- get_derivs_and_plot(modobj = this_mod,smooth_var = smooth_var,low_color = "white",hi_color = level_colors[fcount])
        
        if (fcount != num_levels & fcount != 1){
          # get rid of redundant junk
          this_d$theme$axis.title = element_blank()
          this_d$theme$axis.text.x = element_blank()
          this_d$theme$axis.ticks=element_blank()
          this_d$theme$legend.background=element_blank()
          this_d$theme$legend.box.background = element_blank()
          this_d$theme$legend.key = element_blank()
          this_d$theme$legend.title = element_blank()
          this_d$theme$legend.text = element_blank()
        }
        if (fcount == 1) {
         this_d$theme$axis.title = element_blank()
         this_d$theme$axis.text.x = element_blank()
         this_d$theme$axis.ticks=element_blank()
         this_d$theme$legend.background=element_blank()
         this_d$theme$legend.box.background = element_blank()
         this_d$theme$legend.key = element_blank()
         this_d$theme$legend.text = element_blank()
        }
        if (fcount == num_levels) {
         this_d$theme$legend.background=element_blank()
         this_d$theme$legend.box.background = element_blank()
         this_d$theme$legend.key = element_blank()
         this_d$theme$legend.title = element_blank()
        }
        this_d$labels$fill=NULL
        plotlist[fcount+1] = list(this_d)
      }
      pg<-plot_grid(rel_heights = c(16*num_levels,rep(num_levels,num_levels-1),3*num_levels),plotlist = plotlist,align = "v",axis = "lr",ncol = 1)
      final_plot <- pg
      print(final_plot)
  } else {
    # No need to split
    d1 <- get_derivs_and_plot(modobj = modobj,smooth_var = smooth_var)
    scatter <- list(p1)
    bar <- list(d1)
    allplots <- c(scatter,bar)
    pg<-plot_grid(rel_heights = c(16,3),plotlist = allplots,align = "v",axis = "lr",ncol = 1)
    final_plot <- pg
    print(final_plot)
  }

  }    else {
    # No derivative plot
    p1<- p1+theme(text = element_text(size=font_size),
                axis.text = element_text(size = font_size),
                legend.text = element_text(size = font_size),
                panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                panel.background = element_blank(),
                plot.background = element_blank())
    final_plot <- p1
    print(final_plot)
  }
  
  if (check_diagnostics == T) {
    cp <- check(b,
    a.qq = list(method = "tnorm",
                a.cipoly = list(fill = "light blue")),
    a.respoi = list(size = 0.5),
    a.hist = list(bins = 10))
    print(cp)
  }
  return(final_plot)
}

# function to implement visualize_model_hemis for all tracts
visualize_tractprofiles <- function(tract, scalar, df) {
  if(str_detect(tract, "Forceps")) {
    tract_data <- df %>% filter(tractID == tract)  
     
    tract_model <- gamm(as.formula(sprintf("%s ~ s(nodeID, k=24, fx = F)", scalar)), random = list(subjectID=~1), data=tract_data)
    print(paste("Finished fitting GAM for", tract))
   
    plot <- visualize_model_hemis(tract_model, both_hemis = F, smooth_var = "nodeID", int_var = NULL, group_var="subjectID", plabels = NULL,
                          check_diagnostics = F,derivative_plot = F, custom_color = "#831D69FF") + 
      labs(title = tract) + theme(plot.title = element_text(size=14, hjust=0.5))
    print(paste("Finished plotting", tract))
  } else {
    tract_data_lh <- df %>% filter(tractID == tract) %>% filter(str_detect(tract_hemi, "Left"))
    tract_data_rh <- df %>% filter(tractID == tract) %>% filter(str_detect(tract_hemi, "Right"))
    
     
    tract_model_lh <- gamm(as.formula(sprintf("%s ~ s(nodeID, k=24, fx = F)", scalar)), random = list(subjectID=~1), data=tract_data_lh)
    print(paste("Finished fitting GAM for left", tract))
    tract_model_rh <- gamm(as.formula(sprintf("%s ~ s(nodeID, k=24, fx = F)", scalar)), random = list(subjectID=~1), data=tract_data_rh)
    print(paste("Finished fitting GAM for right", tract))
    
    plot <- visualize_model_hemis(tract_model_lh$gam, tract_model_rh$gam, both_hemis = T, smooth_var = "nodeID", int_var = NULL, group_var="subjectID", plabels = NULL,
                          check_diagnostics = F,derivative_plot = F, custom_color = "#831D69FF", custom_color_rh = "#FA9BAC") + 
      labs(title = tract) + theme(plot.title = element_text(size=14, hjust=0.5))
    print(paste("Finished plotting", tract))
    
  }
  return(plot)
  
}


# function for plotting tractprofiles mean and sd FA/MD
plot_mean_sd <- function(tract, scalar, custom_color, custom_color_rh=NULL) {
  df_profiles <- get(paste0("tract_profiles_", scalar, "_summary"))
  df <- df_profiles %>% filter(tractID == tract)
  
  if(is.null(custom_color_rh)) {
    plot <- ggplot(data = df, aes_string(x = "nodeID", y = scalar)) +
    geom_ribbon(data = df, aes(x = nodeID , ymin = ymin ,ymax = ymax), alpha = .5, linetype = 0, fill= custom_color) + 
    geom_line(data = df, aes(x = nodeID, y = mean_scalar), size = line_size, color = custom_color) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.title.y = element_blank())
    
  } else {
    plot <- ggplot(data = df, aes(x = nodeID, y = mean_scalar, group = hemi)) +
    geom_ribbon(data = df, aes(x = nodeID , ymin = ymin ,ymax = ymax, fill = hemi), alpha = .5, linetype = 0) + 
    geom_line(data = df, aes(x = nodeID, y = mean_scalar, color = hemi),size = 1) +
    scale_fill_manual(values = c("Left" = custom_color, "Right" = custom_color_rh)) +
    scale_color_manual(values = c("Left" = custom_color, "Right" = custom_color_rh)) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract)
  }
  return(plot)
}

# coefficient of variation function
cv <- function(x, na.rm = FALSE)  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)

# function for plotting tractprofiles cv
plot_cv <- function(tract, scalar, custom_color, custom_color_rh=NULL) {
  df_profiles <- get(paste0("tract_profiles_", scalar, "_summary"))
  df <- df_profiles %>% filter(tractID == tract)
  
  if(is.null(custom_color_rh)) {
    plot <- ggplot(data = df, aes_string(x = "nodeID", y = scalar)) +
    geom_line(data = df, aes(x = nodeID, y = cv_scalar), size = line_size, color = custom_color) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.title.y = element_blank())
    
  } else {
    plot <- ggplot(data = df, aes(x = nodeID, y = cv_scalar, group = hemi)) +
    geom_line(data = df, aes(x = nodeID, y = cv_scalar, color = hemi),size = 1) +
    scale_fill_manual(values = c("Left" = custom_color, "Right" = custom_color_rh)) +
    scale_color_manual(values = c("Left" = custom_color, "Right" = custom_color_rh)) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract)
  }
  return(plot)
}


arrange_by_orientation <- function(list_figures, age_effect_type) {
  
  
  AP_plots <- ggarrange(list_figures$`Anterior Thalamic Radiation`, 
                           list_figures$`Cingulum Cingulate`, 
                           list_figures$`Inferior Fronto-occipital Fasciculus`, 
                           list_figures$`Inferior Longitudinal Fasciculus`, 
                           list_figures$`Superior Longitudinal Fasciculus`, ncol=5, nrow = 1)
  AP_plots_final <- annotate_figure(AP_plots, top = text_grob("Anterior - Posterior", 
                 color = "black", face = "bold", size = 14))
  
  
  AP_frontotemp_plots <- ggarrange(list_figures$`Arcuate Fasciculus`,
                                      list_figures$`Uncinate Fasciculus`, ncol=2, nrow = 1)
  AP_frontotemp_plots_final <- annotate_figure(AP_frontotemp_plots, top = text_grob("Anterior - Posterior (Frontal - Temporal)", 
                 color = "black", face = "bold", size = 14))
  
   
  RL_plots <- ggarrange(list_figures$`Forceps Major`, list_figures$`Forceps Minor`, 
                           ncol=2, nrow = 1)
  RL_plots_final <- annotate_figure(RL_plots, top = text_grob("Right - Left", 
                 color = "black", face = "bold", size = 14))
  
  
  SI_plots <- ggarrange(list_figures$`Corticospinal Tract`, list_figures$`Posterior Arcuate`, 
                           list_figures$`Vertical Occipital Fasciculus`, common.legend=TRUE, legend=c("right"),
                           ncol=3, nrow = 1)
  SI_plots_final <- annotate_figure(SI_plots, top = text_grob("Superior - Inferior", 
                 color = "black", face = "bold", size = 14))
  
  
  tractprofiles_plot <- ggarrange(AP_plots_final, ggarrange(AP_frontotemp_plots_final, RL_plots_final, ncol = 2), SI_plots_final, nrow = 3)
  
  tractprofiles_plot_final <- annotate_figure(tractprofiles_plot, top = text_grob(age_effect_type, 
               color = "black", face = "italic", size = 18))
   return(tractprofiles_plot_final)
}

```

## Fractional Anisotropy
Mean FA across subjects
```{r mean sd dti_fa, fig.width = 15, fig.height = 10}
tract_profiles_fa <- tract_profiles %>% select(-dti_md) 
tract_profiles_fa_summary <- tract_profiles_fa %>% 
  group_by(tractID, tract_hemi, nodeID, hemi) %>% 
  summarise(mean_scalar = mean(dti_fa, na.rm=T),
            sd = sd(dti_fa, na.rm=T),
            ymin = mean_scalar - sd,
            ymax = mean_scalar + sd)  

# plot
mean_sd_fa <- lapply(unique(tract_profiles$tractID), plot_mean_sd, scalar="fa", custom_color = "#831D69FF", custom_color_rh = "#FA9BAC")
names(mean_sd_fa) <- unique(tract_profiles$tractID)

# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
x.grob <- textGrob("Position on Tract (Node ID)", 
                   gp=gpar(col="black", fontsize=14))
y.grob <- textGrob("Fractional Anisotropy", 
                   gp=gpar(col="black", fontsize=14), rot=90)

space.grob <- textGrob(" ", 
                   gp=gpar(col="black", fontsize=28), rot=90)

tractprofiles_fa_mean_final <- arrange_by_orientation(mean_sd_fa, "Mean +/- 1 SD")
grid.arrange(arrangeGrob(tractprofiles_fa_mean_final, left = y.grob, bottom = x.grob, right = space.grob))

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/tract_profiles_fa_mean.png", grid.arrange(arrangeGrob(tractprofiles_fa_mean_final, left = y.grob, bottom = x.grob, right = space.grob)), width = 15, height = 10, units = "in")

```

FA tract profiles modeled using GAMs for each tract:
`gamm(FA ~ s(nodeID, k=24, fx = F), random = list(subjectID=~1), data=tract_data)`
```{r fit nodeID gams fa, cache=TRUE, results = 'hide', include = FALSE}
tractprofiles_fa_plots <- lapply(unique(tract_profiles$tractID), visualize_tractprofiles, df = tract_profiles, scalar = "dti_fa")
names(tractprofiles_fa_plots) <- unique(tract_profiles$tractID)
```

```{r tractprofiles fa plots, fig.width=15, fig.height=10}
# plot by AP, AP_frontal_temporal, RL (forceps), and SI
tractprofiles_fa_final <- arrange_by_orientation(tractprofiles_fa_plots, "")
grid.arrange(arrangeGrob(tractprofiles_fa_final, left = y.grob, bottom = x.grob, right = space.grob))

 
#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/tract_profiles_dti_fa.png", grid.arrange(arrangeGrob(tractprofiles_fa_final, left = y.grob, bottom = x.grob, right = space.grob)), width = 15, height = 10, units = "in")
```

## Mean Diffusivity
Mean MD across subjects
```{r mean sd dti_md, fig.width = 15, fig.height = 10}
tract_profiles_dti_md <- tract_profiles %>% select(-dti_fa) 
sqrt_n <- length(unique(tract_profiles_dti_md$subjectID))
tract_profiles_dti_md_summary <- tract_profiles_dti_md %>% 
  group_by(tractID, tract_hemi, nodeID, hemi) %>% 
  summarise(mean_scalar = mean(dti_md, na.rm=T),
            sd = sd(dti_md, na.rm=T),
            #se = sd/sqrt_n,
            #ymin = mean_scalar - se,
            #ymax = mean_scalar + se,
            ymin = mean_scalar - sd,
            ymax = mean_scalar + sd,
            cv_scalar = cv(dti_md, na.rm=T))  

# plot
mean_sd_dti_md <- lapply(unique(tract_profiles$tractID), plot_mean_sd, scalar="dti_md", custom_color = "#831D69FF", custom_color_rh = "#FA9BAC")
names(mean_sd_dti_md) <- unique(tract_profiles$tractID)

# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
x.grob <- textGrob("Position on Tract (Node ID)", 
                   gp=gpar(col="black", fontsize=14))
y.grob <- textGrob("Mean Diffusivity", 
                   gp=gpar(col="black", fontsize=14), rot=90)

space.grob <- textGrob(" ", 
                   gp=gpar(col="black", fontsize=28), rot=90)

# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
tractprofiles_md_mean_final <- arrange_by_orientation(mean_sd_dti_md, "Mean +/- 1 SD")
grid.arrange(arrangeGrob(tractprofiles_md_mean_final, left = y.grob, bottom = x.grob, right = space.grob))


#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/tract_profiles_dti_md_mean.png", grid.arrange(arrangeGrob(tractprofiles_md_mean_final, left = y.grob, bottom = x.grob, right = space.grob)), width = 15, height = 10, units = "in")


```

```{r cv dti_md, fig.width = 15, fig.height = 10}



cv_dti_md <- lapply(unique(tract_profiles$tractID), plot_cv, scalar="dti_md", custom_color = "#831D69FF", custom_color_rh = "#FA9BAC")
names(cv_dti_md) <- unique(tract_profiles$tractID)


# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
x.grob <- textGrob("Position on Tract (Node ID)", 
                   gp=gpar(col="black", fontsize=14))
y.grob <- textGrob("Coefficient of Variation", 
                   gp=gpar(col="black", fontsize=14), rot=90)

space.grob <- textGrob(" ", 
                   gp=gpar(col="black", fontsize=28), rot=90)

# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
tractprofiles_md_cv_final <- arrange_by_orientation(cv_dti_md, "Coefficient of Variation for Mean Diffusivity")
grid.arrange(arrangeGrob(tractprofiles_md_cv_final, left = y.grob, bottom = x.grob, right = space.grob))

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/tract_profiles_dti_md_cv.png", grid.arrange(arrangeGrob(tractprofiles_md_cv_final, left = y.grob, bottom = x.grob, right = space.grob)), width = 15, height = 10, units = "in")


```

MD tract profiles modeled using GAMs for each tract:
`gamm(MD ~ s(nodeID, k=24, fx = F), random = list(subjectID=~1), data=tract_data)`
```{r fit nodeID gams md, cache = TRUE, results = 'hide', include = FALSE}
tractprofiles_md_plots <- lapply(unique(tract_profiles$tractID), visualize_tractprofiles, df = tract_profiles, scalar = "dti_md")
names(tractprofiles_md_plots) <- unique(tract_profiles$tractID)
```

```{r tractprofiles md plots, fig.width=15, fig.height=10}
# plot by AP, AP_frontal_temporal, RL (forceps), and SI
x.grob <- textGrob("Position on Tract (Node ID)", 
                   gp=gpar(col="black", fontsize=14))
y.grob <- textGrob("Mean Diffusivity", 
                   gp=gpar(col="black", fontsize=14), rot=90)

space.grob <- textGrob(" ", 
                   gp=gpar(col="black", fontsize=28), rot=90)
 
# plot by AP, AP_frontal_temporal, RL (forceps), and SI
tractprofiles_md_final <- arrange_by_orientation(tractprofiles_md_plots, "")
grid.arrange(arrangeGrob(tractprofiles_md_final, left = y.grob, bottom = x.grob, right = space.grob))

 
#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/tract_profiles_dti_md.png", grid.arrange(arrangeGrob(tractprofiles_md_final, left = y.grob, bottom = x.grob, right = space.grob)), width = 15, height = 10, units = "in")
```


```{r md profile boxplots, fig.height = 5, fig.width=5}
# filter for central nodes
central_tract_profiles <- tract_profiles %>% filter(nodeID > 39 & nodeID < 60)
mycolors <- colorRampPalette(paletteer::paletteer_d("PNWColors::Sailboat"))(12)
 
boxplot <- central_tract_profiles %>%
  ggplot(aes(x = reorder(tractID, dti_md, fun = mean), y=dti_md, fill=tractID)) +
    geom_boxplot(aes(color = tractID, fill = tractID), alpha=0.6) +
    scale_fill_manual(values = mycolors) + 
    scale_color_manual(values = mycolors) + 
    theme(
      legend.position="none",
      plot.title = element_text(size=11, hjust=0.5),
      axis.text.x = element_text(angle=45, hjust=0.95, size=8)
    ) +
    ggtitle("Median MD of Deep/Central Nodes") +
    xlab("") + ylab("Mean Diffusivity")




ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/tract_profiles_md_boxplot.png", boxplot, width = 5, height = 5, units = "in")
```


```{r,fig.height = 6, fig.width=15}

tract_profiles <- tract_profiles %>% mutate(node_bin5 = ifelse(nodeID < 21, "nodes 0-20", ifelse(nodeID > 20 & nodeID < 41, "nodes 21-40", 
                                                                                                ifelse(nodeID > 40 & nodeID < 61, "nodes 41-60",
                                                                                                       ifelse(nodeID > 60 & nodeID < 81, "nodes 61-80",
                                                                                                              ifelse(nodeID > 80 & nodeID < 100, "nodes 81-99", NA))))))
 
mycolors <- colorRampPalette(paletteer::paletteer_d("PNWColors::Sailboat"))(5)
boxplot_binned <- tract_profiles %>%
  ggplot(aes(x = reorder(tractID, dti_md, fun = mean), y=dti_md, fill=node_bin5)) + 
    geom_boxplot(aes(color = node_bin5, fill = node_bin5), alpha=0.6) +
    scale_fill_manual(values = mycolors) + 
    scale_color_manual(values = mycolors) + 
    theme(
      legend.position="right",
      plot.title = element_text(size=12, hjust=0.5),
      axis.text.x = element_text(angle=45, hjust=0.95, size=12)
    ) +
    ggtitle("MD Across and Within Tracts") + labs(colour = "Node ID") + guides(fill = FALSE) + 
    xlab("") + ylab("Mean Diffusivity")

boxplot_binned

ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/tract_profiles_bin5_md_boxplot.png", boxplot_binned, height = 6, width = 15, units = "in")

```

```{r, fig.height = 6, fig.width=12}
tract_profiles <- tract_profiles %>% mutate(node_bin3 = ifelse(nodeID < 33, "nodes 0-32", ifelse(nodeID > 32 & nodeID < 66, "nodes 33-65", 
                                                                                                ifelse(nodeID > 65 & nodeID < 100, "nodes 66-99", NA))))
 
mycolors <- colorRampPalette(paletteer::paletteer_d("PNWColors::Sailboat"))(3)
 
 
 
boxplot_binned3 <- tract_profiles %>%
  ggplot(aes(x = reorder(tractID, dti_md, fun = mean), y=dti_md, fill=node_bin3)) +
    geom_boxplot(aes(color = node_bin3, fill = node_bin3), alpha=0.6) +
    scale_fill_manual(values = mycolors) + 
    scale_color_manual(values = mycolors) + 
    theme(
      legend.position="right",
      plot.title = element_text(size=12, hjust=0.5),
      axis.text.x = element_text(angle=45, hjust=0.95, size=12)
    ) +
    ggtitle("MD Across and Within Tracts") + labs(colour = "Node ID") + guides(fill = FALSE) + 
    xlab("") + ylab("Mean Diffusivity")

boxplot_binned3

ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/tract_profiles_bin3_md_boxplot.png", boxplot_binned3, height = 6, width = 15, units = "in")


```

```{r,fig.height = 6, fig.width=12}
mycolors <- colorRampPalette(paletteer::paletteer_d("PNWColors::Sailboat"))(3)

boxplot_facet3 <- ggplot(tract_profiles, aes(x = reorder(tractID, dti_md, fun = mean), y=dti_md, fill=node_bin3)) +
    geom_boxplot(aes(color = node_bin3, fill = node_bin3), alpha=0.6) +
    scale_fill_manual(values = mycolors) + 
    scale_color_manual(values = mycolors) + 
    theme(
      legend.position="right",
      plot.title = element_text(size=12, hjust=0.5),
      axis.text.x = element_text(angle=45, hjust=0.95, size=12)
    ) + facet_wrap(~node_bin3) +
    ggtitle("MD Across and Within Tracts") + labs(colour = "Node ID") + guides(fill = FALSE) + 
    xlab("") + ylab("Mean Diffusivity")

ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/tract_profiles_facet3_md_boxplot.png", boxplot_facet3, height = 6, width = 15, units = "in")

```

```{r,fig.height = 6, fig.width=20}

mycolors <- colorRampPalette(paletteer::paletteer_d("PNWColors::Sailboat"))(5)

boxplot_facet5 <- ggplot(tract_profiles, aes(x = reorder(tractID, dti_md, fun = mean), y=dti_md, fill=node_bin5)) +
    geom_boxplot(aes(color = node_bin5, fill = node_bin5), alpha=0.6) +
    scale_fill_manual(values = mycolors) + 
    scale_color_manual(values = mycolors) + 
    theme(
      legend.position="right",
      plot.title = element_text(size=12, hjust=0.5),
      axis.text.x = element_text(angle=45, hjust=0.95, size=12)
    ) + facet_wrap(~node_bin5, ncol = 5) +
    ggtitle("MD Across and Within Tracts") + labs(colour = "Node ID") + guides(fill = FALSE) + 
    xlab("") + ylab("Mean Diffusivity")


ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/tract_profiles_facet5_md_boxplot.png", boxplot_facet5, height = 6, width = 15, units = "in")

```




```{r ttest between central and peripheral, include=FALSE, echo=FALSE, eval=FALSE}

# get top and bottom 10 nodes, get middle 20 nodes for each tract
central <- tract_profiles %>% group_by(tract_hemi) %>% filter(nodeID < 10 | nodeID > 89)

peripheral <- tract_profiles %>% group_by(tract_hemi) %>% filter(nodeID > 39 & nodeID < 61)

central$tract_seg <- "central"
peripheral$tract_seg <- "peripheral"
central_stats <- central %>% select(tract_seg, dti_md)
peripheral_stats <- peripheral %>% select(tract_seg, dti_md)

ttest_df <- data.frame(rbind(central_stats, peripheral_stats))

var.test(central$dti_md, peripheral$dti_md) # different variances
t.test(central$dti_md, peripheral$dti_md)
t.test(dti_md ~ tract_seg, ttest_df) 
library(rstatix)
ttest_df %>%  cohens_d(dti_md ~ tract_seg, var.equal = FALSE)


mean(central$dti_md, na.rm=T) 
sd(central$dti_md, na.rm=T)
mean(peripheral$dti_md, na.rm=T)
sd(peripheral$dti_md, na.rm=T)

(mean(peripheral$dti_md, na.rm=T) - mean(central$dti_md, na.rm=T) ) / mean(central$dti_md, na.rm=T) 


# lm dist to cortex

central_periph_df <- compute_dist_to_cortex(tract_profiles)

central_periph_lm <- lm(dti_md ~ dist_to_cortex, data=central_periph_df)

summary(central_periph_lm)

coef(central_periph_lm)[2]          -> b  # Coefficient of interest
model.matrix(central_periph_lm)[,2] -> x  # Corresponding column of the design matrix
central_periph_df$dti_md - b * x      -> y 


cor.test(central_periph_df$dti_md, central_periph_df$dist_to_cortex)

```

# 2. Age effects
- Age effect was computed in 3 different ways:
1) Using nodewise GAMs to compute delta adjusted rsq for the age term
2) Using tractwise GAMs, and estimating posterior fitted values from the simulated GAM posterior distribution for each node age each age. Then computing the percent change in FA/MD for each node for each draw of the max and min age
3) Same as 2, but using normalized posterior percent change 
  
```{r functions for plotting age effects}
# function for computing credible intervals for percent change
compute_credible_intervals <- function(age_effect_df) {
  df <- age_effect_df %>% select(contains("draw"))
  median_age_effect <- apply(df, 1, function(x){median(x)})
  age_effect.credible.intervals <- apply(df, 1, function(x){quantile(x, probs = c(0.025, 0.975))}) #compute the credible interval for the percent change at each node 
  age_effect.credible.intervals <- as.data.frame(t(age_effect.credible.intervals))
    
  credible.intervals <- cbind(age_effect_df$nodeID, age_effect_df$tract_hemi, age_effect_df$tractID, age_effect_df$hemi, age_effect.credible.intervals)
  credible.intervals <- cbind(credible.intervals, median_age_effect)
  colnames(credible.intervals) <- c("nodeID","tract_hemi", "tractID", "hemi","lower","upper", "median_percent_change")
  credible.intervals$median_percent_change <- credible.intervals$median_percent_change*100
  credible.intervals$lower <- credible.intervals$lower*100
  credible.intervals$upper <- credible.intervals$upper*100
  
  return(credible.intervals)
}
 

# function for plotting age effect
plot_age_effect <- function(tract, age_effect_df, age_effect_type, custom_color, custom_color_rh=NULL) {
  df <- age_effect_df %>% filter(tractID == tract)
   if(age_effect_type == "adj_rsq") { # delta adj rsq
      # NA out color/fill aes if adj rsq = 0 or if p-value doesn't survive FDR correction (makes the color gray)
      includes_zero <- which(df$s_age.delta.adj.rsq_signed==0 | df$s_age.p.value.fdr > 0.05)
      
      if(str_detect(tract, "Forceps")) {
      df$tract_hemi[includes_zero] <- NA
       plot <- ggplot(data = df, aes(x = nodeID, y = s_age.delta.adj.rsq_signed, colour = tract_hemi, fill = tract_hemi)) +
        geom_point(shape = 21, size=2, alpha = 0.6) + 
        scale_colour_manual(values = custom_color, na.value = "grey50") +
        scale_fill_manual(values = custom_color, na.value = "grey50") + 
        theme(legend.position = "none",
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract)
      
      } else {
        df$hemi[includes_zero] <- NA
        plot <- ggplot(data = df, aes(x = nodeID, y = s_age.delta.adj.rsq_signed, group = hemi, colour = hemi, fill = hemi)) +
        geom_point(shape = 21, size=2, alpha = 0.8) + 
        scale_colour_manual(values = c("Left" = custom_color, "Right" = custom_color_rh), na.value = "grey50") +
        scale_fill_manual(values = c("Left" = custom_color, "Right" = custom_color_rh), na.value = "grey50") +
        theme(legend.position = "none",
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract)
      }
    } else if (age_effect_type == "percent_change") { # percent change or normalized percent change
      
      # NA out hemi if median_percent_change = 0 or if lower/upper CI contains 0 (makes the color gray)
      includes_zero <- which(df$lower <= 0 & df$upper >= 0)
      
 
      if(str_detect(tract, "Forceps")) {

        df$tractID[includes_zero] <- NA
        plot <- ggplot(data = df, aes(x = nodeID, y = median_percent_change, colour = tractID, fill = tractID)) +
        geom_point(shape = 21, size=2, alpha = 0.6) + 
        geom_errorbar(aes(ymin=lower, ymax=upper), width=.2,
              position=position_dodge(0.05), alpha = 0.8) + 
        scale_colour_manual(values = custom_color, na.value = "grey50") +
        scale_fill_manual(values = custom_color, na.value = "grey50") + 
        theme(legend.position = "none",
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract)
 
      } else {
        df$hemi[includes_zero] <- NA
        plot <- ggplot(data = df, aes(x = nodeID, y = median_percent_change, group = hemi, colour = hemi, fill = hemi)) +
        geom_point(shape = 21, size=2, alpha = 0.6) + 
        geom_errorbar(aes(ymin=lower, ymax=upper, colour = hemi), width=.2,
                 position=position_dodge(0.05), alpha = 0.8) + 
        scale_colour_manual(values = c("Left" = custom_color, "Right" = custom_color_rh), na.value = "grey50") +
        scale_fill_manual(values = c("Left" = custom_color, "Right" = custom_color_rh), na.value = "grey50") +
        theme(legend.position = "none",
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract)
      }
       
    } else {
      print("Provide valid age effect type")
    }
  return(plot)
}
```

```{r load age effects, cache=TRUE}

###############################################
### load nodewise GAM results: adjusted Rsq ###
###############################################
# load nodewise GAM results for dti_fa and dti_md
gam_age_fa <- read.csv("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_fa_nocovbat/GAMresults.dti_fa.age_nocovbat.csv")
gam_age_fa <- gam_age_fa %>% select(-X)

gam_age_md <- read.csv("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_md_nocovbat/GAMresults.dti_md.age_nocovbat.csv")
gam_age_md <- gam_age_md %>% select(-X)
 
# determine correct age effect sign for delta adjusted Rsq using equivalent linear model
# load linear models
lm_age_fa <- read.csv("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_fa_nocovbat/LinearModelresults.dti_fa.age_nocovbat.csv")
lm_age_md <- read.csv("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_md_nocovbat/LinearModelresults.dti_md.age_nocovbat.csv")
 
# take absolute value of delta adj rsq  
gam_age_fa$s_age.delta.adj.rsq_signed <- abs(gam_age_fa$s_age.delta.adj.rsq) 
gam_age_md$s_age.delta.adj.rsq_signed <- abs(gam_age_md$s_age.delta.adj.rsq) 
 
# apply the sign of linear model age estimate (t-value) to GAM delta adj rsq
gam_age_fa$s_age.delta.adj.rsq_signed <- gam_age_fa$s_age.delta.adj.rsq_signed * ifelse(lm_age_fa$age.estimate >= 0, 1, -1)
gam_age_md$s_age.delta.adj.rsq_signed <- gam_age_md$s_age.delta.adj.rsq_signed * ifelse(lm_age_md$age.estimate >= 0, 1, -1)
 
# How many nodes have significant age effect after FDR?
dti_fa_significant_count <- length(which(gam_age_fa$s_age.p.value.fdr < 0.05))
md_significant_count <- length(which(gam_age_md$s_age.p.value.fdr < 0.05))
 
total_nodes = 2200

print(paste0("Nodewise GAMs - dti_fa: ", dti_fa_significant_count, " nodes ", round(dti_fa_significant_count / total_nodes * 100, 2), "% with significant age effect after FDR correction"))
print(paste0("Nodewise GAMs - dti_md: ", md_significant_count, " nodes ", round(md_significant_count / total_nodes * 100, 2), "% with significant age effect after FDR correction"))


# Add tractID and nodeID to gam_age_fa
gam_age_fa$tractID <- all_subjects$tractID[c(1:2200)]
gam_age_fa$tract_hemi <- all_subjects$tract_hemi[c(1:2200)]
gam_age_fa$hemi <- all_subjects$hemi[c(1:2200)]
gam_age_fa$nodeID <- all_subjects$nodeID[c(1:2200)]

# Add tractID and nodeID to gam_age_md
gam_age_md$tractID <- all_subjects$tractID[c(1:2200)]
gam_age_md$tract_hemi <- all_subjects$tract_hemi[c(1:2200)]
gam_age_md$hemi <- all_subjects$hemi[c(1:2200)]
gam_age_md$nodeID <- all_subjects$nodeID[c(1:2200)]


############################################################
### load tractwise GAM results: posterior percent change ###
############################################################
percent_change_fa <- readRDS("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_fa/posterior_percentchange/HCPD_posterior_percentchange.RData")
percent_change_fa <- percent_change_fa %>% mutate(tractID = gsub("_"," ", tract_hemi)) %>% 
    mutate(hemi = ifelse(grepl("Left", tract_hemi), "Left", "Right"))
percent_change_fa$tractID <- gsub("Left |Right ", "", percent_change_fa$tractID)


percent_change_md <- readRDS("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_md/posterior_percentchange/HCPD_posterior_percentchange.RData")
percent_change_md <- percent_change_md %>% mutate(tractID = gsub("_"," ", tract_hemi)) %>% 
    mutate(hemi = ifelse(grepl("Left", tract_hemi), "Left", "Right"))
percent_change_md$tractID <- gsub("Left |Right ", "", percent_change_md$tractID)


#######################################################################
### load tractwise GAM results: normalized posterior percent change ###
#######################################################################
norm_percent_change_fa <- readRDS("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_fa/posterior_percentchange/HCPD_posterior_percentchange_normalized.RData")
norm_percent_change_fa <- norm_percent_change_fa %>% mutate(tractID = gsub("_"," ", tract_hemi)) %>% 
    mutate(hemi = ifelse(grepl("Left", tract_hemi), "Left", "Right"))
norm_percent_change_fa$tractID <- gsub("Left |Right ", "", norm_percent_change_fa$tractID)

norm_percent_change_md <- readRDS("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_md/posterior_percentchange/HCPD_posterior_percentchange_normalized.RData")
norm_percent_change_md <- norm_percent_change_md %>% mutate(tractID = gsub("_"," ", tract_hemi)) %>% 
    mutate(hemi = ifelse(grepl("Left", tract_hemi), "Left", "Right"))
norm_percent_change_md$tractID <- gsub("Left |Right ", "", norm_percent_change_md$tractID)
 

######################################################### 
### tractwise GAM results: compute credible intervals ###
######################################################### 

ci_pc_fa <- compute_credible_intervals(percent_change_fa)
ci_pc_md <- compute_credible_intervals(percent_change_md)

ci_norm_pc_fa <- compute_credible_intervals(norm_percent_change_fa)
ci_norm_pc_md <- compute_credible_intervals(norm_percent_change_md)


######################################################
### Fix tract orientations for all age effect df's ###
######################################################

fix_tract_orientations <- function(df) {
  # flip tracts
  to_reorient <- df %>%
    group_by(tractID) %>%
    mutate(nodeID = ifelse(tractID %in% c("Cingulum Cingulate", "Inferior Fronto-occipital Fasciculus", "Inferior Longitudinal Fasciculus", "Uncinate Fasciculus", "Corticospinal Tract"), max(nodeID) - nodeID, nodeID))
  
  # label main orientation
  reoriented <- to_reorient %>% 
    mutate(main_orientation = case_when(
      tractID %in% c("Anterior Thalamic Radiation", "Cingulum Cingulate", "Inferior Fronto-occipital Fasciculus",
                     "Inferior Longitudinal Fasciculus", "Superior Longitudinal Fasciculus") ~ "AP",
      tractID %in% c("Arcuate Fasciculus", "Uncinate Fasciculus") ~ "AP_frontal_temporal",
      tractID %in% c("Forceps Minor", "Forceps Major") ~ "RL",
      tractID %in% c("Corticospinal Tract", "Posterior Arcuate","Vertical Occipital Fasciculus") ~ "SI",
      TRUE ~ NA_character_
    ))
  
  reoriented$main_orientation <- as.factor(reoriented$main_orientation)
  
  return(reoriented)
}

gam_age_fa <- fix_tract_orientations(gam_age_fa)
gam_age_md <- fix_tract_orientations(gam_age_md)

ci_pc_fa <- fix_tract_orientations(ci_pc_fa)
ci_pc_md <- fix_tract_orientations(ci_pc_md)

ci_norm_pc_fa <- fix_tract_orientations(ci_norm_pc_fa)
ci_norm_pc_md <- fix_tract_orientations(ci_norm_pc_md)



gam_age_md <- gam_age_md %>% mutate(tract_node_noHemi = paste0(tractID, "_", nodeID))
gam_age_md_bilat <- gam_age_md %>% group_by(tract_node_noHemi) %>% summarise(s_age.delta.adj.rsq_signed = mean(s_age.delta.adj.rsq_signed))
gam_age_md_bilat <- gam_age_md_bilat %>% mutate(tractID = gsub("_[0-9]+", "",tract_node_noHemi))

gam_age_md_bilat <- gam_age_md_bilat %>% mutate(nodeID = str_extract(tract_node_noHemi, "[0-9]+"))
gam_age_md_bilat$nodeID <- as.integer(gam_age_md_bilat$nodeID)
gam_age_md_bilat <- gam_age_md_bilat %>% arrange(tractID, nodeID)

#write.csv(gam_age_md_bilat, "/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_md_nocovbat/GAM_bilateral_signedresults.dti_md.age_nocovbat.csv")

 

#write.csv(gam_age_fa, "/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_fa/GAMresults.dti_fa.age_reoriented.csv")
#write.csv(gam_age_md, "/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_md/GAMresults.dti_md.age_reoriented.csv")
 
#saveRDS(ci_pc_fa, "/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_fa/posterior_percentchange/HCPD_posterior_percentchange_reoriented.RData")
#saveRDS(ci_pc_md, "/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_md/posterior_percentchange/HCPD_posterior_percentchange_reoriented.RData")
 
#saveRDS(ci_norm_pc_fa, "/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_fa/posterior_percentchange/HCPD_posterior_percentchange_normalized_reoriented.RData")
#saveRDS(ci_norm_pc_md, "/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_md/posterior_percentchange/HCPD_posterior_percentchange_normalized_reoriented.RData")
 

bilat_gm_dist <- read.csv("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/all_subjects/across_subjects_bilateral_distances_to_gm.csv")
```

## Fractional Anisotropy
### 1. Nodewise GAMs: Delta Adjusted Rsq
- Delta adjust Rsq was computed using nodewise gam
`gam(fa_node_i_tract_i ~ s(age) + covariates))`
```{r fa nodeID vs. delta adj rsq, fig.width=16, fig.height=10}
# plot
age_rsq_fa_plots <- lapply(unique(gam_age_fa$tractID), plot_age_effect, age_effect_df = gam_age_fa, age_effect_type = "adj_rsq", custom_color = "#831D69FF", custom_color_rh = "#FA9BAC")
names(age_rsq_fa_plots) <- unique(gam_age_fa$tractID)
 
# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
x.grob <- textGrob("Position on Tract (Node ID)", 
                     gp=gpar(col="black", fontsize=14))
y.grob <- textGrob("Fractional Anisotropy Delta Adj Rsq", 
                   gp=gpar(col="black", fontsize=14), rot=90)

space.grob <- textGrob(" ", 
                     gp=gpar(col="black", fontsize=14), rot=90)
  

age_rsq_fa_plots_final <- arrange_by_orientation(age_rsq_fa_plots, "FA Delta Adjusted Rsq") 
grid.arrange(arrangeGrob(age_rsq_fa_plots_final, left = y.grob, bottom = x.grob, right = space.grob))


#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/age_effect_rsq_fa.png", grid.arrange(arrangeGrob(age_rsq_fa_plots_final, left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 10, units = "in")
```

### 2. Tractwise GAMs: Posterior percent change
- GAM used: `gam(tract_i_fa ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr') + ti(age, nodeID, k=c(3, 24), fx = F) + s(subjectID, bs = 're) , method = REML)`
- Estimated posterior fitted values from simulated GAM posterior distribution for each node at each age
- Computed percent change for each node for each draw of the max and min age: `(Y_old - Y_young)/(Y_young)`
- Computed credible intervals on the node-wise percent change from the posterior percent change values
```{r fa nodeID vs. percent change,  fig.width = 16, fig.height = 10}
 
percent_change_fa_plots <- lapply(unique(all_subjects$tractID), plot_age_effect, age_effect_df = ci_pc_fa, age_effect_type = "percent_change", custom_color = "#831D69FF", custom_color_rh = "#FA9BAC")
names(percent_change_fa_plots) <- unique(all_subjects$tractID)
 
 
# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
x.grob <- textGrob("Position on Tract (Node ID)", 
                     gp=gpar(col="black", fontsize=14))
y.grob <- textGrob("Fractional Anisotropy Percent Change", 
                   gp=gpar(col="black", fontsize=14), rot=90)

space.grob <- textGrob(" ", 
                     gp=gpar(col="black", fontsize=14), rot=90)
age_pc_fa_plots_final <- arrange_by_orientation(percent_change_fa_plots, "FA Percent Change") 

grid.arrange(arrangeGrob(age_pc_fa_plots_final, left = y.grob, bottom = x.grob, right = space.grob))

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/age_effect_pc_fa.png", grid.arrange(arrangeGrob(age_pc_fa_plots_final, left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 10, units = "in")

```

### 3. Tractwise GAMs: Normalized posterior percent change
- Same as `2` but computed normalized posterior percent change instead: `(Y_old - Y_young)/(mean_tract_i_node_i_fa)`
```{r fa nodeID vs. normalized percent change, fig.width = 16, fig.height=10}
norm_percent_change_fa_plots <- lapply(unique(all_subjects$tractID), plot_age_effect, age_effect_df = ci_norm_pc_fa, age_effect_type = "percent_change", custom_color = "#831D69FF", custom_color_rh = "#FA9BAC")
names(norm_percent_change_fa_plots) <- unique(all_subjects$tractID)

 

# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
x.grob <- textGrob("Position on Tract (Node ID)", 
                     gp=gpar(col="black", fontsize=14))
y.grob <- textGrob("Fractional Anisotropy Normalized Percent Change", 
                   gp=gpar(col="black", fontsize=14), rot=90)

space.grob <- textGrob(" ", 
                     gp=gpar(col="black", fontsize=14), rot=90)
age_norm_pc_plots_final <- arrange_by_orientation(norm_percent_change_fa_plots, "FA Normalized Percent Change") 

grid.arrange(arrangeGrob(age_norm_pc_plots_final, left = y.grob, bottom = x.grob, right = space.grob))


#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/age_effect_norm_pc_fa.png", grid.arrange(arrangeGrob(age_norm_pc_plots_final, left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 10, units = "in")
```

## Mean Diffusivity
### 1. Nodewise GAMs: Delta Adjusted Rsq
- Delta adjust Rsq was computed using nodewise gam
`gam(md_node_i_tract_i ~ s(age) + covariates))`
```{r md nodeID vs. delta adj rsq, fig.width=16, fig.height=10}
# plot
age_rsq_md_plots <- lapply(unique(gam_age_md$tractID), plot_age_effect, age_effect_df = gam_age_md, age_effect_type = "adj_rsq", custom_color = "#831D69FF", custom_color_rh = "#FA9BAC")
names(age_rsq_md_plots) <- unique(gam_age_md$tractID)
 
# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
x.grob <- textGrob("Position on Tract (Node ID)", 
                     gp=gpar(col="black", fontsize=14))
y.grob <- textGrob("Mean Diffusivity Delta Adj Rsq", 
                   gp=gpar(col="black", fontsize=14), rot=90)

space.grob <- textGrob(" ", 
                     gp=gpar(col="black", fontsize=14), rot=90)
age_rsq_md_plots_final <- arrange_by_orientation(age_rsq_md_plots, "MD Delta Adjusted Rsq") 
grid.arrange(arrangeGrob(age_rsq_md_plots_final, left = y.grob, bottom = x.grob, right = space.grob))

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/age_effect_rsq_md.png", grid.arrange(arrangeGrob(age_rsq_md_plots_final, left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 10, units = "in")
```

### 2. Tractwise GAMs: Posterior percent change
- GAM used: `gam(tract_i_md ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr') + ti(age, nodeID, k=c(3, 24), fx = F) + s(subjectID, bs = 're) , method = REML)`
- Estimated posterior fitted values from simulated GAM posterior distribution for each node at each age
- Computed percent change for each node for each draw of the max and min age: `(Y_old - Y_young)/(Y_young)`
- Computed credible intervals on the node-wise percent change from the posterior percent change values
```{r md nodeID vs. percent change,  fig.width = 16, fig.height = 10}
percent_change_md_plots <- lapply(unique(all_subjects$tractID), plot_age_effect, age_effect_df = ci_pc_md, age_effect_type = "percent_change", custom_color = "#831D69FF", custom_color_rh = "#FA9BAC")
names(percent_change_md_plots) <- unique(all_subjects$tractID)

 
# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
x.grob <- textGrob("Position on Tract (Node ID)", 
                     gp=gpar(col="black", fontsize=14))
y.grob <- textGrob("Mean Diffusivity Percent Change", 
                   gp=gpar(col="black", fontsize=14), rot=90)

space.grob <- textGrob(" ", 
                     gp=gpar(col="black", fontsize=14), rot=90)
age_pc_md_plots_final <- arrange_by_orientation(percent_change_md_plots, "MD Percent Change") 

grid.arrange(arrangeGrob(age_pc_md_plots_final, left = y.grob, bottom = x.grob, right = space.grob))
 
#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/age_effect_pc_md.png", grid.arrange(arrangeGrob(age_pc_md_plots_final, left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 10, units = "in")

```

### 3. Tractwise GAMs: Normalized posterior percent change
- Same as `2` but computed normalized posterior percent change instead: `(Y_old - Y_young)/(mean_tract_i_node_i_md)`
```{r md nodeID vs. normalized percent change, fig.width = 16, fig.height=10}
norm_percent_change_md_plots <- lapply(unique(all_subjects$tractID), plot_age_effect, age_effect_df = ci_norm_pc_md, age_effect_type = "percent_change", custom_color = "#831D69FF", custom_color_rh = "#FA9BAC")
names(norm_percent_change_md_plots) <- unique(all_subjects$tractID)

 

# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
x.grob <- textGrob("Position on Tract (Node ID)", 
                     gp=gpar(col="black", fontsize=14))
y.grob <- textGrob("Mean Diffusivity Normalized Percent Change", 
                   gp=gpar(col="black", fontsize=14), rot=90)

space.grob <- textGrob(" ", 
                     gp=gpar(col="black", fontsize=14), rot=90)
age_norm_pc_plots_final <- arrange_by_orientation(norm_percent_change_md_plots, "MD Normalized Percent Change") 

grid.arrange(arrangeGrob(age_norm_pc_plots_final, left = y.grob, bottom = x.grob, right = space.grob))

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/age_effect_norm_pc_md.png", grid.arrange(arrangeGrob(age_norm_pc_plots_final, left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 10, units = "in")
```

## Compare age effects to an equivalent linear model
- From Jason Yeatman:
```
How do the GAM results compare to a simpler model where you just fit a linear model at each node and look at the slope of development. Are the results pretty similar? 
One issue that we've been having with GAMs is that small differences in model specification have dramatic impacts on the results so I always try to confirm with the simple approach.
```
- I checked with MD only
- Output looks very similar to delta adj rsq and percent change! 
```{r md nodeID vs. slope of age effect from lm, fig.width = 16, fig.height=10}

# Add tractID and nodeID to lm_age_md
lm_age_md$tractID <- all_subjects$tractID[c(1:2200)]
lm_age_md$tract_hemi <- all_subjects$tract_hemi[c(1:2200)]
lm_age_md$hemi <- all_subjects$hemi[c(1:2200)]
lm_age_md$nodeID <- all_subjects$nodeID[c(1:2200)]
lm_age_md <- fix_tract_orientations(lm_age_md)
 
plot_age_effect_lm <- function(tract, age_effect_df, custom_color, custom_color_rh=NULL) {
  df <- age_effect_df %>% filter(tractID == tract)
  # NA out color/fill aes if adj rsq = 0 or if p-value doesn't survive FDR correction (makes the color gray)
  includes_zero <- which(df$age.p.value.fdr > 0.05)
  
  if(str_detect(tract, "Forceps")) {
  df$tract_hemi[includes_zero] <- NA
   plot <- ggplot(data = df, aes(x = nodeID, y = age.estimate, colour = tract_hemi, fill = tract_hemi)) +
    geom_point(shape = 21, size=2, alpha = 0.6) + 
    scale_colour_manual(values= custom_color, na.value = "grey50") + # allows for dynamic updating of tractID value
    scale_fill_manual(values=custom_color, na.value = "grey50") +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract)
  
  } else {
    df$hemi[includes_zero] <- NA
    plot <- ggplot(data = df, aes(x = nodeID, y = age.estimate, group = hemi, colour = hemi, fill = hemi)) +
    geom_point(shape = 21, size=2, alpha = 0.8) + 
    scale_colour_manual(values = c("Left" = custom_color, "Right" = custom_color_rh), na.value = "grey50") +
    scale_fill_manual(values = c("Left" = custom_color, "Right" = custom_color_rh), na.value = "grey50") +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          plot.title = element_text(size=14, hjust=0.5)) + labs(title = tract)
  }

  return(plot)
}

lm_md_plots <- lapply(unique(all_subjects$tractID), plot_age_effect_lm, age_effect_df = lm_age_md, custom_color = "#831D69FF", custom_color_rh = "#FA9BAC")
names(lm_md_plots) <- unique(all_subjects$tractID)

 

# arrange by AP, AP_frontal_temporal, RL (forceps), and SI
x.grob <- textGrob("Position on Tract (Node ID)", 
                     gp=gpar(col="black", fontsize=14))
y.grob <- textGrob("Mean Diffusivity Slope of Development (from linear model)", 
                   gp=gpar(col="black", fontsize=14), rot=90)

space.grob <- textGrob(" ", 
                     gp=gpar(col="black", fontsize=14), rot=90)
age_lm_plots_final <- arrange_by_orientation(lm_md_plots, "MD - Slope of Age Term (from linear model)") 

grid.arrange(arrangeGrob(age_lm_plots_final, left = y.grob, bottom = x.grob, right = space.grob))


##ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/age_effect_lm_md.png", grid.arrange(arrangeGrob(age_lm_plots_final, left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 10, units = "in")
```

# 3. Age effect vs. relative node position
- Are there systematic developmental changes happening along tracts?
```{r relative position function}
# relative position to end of tract: computed by difference between nodeID and 0 or 99. If nodeID < 50, then get abs diff to 0. If nodeID >= 50, then get abs diff to 99
# also, create column called tract_segment to label which half of the tract the node is in (i.e. anterior vs posterior)

compute_dist_to_cortex <- function(df) {
  df <- df %>% mutate(dist_to_cortex = 
           case_when(
             nodeID < 50 & !(tractID %in% c("Anterior Thalamic Radiation", "Corticospinal Tract")) ~ abs(nodeID - 0),
             nodeID >= 50 & !(tractID %in% c("Anterior Thalamic Radiation", "Corticospinal Tract")) ~ abs(nodeID - 99),
             tractID %in% c("Anterior Thalamic Radiation", "Corticospinal Tract") ~ nodeID )) %>%
  mutate(tract_segment = 
           case_when(
             nodeID < 50 & tractID %in% c("Cingulum Cingulate", "Inferior Fronto-occipital Fasciculus", "Inferior Longitudinal Fasciculus", "Superior Longitudinal Fasciculus") ~ "Anterior", 
             nodeID >= 50 & tractID %in% c("Cingulum Cingulate", "Inferior Fronto-occipital Fasciculus", "Inferior Longitudinal Fasciculus", "Superior Longitudinal Fasciculus") ~ "Posterior", 
             
             nodeID < 50 & tractID %in% c("Arcuate Fasciculus", "Uncinate Fasciculus") ~ "Anterior_frontal", 
             nodeID >= 50 & tractID %in% c("Arcuate Fasciculus", "Uncinate Fasciculus") ~ "Posterior_temporal",  
             
             nodeID < 50 & tractID %in% c("Forceps Minor", "Forceps Major") ~ "Right", 
             nodeID >= 50 & tractID %in% c("Forceps Minor", "Forceps Major") ~ "Left",
             
             nodeID < 50 & tractID %in% c("Posterior Arcuate", "Vertical Occipital Fasciculus") ~ "Superior", 
             nodeID >= 50 & tractID %in% c("Posterior Arcuate", "Vertical Occipital Fasciculus") ~ "Inferior",
             
             nodeID < 50 & tractID %in% c("Anterior Thalamic Radiation") ~ "Anterior", 
             nodeID >= 50 & tractID %in% c("Anterior Thalamic Radiation") ~ "Posterior_thalamus",
             
             nodeID < 50 & tractID %in% c("Corticospinal Tract") ~ "Superior_motor", 
             nodeID >= 50 & tractID %in% c("Corticospinal Tract") ~ "Inferior_brainstem",
             
           ))
  return(df)

}


plot_dist_to_cortex <- function(tract, age_effect_df, age_effect_type, custom_color_seg1, custom_color_seg2) {
  df <- age_effect_df %>% filter(tractID == tract) %>% arrange(tract_segment)
   if(age_effect_type == "adj_rsq") { # delta adj rsq
      # NA out color/fill aes if adj rsq = 0 or if p-value doesn't survive FDR correction (makes the color gray)
      includes_zero <- which(df$s_age.delta.adj.rsq_signed==0 | df$s_age.p.value.fdr > 0.05)
      
     if(str_detect(tract, "Forceps")) {
        
       df$tract_segment[includes_zero] <- NA
       seg1 <- "Left"
       seg2 <- "Right"
        
       plot <- ggplot(data = df, aes(x = dist_to_cortex, y = s_age.delta.adj.rsq_signed, colour = tract_segment, fill = tract_segment, group = tract_segment)) +
        geom_point(shape = 21, size=2, alpha = 0.7) + 
         scale_colour_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + # allows for dynamic updating 
        scale_fill_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + theme_classic() + 
        theme(legend.position = "bottom",
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=12, hjust=0.5)) + labs(title = tract) + labs(title = tract, colour = "") + guides(fill=FALSE)
    
       return(plot)
      
      } else {
        
       df$tract_segment[includes_zero] <- NA
       seg1 <- unique(df$tract_segment)[which(!is.na(unique(df$tract_segment)))][1] 
       seg2 <- unique(df$tract_segment)[which(!is.na(unique(df$tract_segment)))][2] 
        
       plot_lh <- df %>% filter(str_detect(df$tract_hemi, "Left")) %>%
        ggplot(aes(x = dist_to_cortex, y = s_age.delta.adj.rsq_signed, group = tract_segment, colour = tract_segment, fill = tract_segment)) +
        geom_point(shape = 21, size=2, alpha = 0.7) + 
        scale_colour_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + # allows for dynamic updating of tractID value
        scale_fill_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + theme_classic() + 
        theme(legend.position = "bottom",
              axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            plot.title = element_text(size=12, hjust=0.5)) +  
        labs(title = paste("Left", tract), colour = "") + guides(fill=FALSE)
      
      plot_rh <-  df %>% filter(str_detect(df$tract_hemi, "Right")) %>%
        ggplot(aes(x = dist_to_cortex, y = s_age.delta.adj.rsq_signed, group = tract_segment, colour = tract_segment, fill = tract_segment)) +
        geom_point(shape = 21, size=2, alpha = 0.7) + 
        scale_colour_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + # allows for dynamic updating of tractID value
        scale_fill_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + theme_classic() + 
        theme(legend.position = "bottom",
            axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            plot.title = element_text(size=12, hjust=0.5)) + labs(title = paste("Right", tract), colour = "") + guides(fill=FALSE)
      
        return(list(plot_lh, plot_rh))
        
        
      }
    } else if (age_effect_type == "percent_change") { # percent change or normalized percent change
      
      # NA out hemi if median_percent_change = 0 or if lower/upper CI contains 0 (makes the color gray)
      includes_zero <- which(df$lower <= 0 & df$upper >= 0)
      
 
      if(str_detect(tract, "Forceps")) {
        df$tractID[includes_zero] <- NA
        seg1 <- "Left"
        seg2 <- "Right"
        plot <- ggplot(data = df, aes(x = dist_to_cortex, y = median_percent_change, colour = tract_segment, fill = tract_segment, group = tract_segment)) +
        geom_point(shape = 21, size=2, alpha = 0.7) + 
        geom_errorbar(aes(ymin=lower, ymax=upper), width=1,
              position=position_dodge(0.05), alpha = 0.8) + 
        scale_colour_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + # allows for dynamic updating 
        scale_fill_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + theme_classic() + 
        theme(legend.position = "bottom",
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=12, hjust=0.5)) + labs(title = tract, colour = "") + guides(fill=FALSE)
        
        return(plot)
 
        } else {
        seg1 <- unique(df$tract_segment)[1]
        seg2 <- unique(df$tract_segment)[2]
        df$tract_segment[includes_zero] <- NA
        
        plot_lh <- df %>% filter(str_detect(df$tract_hemi, "Left")) %>%
          ggplot(aes(x = dist_to_cortex, y = median_percent_change, group = tract_segment, colour = tract_segment, fill = tract_segment)) +
        geom_point(shape = 21, size=2, alpha = 0.7) + 
        geom_errorbar(aes(ymin=lower, ymax=upper, colour = tract_segment), width=1,
                 position=position_dodge(0.05), alpha = 0.8) + 
        scale_colour_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + # allows for dynamic updating of tractID value
        scale_fill_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + theme_classic() + 
        theme(legend.position = "bottom",
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=12, hjust=0.5)) + labs(title = paste("Left", tract), colour = "") + guides(fill=FALSE)
        
        plot_rh <- df %>% filter(str_detect(df$tract_hemi, "Right")) %>%
          ggplot(aes(x = dist_to_cortex, y = median_percent_change, group = tract_segment, colour = tract_segment, fill = tract_segment)) +
        geom_point(shape = 21, size=2, alpha = 0.7) + 
        geom_errorbar(aes(ymin=lower, ymax=upper, colour = tract_segment), width=1,
                 position=position_dodge(0.05), alpha = 0.8) + 
        scale_colour_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + # allows for dynamic updating of tractID value
        scale_fill_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + theme_classic() + 
        theme(legend.position = "bottom",
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=12, hjust=0.5)) + labs(title = paste("Right", tract), colour = "") + guides(fill=FALSE)
        
        return(list(plot_lh, plot_rh))
      } 
       
    } else {
      print("Provide valid age effect type")
    }
  
}



arrange_dist_to_cortex <- function(list_figures, age_effect_type) {
  
  ATR_plots <- ggarrange(list_figures$`Anterior Thalamic Radiation`[[1]], list_figures$`Anterior Thalamic Radiation`[[2]], nrow = 2, common.legend=TRUE, legend = "bottom")
  
  AP_plots <- ggarrange(list_figures$`Cingulum Cingulate`[[1]], 
                           list_figures$`Inferior Fronto-occipital Fasciculus`[[1]], 
                           list_figures$`Inferior Longitudinal Fasciculus`[[1]], 
                           list_figures$`Superior Longitudinal Fasciculus`[[1]], 
                           list_figures$`Cingulum Cingulate`[[2]], 
                           list_figures$`Inferior Fronto-occipital Fasciculus`[[2]], 
                           list_figures$`Inferior Longitudinal Fasciculus`[[2]], 
                           list_figures$`Superior Longitudinal Fasciculus`[[2]], ncol=4, nrow = 2, common.legend=TRUE, legend = "bottom")
  
  AP_ATR_plots <- ggarrange(ATR_plots, AP_plots, ncol=2, widths = c(1,4)) 
  AP_plots_final <- annotate_figure(AP_ATR_plots, top = text_grob("Anterior - Posterior", 
                 color = "black", face = "bold", size = 12))
  
  
  AP_frontotemp_plots <- ggarrange(list_figures$`Arcuate Fasciculus`[[1]], list_figures$`Uncinate Fasciculus`[[1]],
                                   list_figures$`Arcuate Fasciculus`[[2]], list_figures$`Uncinate Fasciculus`[[2]], ncol=2, nrow = 2, common.legend=TRUE, legend = "bottom")
  AP_frontotemp_plots_final <- annotate_figure(AP_frontotemp_plots, top = text_grob("Anterior - Posterior (Frontal - Temporal)", 
                 color = "black", face = "bold", size = 12))
  
   
  RL_plots <- ggarrange(list_figures$`Forceps Major`, list_figures$`Forceps Minor`, 
                           ncol=2, nrow = 1, common.legend=TRUE, legend = "bottom")
  RL_plots_final <- annotate_figure(RL_plots, top = text_grob("Right - Left", 
                 color = "black", face = "bold", size = 12))
  
  CST_plots <- ggarrange(list_figures$`Corticospinal Tract`[[1]],list_figures$`Corticospinal Tract`[[2]], common.legend=TRUE, legend=c("bottom"), ncol=1, nrow = 2)
  SI_plots <- ggarrange(list_figures$`Posterior Arcuate`[[1]], 
                        list_figures$`Vertical Occipital Fasciculus`[[1]], 
                        list_figures$`Posterior Arcuate`[[2]], 
                        list_figures$`Vertical Occipital Fasciculus`[[2]], common.legend=TRUE, legend=c("bottom"), ncol=2, nrow = 2)
  
  SI_CST_plots <- ggarrange(CST_plots, SI_plots, widths = c(1, 2))
  SI_plots_final <- annotate_figure(SI_CST_plots, top = text_grob("Superior - Inferior", 
                 color = "black", face = "bold", size = 12))
  
  
  tractprofiles_plot <- ggarrange(AP_plots_final, ggarrange(AP_frontotemp_plots_final, RL_plots_final, ncol = 2), SI_plots_final, nrow = 3)
  
  tractprofiles_plot_final <- annotate_figure(tractprofiles_plot, top = text_grob(age_effect_type, 
               color = "black", face = "italic", size = 18))
   return(tractprofiles_plot_final)
}

```

```{r compute relative position to end of tract}
gam_age_fa <- compute_dist_to_cortex(gam_age_fa)  
gam_age_md <- compute_dist_to_cortex(gam_age_md)   
ci_pc_fa <- compute_dist_to_cortex(ci_pc_fa)   
ci_pc_md <- compute_dist_to_cortex(ci_pc_md)   
ci_norm_pc_fa <- compute_dist_to_cortex(ci_norm_pc_fa)   
ci_norm_pc_md <- compute_dist_to_cortex(ci_norm_pc_md)   
```

```{r plot relative position fa adj rsq, fig.height = 20, fig.width = 16}
 
custom_color_seg1 = "#4CC3FFFF"
custom_color_seg2 = "#FF9932FF"
 
rsq_fa_dist_plots <- lapply(unique(gam_age_fa$tractID), plot_dist_to_cortex, age_effect_df = gam_age_fa, age_effect_type = "adj_rsq", custom_color_seg1, custom_color_seg2)
names(rsq_fa_dist_plots) <- unique(gam_age_fa$tractID)
                                   
# FA delta adj rsq
x.grob <- textGrob("Distance to Cortex", 
                     gp=gpar(col="black", fontsize=12))
y.grob <- textGrob("Fractional Anisotropy Delta Adjusted Rsq", 
                   gp=gpar(col="black", fontsize=12), rot=90)
space.grob <- textGrob(" ", 
                     gp=gpar(col="black", fontsize=28), rot=90)
 
grid.arrange(arrangeGrob(arrange_dist_to_cortex(rsq_fa_dist_plots, "FA Delta Adjusted Rsq"), left = y.grob, bottom = x.grob, right = space.grob))

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/dist_to_cort_rsq_fa.png", grid.arrange(arrangeGrob(arrange_dist_to_cortex(rsq_fa_dist_plots, "FA Delta Adjusted Rsq"), left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 20, units = "in")
 
```

```{r plot relative position fa pc, fig.height = 20, fig.width = 16}

pc_fa_dist_plots <- lapply(unique(ci_pc_fa$tractID), plot_dist_to_cortex, age_effect_df = ci_pc_fa, age_effect_type = "percent_change", custom_color_seg1, custom_color_seg2)
names(pc_fa_dist_plots) <- unique(ci_pc_fa$tractID)

# FA Percent Change  
y.grob <- textGrob("Fractional Anisotropy Percent Change", 
                   gp=gpar(col="black", fontsize=12), rot=90)
arrange_dist_to_cortex(pc_fa_dist_plots, "FA Percent Change")
#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/dist_to_cort_pc_fa.png", grid.arrange(arrangeGrob(arrange_dist_to_cortex(pc_fa_dist_plots, "FA Percent Change"), left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 20, units = "in")

```

```{r plot relative position fa norm pc, fig.height = 20, fig.width = 16}
norm_pc_fa_dist_plots <- lapply(unique(ci_norm_pc_fa$tractID), plot_dist_to_cortex, age_effect_df = ci_norm_pc_fa, age_effect_type = "percent_change", custom_color_seg1, custom_color_seg2)
names(norm_pc_fa_dist_plots) <- unique(ci_norm_pc_fa$tractID)

# FA Normalized Percent Change 
y.grob <- textGrob("Fractional Anisotropy Normalized Percent Change", 
                   gp=gpar(col="black", fontsize=12), rot=90)
arrange_dist_to_cortex(norm_pc_fa_dist_plots, "FA Normalized Percent Change")

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/dist_to_cort_norm_pc_fa.png", grid.arrange(arrangeGrob(arrange_dist_to_cortex(norm_pc_fa_dist_plots, "FA Normalized Percent Change"), left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 20, units = "in")

```

```{r plot relative position md adj rsq, fig.height = 20, fig.width = 16}
rsq_md_dist_plots <- lapply(unique(gam_age_md$tractID), plot_dist_to_cortex, age_effect_df = gam_age_md, age_effect_type = "adj_rsq", custom_color_seg1, custom_color_seg2)
names(rsq_md_dist_plots) <- unique(gam_age_md$tractID)

# MD delta adj rsq
y.grob <- textGrob("Mean Diffusivity Delta Adjusted Rsq", 
                   gp=gpar(col="black", fontsize=12), rot=90)
arrange_dist_to_cortex(rsq_md_dist_plots, "MD Delta Adjusted Rsq")

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/dist_to_cort_rsq_md.png", grid.arrange(arrangeGrob(arrange_dist_to_cortex(rsq_md_dist_plots, "MD Delta Adjusted Rsq"), left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 20, units = "in")

```

```{r plot relative position md pc, fig.height = 20, fig.width = 16}
pc_md_dist_plots <- lapply(unique(ci_pc_md$tractID), plot_dist_to_cortex, age_effect_df = ci_pc_md, age_effect_type = "percent_change", custom_color_seg1, custom_color_seg2)
names(pc_md_dist_plots) <- unique(ci_pc_md$tractID)

# MD Percent Change  
y.grob <- textGrob("Mean Diffusivity Percent Change", 
                   gp=gpar(col="black", fontsize=12), rot=90)
arrange_dist_to_cortex(pc_md_dist_plots, "MD Percent Change")

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/dist_to_cort_pc_md.png", grid.arrange(arrangeGrob(arrange_dist_to_cortex(pc_md_dist_plots, "MD Percent Change"), left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 20, units = "in")

```

```{r plot relative position md norm pc, fig.height = 20, fig.width = 16}
norm_pc_md_dist_plots <- lapply(unique(ci_norm_pc_md$tractID), plot_dist_to_cortex, age_effect_df = ci_norm_pc_md, age_effect_type = "percent_change", custom_color_seg1, custom_color_seg2)
names(norm_pc_md_dist_plots) <- unique(ci_norm_pc_md$tractID)
 
# MD Normalized Percent Change 
y.grob <- textGrob("Mean Diffusivity Normalized Percent Change", 
                   gp=gpar(col="black", fontsize=12), rot=90)
arrange_dist_to_cortex(norm_pc_md_dist_plots, "MD Normalized Percent Change")

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/dist_to_cort_norm_pc_md.png", grid.arrange(arrangeGrob(arrange_dist_to_cortex(norm_pc_md_dist_plots, "MD Normalized Percent Change"), left = y.grob, bottom = x.grob, right = space.grob)), width = 16, height = 20, units = "in")
```


```{r plot dist to cortex md adj rsq, fig.height = 20, fig.width = 16}
bilat_gm_dist <- read.csv("/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/all_subjects/across_subjects_bilateral_distances_to_gm.csv")


bilat_gm_dist <- bilat_gm_dist %>% mutate(
    
    tractID = case_when(
      bundle_name == "ATR" ~ "Anterior Thalamic Radiation",
      bundle_name == "CGC" ~ "Cingulum Cingulate",
      bundle_name == "CST" ~ "Corticospinal Tract",
      bundle_name == "IFO" ~ "Inferior Fronto-occipital Fasciculus",
      bundle_name == "ILF" ~ "Inferior Longitudinal Fasciculus",
      bundle_name == "SLF" ~ "Superior Longitudinal Fasciculus",
      bundle_name == "ARC" ~ "Arcuate Fasciculus",
      bundle_name == "UNC" ~ "Uncinate Fasciculus",
      bundle_name == "FA" ~ "Forceps Minor",
      bundle_name == "FP" ~ "Forceps Major",
      bundle_name == "pARC" ~ "Posterior Arcuate",
      bundle_name == "VOF" ~ "Vertical Occipital Fasciculus",
      TRUE ~ bundle_name
    )
  )
bilat_gm_dist <- bilat_gm_dist %>% relocate(tractID)
bilat_gm_dist <- bilat_gm_dist[,-2]


bilat_gm_dist_long <- bilat_gm_dist %>% pivot_longer(names_to="dist_to_gm", cols = names(bilat_gm_dist)[c(2:101)])
bilat_gm_dist_long <- bilat_gm_dist_long %>% mutate(tract_node_noHemi = paste0(tractID, "_", gsub("_dist", "", dist_to_gm)))
bilat_gm_dist_long$tract_node_noHemi <- gsub("_node", "", bilat_gm_dist_long$tract_node_noHemi)
 
bilat_gm_dist_long$nodeID <- gsub("_dist", "", bilat_gm_dist_long$dist_to_gm)
bilat_gm_dist_long$nodeID <- gsub("node_", "", bilat_gm_dist_long$nodeID)
bilat_gm_dist_long$nodeID <- as.numeric(bilat_gm_dist_long$nodeID)

bilat_gm_dist_long <- fix_tract_orientations(bilat_gm_dist_long)

bilat_gm_dist_toplot <- merge(bilat_gm_dist_long, gam_age_md_bilat, by = "tract_node_noHemi")
bilat_gm_dist_toplot <- bilat_gm_dist_toplot[,-3]
bilat_gm_dist_toplot <- bilat_gm_dist_toplot %>% rename(dist_to_gm = value)  %>% rename(tractID=tractID.x) 

 
bilat_gm_dist_toplot <- bilat_gm_dist_toplot[,-7]

bilat_gm_dist_toplot <- bilat_gm_dist_toplot %>% 
  mutate(tract_segment = 
           case_when(
             nodeID < 50 & tractID %in% c("Cingulum Cingulate", "Inferior Fronto-occipital Fasciculus", "Inferior Longitudinal Fasciculus", "Superior Longitudinal Fasciculus") ~ "Anterior", 
             nodeID >= 50 & tractID %in% c("Cingulum Cingulate", "Inferior Fronto-occipital Fasciculus", "Inferior Longitudinal Fasciculus", "Superior Longitudinal Fasciculus") ~ "Posterior", 
             
             nodeID < 50 & tractID %in% c("Arcuate Fasciculus", "Uncinate Fasciculus") ~ "Anterior_frontal", 
             nodeID >= 50 & tractID %in% c("Arcuate Fasciculus", "Uncinate Fasciculus") ~ "Posterior_temporal",  
             
             nodeID < 50 & tractID %in% c("Forceps Minor", "Forceps Major") ~ "Right", 
             nodeID >= 50 & tractID %in% c("Forceps Minor", "Forceps Major") ~ "Left",
             
             nodeID < 50 & tractID %in% c("Posterior Arcuate", "Vertical Occipital Fasciculus") ~ "Superior", 
             nodeID >= 50 & tractID %in% c("Posterior Arcuate", "Vertical Occipital Fasciculus") ~ "Inferior",
             
             nodeID < 50 & tractID %in% c("Anterior Thalamic Radiation") ~ "Anterior", 
             nodeID >= 50 & tractID %in% c("Anterior Thalamic Radiation") ~ "Posterior_thalamus",
             
             nodeID < 50 & tractID %in% c("Corticospinal Tract") ~ "Superior_motor", 
             nodeID >= 50 & tractID %in% c("Corticospinal Tract") ~ "Inferior_brainstem",
             
           ))

bilat_gm_dist_toplot %>% filter(tractID=="Corticospinal Tract")
```

```{r, fig.height=15, fig.width=20}
plot_dist_to_gm <- function(tract, age_effect_df, age_effect_type, custom_color_seg1, custom_color_seg2) {
  df <- age_effect_df %>% filter(tractID == tract) %>% arrange(tract_segment)
      
     if(str_detect(tract, "Forceps")) {
        
  
       seg1 <- "Left"
       seg2 <- "Right"
        
       plot <- ggplot(data = df, aes(x = dist_to_gm, y = s_age.delta.adj.rsq_signed, colour = tract_segment, fill = tract_segment, group = tract_segment)) +
        geom_point(shape = 21, size=2, alpha = 0.7) + 
         scale_colour_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + # allows for dynamic updating 
        scale_fill_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + theme_classic() + 
        theme(legend.position = "bottom",
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              plot.title = element_text(size=12, hjust=0.5)) + labs(title = tract) + labs(title = tract, colour = "") + guides(fill=FALSE)
   
       return(plot)
      
      } else {
        
        
       seg1 <- unique(df$tract_segment)[which(!is.na(unique(df$tract_segment)))][1] 
       seg2 <- unique(df$tract_segment)[which(!is.na(unique(df$tract_segment)))][2] 
        
       plot <- df %>%  
        ggplot(aes(x = dist_to_gm, y = s_age.delta.adj.rsq_signed, group = tract_segment, colour = tract_segment, fill = tract_segment)) +
        geom_point(shape = 21, size=2, alpha = 0.7) + 
        scale_colour_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + # allows for dynamic updating of tractID value
        scale_fill_manual(values=c(custom_color_seg1, custom_color_seg2) |> `names<-`(c(seg1, seg2)), na.value = "grey50") + theme_classic() + 
        theme(legend.position = "bottom",
              axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            plot.title = element_text(size=12, hjust=0.5)) +  
        labs(title = tract, colour = "") + guides(fill=FALSE)
      
      
      
        return(plot)
      }
}


dist_plot <- lapply(unique(bilat_gm_dist_toplot$tractID), plot_dist_to_gm, age_effect_df = bilat_gm_dist_toplot, age_effect_type = "adj_rsq", "#99C9C7", '#E69982')
names(dist_plot) <- unique(bilat_gm_dist_toplot$tractID)


dist_plot_oriented <- arrange_by_orientation(dist_plot, "Distance to Cortex")



x.grob <- textGrob("Distance to Cortex", 
                     gp=gpar(col="black", fontsize=12))
y.grob <- textGrob("Age Effect (Delta Adj Rsq)", 
                   gp=gpar(col="black", fontsize=12), rot=90)

space.grob <- textGrob(" ", 
                     gp=gpar(col="black", fontsize=12), rot=90)

grid.arrange(arrangeGrob(dist_plot_oriented, left = y.grob, bottom = x.grob, right = space.grob))
```

Correlations between MD age effect and relative position to tract end
(Spearman's, since relative position is directly related to nodeID)
```{r correlations between age effect and position}

# make a dataframe that has column tract, correlation, and p-value
correlate_age_effect <- function(tract, age_effect_df, age_effect_type, hemi_avg = FALSE) {
  df <- age_effect_df %>% filter(tractID == tract)
  if (age_effect_type == "adj_rsq") {
    age_effect_var <- "s_age.delta.adj.rsq_signed"
  } else {
    age_effect_var <- "median_percent_change"
  }
  if (hemi_avg == TRUE & !str_detect(tract, "Forceps")) {
  
    df_avg <- df %>% group_by(nodeID) %>% summarise(mean=mean(get(age_effect_var)))
    names(df_avg) <- c("nodeID", age_effect_var)
   
    df <- right_join(df[c(1:100), c("tractID", "nodeID", "main_orientation", "dist_to_cortex", "tract_segment")], df_avg, by = "nodeID")
    
  }
  
 
  if(age_effect_type == "adj_rsq") {
    cor_dist_cortex <- cor.test(df$dist_to_cortex, df$s_age.delta.adj.rsq_signed, method = "spearman")$estimate
    pval_dist_cortex <- cor.test(df$dist_to_cortex, df$s_age.delta.adj.rsq_signed, method = "spearman")$p.value
    cor_dist_along_tract <- cor.test(df$nodeID, df$s_age.delta.adj.rsq_signed, method = "spearman")$estimate
    pval_dist_along_tract <- cor.test(df$nodeID, df$s_age.delta.adj.rsq_signed, method = "spearman")$p.value
 
  } else if(age_effect_type == "percent_change") {
    cor_dist_cortex <- cor.test(df$dist_to_cortex, df$median_percent_change, method = "spearman")$estimate
    pval_dist_cortex <- cor.test(df$dist_to_cortex, df$median_percent_change, method = "spearman")$p.value
    cor_dist_along_tract <- cor.test(df$nodeID, df$median_percent_change, method = "spearman")$estimate
    pval_dist_along_tract <- cor.test(df$nodeID, df$median_percent_change, method = "spearman")$p.value
  } else {
    print("Provide valid age effect type")
  }
  cor_df <- data.frame(tract, cor_dist_cortex, pval_dist_cortex, cor_dist_along_tract, pval_dist_along_tract)
  rownames(cor_df) <- NULL
  
  cor_df <- cor_df %>% mutate(pval_dist_cortex_sig = ifelse(pval_dist_cortex < 0.05, 1, 0)) %>%
    mutate(pval_dist_along_tract_sig = ifelse(pval_dist_along_tract < 0.05, 1, 0))
  cor_df <- cor_df %>% select(tract, cor_dist_cortex, pval_dist_cortex, pval_dist_cortex_sig, cor_dist_along_tract, pval_dist_along_tract, pval_dist_along_tract_sig)
    
  
  return(cor_df)
}
 
dist_age_effect_fdr <- function(cor_df) {
  
  cor_df$pval_dist_cortex_fdr <- p.adjust(cor_df$pval_dist_cortex, method=c("fdr"))
  cor_df$pval_dist_along_tract_fdr <- p.adjust(cor_df$pval_dist_along_tract, method=c("fdr"))
  cor_df <- cor_df %>% mutate(pval_dist_cortex_sig = ifelse(pval_dist_cortex_fdr < 0.05, 1, 0)) %>%
    mutate(pval_dist_along_tract_sig = ifelse(pval_dist_along_tract_fdr < 0.05, 1, 0))
  cor_df <- cor_df %>% select(tract, cor_dist_cortex, pval_dist_cortex, pval_dist_cortex_fdr, pval_dist_cortex_sig, cor_dist_along_tract, pval_dist_along_tract, pval_dist_along_tract_fdr, pval_dist_along_tract_sig)
  
  return(cor_df)
  
}

wrapper_correlate_age_effect <- function(age_effect_df, age_effect_type, tracts_exclude, hemi_avg=FALSE) {
  cor_dist <- lapply(unique(age_effect_df$tractID), correlate_age_effect, age_effect_df = age_effect_df, age_effect_type = age_effect_type, hemi_avg)
  cor_dist <- bind_rows(cor_dist)
  cor_dist <- dist_age_effect_fdr(cor_dist)
  mean_cor  <- cor_dist %>% filter(!tract %in% tracts_exclude) %>% 
                                                            summarise(mean_cor_dist_cortex = mean(cor_dist_cortex), 
                                                            mean_cor_dist_along_tract = mean(cor_dist_along_tract))
  return(list(cor_dist, mean_cor))
}

cor_dist_adj_rsq_fa <- wrapper_correlate_age_effect(gam_age_fa, age_effect_type = "adj_rsq", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"))
cor_dist_adj_rsq_md <- wrapper_correlate_age_effect(gam_age_md, age_effect_type = "adj_rsq", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"))
cor_dist_pc_fa <- wrapper_correlate_age_effect(ci_pc_fa, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"))
cor_dist_pc_md <- wrapper_correlate_age_effect(ci_pc_md, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"))
cor_dist_norm_pc_fa <- wrapper_correlate_age_effect(ci_norm_pc_fa, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"))
cor_dist_norm_pc_md <- wrapper_correlate_age_effect(ci_norm_pc_md, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"))
 

kable(cor_dist_adj_rsq_md[[1]][,c(1:3)], caption="Correlation between Age Effect (Delta Adjusted Rsq) and Position Along Tract for MD", align="ccc")  %>%
  kable_styling(font_size = 20, bootstrap_options = "hover")
```

Mean correlation across cortico-cortical tracts with all nodes and after excluding nodes from each end:
```{r correlations between age effect and distance exclude nodes}

# make a dataframe that has column tract, correlation, and p-value
correlate_age_effect_exclude <- function(tract, age_effect_df, age_effect_type, nodes_exclude) {
  df <- age_effect_df %>% filter(tractID == tract)
  to_exclude_lowest <- seq(nodes_exclude-nodes_exclude, nodes_exclude-1, 1)
  to_exclude_highest <- seq(100-nodes_exclude, 100-nodes_exclude+nodes_exclude-1, 1)
  df <- df %>% filter(!(nodeID %in% to_exclude_lowest | nodeID %in% to_exclude_highest))
 
  if(age_effect_type == "adj_rsq") {
    cor_dist_cortex <- cor.test(df$dist_to_cortex, df$s_age.delta.adj.rsq_signed, method = "spearman")$estimate
    pval_dist_cortex <- cor.test(df$dist_to_cortex, df$s_age.delta.adj.rsq_signed, method = "spearman")$p.value
    cor_dist_along_tract <- cor.test(df$nodeID, df$s_age.delta.adj.rsq_signed, method = "spearman")$estimate
    pval_dist_along_tract <- cor.test(df$nodeID, df$s_age.delta.adj.rsq_signed, method = "spearman")$p.value
    
  } else if(age_effect_type == "percent_change") {
    cor_dist_cortex <- cor.test(df$dist_to_cortex, df$median_percent_change, method = "spearman")$estimate
    pval_dist_cortex <- cor.test(df$dist_to_cortex, df$median_percent_change, method = "spearman")$p.value
    cor_dist_along_tract <- cor.test(df$nodeID, df$median_percent_change, method = "spearman")$estimate
    pval_dist_along_tract <- cor.test(df$nodeID, df$median_percent_change, method = "spearman")$p.value
  } else {
    print("Provide valid age effect type")
  }
  cor_df <- data.frame(tract, cor_dist_cortex, pval_dist_cortex, cor_dist_along_tract, pval_dist_along_tract)
  rownames(cor_df) <- NULL
  
  cor_df <- cor_df %>% mutate(pval_dist_cortex_sig = ifelse(pval_dist_cortex < 0.05, 1, 0)) %>%
    mutate(pval_dist_along_tract_sig = ifelse(pval_dist_along_tract < 0.05, 1, 0))
  cor_df <- cor_df %>% select(tract, cor_dist_cortex, pval_dist_cortex, pval_dist_cortex_sig, cor_dist_along_tract, pval_dist_along_tract, pval_dist_along_tract_sig)
    
  
  return(cor_df)
}
 
wrapper_correlate_age_effect_excl <- function(age_effect_df, age_effect_type, tracts_exclude, nodes_exclude) {
  cor_dist <- lapply(unique(age_effect_df$tractID), correlate_age_effect_exclude, age_effect_df = age_effect_df, age_effect_type = age_effect_type, nodes_exclude = nodes_exclude)
  cor_dist <- bind_rows(cor_dist)
  cor_dist <- dist_age_effect_fdr(cor_dist)
  mean_cor  <- cor_dist %>% filter(!tract %in% tracts_exclude) %>% 
                                                            summarise(mean_cor_dist_cortex = mean(cor_dist_cortex), 
                                                            mean_cor_dist_along_tract = mean(cor_dist_along_tract))
  return(list(cor_dist, mean_cor))
}

cor_dist_adj_rsq_excl_fa <- wrapper_correlate_age_effect_excl(gam_age_fa, age_effect_type = "adj_rsq", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), nodes_exclude = 10)
cor_dist_adj_rsq_excl_md <- wrapper_correlate_age_effect_excl(gam_age_md, age_effect_type = "adj_rsq", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), nodes_exclude = 10)
cor_dist_pc_excl_fa <- wrapper_correlate_age_effect_excl(ci_pc_fa, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), nodes_exclude = 10)
cor_dist_pc_excl_md <- wrapper_correlate_age_effect_excl(ci_pc_md, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), nodes_exclude = 10)
cor_dist_norm_pc_excl_fa <- wrapper_correlate_age_effect_excl(ci_norm_pc_fa, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), nodes_exclude = 10)
cor_dist_norm_pc_excl_md <- wrapper_correlate_age_effect_excl(ci_norm_pc_md, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), nodes_exclude = 10)
 
cor_dist_adj_rsq_excl20_fa <- wrapper_correlate_age_effect_excl(gam_age_fa, age_effect_type = "adj_rsq", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), nodes_exclude = 20)
cor_dist_adj_rsq_excl20_md <- wrapper_correlate_age_effect_excl(gam_age_md, age_effect_type = "adj_rsq", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), nodes_exclude = 20)
cor_dist_pc_excl20_fa <- wrapper_correlate_age_effect_excl(ci_pc_fa, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), nodes_exclude = 20)
cor_dist_pc_excl20_md <- wrapper_correlate_age_effect_excl(ci_pc_md, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), nodes_exclude = 20)
cor_dist_norm_pc_excl20_fa <- wrapper_correlate_age_effect_excl(ci_norm_pc_fa, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), nodes_exclude = 20)
cor_dist_norm_pc_excl20_md <- wrapper_correlate_age_effect_excl(ci_norm_pc_md, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), nodes_exclude = 20)


# make a table to compare correlations
mean_cor_compare <- data.frame(rbind(c(cor_dist_adj_rsq_md[[2]][1,1], cor_dist_adj_rsq_excl_md[[2]][1,1], cor_dist_adj_rsq_excl20_md[[2]][1,1]), c(cor_dist_pc_md[[2]][1,1], cor_dist_pc_excl_md[[2]][1,1], cor_dist_pc_excl20_md[[2]][1,1]), c(cor_dist_norm_pc_md[[2]][1,1], cor_dist_norm_pc_excl_md[[2]][1,1], cor_dist_norm_pc_excl20_md[[2]][1,1]))) %>% setNames(c("All Nodes", "Exclude 10 nodes from each end", "Exclude 20 nodes from each end"))
rownames(mean_cor_compare) <-  c("Delta Adj Rsq", "Posterior Percent Change", "Normalized Posterior Percent Change")


mean_cor_compare$notes <- c("CC, IFO, ILF, UNC, forceps major drop in correlation (IFO flips direction); but SLF, ARC, forceps minor, pARC and VOF are the same or better", "forceps maj reverses direction, IFO decreases; but forceps minor, CC, ARC, ILF, pARC, SLF, UNC, VOF increase in corr or stays approx the same ", "")
 
kable(mean_cor_compare, caption="Mean Correlation of Age Effect and Distance to Cortex")  %>%
  kable_styling(font_size = 20)
```

Mean correlations that were computed after averaging age effects of left and right tracts:
```{r correlations between age effect and distance average hemispheres}
 
cor_dist_adj_rsq_fa_hemi <- wrapper_correlate_age_effect(gam_age_fa, age_effect_type = "adj_rsq", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), hemi_avg = TRUE)
cor_dist_adj_rsq_md_hemi <- wrapper_correlate_age_effect(gam_age_md, age_effect_type = "adj_rsq", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), hemi_avg = TRUE)
cor_dist_pc_fa_hemi <- wrapper_correlate_age_effect(ci_pc_fa, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), hemi_avg = TRUE)
cor_dist_pc_md_hemi <- wrapper_correlate_age_effect(ci_pc_md, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), hemi_avg = TRUE)
cor_dist_norm_pc_fa_hemi <- wrapper_correlate_age_effect(ci_norm_pc_fa, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), hemi_avg = TRUE)
cor_dist_norm_pc_md_hemi <- wrapper_correlate_age_effect(ci_norm_pc_md, age_effect_type = "percent_change", tracts_exclude = c("Anterior Thalamic Radiation", "Corticospinal Tract"), hemi_avg = TRUE)
 
 
# make a table to compare correlations
hemi_cor_compare <- data.frame(rbind(c(cor_dist_adj_rsq_md[[2]][1,1], cor_dist_adj_rsq_md_hemi[[2]][1,1]), c(cor_dist_pc_md[[2]][1,1], cor_dist_pc_md_hemi[[2]][1,1]), c(cor_dist_norm_pc_md[[2]][1,1], cor_dist_norm_pc_md_hemi[[2]][1,1]))) %>% setNames(c("LH/RH age effects as separate datapoints", "Averaged LH/RH age effects"))
rownames(hemi_cor_compare) <-  c("Delta Adj Rsq", "Posterior Percent Change", "Normalized Posterior Percent Change")

kable(hemi_cor_compare, caption="Mean Correlation of Age Effect and Relative Positive to Tract End - Average Both Hemispheres", align = "cc")  %>%
  kable_styling(font_size = 20, bootstrap_options = "hover")
```

# 4. Developmental trajectories along each tract
- Plot trajectories for each node and color by nodeID. Each line corresponds to a unique node. 
- Note that the labels for tract orientations correspond to `low_nodeID - high_nodeID`. i.e. Anterior-posterior corresponds to anterior regions having low node IDs (blue) and posterior regions having high nodeIDs (blue)
- I used nodewise GAMs and estimated zero-centered smooths using `smooth_estimates`
(I don't know how to use `smooth_estimates` for tractwise GAMs - can't figure out how to navigate multiple smooths and a tensor product lol)
```{r nodewise GAMs zero centered smooths, fig.height = 5, fig.width=15}

cohortfile <- cohortfile %>% select(subjectID, age, sex, mean_fd)
gam_df <- merge(tract_profiles, cohortfile, by="subjectID")
gam_df <- gam_df %>% mutate(tract_node = paste0(tract_hemi, "_", nodeID))
gam_df$sex <- as.factor(gam_df$sex)
 
# create df with column tract_node and another column (dti_metric) that contains the values (nrow = N * n_tract_nodes)
# fa
nodewise_gam_fa <- gam_df %>% select(dti_fa, age, sex, mean_fd, tract_node)
nodewise_smoothfits_fa <- nodewise_gam_fa %>% group_by(tract_node) %>% 
  do(fit = smooth_estimates(gam(dti_fa ~ s(age,k=3,fx=F),data=.))) %>% unnest(fit)
nodewise_smoothfits_fa <- nodewise_smoothfits_fa %>% mutate(tractID = gsub("Left |Right ", "", gsub("_", " ", gsub("_[0-9]+", "", tract_node)))) %>% 
  mutate(nodeID = str_extract(tract_node, "[0-9]+")) %>%
  mutate(hemi = str_extract(tract_node, "Left|Right"))

orientations <- tract_profiles[,c("tractID", "main_orientation")]
orientations <- orientations[!duplicated(orientations),]
nodewise_smoothfits_fa <- merge(nodewise_smoothfits_fa,  orientations, by = "tractID")
                                                                
# md
nodewise_gam_md <- gam_df %>% select(dti_md, age, sex, mean_fd, tract_node)
nodewise_smoothfits_md <- nodewise_gam_md %>% group_by(tract_node) %>% 
  do(fit = smooth_estimates(gam(dti_md ~ s(age,k=3,fx=F),data=.))) %>% unnest(fit)
nodewise_smoothfits_md <- nodewise_smoothfits_md %>% mutate(tractID = gsub("Left |Right ", "", gsub("_", " ", gsub("_[0-9]+", "", tract_node)))) %>% 
  mutate(nodeID = str_extract(tract_node, "[0-9]+")) %>%
   mutate(hemi = str_extract(tract_node, "Left|Right"))
nodewise_smoothfits_md <- merge(nodewise_smoothfits_md,  orientations, by = "tractID")

 
 
 # plot trajectory for each tract and color by node
 # maybe compute average trajectory for each tract end?
```

```{r function plot zero centered dev traj tract ends}

plot_tract_ends_zeroed <- function(orientation, df, scalar, mycolors, ylim1, ylim2) {
  if (orientation == "CP") {
    title1 <- "Peripheral"
    title2 <- "Central"
    plot1 <- df %>% filter(tractID != "Anterior Thalamic Radiation") %>% 
      filter(tractID != "Corticospinal Tract") %>% 
      filter(nodeID < 10 | nodeID > 90) %>% group_by(tract_node) %>%
      ggplot(aes(x = age, y = est, group = tract_node)) +
        geom_line(aes(colour = tractID), size = 0.8, alpha = 0.7) + 
        theme_classic() +                         
        theme(text=element_text(size=12),
              legend.position = "right",
              plot.title = element_text(hjust=0.5)) + 
        scale_colour_manual(values = mycolors) +
        labs(y = paste("Fitted", scalar), x = "Age", title=title1) + ylim(ylim1, ylim2)
    
     plot2 <- df %>% filter(tractID != "Anterior Thalamic Radiation") %>% 
      filter(tractID != "Corticospinal Tract") %>% 
      filter(nodeID > 39 & nodeID < 60) %>% group_by(tract_node) %>%
      ggplot(aes(x = age, y = est, group = tract_node)) +
        geom_line(aes(colour = tractID), size = 0.8, alpha = 0.7) + 
        theme_classic() +                         
        theme(text=element_text(size=12),
              legend.position = "right",
              plot.title = element_text(hjust=0.5)) + 
        scale_colour_manual(values = mycolors) +
        labs(y = paste("Fitted", scalar), x = "Age", title=title2) + ylim(ylim1, ylim2)
    
  } else {
    if (orientation == "AP") {
      title1 <- "Anterior"
      title2 <- "Posterior"
    } else if (orientation == "AP_frontal_temporal") {
      title1 <- "Frontal"
      title2 <- "Temporal"
    } else if (orientation == "SI") {
      title1 <- "Superior"
      title2 <- "Inferior"
    } else {
      print("Provide valid orientation (AP, AP_frontal_temporal, SI, CP)")
    }
    plot1 <- df %>% filter(main_orientation == orientation) %>% 
      filter(tractID != "Anterior Thalamic Radiation") %>%
      filter(tractID != "Corticospinal Tract") %>%
      filter(nodeID < 10) %>% group_by(tract_node) %>%
      ggplot( aes(x = age, y = est, group = tract_node)) +
      geom_line(aes(colour = tractID), size = 0.8, alpha = 0.7) + theme_classic() +                        
      theme(text=element_text(size=12),
            legend.position = "right",
          plot.title = element_text(hjust=0.5)) + 
      scale_colour_manual(values = mycolors) +
      labs(y = paste("Fitted", scalar), x = "Age", title=title1) + ylim(ylim1, ylim2)
    
    plot2 <- df %>% filter(main_orientation == orientation) %>% 
      filter(tractID != "Anterior Thalamic Radiation") %>%
      filter(tractID != "Corticospinal Tract") %>%
      filter(nodeID > 89) %>% group_by(tract_node) %>%
      ggplot( aes(x = age, y = est, group = tract_node)) +
      geom_line(aes(colour = tractID), size = 0.8, alpha = 0.7) + theme_classic() +                        
      theme(text=element_text(size=12),
            legend.position = "right",
          plot.title = element_text(hjust=0.5)) + 
      scale_colour_manual(values = mycolors) +
      labs(y = paste("Fitted", scalar), x = "Age", title=title2) + ylim(ylim1, ylim2)
  }
  return(ggarrange(plot1, plot2, ncol=2, common.legend=T, legend="right"))
  
  
}

 
plot_nodes_zeroed_hemi <- function(tract, df, scalar, ylim1, ylim2) {
  
  df$nodeID <- as.numeric(df$nodeID)
  
  if(str_detect(tract, "Forceps")) {
      plot <- df %>% filter(tractID == tract) %>% 
      group_by(tract_node) %>%
      ggplot(aes(x = age, y = est, group = tract_node)) +
        geom_line(aes(colour = nodeID), size = 0.8, alpha = 0.6) + 
        theme_classic() +                         
        theme(text=element_text(size=12),
              legend.position = "none",
              plot.title = element_text(hjust=0.5)) + 
        paletteer::scale_colour_paletteer_c("grDevices::RdYlBu", direction=-1) +
        labs(x="", y="", title=tract) + ylim(ylim1, ylim2)
    
       return(plot)
  } else {
    plot1 <- df %>% filter(tractID == tract) %>% 
      filter(hemi == "Left") %>%
      group_by(tract_node) %>%
      ggplot(aes(x = age, y = est, group = tract_node)) +
        geom_line(aes(colour = nodeID), size = 0.8, alpha = 0.6) + 
        theme_classic() +                         
        theme(text=element_text(size=12),
              legend.position = "none",
              plot.title = element_text(hjust=0.5)) + 
        paletteer::scale_colour_paletteer_c("grDevices::RdYlBu", direction=-1) +
        labs(x="", y="", title=paste("Left", tract)) + ylim(ylim1, ylim2)
    
    plot2 <- df %>% filter(tractID == tract) %>% 
      filter(hemi == "Right") %>%
      group_by(tract_node) %>%
      ggplot(aes(x = age, y = est, group = tract_node)) +
        geom_line(aes(colour = nodeID), size = 0.8, alpha = 0.6) + 
        theme_classic() +                         
        theme(text=element_text(size=12),
              legend.position = "none",
              plot.title = element_text(hjust=0.5)) + 
        paletteer::scale_colour_paletteer_c("grDevices::RdYlBu", direction=-1) +
        labs(x="", y="", title=paste("Right", tract)) + ylim(ylim1, ylim2)
    return(plot_grid(plot1, plot2, nrow=2))
  }
  
}


plot_nodes_zeroed <- function(tract, df, scalar, ylim1, ylim2) {
  
  df$nodeID <- as.numeric(df$nodeID)

  plot <- df %>% filter(tractID == tract) %>% 
  group_by(tract_node) %>%
  ggplot(aes(x = age, y = est, group = tract_node)) +
    geom_line(aes(colour = nodeID), size = 0.8, alpha = 0.6) + 
    theme_classic() +                         
    theme(text=element_text(size=12),
          legend.position = "none",
          plot.title = element_text(hjust=0.5)) + 
    paletteer::scale_colour_paletteer_c("grDevices::RdYlBu", direction=-1) +
    labs(x="", y="", title=tract) + ylim(ylim1, ylim2)

   return(plot)
}
 

```

```{r plot fa nodewise dev traj zero centered all, fig.height=24, fig.width=22}
 
devtraj_zeroed_fa <- lapply(unique(nodewise_smoothfits_fa$tractID), plot_nodes_zeroed, df = nodewise_smoothfits_fa, scalar = "FA", -0.05, 0.03) 
names(devtraj_zeroed_fa) <- unique(nodewise_smoothfits_fa$tractID)
 
devtraj_zeroed_fa_plots <- arrange_by_orientation(devtraj_zeroed_fa, "Fractional Anisotropy")
 

x.grob <- textGrob("Age (years)", 
                     gp=gpar(col="black", fontsize=12))
y.grob <- textGrob("Zero-centered FA", 
                   gp=gpar(col="black", fontsize=12), rot=90)

space.grob <- textGrob(" ", 
                     gp=gpar(col="black", fontsize=12), rot=90)

grid.arrange(arrangeGrob(devtraj_zeroed_fa_plots, left = y.grob, bottom = x.grob, right = space.grob))

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/devtraj_zeroed_fa.png", grid.arrange(arrangeGrob(devtraj_zeroed_fa_plots, left = y.grob, bottom = x.grob, right = space.grob)), width = 22, height = 24, units = "in")
```

```{r plot md nodewise dev traj zero centered, fig.height=10, fig.width=25, echo = FALSE, include = FALSE}
# Define the number of colors you want
nb.cols <- 7
mycolors <- colorRampPalette(paletteer::paletteer_d("vapoRwave::floralShoppe"))(nb.cols)
AP_tract_ends_centered_md <- plot_tract_ends_zeroed("AP", nodewise_smoothfits_md, "MD", mycolors, -0.00005, 0.00006) 

nb.cols <- 5
mycolors <- colorRampPalette(paletteer::paletteer_d("vapoRwave::floralShoppe"))(nb.cols)
AP_frontotemp_tract_ends_centered_md <- plot_tract_ends_zeroed("AP_frontal_temporal", nodewise_smoothfits_md, "MD", mycolors, -0.00005, 0.00006) 

SI_tract_ends_centered_md <- plot_tract_ends_zeroed("SI", nodewise_smoothfits_md, "MD", mycolors, -0.00005, 0.00006) 
nb.cols <- 10
mycolors <- colorRampPalette(paletteer::paletteer_d("vapoRwave::floralShoppe"))(nb.cols)
 
CP_tract_ends_centered_md <- plot_tract_ends_zeroed("CP", nodewise_smoothfits_md, "MD", mycolors, -0.00005, 0.00006) 

ggarrange(AP_tract_ends_centered_md, AP_frontotemp_tract_ends_centered_md, SI_tract_ends_centered_md, CP_tract_ends_centered_md, nrow=2, ncol =2)



#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/fitted_val_tract_ends_centered_md.png", ggarrange(AP_tract_ends_centered_md, AP_frontotemp_tract_ends_centered_md, SI_tract_ends_centered_md, CP_tract_ends_centered_md, nrow=2, ncol =2) + bgcolor("white"), width = 25, height = 10, units = "in")
```

We see dissociable trajectories along each tract! 
- For A-P tracts, lower node IDs (blue; anterior segments) appear to have more steeply decreasing trajectories. Higher node IDs (red; posterior segments) appear to flatten out around mid-adolescence. Deeper tract regions (yellow) also have distinct trajectories that appear flatter than either end of the tract. 
- For R-L tracts, we also see distinct trajectories in deep white matter (yellow) compared to peripheral (red/blue), though forceps major seems to have flatter trajectories in left portions (blue) of its tract
- For S-I tracts, deeper white matter (yellow and red for CST; yellow for the other tracts) have flatter developmental trajectories compared to that at tract ends (blue for CST; red and blue for other tracts)

```{r plot md nodewise dev traj zero centered all, fig.height=20, fig.width=22}
 
devtraj_zeroed_md <- lapply(unique(nodewise_smoothfits_md$tractID), plot_nodes_zeroed, df = nodewise_smoothfits_md, scalar = "MD", -0.00005, 0.00006) 
names(devtraj_zeroed_md) <- unique(nodewise_smoothfits_md$tractID)
 
devtraj_zeroed_md_plots <- arrange_by_orientation(devtraj_zeroed_md, "Mean Diffusivity")
 

x.grob <- textGrob("Age (years)", 
                     gp=gpar(col="black", fontsize=12))
y.grob <- textGrob("Zero-centered MD", 
                   gp=gpar(col="black", fontsize=12), rot=90)

space.grob <- textGrob(" ", 
                     gp=gpar(col="black", fontsize=12), rot=90)


grid.arrange(arrangeGrob(devtraj_zeroed_md_plots, left = y.grob, bottom = x.grob, right = space.grob))

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/devtraj_zeroed_md.png", grid.arrange(arrangeGrob(devtraj_zeroed_md_plots, left = y.grob, bottom = x.grob, right = space.grob)), width = 22, height = 20, units = "in")
```


  
I also tried looking at developmental trajectories of nodes using tractwise GAMs. So I fit my GAM for each tract:
`gamm(MD ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr') + ti(age, nodeID, k=c(3, 24), fx = F) + sex + mean_fd, random = list(subjectID=~1), data = df)`, where I used bs='cr' for the smooth spline basis to match the ti() default spline basis. Then I used `fitted_values` to get the fitted MD for each node at each age.
```{r functions for fitted values on tractwise gams, include = FALSE, echo = FALSE}

## Developmental trajectories at each end of the tract
#- fitted values of nodes at each end of the tracts
#- refit GAMs to get smooth fits I think: age vs. DTI metric


fit_tensor_gam <- function(tract_name, scalar) {
  df <- gam_df %>% filter(tract_hemi == tract_name)
  df$subjectID <- as.factor(df$subjectID)
  te_gam <- gamm(as.formula(sprintf("%s ~ s(age, k = 3, fx = T, bs = 'cr') + s(nodeID, k = 24, fx = F, bs = 'cr') + 
                                  ti(age, nodeID, k=c(3, 24), fx = F) + sex + mean_fd", scalar)), random = list(subjectID=~1), data = df) 
 
  print(summary(te_gam$gam))
  print(k.check(te_gam$gam))
  print(paste(tract_name, "done"))
  te_gam <- te_gam$gam
  
  return(te_gam)  # return the model object 
}

fitted_all_ages <- function(tract_name, te_model) {
  df <- gam_df %>% filter(tract_hemi == tract_name)
  # Predict trends for each node, ages 8-22
  pdat <- with(df,
                data.frame(age = rep(8:22, each = 100),
                           nodeID = rep(seq(from = 0, to = 99, by = 1), times = 15), # could maybe do a version of this where by=1
                           mean_fd = mean(te_model$model$mean_fd),
                           subjectID = "subID", 
                           sex = "M"))
  newdf <- expand.grid(sex="M", age=c(8:22), mean_fd = mean(te_model$model$mean_fd))
  pdat <- cbind(fitted_values(te_model,data = pdat), c(rep(tract_name, nrow(newdf))))
  return(pdat)
}


plot_tract_ends <- function(orientation, df, scalar, mycolors, ylim1, ylim2) {
  if (orientation == "CP") {
    title1 <- "Peripheral"
    title2 <- "Central"
    plot1 <- df %>% filter(tractID != "Anterior Thalamic Radiation") %>% 
      filter(tractID != "Corticospinal Tract") %>% 
      filter(nodeID < 20 | nodeID > 80) %>% group_by(tract_node) %>%
      ggplot(aes(x = age, y = fitted, group = tract_node)) +
        geom_line(aes(colour = tractID), size = 0.8, alpha = 0.7) + 
        theme_classic() +                         
        theme(text=element_text(size=16),
              legend.position = "right",
              plot.title = element_text(hjust=0.5)) + 
        scale_colour_manual(values = mycolors) +
        labs(y = paste("Fitted", scalar), x = "Age", title=title1) + ylim(ylim1, ylim2)
    
     plot2 <- df %>% filter(tractID != "Anterior Thalamic Radiation") %>% 
      filter(tractID != "Corticospinal Tract") %>% 
      filter(nodeID > 39 & nodeID < 60) %>% group_by(tract_node) %>%
      ggplot(aes(x = age, y = fitted, group = tract_node)) +
        geom_line(aes(colour = tractID), size = 0.8, alpha = 0.7) + 
        theme_classic() +                         
        theme(text=element_text(size=16),
              legend.position = "right",
              plot.title = element_text(hjust=0.5)) + 
        scale_colour_manual(values = mycolors) +
        labs(y = paste("Fitted", scalar), x = "Age", title=title2) + ylim(ylim1, ylim2)
    
  } else {
    if (orientation == "AP") {
      title1 <- "Anterior"
      title2 <- "Posterior"
    } else if (orientation == "AP_frontal_temporal") {
      title1 <- "Frontal"
      title2 <- "Temporal"
    } else if (orientation == "SI") {
      title1 <- "Superior"
      title2 <- "Inferior"
    } else {
      print("Provide valid orientation (AP, AP_frontal_temporal, SI, CP)")
    }
    plot1 <- df %>% filter(main_orientation == orientation) %>% 
      filter(tractID != "Anterior Thalamic Radiation") %>%
      filter(tractID != "Corticospinal Tract") %>%
      filter(nodeID < 10) %>% group_by(tract_node) %>%
      ggplot( aes(x = age, y = fitted, group = tract_node)) +
      geom_line(aes(colour = tractID), size = 0.8, alpha = 0.7) + theme_classic() +                        
      theme(text=element_text(size=16),
            legend.position = "right",
          plot.title = element_text(hjust=0.5)) + 
      scale_colour_manual(values = mycolors) +
      labs(y = paste("Fitted", scalar), x = "Age", title=title1) + ylim(ylim1, ylim2)
    
    plot2 <- df %>% filter(main_orientation == orientation) %>% 
      filter(tractID != "Anterior Thalamic Radiation") %>%
      filter(tractID != "Corticospinal Tract") %>%
      filter(nodeID > 89) %>% group_by(tract_node) %>%
      ggplot( aes(x = age, y = fitted, group = tract_node)) +
      geom_line(aes(colour = tractID), size = 0.8, alpha = 0.7) + theme_classic() +                        
      theme(text=element_text(size=16),
            legend.position = "right",
          plot.title = element_text(hjust=0.5)) + 
      scale_colour_manual(values = mycolors) +
      labs(y = paste("Fitted", scalar), x = "Age", title=title2) + ylim(ylim1, ylim2)
  }
  return(ggarrange(plot1, plot2, ncol=2, common.legend=T, legend="right"))
}
```

 

```{r fit gams fa for fitted values, cache=TRUE, include=FALSE, echo=FALSE, eval=FALSE}

cohortfile <- cohortfile %>% select(subjectID, age, sex, mean_fd)
gam_df <- merge(tract_profiles, cohortfile, by="subjectID")
gam_df <- gam_df %>% mutate(tract_node = paste0(tract_hemi, "_", nodeID))
gam_df$sex <- as.factor(gam_df$sex)
 
 
# fit GAMs
tractwise_te_models_fa <- lapply(unique(gam_df$tract_hemi), fit_tensor_gam, scalar="dti_fa")
names(tractwise_te_models_fa) <- unique(gam_df$tract_hemi)
tractwise_te_models_fa$Left_Anterior_Thalamic_Radiation

# compute fitted values
fitted_all_ages_dfs_fa <- lapply(1:length(unique(gam_df$tract_hemi)), 
                              function(x) fitted_all_ages(unique(gam_df$tract_hemi)[x], tractwise_te_models_fa[[unique(gam_df$tract_hemi)[x]]]))
names(fitted_all_ages_dfs_fa) <- unique(gam_df$tract_hemi)
fitted_all_ages_dfs_fa <- lapply(fitted_all_ages_dfs_fa, setNames, nm=c("age", "nodeID", "mean_fd", "subjectID", "sex", "fitted", "se", "lower", "upper", "tract_hemi"))
 

# for now, keep nodes 0-19, 40-59, 80-99
fitted_fa <- bind_rows(fitted_all_ages_dfs_fa)
orientations <- tract_profiles[,c("tractID", "tract_hemi", "main_orientation")]
orientations <- orientations[!duplicated(orientations),]
fitted_fa <- left_join(fitted_fa, orientations, by = "tract_hemi")
fitted_fa <- fitted_fa %>% mutate(tract_node = paste0(tract_hemi, "_", nodeID))
```


```{r plot fa fitted values for tract ends, fig.height=10, fig.width=20, include=FALSE, echo=FALSE, eval=FALSE}


nb.cols <- 7
mycolors <- colorRampPalette(paletteer::paletteer_d("vapoRwave::floralShoppe"))(nb.cols)
AP_tract_ends_fa <- plot_tract_ends("AP", fitted_fa, "FA", mycolors, 0, 1) 

nb.cols <- 5
mycolors <- colorRampPalette(paletteer::paletteer_d("vapoRwave::floralShoppe"))(nb.cols)
AP_frontotemp_tract_ends_fa <- plot_tract_ends("AP_frontal_temporal", fitted_fa, "FA", mycolors, 0, 1) 
SI_tract_ends_fa <- plot_tract_ends("SI", fitted_fa, "FA", mycolors, 0, 1) 

nb.cols <- 10
mycolors <- colorRampPalette(paletteer::paletteer_d("vapoRwave::floralShoppe"))(nb.cols)
 
CP_tract_ends_fa <- plot_tract_ends("CP", fitted_fa, "FA", mycolors, 0, 1) 

ggarrange(AP_tract_ends_fa, AP_frontotemp_tract_ends_fa, SI_tract_ends_fa, CP_tract_ends_fa, nrow=2, ncol =2)
  

##ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/fitted_val_tract_ends_fa.png", 
#       ggarrange(AP_tract_ends_fa, AP_frontotemp_tract_ends_fa, SI_tract_ends_fa, CP_tract_ends_fa, nrow=2, ncol =2) + bgcolor("white"), width = 20, height = #10, units = "in")

```



```{r fit gams md for fitted values, cache = TRUE, include = FALSE}

# fit GAMs
tractwise_te_models_md <- lapply(unique(gam_df$tract_hemi), fit_tensor_gam, scalar="dti_md")
names(tractwise_te_models_md) <- unique(gam_df$tract_hemi)
tractwise_te_models_md$Left_Anterior_Thalamic_Radiation


# compute fitted values
fitted_all_ages_dfs_md <- lapply(1:length(unique(gam_df$tract_hemi)), 
                              function(x) fitted_all_ages(unique(gam_df$tract_hemi)[x], tractwise_te_models_md[[unique(gam_df$tract_hemi)[x]]]))
names(fitted_all_ages_dfs_md) <- unique(gam_df$tract_hemi)
fitted_all_ages_dfs_md <- lapply(fitted_all_ages_dfs_md, setNames, nm=c("age", "nodeID", "mean_fd", "subjectID", "sex", "fitted", "se", "lower", "upper", "tract_hemi"))

# for now, keep nodes 0-9, 45-55, 90-99
fitted_md <- bind_rows(fitted_all_ages_dfs_md)
orientations <- tract_profiles[,c("tractID", "tract_hemi", "main_orientation")]
orientations <- orientations[!duplicated(orientations),]
fitted_md <- left_join(fitted_md, orientations, by = "tract_hemi")
fitted_md <- fitted_md %>% mutate(tract_node = paste0(tract_hemi, "_", nodeID))
saveRDS(fitted_md, "/cbica/projects/luo_wm_dev/output/HCPD/tract_profiles/GAM/dti_md/dev_traj_fitted_md.RData")

```
 
 
It's definitely easier to see the differences in developmental trajectories with the zero-centered smooths above, but this is also kind of interesting... There's a super stark difference between central (nodes 40-59) and peripheral (nodes 0-19 and 80-99) tract regions, where deeper white matter have super distinct MD profiles between tracts.  
```{r  plot md fitted values for tract ends, fig.height=10, fig.width=20}
# Define the number of colors you want
nb.cols <- 7
mycolors <- colorRampPalette(paletteer::paletteer_d("LaCroixColoR::Pamplemousse"))(nb.cols)
AP_tract_ends_md <- plot_tract_ends("AP", fitted_md, "MD", mycolors, 0.00045, 0.0008) 

nb.cols <- 5
mycolors <- colorRampPalette(paletteer::paletteer_d("LaCroixColoR::Pamplemousse"))(nb.cols)
AP_frontotemp_tract_ends_md <- plot_tract_ends("AP_frontal_temporal", fitted_md, "MD", mycolors, 0.00045, 0.0008) 

SI_tract_ends_md <- plot_tract_ends("SI", fitted_md, "MD", mycolors, 0.00045, 0.0008) 
nb.cols <- 10
mycolors <- colorRampPalette(paletteer::paletteer_d("LaCroixColoR::Pamplemousse"))(nb.cols)
 
CP_tract_ends_md <- plot_tract_ends("CP", fitted_md, "MD", mycolors, 0.00045, 0.0008) 

ggarrange(AP_tract_ends_md, AP_frontotemp_tract_ends_md, SI_tract_ends_md, CP_tract_ends_md, nrow=2, ncol =2)

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/fitted_val_tract_ends_md.png", ggarrange(AP_tract_ends_md, AP_frontotemp_tract_ends_md, SI_tract_ends_md, CP_tract_ends_md, nrow=2, ncol =2) + bgcolor("white"), width = 20, height = 10, units = "in")

```


```{r md plot central, fig.height=13, fig.width=11}

mycolors <- colorRampPalette(paletteer::paletteer_d("PNWColors::Sailboat"))(5)[c(1,3)] 
somato_fitted <- fitted_md %>% filter(tractID == "Corticospinal Tract" | tractID == "Vertical Occipital Fasciculus") %>%
    filter(nodeID > 39 & nodeID < 60) %>% group_by(tract_node) %>%
    ggplot(aes(x = age, y = fitted, group = tract_node)) +
      geom_line(aes(colour = tractID), size = 0.8, alpha = 0.7) + 
      theme_classic() +                         
      theme(text=element_text(size=16),
            legend.position = "bottom", legend.text = element_text(size=10),legend.title=element_blank(),
            plot.title = element_text(hjust=0.5)) + #Top, left,bottom, bottom
      scale_colour_manual(values = mycolors) +
      labs(y = paste("Fitted MD"), x = "Age", title="") + ylim(0.00035, 0.0007) + guides(col = guide_legend(nrow = 2))

mycolors <- colorRampPalette(paletteer::paletteer_d("PNWColors::Sailboat"))(8)[c(3, 5, 7)]
slf_arc_fitted <- fitted_md %>% filter(tractID == "Superior Longitudinal Fasciculus" | 
                       tractID == "Arcuate Fasciculus" |
                       tractID == "Posterior Arcuate") %>%
    filter(nodeID > 39 & nodeID < 60) %>% group_by(tract_node) %>%
    ggplot(aes(x = age, y = fitted, group = tract_node)) +
      geom_line(aes(colour = tractID), size = 0.8, alpha = 0.7) + 
      theme_classic() +                         
      theme(text=element_text(size=16),
            legend.position = "bottom", legend.text = element_text(size=10),legend.title=element_blank(),
            plot.title = element_text(hjust=0.5)) + 
      scale_colour_manual(values = mycolors) +
      labs(y = paste("Fitted MD"), x = "Age", title="") + ylim(0.00035, 0.0007) + guides(col = guide_legend(nrow = 2))


mycolors <- colorRampPalette(paletteer::paletteer_d("PNWColors::Sailboat"))(15)[c(6, 8, 10, 12)]
frontal_fitted <- fitted_md %>% filter(tractID == "Anterior Thalamic Radiation" | tractID == "Cingulum Cingulate" | 
                       tractID == "Inferior Fronto-occipital Fasciculus" | tractID == "Inferior Longitudinal Fasciculus" |
                       tractID == "Uncinate") %>%
    filter(nodeID > 39 & nodeID < 60) %>% group_by(tract_node) %>%
    ggplot(aes(x = age, y = fitted, group = tract_node)) +
      geom_line(aes(colour = tractID), size = 0.8, alpha = 0.7) + 
      theme_classic() +                         
      theme(text=element_text(size=16),
            legend.position = "bottom", legend.text = element_text(size=10),legend.title=element_blank(),
            plot.title = element_text(hjust=0.5)) + 
      scale_colour_manual(values = mycolors) +
      labs(y = paste("Fitted MD"), x = "Age", title="") + ylim(0.00035, 0.0007) + guides(col = guide_legend(nrow = 2))
 
mycolors <- colorRampPalette(paletteer::paletteer_d("PNWColors::Sailboat"))(12)[c(1, 10)]
forceps_fitted <- fitted_md %>% filter(tractID == "Forceps Major" | tractID == "Forceps Minor") %>%
    filter(nodeID > 39 & nodeID < 60) %>% group_by(tract_node) %>%
    ggplot(aes(x = age, y = fitted, group = tract_node)) +
      geom_line(aes(colour = tractID), size = 0.8, alpha = 0.7) + 
      theme_classic() +                         
      theme(text=element_text(size=16),
            legend.position = "bottom", legend.text = element_text(size=10),legend.title=element_blank(),
            plot.title = element_text(hjust=0.5)) + 
      scale_colour_manual(values = mycolors) +
      labs(y = paste("Fitted MD"), x = "Age", title="") + ylim(0.00035, 0.0007) + guides(col = guide_legend(nrow = 2))

gA = ggplotGrob(somato_fitted)
gB = ggplotGrob(slf_arc_fitted)
gC = ggplotGrob(frontal_fitted)
gD = ggplotGrob(forceps_fitted)
gA$heights <- gB$heights
gC$heights <- gB$heights
gD$heights <- gB$heights

grid.arrange(gA, gB, gC, gD)


 
#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/fitted_val_tract_ends_md_arranged.png", grid.arrange(gA, gB, gC, gD), width = 11, height = 13, units = "in")

```


What do central vs. peripheral nodes look like for specific tracts? 
```{r md unpack central peripheral, fig.height=20, fig.width=13}
scalar <- "MD"
df <- fitted_md
title1 <- "Peripheral"
title2 <- "Central"
ylim1 <- 0.00045
ylim2 <- 0.0008

plot_central_peripheral <- function(tract) {
  plot1 <- df %>% filter(tractID == tract) %>% 
  filter(nodeID < 10 | nodeID > 90) %>% group_by(tract_node) %>%
  ggplot(aes(x = age, y = fitted, group = tract_node)) +
    geom_line(aes(colour = tractID), size = 0.9, alpha = 0.7) + 
    theme_classic() +                         
    theme(text=element_text(size=16),
          legend.position = "bottom",
          plot.title = element_text(hjust=0.5)) + 
    scale_colour_manual(values = "#4CC3FFFF") +
    labs(y = paste("Fitted", scalar), x = "Age", title=title1) + ylim(ylim1, ylim2)

 plot2 <- df %>% filter(tractID == tract) %>% 
  filter(nodeID > 39 & nodeID < 60) %>% group_by(tract_node) %>%
  ggplot(aes(x = age, y = fitted, group = tract_node)) +
    geom_line(aes(colour = tractID), size = 0.9, alpha = 0.7) + 
    theme_classic() +                         
    theme(text=element_text(size=16),
          legend.position = "bottom",
          plot.title = element_text(hjust=0.5)) + 
    scale_colour_manual(values = "#FF9932FF") +
    labs(y = paste("Fitted", scalar), x = "Age", title=title2) + ylim(ylim1, ylim2)
 
 return(plot_grid(plot1, plot2))
}

central_peripheral_plots <- lapply(c("Arcuate Fasciculus", "Cingulum Cingulate", "Forceps Major", "Forceps Minor", "Inferior Fronto-occipital Fasciculus", "Inferior Longitudinal Fasciculus", "Superior Longitudinal Fasciculus", "Uncinate Fasciculus", "Posterior Arcuate","Vertical Occipital Fasciculus"), plot_central_peripheral)

names(central_peripheral_plots) <- c("Arcuate Fasciculus", "Cingulum Cingulate", "Forceps Major", "Forceps Minor", "Inferior Fronto-occipital Fasciculus", "Inferior Longitudinal Fasciculus", "Superior Longitudinal Fasciculus", "Uncinate Fasciculus", "Posterior Arcuate","Vertical Occipital Fasciculus")

central_peripheral_specific <- ggarrange(plot_grid(plotlist=central_peripheral_plots[c(1:4)], ncol=2), plot_grid(plotlist=central_peripheral_plots[c(5:8)], ncol=2), plot_grid(plotlist=central_peripheral_plots[c(9:10)], ncol=2), heights = c(1,1,0.5), nrow=3)

central_peripheral_specific

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/fitted_val_tract_ends_cp_md.png", central_peripheral_specific, width = 13, height = 20, units = "in")
 
```

For some of the tracts, there's kind of two chunks of developmental trajectories - could they correspond to each end of the tract? 
```{r md anterior posterior specific tracts, fig.height = 15, fig.width=10}

scalar <- "MD"
df <- fitted_md
title1 <- "Anterior"
title2 <- "Posterior"
ylim1 <- 0.00045
ylim2 <- 0.0008

  

plot_anterior_posterior <- function(tract) {
  plot1 <- df %>% filter(tractID == tract) %>% 
  filter(nodeID < 10) %>% group_by(tract_node) %>%
  ggplot(aes(x = age, y = fitted, group = tract_node)) +
    geom_line(aes(colour = tractID), size = 0.9, alpha = 0.7) + 
    theme_classic() +                         
    theme(text=element_text(size=16),
          legend.position = "bottom",
          plot.title = element_text(hjust=0.5)) + 
    scale_colour_manual(values = "#4CC3FFFF") +
    labs(y = paste("Fitted", scalar), x = "Age", title=title1) + ylim(ylim1, ylim2)

 plot2 <- df %>% filter(tractID == tract) %>% 
  filter(nodeID > 89) %>% group_by(tract_node) %>%
  ggplot(aes(x = age, y = fitted, group = tract_node)) +
    geom_line(aes(colour = tractID), size = 0.9, alpha = 0.7) + 
    theme_classic() +                         
    theme(text=element_text(size=16),
          legend.position = "bottom",
          plot.title = element_text(hjust=0.5)) + 
    scale_colour_manual(values = "#FF9932FF") +
    labs(y = paste("Fitted", scalar), x = "Age", title=title2) + ylim(ylim1, ylim2)
 
 return(plot_grid(plot1, plot2))
}

anterior_posterior_specific <- ggarrange(plot_anterior_posterior("Inferior Longitudinal Fasciculus"),
plot_anterior_posterior("Uncinate Fasciculus"),
plot_anterior_posterior("Inferior Fronto-occipital Fasciculus"), nrow=3)

anterior_posterior_specific

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/fitted_val_tract_ends_ap_md.png", anterior_posterior_specific, width = 10, height = 15, units = "in")
```


```{r scrap2, include=FALSE, echo = FALSE, eval = FALSE}

## PCA of nodal developmental trajectories for each tract
#- is there an axis that best captures variance in developmental trajectories along tracts?

# fitted md
fitted_md

tract_profiles


fitted_md_anterior_thalamic <- fitted_md %>% filter(tract_hemi == "Left_Anterior_Thalamic_Radiation") %>% select(tract_node, age, fitted)
fitted_md_anterior_thalamic.long <- fitted_md_anterior_thalamic %>% pivot_wider(names_from="tract_node",values_from="fitted")

fitted_md_anterior_thalamic.long <- fitted_md_anterior_thalamic.long %>% select(-age) #ages by regions matrix for PCA
names(fitted_md_anterior_thalamic.long) <- gsub("Left_Anterior_Thalamic_Radiation_", "node_", names(fitted_md_anterior_thalamic.long))

smooths.pca <- prcomp(fitted_md_anterior_thalamic.long, retx = T) #PCA
 
summary(smooths.pca)$importance[2,1] # pc1 explains 99.5% of variance

 
smooths.pca$x # pc scores
library(factoextra)
fviz_eig(smooths.pca)
 
fviz_pca_var(smooths.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

pc1 <- as.data.frame(smooths.pca$rotation[,1])
pc1 <- cbind(rownames(pc1), pc1)
names(pc1) <- c("node","loading")


df <- data.frame(smooths.pca$rotation)
df$nodeID <- rownames(df)
ggplot(data=df, aes(x=nodeID, y=PC1)) +
  geom_point()

df_age <- data.frame(smooths.pca$x)
df_age$age <- c(8:22) 
ggplot(data=df_age, aes(x=age, y=PC1)) +
  geom_point()

# look at loadings / correlation between 



# nodewise  

zeroed_md_anterior_thalamic <- nodewise_smoothfits_md %>% filter(tractID == "Anterior Thalamic Radiation" & hemi == "Left") %>% select(tract_node, age, est)
zeroed_md_anterior_thalamic.long_age <- zeroed_md_anterior_thalamic %>% pivot_wider(names_from="tract_node",values_from="est")

zeroed_md_anterior_thalamic.long <- zeroed_md_anterior_thalamic.long_age %>% select(-age) #ages by regions matrix for PCA
names(zeroed_md_anterior_thalamic.long) <- gsub("Left_Anterior_Thalamic_Radiation_", "node_", names(zeroed_md_anterior_thalamic.long))

smooths.pca <- prcomp(zeroed_md_anterior_thalamic.long, scale=TRUE) #PCA
 
summary(smooths.pca)# pc1 explains 99.7% of variance



pc1 <- as.data.frame(smooths.pca$rotation[,1])
pc1 <- cbind(rownames(pc1), pc1)
names(pc1) <- c("node","loading")


df <- data.frame(smooths.pca$rotation)
df$nodeID <- rownames(df)

df %>% 
  ggplot( aes(x=nodeID, y=PC1)) +
  geom_point() + theme_classic()

df %>% filter(nodeID %in% c("node_1",  "node_2",  "node_3",  "node_4", "node_13", "node_24", "node_35", "node_46", "node_57", "node_68", "node_79", "node_90")) %>%
  ggplot( aes(x=nodeID, y=PC1)) +
  geom_point() + theme_classic()
which(df$PC1 < -0.125) # 1  2  3  4 13 24 35 46 57 68 79 90

# smooths.pca$x = age by PC's (scores)
df_age <- data.frame(smooths.pca$x)
df_age$age <- zeroed_md_anterior_thalamic.long_age$age
ggplot(data=df_age, aes(x=age, y=PC1)) +
  geom_point() + theme_classic()

 

cor.test(smooths.pca$x[,1], zeroed_md_anterior_thalamic.long$node_0)

lapply(zeroed_md_anterior_thalamic.long)

correlations <- apply(zeroed_md_anterior_thalamic.long, 2, function(column) cor(smooths.pca$x[,2], column))


PC1 <- as.data.frame(smooths.pca$rotation[,1]) #first principal component loadings
PC1$label <- row.names(PC1) 
colnames(PC1) <- c("PC1","label")
PC1$ranked <- rank(PC1$PC1)


 
#orrelations <- data.frame(correlations)
#correlations$nodeID <- rownames(correlations)
#correlations$node <- gsub("node_", "", correlations$nodeID)
#correlations$node <- as.numeric(correlations$node)
#correlations$ranked <- rank(correlations$correlations)


PC1$nodeID <- gsub("node_", "", PC1$label)
PC1$nodeID <- as.numeric(PC1$nodeID)
ggplot(data=PC1, aes(x=nodeID, y=ranked)) +
  geom_point() + theme_classic()
```
 

```{r function nodewise pca, include=FALSE, echo = FALSE, eval=FALSE}
# i think PC2 might be interesting?
# PC1 might be just capturing the fact that all the nodes are decreasing with age
 
pca_analysis <- function(tract, df) {
  print(tract)
  if (str_detect(tract,"Forceps")) {
  zeroed <- df %>% filter(tractID == tract) %>% select(nodeID, age, est)
  zeroed.long_age <- zeroed %>% pivot_wider(names_from="nodeID",values_from="est")
  zeroed.long <- zeroed.long_age %>% select(-age) #ages by regions matrix for PCA
    
  smooths.pca <- prcomp(zeroed.long, scale=TRUE) #PCA
   
  PC1 <- as.data.frame(smooths.pca$rotation[,1]) #first principal component loadings
  PC1$nodeID <- row.names(PC1)
  PC1$nodeID <- as.numeric(PC1$nodeID)
  colnames(PC1) <- c("PC1","nodeID")
  PC1$ranked <- rank(PC1$PC1)
  
  PC2 <- as.data.frame(smooths.pca$rotation[,2]) #second principal component loadings
  PC2$nodeID <- row.names(PC2)
  PC2$nodeID <- as.numeric(PC2$nodeID)
  colnames(PC2) <- c("PC2","nodeID")
  PC2$ranked <- rank(PC2$PC2)
  
  pca_output <- list(PC1, PC2, summary(smooths.pca))
  names(pca_output) <- c("PC1", "PC2", "summary")
  
  } else {
  lh_zeroed <- df %>% filter(tractID == tract & hemi == "Left") %>% select(nodeID, age, est)
  lh_zeroed.long_age <- lh_zeroed %>% pivot_wider(names_from="nodeID",values_from="est")
  lh_zeroed.long <- lh_zeroed.long_age %>% select(-age) #ages by regions matrix for PCA
    
  lh_smooths.pca <- prcomp(lh_zeroed.long, scale=TRUE) #PCA
   
  summary(lh_smooths.pca) # may want to save
   
  lh_PC1 <- as.data.frame(lh_smooths.pca$rotation[,1]) #first principal component loadings
  lh_PC1$nodeID <- row.names(lh_PC1)
  lh_PC1$nodeID <- as.numeric(lh_PC1$nodeID)
  colnames(lh_PC1) <- c("lh_PC1","nodeID")
  lh_PC1$ranked <- rank(lh_PC1$lh_PC1)
  
  lh_PC2 <- as.data.frame(lh_smooths.pca$rotation[,2]) #second principal component loadings
  lh_PC2$nodeID <- row.names(lh_PC2)
  lh_PC2$nodeID <- as.numeric(lh_PC2$nodeID)
  colnames(lh_PC2) <- c("lh_PC2","nodeID")
  lh_PC2$ranked <- rank(lh_PC2$lh_PC2)

  
  rh_zeroed <- df %>% filter(tractID == tract & hemi == "Right") %>% select(nodeID, age, est)
  rh_zeroed.long_age <- rh_zeroed %>% pivot_wider(names_from="nodeID",values_from="est")
  rh_zeroed.long <- rh_zeroed.long_age %>% select(-age) #ages by regions matrix for PCA
    
  rh_smooths.pca <- prcomp(rh_zeroed.long, scale=TRUE) #PCA
   
  summary(rh_smooths.pca) # may want to save
   
  rh_PC1 <- as.data.frame(rh_smooths.pca$rotation[,1]) #first principal component loadings
  rh_PC1$nodeID <- row.names(rh_PC1)
  rh_PC1$nodeID <- as.numeric(rh_PC1$nodeID)
  colnames(rh_PC1) <- c("rh_PC1","nodeID")
  rh_PC1$ranked <- rank(rh_PC1$rh_PC1)
  
  rh_PC2 <- as.data.frame(rh_smooths.pca$rotation[,2]) #second principal component loadings
  rh_PC2$nodeID <- row.names(rh_PC2)
  rh_PC2$nodeID <- as.numeric(rh_PC2$nodeID)
  colnames(rh_PC2) <- c("rh_PC2","nodeID")
  rh_PC2$ranked <- rank(rh_PC2$rh_PC2)
  
  pca_output <- list(lh_PC1, lh_PC2, rh_PC1, rh_PC2, summary(lh_smooths.pca),  summary(rh_smooths.pca))
  names(pca_output) <- c("lh_PC1", "lh_PC2", "rh_PC1", "rh_PC2", "summary_lh", "summary_rh")
  
  }
  return(pca_output)
  
}


pca_md_nodewise <- lapply(unique(nodewise_smoothfits_md$tractID), pca_analysis, df = nodewise_smoothfits_md)
names(pca_md_nodewise) <- unique(nodewise_smoothfits_md$tractID)
 
# PC1
variance_explained_PC1 <- matrix(NA, nrow = length(pca_md_nodewise), ncol = 2)

for (i in seq_along(pca_md_nodewise)) {
  tract <- names(pca_md_nodewise)[i]
  if(str_detect(tract, "Forceps")) {
  variance_explained_PC1[i, 1] <- pca_md_nodewise[[tract]]$summary$importance[2, 1]
  variance_explained_PC1[i, 2] <- NA
  } else {
  
  variance_explained_PC1[i, 1] <- pca_md_nodewise[[tract]]$summary_lh$importance[2, 1]
  variance_explained_PC1[i, 2] <- pca_md_nodewise[[tract]]$summary_rh$importance[2, 1]
  }
  
}
rownames(variance_explained_PC1) <- names(pca_md_nodewise)
variance_explained_PC1 <- data.frame(variance_explained_PC1) %>% setNames(c("var_expl_lh", "var_expl_rh"))


# PC2
variance_explained_PC2 <- matrix(NA, nrow = length(pca_md_nodewise), ncol = 2)

for (i in seq_along(pca_md_nodewise)) {
  tract <- names(pca_md_nodewise)[i]
  if(str_detect(tract, "Forceps")) {
  
  variance_explained_PC2[i, 1] <- pca_md_nodewise[[tract]]$summary$importance[2, 2]
  variance_explained_PC2[i, 2] <- NA
  } else {
  tract <- names(pca_md_nodewise)[i]
  variance_explained_PC2[i, 1] <- pca_md_nodewise[[tract]]$summary_lh$importance[2, 2]
  variance_explained_PC2[i, 2] <- pca_md_nodewise[[tract]]$summary_rh$importance[2, 2]
  }
}
rownames(variance_explained_PC2) <- names(pca_md_nodewise)
variance_explained_PC2 <- data.frame(variance_explained_PC2) %>% setNames(c("var_expl_lh", "var_expl_rh"))
 
variance_explained_PC1
variance_explained_PC2
```


```{r visualize pcs, fig.height = 20, fig.width = 20, eval=FALSE, include=FALSE, echo=FALSE}
plot_pcs <- function(tract, df, pc_number){
  if (str_detect(tract, "Forceps")) {
    PC <- df[[tract]][[paste0("", pc_number)]] 
    PC_plot <- ggplot(data=PC, aes(x=nodeID, y=ranked, group=nodeID)) +
    geom_point() + 
    #paletteer::scale_colour_paletteer_c("grDevices::RdYlBu", direction=-1) +
    #paletteer::scale_fill_paletteer_c("grDevices::RdYlBu", direction=-1) +
    theme_classic() + labs(title=paste("Left", tract, pc_number)) 
    return(plot_grid(PC_plot))
  } else {
    lh_PC <- df[[tract]][[paste0("lh_", pc_number)]] 
    lh_PC_plot <- ggplot(data=lh_PC, aes(x=nodeID, y=ranked)) +
    geom_point() + 
    #paletteer::scale_colour_paletteer_c("grDevices::RdYlBu", direction=-1) +
    #paletteer::scale_fill_paletteer_c("grDevices::RdYlBu", direction=-1) +
    theme_classic() + labs(title=paste("Left", tract, pc_number))
    
    rh_PC <- df[[tract]][[paste0("rh_", pc_number)]] 
    rh_PC_plot <- ggplot(data=rh_PC, aes(x=nodeID, y=ranked)) +
    geom_point() + 
     # paletteer::scale_colour_paletteer_c("grDevices::RdYlBu", direction=-1) +
    #paletteer::scale_fill_paletteer_c("grDevices::RdYlBu", direction=-1) +
      theme_classic() + labs(title=paste("Right", tract, pc_number))
    return(plot_grid(lh_PC_plot, rh_PC_plot, nrow=2))
  }
  
  
}


pc1_plots_md <- lapply(unique(nodewise_smoothfits_md$tractID), plot_pcs, df = pca_md_nodewise, pc_number = "PC1")
names(pc1_plots_md) <- unique(nodewise_smoothfits_md$tractID)

pc2_plots_md <- lapply(unique(nodewise_smoothfits_md$tractID), plot_pcs, df = pca_md_nodewise, pc_number = "PC2")
names(pc2_plots_md) <- unique(nodewise_smoothfits_md$tractID)

 



arrange_by_orientation_devtraj(pc2_plots_md, "MD PC2")
 

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/pc1_nodewise_devtraj_md.png", arrange_by_orientation_devtraj(pc1_plots_md, "MD PC1"), width = 20, height = 20, units = "in")

#ggsave("/Users/audluo/PennLINC/luowm_local/output/tract_profiles_testing/HCPD/pc2_nodewise_devtraj_md.png", arrange_by_orientation_devtraj(pc2_plots_md, "MD PC2"), width = 20, height = 20, units = "in")

```


Other things to look at:
- double check posterior-anterior and inferior-superior gradients
- age of first developmental slowing
- compute Euclidean distance to cortex; look at development vs. distance to cortex (instead of relative node position)
- PCA on nodal age effect data? 
 